### 2016-04-08 paper/paper.tex
Add ieeetran template with authors!

% Copied from:
% http://mirror.its.dal.ca/ctan/macros/latex/contrib/IEEEtran/bare_conf.tex
\documentclass[conference]{IEEEtran}
% Some Computer Society conferences also require the compsoc mode option,
% but others use the standard conference format.
%
% If IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it like:
% \documentclass[conference]{../sty/IEEEtran}

% Some very useful LaTeX packages include:
% (uncomment the ones you want to load)

% *** MISC UTILITY PACKAGES ***
%
%\usepackage{ifpdf}
% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
% compilation based on whether the output is pdf or dvi.
% usage:
% \ifpdf
%   % pdf code
% \else
%   % dvi code
% \fi
% The latest version of ifpdf.sty can be obtained from:
% http://www.ctan.org/pkg/ifpdf
% Also, note that IEEEtran.cls V1.7 and later provides a builtin
% \ifCLASSINFOpdf conditional that works the same way.
% When switching from latex to pdflatex and vice-versa, the compiler may
% have to be run twice to clear warning/error messages.

% *** CITATION PACKAGES ***
%
%\usepackage{cite}
% cite.sty was written by Donald Arseneau
% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
% \cite{} output to follow that of the IEEE. Loading the cite package will
% result in citation numbers being automatically sorted and properly
% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
% \cite will automatically add leading space, if needed. Use cite.sty's
% noadjust option (cite.sty V3.8 and later) if you want to turn this off
% such as if a citation ever needs to be enclosed in parenthesis.
% cite.sty is already installed on most LaTeX systems. Be sure and use
% version 5.0 (2009-03-20) and later if using hyperref.sty.
% The latest version can be obtained at:
% http://www.ctan.org/pkg/cite
% The documentation is contained in the cite.sty file itself.

% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf
  % \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../pdf/}{../jpeg/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  % \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.eps}
\fi
% graphicx was written by David Carlisle and Sebastian Rahtz. It is
% required if you want graphics, photos, etc. graphicx.sty is already
% installed on most LaTeX systems. The latest version and documentation
% can be obtained at:
% http://www.ctan.org/pkg/graphicx
% Another good source of documentation is "Using Imported Graphics in
% LaTeX2e" by Keith Reckdahl which can be found at:
% http://www.ctan.org/pkg/epslatex
%
% latex, and pdflatex in dvi mode, support graphics in encapsulated
% postscript (.eps) format. pdflatex in pdf mode supports graphics
% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
% not a bitmapped formats (.jpeg, .png). The IEEE frowns on bitmapped formats
% which can result in "jaggedy"/blurry rendering of lines and letters as
% well as large increases in file sizes.
%
% You can find documentation about the pdfTeX application at:
% http://www.tug.org/applications/pdftex

% *** MATH PACKAGES ***
%
%\usepackage{amsmath}
% A popular package from the American Mathematical Society that provides
% many useful and powerful commands for dealing with mathematics.
%
% Note that the amsmath package sets \interdisplaylinepenalty to 10000
% thus preventing page breaks from occurring within multiline equations. Use:
%\interdisplaylinepenalty=2500
% after loading amsmath to restore such page breaks as IEEEtran.cls normally
% does. amsmath.sty is already installed on most LaTeX systems. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/pkg/amsmath

% *** SPECIALIZED LIST PACKAGES ***
%
%\usepackage{algorithmic}
% algorithmic.sty was written by Peter Williams and Rogerio Brito.
% This package provides an algorithmic environment fo describing algorithms.
% You can use the algorithmic environment in-text or within a figure
% environment to provide for a floating algorithm. Do NOT use the algorithm
% floating environment provided by algorithm.sty (by the same authors) or
% algorithm2e.sty (by Christophe Fiorio) as the IEEE does not use dedicated
% algorithm float types and packages that provide these will not provide
% correct IEEE style captions. The latest version and documentation of
% algorithmic.sty can be obtained at:
% http://www.ctan.org/pkg/algorithms
% Also of interest may be the (relatively newer and more customizable)
% algorithmicx.sty package by Szasz Janos:
% http://www.ctan.org/pkg/algorithmicx

% *** ALIGNMENT PACKAGES ***
%
%\usepackage{array}
% Frank Mittelbach's and David Carlisle's array.sty patches and improves
% the standard LaTeX2e array and tabular environments to provide better
% appearance and additional user controls. As the default LaTeX2e table
% generation code is lacking to the point of almost being broken with
% respect to the quality of the end results, all users are strongly
% advised to use an enhanced (at the very least that provided by array.sty)
% set of table tools. array.sty is already installed on most systems. The
% latest version and documentation can be obtained at:
% http://www.ctan.org/pkg/array

% IEEEtran contains the IEEEeqnarray family of commands that can be used to
% generate multiline equations as well as matrices, tables, etc., of high
% quality.

% *** SUBFIGURE PACKAGES ***
%\ifCLASSOPTIONcompsoc
%  \usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
%\else
%  \usepackage[caption=false,font=footnotesize]{subfig}
%\fi
% subfig.sty, written by Steven Douglas Cochran, is the modern replacement
% for subfigure.sty, the latter of which is no longer maintained and is
% incompatible with some LaTeX packages including fixltx2e. However,
% subfig.sty requires and automatically loads Axel Sommerfeldt's caption.sty
% which will override IEEEtran.cls' handling of captions and this will result
% in non-IEEE style figure/table captions. To prevent this problem, be sure
% and invoke subfig.sty's "caption=false" package option (available since
% subfig.sty version 1.3, 2005/06/28) as this is will preserve IEEEtran.cls
% handling of captions.
% Note that the Computer Society format requires a larger sans serif font
% than the serif footnote size font used in traditional IEEE formatting
% and thus the need to invoke different subfig.sty package options depending
% on whether compsoc mode has been enabled.
%
% The latest version and documentation of subfig.sty can be obtained at:
% http://www.ctan.org/pkg/subfig

% *** FLOAT PACKAGES ***
%
%\usepackage{fixltx2e}
% fixltx2e, the successor to the earlier fix2col.sty, was written by
% Frank Mittelbach and David Carlisle. This package corrects a few problems
% in the LaTeX2e kernel, the most notable of which is that in current
% LaTeX2e releases, the ordering of single and double column floats is not
% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
% single column figure to be placed prior to an earlier double column
% figure.
% Be aware that LaTeX2e kernels dated 2015 and later have fixltx2e.sty's
% corrections already built into the system in which case a warning will
% be issued if an attempt is made to load fixltx2e.sty as it is no longer
% needed.
% The latest version and documentation can be found at:
% http://www.ctan.org/pkg/fixltx2e

%\usepackage{stfloats}
% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
% the ability to do double column floats at the bottom of the page as well
% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
% LaTeX2e). It also provides a command:
%\fnbelowfloat
% to enable the placement of footnotes below bottom floats (the standard
% LaTeX2e kernel puts them above bottom floats). This is an invasive package
% which rewrites many portions of the LaTeX2e float routines. It may not work
% with other packages that modify the LaTeX2e float routines. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/pkg/stfloats
% Do not use the stfloats baselinefloat ability as the IEEE does not allow
% \baselineskip to stretch. Authors submitting work to the IEEE should note
% that the IEEE rarely uses double column equations and that authors should try
% to avoid such use. Do not be tempted to use the cuted.sty or midfloat.sty
% packages (also by Sigitas Tolusis) as the IEEE does not format its papers in
% such ways.
% Do not attempt to use stfloats with fixltx2e as they are incompatible.
% Instead, use Morten Hogholm'a dblfloatfix which combines the features
% of both fixltx2e and stfloats:
%
% \usepackage{dblfloatfix}
% The latest version can be found at:
% http://www.ctan.org/pkg/dblfloatfix

% *** PDF, URL AND HYPERLINK PACKAGES ***

\usepackage{url}
% url.sty was written by Donald Arseneau. It provides better support for
% handling and breaking URLs. url.sty is already installed on most LaTeX
% systems. The latest version and documentation can be obtained at:
% http://www.ctan.org/pkg/url
% Basically, \url{my_url_here}.

% *** Do not adjust lengths that control margins, column widths, etc. ***
% *** Do not use packages that alter fonts (such as pslatex).         ***
% There should be no need to do such things with IEEEtran.cls V1.6 and later.
% (Unless specifically asked to do so by the journal or conference you plan
% to submit to, of course. )

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\begin{document}
%
% paper title
% Titles are generally capitalized except for words such as a, an, and, as,
% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
% not capitalized unless they are the first or last word of the title.
% Linebreaks \\ can be used within to get better formatting as desired.
% Do not put math or special symbols in the title.
\title{Visualizing Project Evolution Through Abstract Syntax Tree Analysis}

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
%\author{\IEEEauthorblockN{Michael D. Feist \and Eddie Antonio Santos \and
%  Ian Watts \and Abram Hindle}
%\IEEEauthorblockA{Department of Computing Science\\
%University of Alberta\\
%Edmonton, Canada\\
%\{mdfeist,easantos,watts1,hindle1\}@ualberta.ca}}

% conference papers do not typically use \thanks and this command
% is locked out in conference mode. If really needed, such as for
% the acknowledgment of grants, issue a \IEEEoverridecommandlockouts
% after \documentclass

% for over three affiliations, or if they all won't fit within the width
% of the page, use this alternative format:
%
\author{\IEEEauthorblockN{Michael D. Feist,
Eddie Antonio Santos,
Ian Watts,
Abram Hindle}
\IEEEauthorblockA{Department of Computing Science\\
University of Alberta\\
Edmonton, Canada\\
\{mdfeist,easantos,watts1,hindle1\}@ualberta.ca}}

% use for special paper notices
%\IEEEspecialpapernotice{(Invited Paper)}

% make the title area
\maketitle

% As a general rule, do not put math, special symbols or citations
% in the abstract
\begin{abstract}
Analysing the changes in software repositories over time is difficult since changes come from different developers and affect different parts of a repository.
Individual commits provide the changed files, code diffs and a message.
However, when many commits are viewed as a group, the details are lost.
Currently there is a lack of tools for tracking the changes in a repository beyond simply looking at the number of commits of lines changed.
The commit information can be misleading since lines of code gives no indication of what was actually being worked on without careful examination of the changed code.
Providing a more comprehensive view of this information could provide deeper insights into how software projects evolve since changes to design and features are not clearly visible.

We present a novel tool for visualizing Java projects.
It provides more detailed information regarding all of the changes along the master branch of a project.
It visualizes changes in type usage over time for projects and developers and derives detailed statistics using the differences between Abstract Syntax Trees (ASTs).
We are tracking additions and deletions of declarations and invocations for each type.
The AST diff between commits is a reentrant algorithm that can be incrementally calculated as new commits are made.

Using this tool we show the changes in type coverage and other usage statistics over time in several popular repositories.
Type coverage information from the AST is compared to file coverage to check for correlation or if there is unique information provided by type information.
Potential uses for such a tool for managers and developers is discussed.
\end{abstract}

% no keywords

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle

\section{Introduction}
% no \IEEEPARstart
This demo file is intended to serve as a ``starter file''
for IEEE conference papers produced under \LaTeX\ using
IEEEtran.cls version 1.8b and later.
% You must have at least 2 lines in the paragraph with the drop letter
% (should never be an issue)
I wish you the best of success.

\hfill mds

\hfill August 26, 2015

\subsection{Subsection Heading Here}
Subsection text here.

\subsubsection{Subsubsection Heading Here}
Subsubsection text here.

% An example of a floating figure using the graphicx package.
% Note that \label must occur AFTER (or within) \caption.
% For figures, \caption should occur after the \includegraphics.
% Note that IEEEtran v1.7 and later has special internal code that
% is designed to preserve the operation of \label within \caption
% even when the captionsoff option is in effect. However, because
% of issues like this, it may be the safest practice to put all your
% \label just after \caption rather than within \caption{}.
%
% Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
% option should be used if it is desired that the figures are to be
% displayed while in draft mode.
%
%\begin{figure}[!t]
%\centering
%\includegraphics[width=2.5in]{myfigure}
% where an .eps filename suffix will be assumed under latex,
% and a .pdf suffix will be assumed for pdflatex; or what has been declared
% via \DeclareGraphicsExtensions.
%\caption{Simulation results for the network.}
%\label{fig_sim}
%\end{figure}

% Note that the IEEE typically puts floats only at the top, even when this
% results in a large percentage of a column being occupied by floats.

% An example of a double column floating figure using two subfigures.
% (The subfig.sty package must be loaded for this to work.)
% The subfigure \label commands are set within each subfloat command,
% and the \label for the overall figure must come after \caption.
% \hfil is used as a separator to get equal spacing.
% Watch out that the combined width of all the subfigures on a
% line do not exceed the text width or a line break will occur.
%
%\begin{figure*}[!t]
%\centering
%\subfloat[Case I]{\includegraphics[width=2.5in]{box}%
%\label{fig_first_case}}
%\hfil
%\subfloat[Case II]{\includegraphics[width=2.5in]{box}%
%\label{fig_second_case}}
%\caption{Simulation results for the network.}
%\label{fig_sim}
%\end{figure*}
%
% Note that often IEEE papers with subfigures do not employ subfigure
% captions (using the optional argument to \subfloat[]), but instead will
% reference/describe all of them (a), (b), etc., within the main caption.
% Be aware that for subfig.sty to generate the (a), (b), etc., subfigure
% labels, the optional argument to \subfloat must be present. If a
% subcaption is not desired, just leave its contents blank,
% e.g., \subfloat[].

% An example of a floating table. Note that, for IEEE style tables, the
% \caption command should come BEFORE the table and, given that table
% captions serve much like titles, are usually capitalized except for words
% such as a, an, and, as, at, but, by, for, in, nor, of, on, or, the, to
% and up, which are usually not capitalized unless they are the first or
% last word of the caption. Table text will default to \footnotesize as
% the IEEE normally uses this smaller font for tables.
% The \label must come after \caption as always.
%
%\begin{table}[!t]
%% increase table row spacing, adjust to taste
%\renewcommand{\arraystretch}{1.3}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
%\caption{An Example of a Table}
%\label{table_example}
%\centering
%% Some packages, such as MDW tools, offer better commands for making tables
%% than the plain LaTeX2e tabular which is used here.
%\begin{tabular}{|c||c|}
%\hline
%One & Two\\
%\hline
%Three & Four\\
%\hline
%\end{tabular}
%\end{table}

% Note that the IEEE does not put floats in the very first column
% - or typically anywhere on the first page for that matter. Also,
% in-text middle ("here") positioning is typically not used, but it
% is allowed and encouraged for Computer Society conferences (but
% not Computer Society journals). Most IEEE journals/conferences use
% top floats exclusively.
% Note that, LaTeX2e, unlike IEEE journals/conferences, places
% footnotes above bottom floats. This can be corrected via the
% \fnbelowfloat command of the stfloats package.

\section{Conclusion}
The conclusion goes here.

% conference papers do not normally have an appendix

% use section* for acknowledgment
\section*{Acknowledgment}

The authors would like to thank...

% trigger a \newpage just before the given reference
% number - used to balance the columns on the last page
% adjust value as needed - may need to be readjusted if
% the document is modified later
%\IEEEtriggeratref{8}
% The "triggered" command can be changed if desired:
%\IEEEtriggercmd{\enlargethispage{-5in}}

% references section

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://mirror.ctan.org/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
%\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
%\bibliography{IEEEabrv,../bib/paper}
%
% <OR> manually copy in the resultant .bbl file
% set second argument of \begin to the number of references
% (used to reserve space for the reference number labels box)
\begin{thebibliography}{1}

\bibitem{IEEEhowto:kopka}
H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, 3rd~ed.\hskip 1em plus
  0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.

\end{thebibliography}

% that's all folks
\end{document}

### 2016-04-08 paper/paper.tex
Re-add bibliography.

@@
 % \documentclass[conference]{../sty/IEEEtran}
+\usepackage{changepage}
 
@@
 \section{Introduction}
-% no \IEEEPARstart
-This demo file is intended to serve as a ``starter file''
-for IEEE conference papers produced under \LaTeX\ using
-IEEEtran.cls version 1.8b and later.
-% You must have at least 2 lines in the paragraph with the drop letter
-% (should never be an issue)
-I wish you the best of success.
 
-\hfill mds
+The history and evolution of software repositories are often tracked by version control systems.
+The information is generally kept in the form of commits, which show the differences between each subsequent version of the software.
+These services become vital for large projects since individual changes would be impossible to track manually.
+Large projects have thousands of individual changes made by many different developers.
+With more changes to look through, it becomes more difficult to see how a project is evolving over time since there is so much information to look through.
+Most hosting services for git repositories provide some basic tools for analysing projects over time, but too much information is lost due to the lack of fine grained details.
+The information regarding contributions from developers are also misleading since the number of commits and lines of code do not necessarily reflect meaningful changes to a repository.
 
-\hfill August 26, 2015
+In Java, a type can be a primitive type or an object reference.
+Primitive types are those that are built into the language such as an int or char.
+Object references point to instances of classes in memory which are defined in the project or through an included library.
+The types in a project represent the functionality available for a developer to work with.
+A large or complex project may have many different types, whereas a small project might only use a few.
+Different parts of a program are going to use different types and libraries.
+When writing or contributing to a project, the author’s choice of which part of the program to edit will in turn affect what types they end up working with.
 
-\subsection{Subsection Heading Here}
-Subsection text here.
+Having advanced tools for analysing repositories would benefit many different parts of the software development process.
+Software management requires a high level understanding of the state of a repository and the changes being made which affect the functionality of the software.
+Tracking developer contributions in a less misleading way is also important for management since important contributions may be small and focused.
+Line difference only provide meaningful info when they are analysed by someone with familiarity with the source code being changed.
+Providing tools which bring more meaning to line differences would allow people less familiar with the code to understand changes being made.
+When large changes are made that cover many lines of code, it would also be useful to see how the change affects the software rather than a several thousand line diff.
+Seeing how a project has changed over time is also vital for management since evaluating the state of completion of software can be very difficult.
 
-\subsubsection{Subsubsection Heading Here}
-Subsubsection text here.
+Type information is useful for seeing what libraries and functions are being used.
+When a part of code becomes deprecated, continuing to use it may lead to design or  security flaws.
+Viewing the type information of a change can show when deprecated classes are added or removed from a project as well as how many remain.
 
-% An example of a floating figure using the graphicx package.
-% Note that \label must occur AFTER (or within) \caption.
-% For figures, \caption should occur after the \includegraphics.
-% Note that IEEEtran v1.7 and later has special internal code that
-% is designed to preserve the operation of \label within \caption
-% even when the captionsoff option is in effect. However, because
-% of issues like this, it may be the safest practice to put all your
-% \label just after \caption rather than within \caption{}.
-%
-% Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
-% option should be used if it is desired that the figures are to be
-% displayed while in draft mode.
-%
-%\begin{figure}[!t]
-%\centering
-%\includegraphics[width=2.5in]{myfigure}
-% where an .eps filename suffix will be assumed under latex,
-% and a .pdf suffix will be assumed for pdflatex; or what has been declared
-% via \DeclareGraphicsExtensions.
-%\caption{Simulation results for the network.}
-%\label{fig_sim}
-%\end{figure}
+\section{Related Work}
 
-% Note that the IEEE typically puts floats only at the top, even when this
-% results in a large percentage of a column being occupied by floats.
+Robert Dyer \textit{et al.} \cite{Dyer:2014:MBA:2568225.2568295} mined AST nodes to study the use of new Java language features over time. (More in challenge paper)
 
-% An example of a double column floating figure using two subfigures.
-% (The subfig.sty package must be loaded for this to work.)
-% The subfigure \label commands are set within each subfloat command,
-% and the \label for the overall figure must come after \caption.
-% \hfil is used as a separator to get equal spacing.
-% Watch out that the combined width of all the subfigures on a
-% line do not exceed the text width or a line break will occur.
-%
-%\begin{figure*}[!t]
-%\centering
-%\subfloat[Case I]{\includegraphics[width=2.5in]{box}%
-%\label{fig_first_case}}
-%\hfil
-%\subfloat[Case II]{\includegraphics[width=2.5in]{box}%
-%\label{fig_second_case}}
-%\caption{Simulation results for the network.}
-%\label{fig_sim}
-%\end{figure*}
-%
-% Note that often IEEE papers with subfigures do not employ subfigure
-% captions (using the optional argument to \subfloat[]), but instead will
-% reference/describe all of them (a), (b), etc., within the main caption.
-% Be aware that for subfig.sty to generate the (a), (b), etc., subfigure
-% labels, the optional argument to \subfloat must be present. If a
-% subcaption is not desired, just leave its contents blank,
-% e.g., \subfloat[].
+Grechanik \textit{et al.} \cite{Grechanik:2010:EIL:1852786.1852801} examined the structure of Java programs mined from 2080 programs.
+This paper examined the breakdown of syntactic structures in open source repositories, however it does not consider per author statistics.
 
-% An example of a floating table. Note that, for IEEE style tables, the
-% \caption command should come BEFORE the table and, given that table
-% captions serve much like titles, are usually capitalized except for words
-% such as a, an, and, as, at, but, by, for, in, nor, of, on, or, the, to
-% and up, which are usually not capitalized unless they are the first or
-% last word of the caption. Table text will default to \footnotesize as
-% the IEEE normally uses this smaller font for tables.
-% The \label must come after \caption as always.
-%
-%\begin{table}[!t]
-%% increase table row spacing, adjust to taste
-%\renewcommand{\arraystretch}{1.3}
-% if using array.sty, it might be a good idea to tweak the value of
-% \extrarowheight as needed to properly center the text within the cells
-%\caption{An Example of a Table}
-%\label{table_example}
-%\centering
-%% Some packages, such as MDW tools, offer better commands for making tables
-%% than the plain LaTeX2e tabular which is used here.
-%\begin{tabular}{|c||c|}
-%\hline
-%One & Two\\
-%\hline
-%Three & Four\\
-%\hline
-%\end{tabular}
-%\end{table}
+Meyerovich and Rabkin \cite{Meyerovich:2013:EAP:2509136.2509515} surveyed developers and examined repositories to learn about language adoption and usage.
+Developers feel that certain language features are more important than others.
 
-% Note that the IEEE does not put floats in the very first column
-% - or typically anywhere on the first page for that matter. Also,
-% in-text middle ("here") positioning is typically not used, but it
-% is allowed and encouraged for Computer Society conferences (but
-% not Computer Society journals). Most IEEE journals/conferences use
-% top floats exclusively.
-% Note that, LaTeX2e, unlike IEEE journals/conferences, places
-% footnotes above bottom floats. This can be corrected via the
-% \fnbelowfloat command of the stfloats package.
+Parnin \textit{et al.} \cite{Parnin:2011:JGA:1985441.1985446} mined repositories to see how Java generics have been used in open source projects and found that generic usages were often introduced by a single developer in a project.
+Generics usage was also fairly narrow, with the primary use being collecting and traversing lists of objects.
+
+Patrick Wagstrom \textit{et al.} \cite{Patrick:Wagstrom:2012} explored the roles that developers take in a networked, social development environments like GitHub.
+These role were defined as their level of contribution as well as the types of issues they dealt with.
+Developers will take on multiple roles in a project and will sometimes fulfill the same role across different projects
+
+L\"{a}mmel \textit{et al.} \cite{Lammel:2011:LAA:1982185.1982471} used ASTs to examine the usage of APIs in Java projects.
+They find what APIs are popular as well as if they are used in a framework like manner.
+
+\section{Methodology}
+
+An Abstract Syntax Tree (AST) is a tree structure which represents a piece of code by breaking its syntactic constructs into a tree structure.
+Each node in the tree represents a syntactically valid chunk of code which can be contained in a parent node or broken up into child nodes.
+This captures the structure of the code in an abstract manner.
+When a change is made to a file, the difference will also appear in the AST.
+This means that the AST of revisions in code repositories can be compared to see what structures a developer is touching.
+
+The ASTs were generated using Spoon \cite{pawlak:hal-01169705}.
+Spoon is a tool for transforming and analysing Java source code.
+It breaks code up into a meta-model including AST.
+The model consists of three parts: structural elements, code elements and references to program elements.
+The Spoon model is convenient because it provides the structure for performing an AST diff while retaining the code elements for analysis afterwards.
+Spoon also preserves all type and library information, which is what we were interested in.
+
+GumTree \cite{falleri:hal-01054552} is an algorithm which is used to compute the difference between two ASTs generated by Spoon.
+For new files and deleted files we did not use GumTree; we simply ran Spoon on the files and treated everything as an addition or deletion respectively.
+
+Once we have the generated ASTs we were able to run through the trees and count the declarations and invocations for a given type.
+For invocations we also recorded the method call.
+These stats were then marked as additions or deletions.
+\\
+\begin{adjustwidth}{2.5em}{0pt}
+\#DECLARE $\vert$ INSERT $\vert$ java.net.URI $\vert$ 2 \\
+\#DECLARE $\vert$ DELETE $\vert$ java.lang.String $\vert$ 2 \\
+\end{adjustwidth}
+
+In order to gather all the changes for a given repository over time we generated AST for each commit and compared the difference.
+We did this by running ‘git log’ over the git repository and listing all the commit ids in increasing time.
+Each commit and its parent were then compared.
+If the commit had no parent then every file in the commit was treated as an addition.
+Since we compared each commit to its parent we were able to account for branching that happened in the master branch of the repository.
+Also we ignored commits that were merges as all changes in the merge should already be accounted for in previous commits.
+
+\section{Case Study}
+
+In the following sections, we present {N} case studies demonstrating the extra insight that TypeV makes available.
+
+\subsection{antlr4}
+
+We chose Antlr4 as it is a substantial Java project, with a commit history spanning several years. We can also see distinct changes throughout its history.
+
+For example, in December of 2011 we can see that ParserATNSimulator has a large amount of additions and deletions in roughly equal proportion (Figure~\ref{fig:parser1}.
+This suggests that ParserATNSimulator is not being removed from the project but instead there is some refactoring being done.
+If we take a closer look at the commit messages we can see that they mention refactoring and reorganization of code (Figure~\ref{fig:parser2}).
+Zooming in on that date,  shows that during that month they were finishing work on a new version of ParserATNSimulator.
+
+If we tried to do the same analyses with the GitHub tools it would be significantly more difficult to achieve the same conclusion.
+First if we only look at line changes this change to ParserATNSimulator is overwhelmed by other changes and hardly shows up on the GitHub tools.
+Then if we did notice an increase in activity we would not know what types were being worked on.
+Finally we could not quickly look at the commit messages for that specific time and type.
+
+At certain times we can can see major increases across all types.
+This shows us when major changes to the code were being done versus code maintenance.
+Usually this happens when a large amount of code is being pulled in from another location.
+
+\subsection{Apache Bookkeeper}
+
+For deprecation analysis, we chose Apache Bookkeeper.
+As an Apache Foundation project it is a well-curated example of an open source software project.
+It also demonstrates an instance of deprecated methods.
+
+Apache Bookkeeper deprecated the method AbstractConfiguration.getLedgerManagerType() in release 4.2.0~\footnote{\url{https://bookkeeper.apache.org/docs/r4.2.0/apidocs/deprecated-list.html}} and replaced it with AbstractConfiguration.getLedgerManagerFactoryClass().
+If we look at getLedgerManagerType() we can see at the time of the start of the deprecation all instances were removed and getLedgerManagerFactoryClass() was added.
+After the deprecation, we see no more changes to getLedgerManagerType() and only some additions of getLedgerManagerFactoryClass().
+
+\section{Evaluation}
+
+\section{Discussion}
 
@@
 % http://www.michaelshell.org/tex/ieeetran/bibtex/
-%\bibliographystyle{IEEEtran}
+\nocite{*}
+\bibliographystyle{IEEEtran}
 % argument is your BibTeX string definitions and bibliography database(s)
-%\bibliography{IEEEabrv,../bib/paper}
-%
-% <OR> manually copy in the resultant .bbl file
-% set second argument of \begin to the number of references
-% (used to reserve space for the reference number labels box)
-\begin{thebibliography}{1}
-
-\bibitem{IEEEhowto:kopka}
-H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, 3rd~ed.\hskip 1em plus
-  0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.
-
-\end{thebibliography}
+\bibliography{bibliography,autogen}
 

### 2016-04-08 server.py
change how outputs are loaded...

@@ -1,6 +1,11 @@
-import sys
-import os
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+
+import glob
 import json
+import os
+import sys
+
 from flask import Flask, request, redirect, send_from_directory, render_template
 
 projects = []
@@ -29,7 +34,7 @@ class Project:
 
     def getAuthors(self):
         authors = set()
-        
+
         for commit in self._commits:
             author = commit.getAuthor()
             authors.add(author)
@@ -113,7 +118,7 @@ class Project:
                         # Skip java.lang.Object#Object()
                         if (t == "java.lang.Object#Object()"):
                             continue
-                            
+
                         types.add(t)
                         json_data["dates"].append(self.getEdit(commit, "REMOVE", t))
 
@@ -130,7 +135,7 @@ class Project:
             output += commit.toStr("\t")
 
         return output
-    
+
     def __repr__(self):
         return self.__str__()
 
@@ -201,14 +206,14 @@ class Commit:
 
         output += tab + "Stats: \n"
         for f in self._files:
-            output += f.toStr(tab + "\t") + "\n" 
+            output += f.toStr(tab + "\t") + "\n"
         output += tab + "######################################################\n"
 
         return output
 
     def __str__(self):
         return self.toStr("")
-    
+
     def __repr__(self):
         return self.__str__()
 
@@ -248,7 +253,7 @@ class File:
 
     def __str__(self):
         return self.toStr("")
-    
+
     def __repr__(self):
         return self.__str__()
 
@@ -292,7 +297,7 @@ class Histogram:
 
     def __str__(self):
         return self.toStr("")
-    
+
     def __repr__(self):
         return self.__str__()
 
@@ -399,21 +404,11 @@ def mergeProjectCSV(csv):
     return os.linesep.join([s for s in csv.splitlines() if s])
 
 
-files = ["out1.out", 
-        "out2.out",
-        "out3.out",
-        "out4.out",
-        "out5.out",
-        "out6.out",
-        "out7.out"]
-
-files = ["test.out"]
-files = ["antlr4.out"]
-#files = ["antlr4.out", "301.out", "Paper_Repos.out"]
-
 
 # set the project root directory as the static folder, you can set others.
 path = os.path.dirname(os.path.realpath(__file__))
+# Files = list of ast-output/*.out
+files = glob.glob(os.path.join(path, 'ast-output', '*.out'))
[…]

### 2016-04-09 paper/paper.tex
Add some figures to latex.
Update tex file.

@@
 % \documentclass[conference]{../sty/IEEEtran}
+
 \usepackage{changepage}
@@
 \ifCLASSINFOpdf
-  % \usepackage[pdftex]{graphicx}
+  \usepackage[pdftex]{graphicx}
+  \usepackage{epstopdf}
   % declare the path(s) where your graphic files are
-  % \graphicspath{{../pdf/}{../jpeg/}}
+ \graphicspath{{diagrams/}}
   % and their extensions so you won't have to specify these with
   % every instance of \includegraphics
-  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
+  \DeclareGraphicsExtensions{.pdf,.jpeg,.png,.eps}
 \else
@@
   % driver is specified.
-  % \usepackage[dvips]{graphicx}
+  \usepackage[dvips]{graphicx}
   % declare the path(s) where your graphic files are
-  % \graphicspath{{../eps/}}
+  \graphicspath{{diagrams/}}
   % and their extensions so you won't have to specify these with
   % every instance of \includegraphics
-  % \DeclareGraphicsExtensions{.eps}
+  \DeclareGraphicsExtensions{.eps}
 \fi
@@
 \begin{abstract}
-Analysing the changes in software repositories over time is difficult since changes come from different developers and affect different parts of a repository.
+What is a developer's contribution to a repository?
+By using only commit counts and number of lines changed, existing tools that analyze code repositories fall short on showing the tangible contributions made by each developer.
 Individual commits provide the changed files, code diffs and a message.
@@
 
-We present a novel tool for visualizing Java projects.
-It provides more detailed information regarding all of the changes along the master branch of a project.
-It visualizes changes in type usage over time for projects and developers and derives detailed statistics using the differences between Abstract Syntax Trees (ASTs).
-We are tracking additions and deletions of declarations and invocations for each type.
-The AST diff between commits is a reentrant algorithm that can be incrementally calculated as new commits are made.
+We present a novel method for visualizing Java projects.
+Instead of simply counting line changes in a commit we extract detailed type information over time by using the differences between Abstract Syntax Trees (ASTs).
+We are then able to track the additions and deletions of declarations and invocations for each type.
 
-Using this tool we show the changes in type coverage and other usage statistics over time in several popular repositories.
-Type coverage information from the AST is compared to file coverage to check for correlation or if there is unique information provided by type information.
-Potential uses for such a tool for managers and developers is discussed.
+Using this tool we examine specific cases in well known repositories where our tool reveals interesting and useful information.
+We then compare type coverage information from the AST compared to file coverage to determine if unique information is provided by type information.
 \end{abstract}
@@
 
-The history and evolution of software repositories are often tracked by version control systems.
-The information is generally kept in the form of commits, which show the differences between each subsequent version of the software.
-These services become vital for large projects since individual changes would be impossible to track manually.
+Source code is often represented as text in a source code file.
+This text is then lexed and parsed by a compiler or interpreter with the intent of eventually executing the instructions within the source code file.
+While programmers may type in characters, their intention is at a much higher level and more abstract.
+Their intent is written in tokens that represent statements, identifiers and types that are used to solve their problem.
+Programmers solve their problem in a way that is readable to other developers and executable by a computer.
+The programmer rarely operates at the character level, or even the line level, so why should our analyses of source code operate at granularities unfamiliar to developers?
+
+The history and evolution of a software project is often tracked by version control systems such as Git in a programming language agnostic way.
+Updates to source code, documentation, and tests is recorded in the form of commits, which show the differences in lines between each subsequent revision of the software.
+Version control systems become vital for large projects since individual changes would be impossible to track manually.
 Large projects have thousands of individual changes made by many different developers.
-With more changes to look through, it becomes more difficult to see how a project is evolving over time since there is so much information to look through.
-Most hosting services for git repositories provide some basic tools for analysing projects over time, but too much information is lost due to the lack of fine grained details.
-The information regarding contributions from developers are also misleading since the number of commits and lines of code do not necessarily reflect meaningful changes to a repository.
+With more changes to peruse, it becomes more difficult to see how a project is evolving over time since there is so much information to look through.
+Most hosting services for Git repositories provide some basic tools for analysing projects over time~\cite{github-graphs,bitbucket-graphs}; but, as we will show, these tools often lack the kind of detail to gather insightful analysis.
+While commits may store individual changes to lines of code, they represent the change in \emph{structure} of the software.
+By visualizing lines changed, or files changed, the current state-of-the-art for software project visualization does not capture the \emph{meaning} behind each commit; lines of code alone does not describe how a repository changed at a particular point in time.
+Thus, information regarding contributions from developers is often misleading since the number of commits and lines of code changed does not necessarily reflect meaningful changes to a repository~\cite{robles2014}.
+
+In this paper, we propose a method that analyses the changes in type usage and method invocations over time, which we claim offers more insight than line-based, file-based, and commit-based methods.
+Our method is novel because none of the state-of-the-art tools has considered analyzing types changes over time or doing AST diffs.
 
@@
 Primitive types are those that are built into the language such as an int or char.
-Object references point to instances of classes in memory which are defined in the project or through an included library.
+Object references point to instances of classes in memory which are defined in the project itself or through an included library.
 The types in a project represent the functionality available for a developer to work with.
 A large or complex project may have many different types, whereas a small project might only use a few.
-Different parts of a program are going to use different types and libraries.
+Different parts of a program use different types and external dependencies.
 When writing or contributing to a project, the author’s choice of which part of the program to edit will in turn affect what types they end up working with.
@@
 Viewing the type information of a change can show when deprecated classes are added or removed from a project as well as how many remain.
+In a more general sense we can see when types are being added or removed from the project as well as what types are heavily used throughout the project.
 
@@
 
-Robert Dyer \textit{et al.} \cite{Dyer:2014:MBA:2568225.2568295} mined AST nodes to study the use of new Java language features over time. (More in challenge paper)
+Dyer \textit{et al.} \cite{Dyer:2014:MBA:2568225.2568295} mined AST nodes to study the use of new Java language features over time.
+They found that all new Java features do get used, but not nearly as often as they could have been.
+New features varied greatly in popularity and adoption rate.
+Developers would also modify old code to use new features as they were released.
+The authors also found the most popular features and how teams adopted new features.
+They did not check to see how much a developer used a feature but instead only checked to see if they used it at all.
 
@@
 
-Patrick Wagstrom \textit{et al.} \cite{Patrick:Wagstrom:2012} explored the roles that developers take in a networked, social development environments like GitHub.
+Wagstrom \textit{et al.} \cite{Patrick:Wagstrom:2012} explored the roles that developers take in a networked, social development environments like GitHub.
 These role were defined as their level of contribution as well as the types of issues they dealt with.
-Developers will take on multiple roles in a project and will sometimes fulfill the same role across different projects
+Developers will take on multiple roles in a project and will sometimes fulfill the same role across different projects.
 
@@
 They find what APIs are popular as well as if they are used in a framework like manner.
+
+Wu \textit{et al.} \cite{wu2004a} explore punctuated changes to software repositories with a tool that generates an Evolution Spectrograph.
+Each file has a representative row in the chart, with each cell in the row representing a section of time.
+This diagram would show changes to incoming and outgoing dependencies by darkening the cell where the change happened and lightening it over time if no changes were made.
+
+In another paper, Wu \textit{et al.} \cite{wu2004} the Evolutionary Spectrogram is used but with greater control of granularity for files and software subsystems such as directory structure.
+Spectrograms are also used to display the cardinality of a developer’s commits.
+The cardinality of a commit is the number of files or subsystems it affects in a system.
+All of the developers changes together show how many changes affect broad sections of the software.
+
+Voinea \textit{et al.} \cite{voinea2006} visualizes CVS repositories using CVSgrab and adding their own enhancements to show changes to files over time.
+Files can be sorted by creation time, activity, or alphabetically and colour coded based on author, file size, file type or by search strings.
+
+Kim \textit{et al.} \cite{Kim:2009:DRS:1555001.1555046} built a tool called Logical Structural Diff (LSdiff) which examines code changes and produces logical rules to represent them.
+Changes which do not match the rules are then pointed out to the developer.
+They also formed a focus group with software engineers who found that LSdiff complements existing tools well and showed promise for discovering inconsistent changes.
+
+The GitHub tools have the strength that they do not rely on what type of code is being committed.
+It does not matter what language or errors exist in the commit GitHub will simply spit out the number of line additions and deletions.
+The issue with this type of analysis is that it does not give a clear breakdown of is being changed only that changes have occurred.
 
@@
 
-In order to gather all the changes for a given repository over time we generated AST for each commit and compared the difference.
+\begin{figure}[!h]
+\centering
+\includegraphics[height=2.5in]{commits_view}
+\caption{The commits increase in time from bottom up.
+Therefore, B and C are compared against A and E is compared with D.
+Since D is a merge commit it is ignored as its changes are recorded in C and B.}
+\end{figure}
+
+In order to gather all the changes for a given repository over time we generated ASTs for each commit and compared the difference.
 We did this by running ‘git log’ over the git repository and listing all the commit ids in increasing time.
@@
 
+\begin{figure}[!h]
+\centering
+\includegraphics[width=\columnwidth]{context}
+\caption{}
+\end{figure}
+
+\section{The Tool}
+For convenience we provide a variety of views to help organize the data.
+These are to provide a finer level of detail.
+The main view shows a stacked bar graph over time for all types.
+Where the blue and orange portions account for the number of additions and deletions respectively.
+The height of the bar shows the total number of changes in that time period with respect to the maximum number of changes in a single period.
+The total size of the bar shows the total number of changes.
+This is calculated by multiplying the bar height by the number of changes during that time period divided by the max number of changes for this type.
+The width of each bar spans over a given time period and where the time period can be changed to span a hour, day, week, or month.
+The types are sorted from most to least used.
+
+If a user mouses over a specific bar they are given some simple stats for the given type and time period.
+This includes the types full name, number of commits, number of authors, number of additions, number of deletions, and period start and end date.
+If the user then clicks on the bar they are given more detailed stats where we list the authors and commits.
+
 \section{Case Study}
 
-In the following sections, we present {N} case studies demonstrating the extra insight that TypeV makes available.
+In the following sections, we present {N} case studies demonstrating the extra insight that TypeV method makes available.
 
@@
 
-We chose Antlr4 as it is a substantial Java project, with a commit history spanning several years. We can also see distinct changes throughout its history.
+We chose Antlr4 because it is a substantial Java project with a commit history spanning several years. We can also see distinct changes throughout its history.
 
-For example, in December of 2011 we can see that ParserATNSimulator has a large amount of additions and deletions in roughly equal proportion (Figure~\ref{fig:parser1}.
+For example, in December of 2011 we can see that the ParserATNSimulator class has a large amount of additions and deletions in roughly equal proportion (Figure~\ref{fig:parser1}.
 This suggests that ParserATNSimulator is not being removed from the project but instead there is some refactoring being done.
@@
 
+\begin{table}[!t]
+\renewcommand{\arraystretch}{1.3}
+\caption{Comparision of Type Coverage and File Coverage in Projects}
+\label{tab:summary}
+\centering
+\begin{tabular}{c|ccc}
+\hline
+\bfseries Project & \bfseries Wilcoxon & \bfseries Pearson (lin) cor & \bfseries Spearman (rank) cor \\
+\hline
+antlr4 & p$\ll$0.01 & 0.943 & 0.907\\
+bookkeeper & 0.313 & 0.984 & 0.972\\
+curator & 0.007 & 0.975 & 0.980\\
+tika & 0.653 & 0.955 & 0.998\\
+\hline
+\end{tabular}
+\end{table}
+
+To determine if type coverage as returned by TypeV gives different information than file coverage alone, we computed the following.
+For four projects (Antlr, Apache Bookkeeper, Apache Curator, Apache Tika), we counted the file coverage (cumulative amount of files touched per author), as well as type coverage (cumulative amount of types touched per author) on a commit-by-commit basis.
+That is, each datum is represents a commit with two dimensions: file coverage and type coverage.
+Data is sorted in ascending order of commit date.
+Table~\ref{tab:summary} lists a summary of the data
+
 \section{Discussion}
+
+\subsection{Threats to Validity}
+
+Authors can pull other projects into the repository and these additions will count towards the author's changes.
+
+We do not look at other branches besides the master; therefore, we might miss some of the work done by an author.
+
+We ignore the merge commits so we do not double count changes; however, if there is a merge conflict and the developer makes changes then these changes will not appear in TypeV.
+We make the assumptions that if this case happens then the changes are minor.
+
+\subsection{Future work}
+
+Currently for generic types we treat every instance of a generic as a different type.
+So for example if a user declares a list of strings and a list of integers these will appear as two different types.
+However, if they declare multiple instances of a list of strings then these instances will count toward the same type.
+In the future we would nice to come up with a way to better handle generic types.
+
+Currently we are only looking at declarations and invocations, but it would be nice to also view what classes these changes were made to.
 
@@
 % http://www.michaelshell.org/tex/ieeetran/bibtex/
-\nocite{*}
+%\nocite{*}
 \bibliographystyle{IEEEtran}

### 2016-04-09 corr.R
More R updates...

@@ -1,13 +1,15 @@
 #!/usr/bin/env Rscript
 
+library(ggplot2)
+
 # Antlr
 antlr <- read.csv('antlr.csv')
 # Turn the Unix Timestamps (in MS) to R dates.
 antlr$Date <- as.Date(as.POSIXct(antlr$Date / 1000, 'UTC', origin="1970-01-01"))
 
 with(antlr, {
-    file.cov <- antlr[Metric == 'file',]
-    type.cov <- antlr[Metric == 'type',]
+    file.cov <- antlr[antlr$Metric == 'file',]
+    type.cov <- antlr[antlr$Metric == 'type',]
 
     # Get significance tests
     print("[antlr] Wilcoxon")
@@ -21,6 +23,13 @@ with(antlr, {
 
     d.type <- density(type.cov$Coverage)
     d.file <- density(file.cov$Coverage)
+
+    ggplot(antlr, aes(antlr$Coverage, colour = antlr$Metric, fill =
+                      antlr$Metric)) +
+     geom_density(alpha = 0.4, size = 1) +
+     labs(x = "Coverage for Antlr4", y = "Density",
+          colour = "Metric", fill = "Metric")
+
 })
 
 rm(antlr)
@@ -45,8 +54,14 @@ with(tika, {
     print("[tika] Spearman (rank) correlation")
     print(cor(type.cov$Coverage, file.cov$Coverage, method = "spearman"))
 
-    d.type <- density(type.cov$Coverage)
-    d.file <- density(file.cov$Coverage)
+    #d.type <- density(type.cov$Coverage)
+    #d.file <- density(file.cov$Coverage)
+
+    ggplot(tika, aes(tika$Coverage, colour = tika$Metric, fill =
+                      tika$Metric)) +
+     geom_density(alpha = 0.4, size = 1) +
+     labs(x = "Coverage for Apache Tika", y = "Density",
+          colour = "Metric", fill = "Metric")
 })
 rm(tika)
 

### 2016-04-09 www/js/generate_table.js
Added Save SVG. Colors are broken.

@@ -1184,6 +1184,22 @@ window.makeCSVLink = function makeCSVLink(data) {
 
 
 /**
+ * Returns the download link for a SVG file
+ */
+window.makeSVGLink = function makeSVGLink(data) {
+  var doctype = '<?xml version="1.0" standalone="no"?>'
+    + '<!DOCTYPE svg PUBLIC "-//W3C//DTD SVG 1.1//EN" "http://www.w3.org/Graphics/SVG/1.1/DTD/svg11.dtd">';
+
+  // serialize our SVG XML to a string.
+  var source = (new XMLSerializer()).serializeToString(d3.select('#dna-table').select('.main-figure').node());
+
+  // create a file blob of our SVG.
+  var blob = new Blob([source], { type: 'image/svg' });
+  return URL.createObjectURL(blob);
+}
+
+
+/**
  * Places the axis on the bottom of the graph on initial render, when the
  * screen is too big.
  */

### 2016-04-09 paper/paper.tex
Update tex.

@@
 
+\begin{figure*}[!t]
+\centering
+\includegraphics[width=\textwidth]{antlr-3-code-freq-edit}
+\caption{}
+\end{figure*}
+
 \section{Introduction}
@@
 
+\begin{figure}[!ht]
+\centering
+\includegraphics[width=\columnwidth]{ParserATNSimulator}
+\caption{}
+\label{fig:parser1}
+\end{figure}
+
 We chose Antlr4 because it is a substantial Java project with a commit history spanning several years. We can also see distinct changes throughout its history.
 
-For example, in December of 2011 we can see that the ParserATNSimulator class has a large amount of additions and deletions in roughly equal proportion (Figure~\ref{fig:parser1}.
+For example, in December of 2011 we can see that the ParserATNSimulator class has a large amount of additions and deletions in roughly equal proportion (Figure~\ref{fig:parser1}).
 This suggests that ParserATNSimulator is not being removed from the project but instead there is some refactoring being done.
@@
 \end{table}
+
+\begin{figure}[!h]
+\centering
+\includegraphics[width=\columnwidth]{antlr-density}
+\caption{}
+\end{figure}
+
+\begin{figure}[!h]
+\centering
+\includegraphics[width=\columnwidth]{tika-density}
+\caption{}
+\end{figure}
 

### 2016-04-09 paper/paper.tex
Update tex. Add figure to compare GH with TypeV.

@@
 
-In the following sections, we present {N} case studies demonstrating the extra insight that TypeV method makes available.
+In the following sections, we present {N} case studies demonstrating the extra insight that the TypeV method makes available.
 
@@
 \includegraphics[width=\columnwidth]{ParserATNSimulator}
-\caption{}
+\caption{Snapshot of TypeV showing changes to ParserATNSimulator in antlr4. Date ranges from approximately the begining of 2011 to the end of April in 2012.}
 \label{fig:parser1}
+
+\end{figure}
+\begin{figure}[!ht]
+\centering
+\includegraphics[width=\columnwidth]{Antlr4-GH-Lines-ParserATNSimulator}
+\caption{GitHub's line difference graph showing changes to antlr4. The time spans approximately the same length as Figure~\ref{fig:parser1}.}
+\label{fig:parser2}
 \end{figure}
@@
 This suggests that ParserATNSimulator is not being removed from the project but instead there is some refactoring being done.
-If we take a closer look at the commit messages we can see that they mention refactoring and reorganization of code (Figure~\ref{fig:parser2}).
+If we take a closer look at the commit messages we can see that they mention refactoring and reorganization of code.
 Zooming in on that date,  shows that during that month they were finishing work on a new version of ParserATNSimulator.
 
-If we tried to do the same analyses with the GitHub tools it would be significantly more difficult to achieve the same conclusion.
+If we tried to do the same analyses with the GitHub tools it would be significantly more difficult to achieve the same conclusion (Figure~\ref{fig:parser2}).
 First if we only look at line changes this change to ParserATNSimulator is overwhelmed by other changes and hardly shows up on the GitHub tools.

### 2016-04-09 paper/diagrams/embed-styles.py
Fix commit view and eps export.

@@ -1,28 +1,0 @@
-#!/usr/bin/env python
-
-from xml.etree import ElementTree
-import sys
-from os import path as p
-
-CSS_FILENAME = p.abspath(p.join(p.dirname(__file__), '..', 'www', 'css',
-                                'graph-styles.css'))
-
-_, filename, out_name = sys.argv
-
-with open(filename, encoding='utf-8') as raw_svg:
-    svg = ElementTree.fromstring(raw_svg.read())
-
-with open(CSS_FILENAME, encoding='utf-8') as css_file:
-    styles = css_file.read()
-
-template = '''<style type="text/css">\
-<![CDATA[
-{0}
-]]> 
-</style>
-'''
-
-style_element = ElementTree.fromstring(template.format(styles))
-svg.insert(0, style_element)
-
-ElementTree.ElementTree(svg).write(out_name)

### 2016-04-10 paper/paper.tex
Fixed issue with long urls.
Update tex.

@@
 
+\usepackage{enumitem}
+\setlist[description]{leftmargin=1cm,labelindent=\parindent}
+
 \usepackage{changepage}
+\usepackage{url}
+\usepackage{breakurl}
+\usepackage[breaklinks]{hyperref}
+\def\UrlBreaks{\do\/\do-}
 
@@
 What is a developer's contribution to a repository?
-By using only commit counts and number of lines changed, existing tools that analyze code repositories fall short on showing the tangible contributions made by each developer.
-Individual commits provide the changed files, code diffs and a message.
-However, when many commits are viewed as a group, the details are lost.
-Currently there is a lack of tools for tracking the changes in a repository beyond simply looking at the number of commits of lines changed.
-The commit information can be misleading since lines of code gives no indication of what was actually being worked on without careful examination of the changed code.
-Providing a more comprehensive view of this information could provide deeper insights into how software projects evolve since changes to design and features are not clearly visible.
-
-We present a novel method for visualizing Java projects.
-Instead of simply counting line changes in a commit we extract detailed type information over time by using the differences between Abstract Syntax Trees (ASTs).
+By using only counting commits and number of lines changed, existing tools that analyze source code repositories fall short on showing the effective contributions made by each developer.
+When many commits are viewed as a group, the details are lost.
+Commit information can be misleading since lines of code give no indication of what was actually being worked on without careful examination of the changed code.
+Providing a semantic view of this information could provide deeper insights into how software projects evolve since changes to design and features are not clearly visible from line changes alone.
+We present TypeV: a method for visualizing Java source code repositories.
+Instead of counting line changes in a commit we extract detailed type information over time by using the differences between abstract syntax trees (ASTs).
 We are then able to track the additions and deletions of declarations and invocations for each type.
-
-Using this tool we examine specific cases in well known repositories where our tool reveals interesting and useful information.
+Furthermore, we can track each author’s type usage over time.
+Using this tool we examine specific cases in well-known repositories where our tool reveals interesting and useful information.
 We then compare type coverage information from the AST compared to file coverage to determine if unique information is provided by type information.
@@
 Source code is often represented as text in a source code file.
-This text is then lexed and parsed by a compiler or interpreter with the intent of eventually executing the instructions within the source code file.
-While programmers may type in characters, their intention is at a much higher level and more abstract.
-Their intent is written in tokens that represent statements, identifiers and types that are used to solve their problem.
-Programmers solve their problem in a way that is readable to other developers and executable by a computer.
+This text is then parsed by a compiler or interpreter with the intent of eventually executing the instructions within the source code file.
+While programmers may type in individual characters, their intention is at a much higher level.
+Their intent is written in tokens that represent statements, identifiers and types that are used to solve their problems.
+Programmers solve problems in a way that is readable to other developers and executable by a computer.
 The programmer rarely operates at the character level, or even the line level, so why should our analyses of source code operate at granularities unfamiliar to developers?
 
-The history and evolution of a software project is often tracked by version control systems such as Git in a programming language agnostic way.
-Updates to source code, documentation, and tests is recorded in the form of commits, which show the differences in lines between each subsequent revision of the software.
+The history and evolution of a software project is often tracked by version control systems, such as Git, in a programming language agnostic way.
+Updates to source code, documentation, and tests are recorded in the form of commits, which show the differences in lines between each subsequent revision of the software.
 Version control systems become vital for large projects since individual changes would be impossible to track manually.
@@
 While commits may store individual changes to lines of code, they represent the change in \emph{structure} of the software.
-By visualizing lines changed, or files changed, the current state-of-the-art for software project visualization does not capture the \emph{meaning} behind each commit; lines of code alone does not describe how a repository changed at a particular point in time.
+By visualizing lines changed, or files changed, the current state-of-the-art for software project visualization rarely capture the \emph{meaning} behind each commit; lines of code alone does not describe how a repository changed at a particular point in time.
 Thus, information regarding contributions from developers is often misleading since the number of commits and lines of code changed does not necessarily reflect meaningful changes to a repository~\cite{robles2014}.
@@
 In this paper, we propose a method that analyses the changes in type usage and method invocations over time, which we claim offers more insight than line-based, file-based, and commit-based methods.
-Our method is novel because none of the state-of-the-art tools has considered analyzing types changes over time or doing AST diffs.
+Of the services we have surveyed, none of the tools visualize project evolution using granular type analysis per author via commit-by-commit abstract syntax tree differencing.
+For simplicity we decided to stick to analyzing Java projects as it allowed us to quickly develop and test our software.
+In the future the approach described later in the paper could be implemented to handle other strongly typed languages.
 
@@
 
+Type information is useful for seeing what libraries and functions are being used.
+For example when a part of code becomes deprecated, continuing to use it may lead to design or even security flaws.
+Viewing the type information of a change can show when deprecated classes are added or removed from a project as well as how many remain over time.
+In a more general sense, we can see when types are being added or removed from the project as well as what types are heavily used throughout the project.
+Even better, we can see \emph{who} uses which types heavily.
+
 Having advanced tools for analysing repositories would benefit many different parts of the software development process.
@@
 Providing tools which bring more meaning to line differences would allow people less familiar with the code to understand changes being made.
-When large changes are made that cover many lines of code, it would also be useful to see how the change affects the software rather than a several thousand line diff.
+When large changes are made that cover many aspects of a project, it would also be useful to see how the change affects the software rather than a several thousand line diff.
 Seeing how a project has changed over time is also vital for management since evaluating the state of completion of software can be very difficult.
 
-Type information is useful for seeing what libraries and functions are being used.
-When a part of code becomes deprecated, continuing to use it may lead to design or security flaws.
-Viewing the type information of a change can show when deprecated classes are added or removed from a project as well as how many remain.
-In a more general sense we can see when types are being added or removed from the project as well as what types are heavily used throughout the project.
+In the following sections of the paper we will introduce a tool that we developed that visualizes changes to types used in a project over time.
+The tool is able to extract type information from a repository by using comparing ASTs.
+We then ask the following questions:
+\\ \\
+\textbf{RQ1:} Can we extract type information from repositories? \\ \\
+\textbf{RQ2:} Does the type information give any interesting or useful insight into the project? \\ \\
+\textbf{RQ3:} How does type coverage compare to file coverage?
 
 \section{Related Work}
+
+Analyzing source code is a popular research topic.
+Most of the focus is on how certain programming languages and features are used.
+However, there is no focus specifically on how types are used throughout a project.
 
@@
 Parnin \textit{et al.} \cite{Parnin:2011:JGA:1985441.1985446} mined repositories to see how Java generics have been used in open source projects and found that generic usages were often introduced by a single developer in a project.
-Generics usage was also fairly narrow, with the primary use being collecting and traversing lists of objects.
+Generics usage was also fairly narrow, with the primary use being collecting and traversing lists of objects
 
@@
 They find what APIs are popular as well as if they are used in a framework like manner.
+
+The aforementioned research is trying to learn how developers are using programming languages by looking at many repositories.
+This type of info can be extracted from individual repositories using our new tool.
+Type info may also provide additional information on how developers use languages.
 
@@
 
+The service EasyBi \cite{EasyBi} provides additional visualizations for Git repositories for the purpose of business decisions.
+EasyBi works by taking the git log file and generating charts and graphs from it.
+It tries to solve the issue of summarizing many commits and has greater controls for granularity and authors but still operates on commits and line differences.
+
+Wagstrom \textit{et al.} \cite{Patrick:Wagstrom:2012} explored the roles that developers take in a networked, social development environments like GitHub.
+These role were defined as their level of contribution as well as the types of issues they dealt with.
+Developers will take on multiple roles in a project and will sometimes fulfill the same role across different projects.
+
 The GitHub tools have the strength that they do not rely on what type of code is being committed.
@@
 
-An Abstract Syntax Tree (AST) is a tree structure which represents a piece of code by breaking its syntactic constructs into a tree structure.
-Each node in the tree represents a syntactically valid chunk of code which can be contained in a parent node or broken up into child nodes.
-This captures the structure of the code in an abstract manner.
-When a change is made to a file, the difference will also appear in the AST.
-This means that the AST of revisions in code repositories can be compared to see what structures a developer is touching.
+The core of our analysis is calculating the differences between abstract syntax trees of subsequent commits (Figure~\ref{fig:context}).
 
-The ASTs were generated using Spoon \cite{pawlak:hal-01169705}.
+An \emph{abstract syntax tree} (AST) is a data structure derived from source code by breaking its syntactic constructs into a tree.
+Each node in the tree represents a syntactically valid chunk of code.
+This captures the structure of the code in an abstract and tractable manner.
+Converting raw source text to an AST is often one of the first steps performed by compilers.
+When a programmer changes code in a file, the difference will appear in the AST.
+This means that the AST of revisions in code repositories can be compared to see what structures a programmer is modifying.
+
+The Java ASTs in our analysis were generated using Spoon~\cite{pawlak:hal-01169705}.
 Spoon is a tool for transforming and analysing Java source code.
-It breaks code up into a meta-model including AST.
+It breaks code up into a meta-model including the AST itself.
 The model consists of three parts: structural elements, code elements and references to program elements.
-The Spoon model is convenient because it provides the structure for performing an AST diff while retaining the code elements for analysis afterwards.
-Spoon also preserves all type and library information, which is what we were interested in.
+The Spoon model is convenient because it provides the infrastructure for performing AST differencing (described later) while retaining the code elements (such as type information) for analysis afterwards.
+Spoon also preserves all type and package (library) information—information vital to our semantic analysis.
 
-GumTree \cite{falleri:hal-01054552} is an algorithm which is used to compute the difference between two ASTs generated by Spoon.
+To determine what has changed between two subsequent commits, we computed their \emph{AST difference}.
+An AST difference (or simply \emph{AST diff}) states which nodes were added and removed from a tree (step two in Figure~\ref{fig:context}).
+For this purpose, we used GumTree~\cite{falleri:hal-01054552}, an algorithm made to compute the difference between two ASTs generated by Spoon.
 For new files and deleted files we did not use GumTree; we simply ran Spoon on the files and treated everything as an addition or deletion respectively.
 
-Once we have the generated ASTs we were able to run through the trees and count the declarations and invocations for a given type.
-For invocations we also recorded the method call.
-These stats were then marked as additions or deletions.
+Once we have the generated ASTs, we were able to traverse through the trees to count the number of added or deleted \emph{declarations} and \emph{invocations} for a given type.
+
+\begin{description}
+\item [declaration] 
+  Stating that use of a field or variable of a given type. For example declaration of a String type variable might look as follows:
   \\
+  \begin{adjustwidth}{2.5em}{0pt}
+  String myString; \\
+  \end{adjustwidth}
+\item [invocation] 
+  A call site for a method that belongs to a type. For example an invocation of the trim method of a String type might look as follows:
+  \\
+  \begin{adjustwidth}{2.5em}{0pt}
+  myString.trim(); \\
+  \end{adjustwidth}
+\end{description}
+
+Our analysis tool would produce the following output:
+
 \begin{adjustwidth}{2.5em}{0pt}
@@
 Since D is a merge commit it is ignored as its changes are recorded in C and B.}
+\label{fig:commits_view}
 \end{figure}
@@
 \caption{}
+\label{fig:context}
 \end{figure}
@@
 \label{fig:parser1}
+\end{figure}
 
-\end{figure}
 \begin{figure}[!ht]

### 2016-04-10 paper/paper.tex
Update tex and some bibs.

@@
 
-Our analysis tool would produce the following output:
+An example of what our analysis tool might produce is as follows:
 
@@
 Since D is a merge commit it is ignored as its changes are recorded in C and B.}
-\label{fig:commits_view}
+\label{fig:commits}
 \end{figure}
 
+This states that two declarations of type java.net.URI were added, and two declarations of type java.lang.String were deleted.
+Note the fully qualified type names: the source text may not literally have the characters java.net.URI, but Spoon was able to infer from the import statements in the Java source text that references to ''URI'' resolve to the fully-qualified type name \texttt{java.net.URI}.
+Thus, TypeV is able to keep track of the actual class being used and is not confused when two classes in separate packages have the same name.
+
 In order to gather all the changes for a given repository over time we generated ASTs for each commit and compared the difference.
-We did this by running ‘git log’ over the git repository and listing all the commit ids in increasing time.
-Each commit and its parent were then compared.
-If the commit had no parent then every file in the commit was treated as an addition.
-Since we compared each commit to its parent we were able to account for branching that happened in the master branch of the repository.
-Also we ignored commits that were merges as all changes in the merge should already be accounted for in previous commits.
+We did this by running \texttt{git log} over the Git repository and listing all the commit IDs in chronological order (Figure~\ref{fig:commits}) Each commit and its parent were then compared, using AST differencing as described above.
+There are a few edges cases, however: If a commit has no parent, then every declaration or invocation in the commit was treated as an addition.
+Since we compared each commit to its parent, as reported by Git, we were able to account for any branching that happened outside the primary branch of the repository (usually named \texttt{master}).
+If a commit has more than one parent, this indicates a merge commit in Git.
+Merge commits contain the changes made in each parent commit combined.
+Therefore, we ignored merge commits since all changes in the merge should already have been accounted for in previous commits.
 

### 2016-04-10 paper/paper.tex
Update paper.

@@
 \caption{}
+\label{fig:GH_code_freq}
 \end{figure*}
@@
 
-In this paper, we propose a method that analyses the changes in type usage and method invocations over time, which we claim offers more insight than line-based, file-based, and commit-based methods.
+In this paper, we propose a method that analyses the changes in type usage and method invocations over time, as well as tool to visualize this analysis, which we call TypeV.
+We claim that type usage offers more insight than line-based, file-based, and commit-based methods of visualization and analysis.
 Of the services we have surveyed, none of the tools visualize project evolution using granular type analysis per author via commit-by-commit abstract syntax tree differencing.
-For simplicity we decided to stick to analyzing Java projects as it allowed us to quickly develop and test our software.
-In the future the approach described later in the paper could be implemented to handle other strongly typed languages.
+For simplicity, we decided to only analyze Java projects, as it allowed us to quickly develop and test our software.
+Since our approach described in this paper can be generally applied to any statically-typed language, we plan to extend it to other languages.
 
-In Java, a type can be a primitive type or an object reference.
-Primitive types are those that are built into the language such as an int or char.
-Object references point to instances of classes in memory which are defined in the project itself or through an included library.
+In Java, a type can be a primitive, provided by the language itself, or object references.
+Primitive types are those that are built into the language such as an \texttt{int} or \texttt{char}.
+Object references point to instances of classes.
+Classes are defined in the project itself or through an included library.
 The types in a project represent the functionality available for a developer to work with.
-A large or complex project may have many different types, whereas a small project might only use a few.
+Java’s syntax encourages the creation of classes, thus we expect substantial Java projects to define many domain-specific types.
 Different parts of a program use different types and external dependencies.
-When writing or contributing to a project, the author’s choice of which part of the program to edit will in turn affect what types they end up working with.
+When writing or contributing to a project, the author’s choice of which part of the program to edit will in turn affect what types they end up working with --- that is, it will affect the author’s \emph{type coverage}.
 
 Type information is useful for seeing what libraries and functions are being used.
-For example when a part of code becomes deprecated, continuing to use it may lead to design or even security flaws.
-Viewing the type information of a change can show when deprecated classes are added or removed from a project as well as how many remain over time.
+For example when a part of code becomes \emph{deprecated}, continuing its use may lead to design flaws or even security vulnerabilities.
+For example, in Apache Nifi, an earlier version of the software used a compromised key-derivation functions; hence, this class was deprecated, but left in for compatibility with older versions of the software~\cite{nifi}.
+Viewing the type information of a change can show when usage of deprecated classes are added or removed from a project as well as how many remain over time.
 In a more general sense, we can see when types are being added or removed from the project as well as what types are heavily used throughout the project.
@@
 
-Having advanced tools for analysing repositories would benefit many different parts of the software development process.
+Having tools for analysing repositories based on type usage and method invocation would benefit many different parts of the software development process.
 Software management requires a high level understanding of the state of a repository and the changes being made which affect the functionality of the software.
-Tracking developer contributions in a less misleading way is also important for management since important contributions may be small and focused.
-Line difference only provide meaningful info when they are analysed by someone with familiarity with the source code being changed.
+Tracking developer contributions in a less misleading way is also important for management since important contributions may be small and focused, but nevertheless, significant.
+Line difference (or simply, \emph{diff}) only provide meaningful information when they are analysed by someone with familiarity with the source code being changed.
 Providing tools which bring more meaning to line differences would allow people less familiar with the code to understand changes being made.
@@
 
-In the following sections of the paper we will introduce a tool that we developed that visualizes changes to types used in a project over time.
-The tool is able to extract type information from a repository by using comparing ASTs.
-We then ask the following questions:
+In the following sections of the paper we will introduce a tool, TypeV, that we developed that visualizes changes to types used in a project over time.
+TypeV is able to extract type information from a repository by computing commit-by-commit abstract syntax tree differencing, which we describe in section~\ref{sec:methodology}.
+We seek to answer the following questions:
 \\ \\
-\textbf{RQ1:} Can we extract type information from repositories? \\ \\
+\textbf{RQ1:} How can we extract type and method usage over time from repositories? \\ \\
 \textbf{RQ2:} Does the type information give any interesting or useful insight into the project? \\ \\
@@
 Parnin \textit{et al.} \cite{Parnin:2011:JGA:1985441.1985446} mined repositories to see how Java generics have been used in open source projects and found that generic usages were often introduced by a single developer in a project.
-Generics usage was also fairly narrow, with the primary use being collecting and traversing lists of objects
+Generics usage was also fairly narrow, with the primary use being collecting and traversing lists of objects.
 
@@
 
+The Gitinspector project \cite{Gitinspector} is an open source tool for the statistical analysis of git repositories.
+It provides an advanced timeline view of changes per author and cumulative statistics but again relies upon the line differences to summarize the work of each author.
+
 Wagstrom \textit{et al.} \cite{Patrick:Wagstrom:2012} explored the roles that developers take in a networked, social development environments like GitHub.
@@
 
+Lee \textit{et al.} \cite{lee2013} developed a tool for visualizing the branching structure of git repositories. This was used to analyze concurrent workflows in popular open source repositories.
+
 The GitHub tools have the strength that they do not rely on what type of code is being committed.
-It does not matter what language or errors exist in the commit GitHub will simply spit out the number of line additions and deletions.
+It does not matter what language or if errors exist in the commit, GitHub will simply spit out the number of line additions and deletions.
 The issue with this type of analysis is that it does not give a clear breakdown of is being changed only that changes have occurred.
@@
 \section{Methodology}
+\label{sec:methodology}
 
@@
 The model consists of three parts: structural elements, code elements and references to program elements.
-The Spoon model is convenient because it provides the infrastructure for performing AST differencing (described later) while retaining the code elements (such as type information) for analysis afterwards.
+''The structural part contains the declarations of the program elements, such as interface, class, variable, method, annotation, and enum declarations.
+The code part contains the executable Java code, such as the one found in method bodies.
+The reference part models the references to program elements (for instance a reference to a type)'' \cite{pawlak:hal-01169705} The Spoon model is convenient because it provides the infrastructure for performing AST differencing (described later) while retaining the code elements (such as type information) for analysis afterwards.
 Spoon also preserves all type and package (library) information—information vital to our semantic analysis.
@@
 For this purpose, we used GumTree~\cite{falleri:hal-01054552}, an algorithm made to compute the difference between two ASTs generated by Spoon.
-For new files and deleted files we did not use GumTree; we simply ran Spoon on the files and treated everything as an addition or deletion respectively.
+For new files and deleted files we did not use GumTree; we were simply able to run Spoon on the files and treated everything as an addition or deletion respectively.
 
@@
 \includegraphics[height=2.5in]{commits_view}
-\caption{The commits increase in time from bottom up.
-Therefore, B and C are compared against A and E is compared with D.
-Since D is a merge commit it is ignored as its changes are recorded in C and B.}
+\caption{Commits increase in time from bottom up.
+Therefore, ASTs from B and C are diff’d against ASTs in A.
+ASTs from E are diff’d against D.
+Since D is a merge commit, its changes are ignored as they are recorded in C and B.}
 \label{fig:commits}
@@
 
-This states that two declarations of type java.net.URI were added, and two declarations of type java.lang.String were deleted.
-Note the fully qualified type names: the source text may not literally have the characters java.net.URI, but Spoon was able to infer from the import statements in the Java source text that references to ''URI'' resolve to the fully-qualified type name \texttt{java.net.URI}.
+This states that two declarations of type \texttt{java.net.URI} were added, and two declarations of type \texttt{java.lang.String} were deleted.
+Note the fully qualified type names: the source text may not literally have the characters \texttt{java.net.URI}, but Spoon was able to infer from the import statements in the Java source text that references to ''URI'' resolve to the fully-qualified type name \texttt{java.net.URI}.
 Thus, TypeV is able to keep track of the actual class being used and is not confused when two classes in separate packages have the same name.
@@
 In order to gather all the changes for a given repository over time we generated ASTs for each commit and compared the difference.
-We did this by running \texttt{git log} over the Git repository and listing all the commit IDs in chronological order (Figure~\ref{fig:commits}) Each commit and its parent were then compared, using AST differencing as described above.
+We did this by running \texttt{git log} over the Git repository and listing all the commit IDs in chronological order (Figure~\ref{fig:commits}).
+Each commit and its parent were then compared, using AST differencing as described above.
 There are a few edges cases, however: If a commit has no parent, then every declaration or invocation in the commit was treated as an addition.
@@
 \section{The Tool}
-For convenience we provide a variety of views to help organize the data.
-These are to provide a finer level of detail.
-The main view shows a stacked bar graph over time for all types.
-Where the blue and orange portions account for the number of additions and deletions respectively.
-The height of the bar shows the total number of changes in that time period with respect to the maximum number of changes in a single period.
-The total size of the bar shows the total number of changes.
-This is calculated by multiplying the bar height by the number of changes during that time period divided by the max number of changes for this type.
-The width of each bar spans over a given time period and where the time period can be changed to span a hour, day, week, or month.
-The types are sorted from most to least used.
 
-If a user mouses over a specific bar they are given some simple stats for the given type and time period.
-This includes the types full name, number of commits, number of authors, number of additions, number of deletions, and period start and end date.
-If the user then clicks on the bar they are given more detailed stats where we list the authors and commits.
+\begin{figure*}[!t]
+\centering
+\includegraphics[width=\textwidth]{Screenshot_antlr4}
+\caption{}
+\label{fig:main-view}
+\end{figure*}
+
+The analyzed data would be difficult to analyze if displayed as its raw text form; instead, we created an HTML5 web visualization application for intended for developers, project managers, and researchers to use~\ref{fig:main-view}.
+
+TypeV provide a variety of views to help organize the data and provide a finer level of detail.
+The main view displays a series of stacked bar graphs which show changes to types over a period of time in a project~\ref{fig:main-view}.
+The blue and orange portions account for the number of additions and deletions in a similar fashion to GitHub’s code frequency graph~\ref{fig:GH_code_freq}.
+The height of each bar shows the number of changes to a type in a smaller unit of time.
+This is calculated by multiplying the bar height by the number of changes during that time period divided by the maximum number of changes attested for this type.
+A bar which fills the row vertically is the period with the largest number of changes to that type out of the visible time periods.
+The other bars in the row are proportional to that maximum number.
+The width of each bar represents a set time period and can be changed to an hour, day, week, or month.
+The rows are sorted so that the graph for the most used types appear the top and are in descending order.
+
+When a user mouses over a specific bar in the graph they are given simple statistics for the type and time period at that location.
+This includes the fully-qualified name of the type, the number of commits, the number of authors, the number of additions, the number of deletions, and the start and end date of that time period.
+If the user clicks on the bar, they are given more detailed statistics which list the authors who made the commits and each commit message~\ref{fig:typev-deets}.
 

### 2016-04-10 paper/paper.tex
Fixed some formatting.

@@
 TypeV is able to extract type information from a repository by computing commit-by-commit abstract syntax tree differencing, which we describe in section~\ref{sec:methodology}.
-We seek to answer the following questions:
-\\ \\
+We seek to answer the following questions: \\
+
 \textbf{RQ1:} How can we extract type and method usage over time from repositories? \\ \\
@@
 
-The analyzed data would be difficult to analyze if displayed as its raw text form; instead, we created an HTML5 web visualization application for intended for developers, project managers, and researchers to use~\ref{fig:main-view}.
+The analyzed data would be difficult to analyze if displayed as its raw text form; instead, we created an HTML5 web visualization application for intended for developers, project managers, and researchers to use (Figure~\ref{fig:main-view}).
 
 TypeV provide a variety of views to help organize the data and provide a finer level of detail.
-The main view displays a series of stacked bar graphs which show changes to types over a period of time in a project~\ref{fig:main-view}.
-The blue and orange portions account for the number of additions and deletions in a similar fashion to GitHub’s code frequency graph~\ref{fig:GH_code_freq}.
+The main view displays a series of stacked bar graphs which show changes to types over a period of time in a project (Figure~\ref{fig:main-view}).
+The blue and orange portions account for the number of additions and deletions in a similar fashion to GitHub’s code frequency graph (Figure~\ref{fig:GH_code_freq}).
 The height of each bar shows the number of changes to a type in a smaller unit of time.
@@
 This includes the fully-qualified name of the type, the number of commits, the number of authors, the number of additions, the number of deletions, and the start and end date of that time period.
-If the user clicks on the bar, they are given more detailed statistics which list the authors who made the commits and each commit message~\ref{fig:typev-deets}.
+If the user clicks on the bar, they are given more detailed statistics which list the authors who made the commits and each commit message (Figure~\ref{fig:typev-deets}).
+
+One issue with analyzing individual authors is that authors can commit from a variety of different computers which can have different emails and usernames attached to the git configuration.
+This means in the git log one author might appear as several different authors.
+There is no perfect way to deal with this problem; however, we allow the user of our software to select and view multiple users at a time.
+Thus allowing the user to view changes from an individual or group of developers.
+
+To allow for a more detailed analyzes of a specific component of the repository we added some basic filtering tools.
+The first is that a user can filter by type name.
+The second allows the user to view changes that fall in a specific start and end date.
+The third allows them to specify the time period of each bar and this includes hour, day, week, and month.
+Finally, they can choose to see only declarations, types used in declarations and invocations, or invocations.
 
@@
 
-In the following sections, we present {N} case studies demonstrating the extra insight that the TypeV method makes available.
+In the following sections, we present case studies demonstrating the extra insight that the TypeV method makes available.
 
@@
 
-Apache Bookkeeper deprecated the method AbstractConfiguration.getLedgerManagerType() in release 4.2.0~\footnote{\url{https://bookkeeper.apache.org/docs/r4.2.0/apidocs/deprecated-list.html}} and replaced it with AbstractConfiguration.getLedgerManagerFactoryClass().
+Apache Bookkeeper deprecated the method AbstractConfiguration.getLedgerManagerType() in release 4.2.0\footnote{\url{https://bookkeeper.apache.org/docs/r4.2.0/apidocs/deprecated-list.html}} and replaced it with AbstractConfiguration.getLedgerManagerFactoryClass().
 If we look at getLedgerManagerType() we can see at the time of the start of the deprecation all instances were removed and getLedgerManagerFactoryClass() was added.
@@
 
-To determine if type coverage as returned by TypeV gives different information than file coverage alone, we computed the following.
-For four projects (Antlr, Apache Bookkeeper, Apache Curator, Apache Tika), we counted the file coverage (cumulative amount of files touched per author), as well as type coverage (cumulative amount of types touched per author) on a commit-by-commit basis.
-That is, each datum is represents a commit with two dimensions: file coverage and type coverage.
+The information regarding which types are used in each change is unique to the AST diff method, so it will be useful to see how changes affect the types in a project rather than just the files involved in a diff.
+We define type coverage as the number of types that have been touched out of the total number of types during either a certain time period or the entire project.
+File coverage is the same measurement, but using the number of source files instead.
+To determine if type coverage gives different information than file coverage, we analyzed four popular Java projects: Antlr, Apache Bookkeeper, Apache Curator and Apache Tika.
+We counted the file coverage and type coverage on a commit-by-commit basis.
+That is, each datum represents a commit with three dimensions: file coverage and type coverage, and its commit date.
 Data is sorted in ascending order of commit date.
-Table~\ref{tab:summary} lists a summary of the data
+Table~\ref{tab:summary} lists a summary of the data provided.
+In general, all measures are strongly correlated, both for Pearson (linear) correlation, and Spearman (rank) correlation.
+As type coverage and file coverage are both cumulative measures, positive correlation is expected.
+However, even when the distributions are significantly different (as with Antlr and Curator), the correlation is still strongly positive, but not 100\%.
+This means that when the coverage for types and files of a commit increase, the relationship is not necessarily 1-to-1.
+This is remarkable in a language like Java, a classes and files are generally in a 1-to-1 relationship.We conclude from this that type coverage yields related, yet different information than file coverage alone.
 
@@
 
+When managing group projects in an introductory software engineering course, it is very important to be able to see the contributions that students are making to the group repository to judge learning, progress and collaboration.
+When supervising 4-5 groups of students, it becomes unfeasible to look through all of the commits of each student.
+Students may take the wrong approach in a certain piece of code which becomes ingrained in the project and very difficult to change.
+Many groups have problems with group members who do not complete their assigned work or make contributions which end up being meaningless.
+Issues like this can be difficult to detect and prevent.
+As a manager, having a tool like TypeV allows one to quickly summarize the work done by each group member, organized by the types they have been working on.
+This can show when intervention or guidance is required, without the large time commitment of looking through commits.
+
+The purpose of deprecated software features is to give developers time to change projects to follow a new or better standard without breaking the code.
+Deprecated features can be marked for removal from a project or have some flaw that would make a different feature the better option.
+Given the time at which a feature was deprecated, seeing the progress of replacement has some important implications.
+In the case of a security flaw, that time could indicated times at which systems or users were vulnerable.
+If a feature is marked for removal, a project using a deprecated type has a limited amount of time to make the changes.
+Removing all instances of deprecated features could also mean a project can be upgraded to use new tools or features that were not possible while the deprecated features were present.
+Here we show examples where deprecated classes were removed f
+
+Bug triaging for large open source projects is a very time consuming task, with hundreds  of bugs being reported daily.
+Each bug must be processed and assigned to a developer that has the experience required to fix the bug.
+In order to be effective, the triager has to have knowledge of what each developer has worked on as well as the parts of the code may be affected by the bugs.
+Being able to directly see what types a developer has worked with over time is important information for triaging.
+If the type causing the bug is known, the developers which have recently used that type can be assigned the bug.
+The commit messages for the affected types are also visible, providing greater context about how the type has been used.
+The developer who is assigned the bug can also use type information to locate other affected pieces of code and know which developers have been using the affected types.
+
 \subsection{Threats to Validity}
 
+Some commands in git present a problem gathering the statistics.
 Authors can pull other projects into the repository and these additions will count towards the author's changes.
+These will show up as massive changes by a single author with many types that they did not actually create.
+Changes of this size are visible in the graph and can be found by inspecting the commit messages.
 
-We do not look at other branches besides the master; therefore, we might miss some of the work done by an author.
+We only look at the master branch of a repository and ignore other branches; therefore, we might miss some of the work done by an author in another branch.
+Going through all of the branches in a repository would take a very long time and contain a large amount of duplicate data.
+Being able to choose which branch to analyze is something that can be added.
 

### 2016-04-10 paper/paper.tex
Updated tex.

@@
 What is a developer's contribution to a repository?
-By using only counting commits and number of lines changed, existing tools that analyze source code repositories fall short on showing the effective contributions made by each developer.
+By only counting commits and number of lines changed, existing tools that analyze source code repositories fall short on showing the effective contributions made by each developer.
 When many commits are viewed as a group, the details are lost.
@@
 \includegraphics[width=\textwidth]{antlr-3-code-freq-edit}
-\caption{}
+\caption{GitHub graph showing code frequency.
+Note that the graph was truncated to better show the details.
+Thus the changes that occur in mid 2015 are much larger than appear.
+The additions reach over 400k and the deletions go over 300k.}
 \label{fig:GH_code_freq}
@@
 Thus, information regarding contributions from developers is often misleading since the number of commits and lines of code changed does not necessarily reflect meaningful changes to a repository~\cite{robles2014}.
+
+Consider tools easily available to an open source software developer: GitHub graphs~\cite{github-graphs}.
+Since GitHub tracks contributions based solely on number of commits, and numbers of lines changes, data may be lacking, or even misleading.
+In the case of the Antlr4 project (Figure~\ref{fig:GH_code_freq}), the graph creates more questions than it answers.
+What happened at those big spikes of line changes in the graph?
+One of the spikes is so large, we had to truncate the graph to be able to show it.
+How did these large changes affect the code base?
+Who committed these large changes?
+Was it a team of contributors or a one particularly prolific programmer?
+Plotting the number of lines changed may be misleading, because a huge spike in additions may be caused a number of reasons, such as the inclusion of automatically-generated documentation, the inclusion of an external library, or perhaps even a huge refactoring effort.
 
@@
 TypeV is able to extract type information from a repository by computing commit-by-commit abstract syntax tree differencing, which we describe in section~\ref{sec:methodology}.
-We seek to answer the following questions: \\
-
-\textbf{RQ1:} How can we extract type and method usage over time from repositories? \\ \\
+We seek to answer the following questions: \\ \\
+\textbf{RQ1:} How can we extract type and method usage over time from repositories \\ \\
 \textbf{RQ2:} Does the type information give any interesting or useful insight into the project? \\ \\
@@
 
-\begin{adjustwidth}{2.5em}{0pt}
-\#DECLARE $\vert$ INSERT $\vert$ java.net.URI $\vert$ 2 \\
-\#DECLARE $\vert$ DELETE $\vert$ java.lang.String $\vert$ 2 \\
-\end{adjustwidth}
+\begin{verbatim}
+#DECLARE | INSERT | java.net.URI | 2
+#DECLARE | DELETE | java.lang.String | 2
+\end{verbatim}
 
@@
 
-To allow for a more detailed analyzes of a specific component of the repository we added some basic filtering tools.
+To allow for a more detailed analysis of a specific component of the repository we added some basic filtering tools.
 The first is that a user can filter by type name.
-The second allows the user to view changes that fall in a specific start and end date.
-The third allows them to specify the time period of each bar and this includes hour, day, week, and month.
-Finally, they can choose to see only declarations, types used in declarations and invocations, or invocations.
+The second allows the user to view changes that fall within a specific start and end date.
+The third allows them to specify the time span represented by each bar: either hour, day, week, or month.
+Finally, they can choose to see only types declarations, types used in declarations and invocations, or only invocations.
 

### 2016-04-10 paper/paper.tex
Minor edits.

@@
 We are then able to track the additions and deletions of declarations and invocations for each type.
-Furthermore, we can track each author’s type usage over time.
+Furthermore, we can track each author's type usage over time.
 Using this tool we examine specific cases in well-known repositories where our tool reveals interesting and useful information.
@@
 \caption{GitHub graph showing code frequency.
-Note that the graph was truncated to better show the details.
-Thus the changes that occur in mid 2015 are much larger than appear.
-The additions reach over 400k and the deletions go over 300k.}
+Note that the graph was truncated to better show the details, thus the changes that occur in mid 2015 are much larger than shown.
+The additions reach over 400k and the deletions are over 300k.}
 \label{fig:GH_code_freq}
@@
 With more changes to peruse, it becomes more difficult to see how a project is evolving over time since there is so much information to look through.
-Most hosting services for Git repositories provide some basic tools for analysing projects over time~\cite{github-graphs,bitbucket-graphs}; but, as we will show, these tools often lack the kind of detail to gather insightful analysis.
+Most hosting services for Git repositories provide some basic tools for analyzing projects over time~\cite{github-graphs,bitbucket-graphs}; however, as we will show, these tools often lack the kind of detail to gather insightful analysis.
 While commits may store individual changes to lines of code, they represent the change in \emph{structure} of the software.
@@
 
-In this paper, we propose a method that analyses the changes in type usage and method invocations over time, as well as tool to visualize this analysis, which we call TypeV.
+In this paper, we propose a method that analyses the changes in type usage and method invocations over time, as well as a tool to visualize this analysis, which we call TypeV.
 We claim that type usage offers more insight than line-based, file-based, and commit-based methods of visualization and analysis.
 Of the services we have surveyed, none of the tools visualize project evolution using granular type analysis per author via commit-by-commit abstract syntax tree differencing.
-For simplicity, we decided to only analyze Java projects, as it allowed us to quickly develop and test our software.
+For simplicity, we analyze only Java projects, as it allowed us to quickly develop and test our software.
 Since our approach described in this paper can be generally applied to any statically-typed language, we plan to extend it to other languages.
@@
 The types in a project represent the functionality available for a developer to work with.
-Java’s syntax encourages the creation of classes, thus we expect substantial Java projects to define many domain-specific types.
+Java's syntax encourages the creation of classes, thus we expect substantial Java projects to define many domain-specific types.
 Different parts of a program use different types and external dependencies.
-When writing or contributing to a project, the author’s choice of which part of the program to edit will in turn affect what types they end up working with --- that is, it will affect the author’s \emph{type coverage}.
+When writing or contributing to a project, the author's choice of which part of the program to edit will in turn affect what types they end up working with --- that is, it will affect the author's \emph{type coverage}.
 
 Type information is useful for seeing what libraries and functions are being used.
-For example when a part of code becomes \emph{deprecated}, continuing its use may lead to design flaws or even security vulnerabilities.
-For example, in Apache Nifi, an earlier version of the software used a compromised key-derivation functions; hence, this class was deprecated, but left in for compatibility with older versions of the software~\cite{nifi}.
+For example, when a part of code becomes \emph{deprecated}, continuing its use may lead to design flaws or even security vulnerabilities.
+In Apache Nifi, an earlier version of the software used a compromised key-derivation functions; hence, this class was deprecated, but left in for compatibility with older versions of the software~\cite{nifi}.
 Viewing the type information of a change can show when usage of deprecated classes are added or removed from a project as well as how many remain over time.
@@
 
-Having tools for analysing repositories based on type usage and method invocation would benefit many different parts of the software development process.
+Having tools for analyzing repositories based on type usage and method invocation would benefit many different parts of the software development process.
 Software management requires a high level understanding of the state of a repository and the changes being made which affect the functionality of the software.
 Tracking developer contributions in a less misleading way is also important for management since important contributions may be small and focused, but nevertheless, significant.
-Line difference (or simply, \emph{diff}) only provide meaningful information when they are analysed by someone with familiarity with the source code being changed.
+Line difference (or simply, \emph{diff}) only provide meaningful information when they are analyzed by someone with familiarity with the source code being changed.
 Providing tools which bring more meaning to line differences would allow people less familiar with the code to understand changes being made.
@@
 We seek to answer the following questions: \\ \\
-\textbf{RQ1:} How can we extract type and method usage over time from repositories \\ \\
+\textbf{RQ1:} How can we extract type and method usage over time from repositories? \\ \\
 \textbf{RQ2:} Does the type information give any interesting or useful insight into the project? \\ \\
@@
 In another paper, Wu \textit{et al.} \cite{wu2004} the Evolutionary Spectrogram is used but with greater control of granularity for files and software subsystems such as directory structure.
-Spectrograms are also used to display the cardinality of a developer’s commits.
+Spectrograms are also used to display the cardinality of a developer's commits.
 The cardinality of a commit is the number of files or subsystems it affects in a system.
@@
 The model consists of three parts: structural elements, code elements and references to program elements.
-''The structural part contains the declarations of the program elements, such as interface, class, variable, method, annotation, and enum declarations.
+According to Spoon's authors: ''The structural part contains the declarations of the program elements, such as interface, class, variable, method, annotation, and enum declarations.
 The code part contains the executable Java code, such as the one found in method bodies.
-The reference part models the references to program elements (for instance a reference to a type)'' \cite{pawlak:hal-01169705} The Spoon model is convenient because it provides the infrastructure for performing AST differencing (described later) while retaining the code elements (such as type information) for analysis afterwards.
+The reference part models the references to program elements (for instance a reference to a type)''~\cite{pawlak:hal-01169705}.
+The Spoon model is convenient because it provides the infrastructure for performing AST differencing (described later) while retaining the code elements (such as type information) for analysis afterwards.
 Spoon also preserves all type and package (library) information—information vital to our semantic analysis.
@@
 \caption{Commits increase in time from bottom up.
-Therefore, ASTs from B and C are diff’d against ASTs in A.
-ASTs from E are diff’d against D.
+Therefore, ASTs from B and C are diff'd against ASTs in A.
+ASTs from E are diff'd against D.
 Since D is a merge commit, its changes are ignored as they are recorded in C and B.}
@@
 \includegraphics[width=\columnwidth]{context}
-\caption{}
+\caption{Shows the pipeline for TypeV. First we list the commits in chronological order. Then we perform AST diffs for each file. Finally, we display the information.}
 \label{fig:context}
@@
 
-\section{The Tool}
+\section{TypeV: The Visualization Tool}
 
@@
 
-The analyzed data would be difficult to analyze if displayed as its raw text form; instead, we created an HTML5 web visualization application for intended for developers, project managers, and researchers to use (Figure~\ref{fig:main-view}).
+The analyzed data would be difficult to analyze if displayed as its raw text form; instead, we created an HTML5 web visualization application which is intended for developers, project managers, and researchers to use (Figure~\ref{fig:main-view}).
+The source code is available online on GitHub,\footnote{\url{https://github.com/mdfeist/TypeV}}.
 
@@
 The main view displays a series of stacked bar graphs which show changes to types over a period of time in a project (Figure~\ref{fig:main-view}).
-The blue and orange portions account for the number of additions and deletions in a similar fashion to GitHub’s code frequency graph (Figure~\ref{fig:GH_code_freq}).
+The blue and orange portions account for the number of additions and deletions in a similar fashion to GitHub's code frequency graph (Figure~\ref{fig:GH_code_freq}).
 The height of each bar shows the number of changes to a type in a smaller unit of time.
@@
 This includes the fully-qualified name of the type, the number of commits, the number of authors, the number of additions, the number of deletions, and the start and end date of that time period.
-If the user clicks on the bar, they are given more detailed statistics which list the authors who made the commits and each commit message (Figure~\ref{fig:typev-deets}).
+If the user clicks on the bar, they are given more detailed statistics which list the authors who made the commits and each commit message.
 
-One issue with analyzing individual authors is that authors can commit from a variety of different computers which can have different emails and usernames attached to the git configuration.
-This means in the git log one author might appear as several different authors.
+One issue with analyzing individual authors is that authors can commit from a variety of different computers which can have different emails and usernames attached to the Git configuration.
+This means in the \texttt{git log} one author might appear as several different users.
 There is no perfect way to deal with this problem; however, we allow the user of our software to select and view multiple users at a time.
@@
 % use section* for acknowledgment
-\section*{Acknowledgment}
-
-The authors would like to thank...
+%\section*{Acknowledgment}
+%The authors would like to thank...
 

### 2016-04-10 paper/paper.tex
Update tex.

@@
 
+\begin{figure}[!h]
+\centering
+\includegraphics[width=\columnwidth]{context}
+\caption{Shows the pipeline for TypeV. First we list the commits in chronological order. Then we perform AST diffs for each file. Finally, we display the information.}
+\label{fig:context}
+\end{figure}
+
 The core of our analysis is calculating the differences between abstract syntax trees of subsequent commits (Figure~\ref{fig:context}).
@@
 The Java ASTs in our analysis were generated using Spoon~\cite{pawlak:hal-01169705}.
-Spoon is a tool for transforming and analysing Java source code.
+Spoon is a tool for transforming and analyzing Java source code.
 It breaks code up into a meta-model including the AST itself.
@@
 The Spoon model is convenient because it provides the infrastructure for performing AST differencing (described later) while retaining the code elements (such as type information) for analysis afterwards.
-Spoon also preserves all type and package (library) information—information vital to our semantic analysis.
+Spoon also preserves all type and package (library) information, information vital to our semantic analysis.
 
@@
 
-Once we have the generated ASTs, we were able to traverse through the trees to count the number of added or deleted \emph{declarations} and \emph{invocations} for a given type.
+Once we generated ASTs, we were able to traverse through the trees to count the number of added or deleted \emph{declarations} and \emph{invocations} for a given type.
 
@@
 
+This states that two declarations of type \texttt{java.net.URI} were added, and two declarations of type \texttt{java.lang.String} were deleted.
+Note the fully qualified type names.
+The source text may not literally have the characters \texttt{java.net.URI}, but Spoon was able to infer from the import statements in the Java source text that references to ''URI'' resolve to the fully-qualified type name \texttt{java.net.URI}.
+Thus, TypeV is able to keep track of the actual class being used and is not confused when two classes in separate packages have the same name.
+
 \begin{figure}[!h]
@@
 
-This states that two declarations of type \texttt{java.net.URI} were added, and two declarations of type \texttt{java.lang.String} were deleted.
-Note the fully qualified type names: the source text may not literally have the characters \texttt{java.net.URI}, but Spoon was able to infer from the import statements in the Java source text that references to ''URI'' resolve to the fully-qualified type name \texttt{java.net.URI}.
-Thus, TypeV is able to keep track of the actual class being used and is not confused when two classes in separate packages have the same name.
-
-In order to gather all the changes for a given repository over time we generated ASTs for each commit and compared the difference.
+In order to gather all the changes for a given repository over time, we generated ASTs for each commit and compared the difference.
 We did this by running \texttt{git log} over the Git repository and listing all the commit IDs in chronological order (Figure~\ref{fig:commits}).
-Each commit and its parent were then compared, using AST differencing as described above.
-There are a few edges cases, however: If a commit has no parent, then every declaration or invocation in the commit was treated as an addition.
+Each commit and its parent were then compared using AST differencing as described above.
+There are a few edge cases; however, if a commit has no parent, then every declaration or invocation in the commit was treated as an addition.
 Since we compared each commit to its parent, as reported by Git, we were able to account for any branching that happened outside the primary branch of the repository (usually named \texttt{master}).
@@
 Therefore, we ignored merge commits since all changes in the merge should already have been accounted for in previous commits.
-
-\begin{figure}[!h]
-\centering
-\includegraphics[width=\columnwidth]{context}
-\caption{Shows the pipeline for TypeV. First we list the commits in chronological order. Then we perform AST diffs for each file. Finally, we display the information.}
-\label{fig:context}
-\end{figure}
 
@@
 
-The analyzed data would be difficult to analyze if displayed as its raw text form; instead, we created an HTML5 web visualization application which is intended for developers, project managers, and researchers to use (Figure~\ref{fig:main-view}).
-The source code is available online on GitHub,\footnote{\url{https://github.com/mdfeist/TypeV}}.
+The data would be difficult to analyze if displayed as its raw text form; instead, we created an HTML5 web visualization application which is intended for developers, project managers, and researchers to use (Figure~\ref{fig:main-view}).
+The source code for TypeV is available online on GitHub,\footnote{\url{https://github.com/mdfeist/TypeV}}.
 
-TypeV provide a variety of views to help organize the data and provide a finer level of detail.
+TypeV provide a variety of views to help organize the data and provide different levels of detail.
 The main view displays a series of stacked bar graphs which show changes to types over a period of time in a project (Figure~\ref{fig:main-view}).
-The blue and orange portions account for the number of additions and deletions in a similar fashion to GitHub's code frequency graph (Figure~\ref{fig:GH_code_freq}).
-The height of each bar shows the number of changes to a type in a smaller unit of time.
-This is calculated by multiplying the bar height by the number of changes during that time period divided by the maximum number of changes attested for this type.
+The blue and orange portions account for the number of additions and deletions in a similar fashion to GitHub’s code frequency graph (Figure~\ref{fig:GH_code_freq}).
+The height of each bar shows the number of changes to a type.
+The width of each bar represents a set time period and can be changed to an hour, day, week, or month.
 A bar which fills the row vertically is the period with the largest number of changes to that type out of the visible time periods.
 The other bars in the row are proportional to that maximum number.
-The width of each bar represents a set time period and can be changed to an hour, day, week, or month.
+The bar height is calculated by multiplying the maximum bar height by the number of changes during that time period divided by the maximum number of changes attested for this type.
 The rows are sorted so that the graph for the most used types appear the top and are in descending order.

### 2016-04-11 paper/paper.tex
Add bookkeeper figure and update tex.

@@
 
-Consider tools easily available to an open source software developer: GitHub graphs~\cite{github-graphs}.
-Since GitHub tracks contributions based solely on number of commits, and numbers of lines changes, data may be lacking, or even misleading.
+Consider tools easily available to an open source software developer; for example GitHub graphs (see Figure~\ref{fig:GH_code_freq})~\cite{github-graphs}.
+Since GitHub tracks contributions based solely on number of commits and numbers of lines changes, data may be lacking, or even misleading.
 In the case of the Antlr4 project (Figure~\ref{fig:GH_code_freq}), the graph creates more questions than it answers.
 What happened at those big spikes of line changes in the graph?
-One of the spikes is so large, we had to truncate the graph to be able to show it.
+One of the spikes is so large, we had to truncate the graph to be able to show details of other changes.
 How did these large changes affect the code base?
 Who committed these large changes?
-Was it a team of contributors or a one particularly prolific programmer?
-Plotting the number of lines changed may be misleading, because a huge spike in additions may be caused a number of reasons, such as the inclusion of automatically-generated documentation, the inclusion of an external library, or perhaps even a huge refactoring effort.
+Was it a team of contributors or one particularly prolific programmer?
+Plotting the number of lines changed may be misleading.
+A huge spike in additions may be caused by numerous reasons, such as the inclusion of automatically-generated documentation, the inclusion of an external library, or perhaps even a huge refactoring effort.
 
@@
 Therefore, we ignored merge commits since all changes in the merge should already have been accounted for in previous commits.
+Thus, any branch that has been merged into the primary branch would have been analyzed and included by TypeV.
 
@@
 \includegraphics[width=\textwidth]{Screenshot_antlr4}
-\caption{}
+\caption{Screenshot showing the main view for TypeV. The data being displayed is from the antlr4 repository.}
 \label{fig:main-view}
@@
 
-When a user mouses over a specific bar in the graph they are given simple statistics for the type and time period at that location.
-This includes the fully-qualified name of the type, the number of commits, the number of authors, the number of additions, the number of deletions, and the start and end date of that time period.
+When a user mouses over a specific bar in the graph they are given information for the type and time period at that location.
+The information includes the fully-qualified name of the type, the number of commits, the number of authors, the number of additions, the number of deletions, and the start and end date of that time period.
 If the user clicks on the bar, they are given more detailed statistics which list the authors who made the commits and each commit message.
@@
 One issue with analyzing individual authors is that authors can commit from a variety of different computers which can have different emails and usernames attached to the Git configuration.
-This means in the \texttt{git log} one author might appear as several different users.
-There is no perfect way to deal with this problem; however, we allow the user of our software to select and view multiple users at a time.
-Thus allowing the user to view changes from an individual or group of developers.
+This may result in one author appearing as several different users in the \texttt{git log}.
+There is no perfect way to deal with this problem; however, we allow the user of our software to select and view multiple users at a time, thus allowing the user to view changes from an individual or group of developers.
 
 To allow for a more detailed analysis of a specific component of the repository we added some basic filtering tools.
-The first is that a user can filter by type name.
-The second allows the user to view changes that fall within a specific start and end date.
-The third allows them to specify the time span represented by each bar: either hour, day, week, or month.
-Finally, they can choose to see only types declarations, types used in declarations and invocations, or only invocations.
+First a user can filter by type name.
+Second the user can view changes that fall within a specific start and end date.
+Third the user can specify the time span represented by each bar: hour, day, week, or month.
+Finally, the user can choose to see only declarations of types, types used in declarations and invocations, or only invocations.
 
@@
 
-In the following sections, we present case studies demonstrating the extra insight that the TypeV method makes available.
+In the following sections we present case studies demonstrating the extra insight that the TypeV method makes available.
 
@@
 \includegraphics[width=\columnwidth]{ParserATNSimulator}
-\caption{Snapshot of TypeV showing changes to ParserATNSimulator in antlr4. Date ranges from approximately the begining of 2011 to the end of April in 2012.}
+\caption{Snapshot of TypeV showing changes to ParserATNSimulator in antlr4. Date ranges from approximately the begining of 2011 to the end of April, 2012.}
 \label{fig:parser1}
@@
 
-If we tried to do the same analyses with the GitHub tools it would be significantly more difficult to achieve the same conclusion (Figure~\ref{fig:parser2}).
+If we tried to do the same analysis with the GitHub tools it would be significantly more difficult to achieve the conclusion achieved using TypeV (Figure~\ref{fig:parser2}).
 First if we only look at line changes this change to ParserATNSimulator is overwhelmed by other changes and hardly shows up on the GitHub tools.
-Then if we did notice an increase in activity we would not know what types were being worked on.
-Finally we could not quickly look at the commit messages for that specific time and type.
+If we did notice an increase in activity we would not know what types were being worked on.
+Finally, we could not quickly look at the commit messages for that specific time and type if we used the GitHub tools.
 
-At certain times we can can see major increases across all types.
-This shows us when major changes to the code were being done versus code maintenance.
-Usually this happens when a large amount of code is being pulled in from another location.
+Using TypeV we can can see that at certain times there were major changes across all types. This shows us when major changes to the code were being done versus code maintenance.
 
 \subsection{Apache Bookkeeper}
+
+\begin{figure}[!ht]
+\centering
+\includegraphics[width=\columnwidth]{bookkeeper_dep}
+\caption{Shows deprecated invocation getLedgerManagerType() and replacement invocation getLedgerManagerFactoryClass() for Apache Bookkeeper.}
+\label{fig:bookkeeper-depr}
+\end{figure}
 
@@
 Apache Bookkeeper deprecated the method AbstractConfiguration.getLedgerManagerType() in release 4.2.0\footnote{\url{https://bookkeeper.apache.org/docs/r4.2.0/apidocs/deprecated-list.html}} and replaced it with AbstractConfiguration.getLedgerManagerFactoryClass().
-If we look at getLedgerManagerType() we can see at the time of the start of the deprecation all instances were removed and getLedgerManagerFactoryClass() was added.
+If we look at getLedgerManagerType() we can see at the time of the start of the deprecation all instances were removed and getLedgerManagerFactoryClass() was added (Figure~\ref{fig:bookkeeper-depr}).
 After the deprecation, we see no more changes to getLedgerManagerType() and only some additions of getLedgerManagerFactoryClass().
@@
 \renewcommand{\arraystretch}{1.3}
-\caption{Comparision of Type Coverage and File Coverage in Projects}
+\caption{Comparison of Type Coverage and File Coverage in Projects}
 \label{tab:summary}
@@
 \hline
-\bfseries Project & \bfseries Wilcoxon & \bfseries Pearson (lin) cor & \bfseries Spearman (rank) cor \\
+\bfseries Project & \bfseries Wilcoxon& \bfseries Pearson (linear) & \bfseries Spearman (rank) \\
+& p-value & Correlation & Correlation \\
 \hline
@@
 
-Currently for generic types we treat every instance of a generic as a different type.
-So for example if a user declares a list of strings and a list of integers these will appear as two different types.
-However, if they declare multiple instances of a list of strings then these instances will count toward the same type.
-In the future we would nice to come up with a way to better handle generic types.
+Currently, we treat every instance of a generic type as a different type.
+For example, considered the \texttt{List<T>} generic class in Java.
+If a user declares a List of Strings and later declares a List of Integers these will appear as two different types in TypeV.
+However, if they declare multiple instances of a List of Strings then these instances will count toward the same type.
+In the future, we want to work on a better way to track and represent generic types.
 
-Currently we are only looking at declarations and invocations, but it would be nice to also view what classes these changes were made to.
+Currently we are only looking at declaration of types and invocation of methods; it would be useful to also view what classes changes are being made to.
+Enabling this ability would also enable analysis of self-invocations.
+
+The methodology used in TypeV is only well-defined for statically-typed languages such as Haskell and Java.
+Dynamically-typed languages like Erlang and JavaScript.
+However, type annotation systems such as TypEr for Erlang~\cite{typer} or statically-typed language dialects such as TypeScript for JavaScript may prove to be sufficient to enable type analysis as we have demonstrated for Java.
 
 \section{Conclusion}
-The conclusion goes here.
+
+In this paper, we introduced TypeV, a source code repository analysis and visualization tool.
+Its use of commit-by-commit abstract syntax tree differencing enable for more insightful visualization than tools that merely use commit counts, line counts, and file changes.
+We answered our research questions as thus:
+
+RQ1: Using abstract syntax trees, we can determine the types affected by each commits in a repository. Type information is not directly available through regular line diffs recorded by Git.
+
+RQ2: The evolution of a project and developer activity can be seen through the changes to types in a project.
+Significant events can be observed with greater understanding by seeing the changes to types in addition to the existing commit information.
+
+RQ3: Type coverage and file coverage are found to be highly correlated but do not always have the same distribution or a one to one ratio as might be expected in Java.
 

### 2016-04-11 server.py
Fix redirection and server paths.

@@ -424,10 +424,9 @@ def send_css(path):
 
 @app.route('/')
 def root():
-    #return app.send_static_file('index.html')
-    return redirect('/projects')
+    return redirect('/projects/')
 
-@app.route('/projects')
+@app.route('/projects/')
 def dashboard():
     return render_template('projects.html', projects=projects)
 

### 2016-04-12 ast.sh
Change all instances of CMPUT664_PROJECT -> TYPEV_PATH.

@@ -15,7 +15,7 @@ then
     echo "#FILE2 | $3"
     echo "#TYPE | $1"
 
-	$TIMEOUT java -cp $CMPUT664_PROJECT/gumtree-spoon-ast-diff/target/gumtree-spoon-ast-diff-0.0.3-SNAPSHOT-jar-with-dependencies.jar fr.inria.sacha.spoon.diffSpoon.DiffSpoonImpl $1 $2 $3
+	$TIMEOUT java -cp $TYPEV_PATH/gumtree-spoon-ast-diff/target/gumtree-spoon-ast-diff-0.0.3-SNAPSHOT-jar-with-dependencies.jar fr.inria.sacha.spoon.diffSpoon.DiffSpoonImpl $1 $2 $3
 
     echo "#STATS_END"
 fi

### 2016-04-12 get_stats.sh

@@ -3,5 +3,5 @@ echo $1
 cd $1
 
 for d in */ ; do
-	bash $CMPUT664_PROJECT/project_stats.sh $d
+	bash $TYPEV_PATH/project_stats.sh $d
 done

### 2016-04-12 tar_tests.sh
Consolidate tar and untar tests into Makefile. Addresses #16.

@@ -1,3 +1,0 @@
-#!/bin/bash
-
-tar -zcvf tests.tar.gz tests/

### 2016-04-12 untar_tests.sh

@@ -1,4 +1,0 @@
-#!/bin/bash
-
-rm tests
-tar -zxvf tests.tar.gz

### 2016-04-12 project_stats.sh
Add basic documentation for how to build and use the tools.

@@ -10,7 +10,7 @@ if [ ! -d "$GIT" ]; then
 	echo "#ERROR | Not git repository."
 	echo "#PROJECT_END"
 	cd ..
-	exit
+	exit -1
 fi
 
 # Create a temporary directory filled with all of this junk.

### 2016-04-15 paper/document.tex
Move document body to document.tex.

--- 2016-04-11 paper/paper.tex
@@
-% Copied from:
-% http://mirror.its.dal.ca/ctan/macros/latex/contrib/IEEEtran/bare_conf.tex
-\documentclass[conference]{IEEEtran}
-% Some Computer Society conferences also require the compsoc mode option,
-% but others use the standard conference format.
-%
-% If IEEEtran.cls has not been installed into the LaTeX system files,
-% manually specify the path to it like:
-% \documentclass[conference]{../sty/IEEEtran}
-
-\usepackage{enumitem}
-\setlist[description]{leftmargin=1cm,labelindent=\parindent}
-
-\usepackage{changepage}
-\usepackage{url}
-\usepackage{breakurl}
-\usepackage[breaklinks]{hyperref}
-\def\UrlBreaks{\do\/\do-}
-
-% Some very useful LaTeX packages include:
-% (uncomment the ones you want to load)
-
-% *** MISC UTILITY PACKAGES ***
-%
-%\usepackage{ifpdf}
-% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
-% compilation based on whether the output is pdf or dvi.
-% usage:
-% \ifpdf
-%   % pdf code
-% \else
-%   % dvi code
-% \fi
-% The latest version of ifpdf.sty can be obtained from:
-% http://www.ctan.org/pkg/ifpdf
-% Also, note that IEEEtran.cls V1.7 and later provides a builtin
-% \ifCLASSINFOpdf conditional that works the same way.
-% When switching from latex to pdflatex and vice-versa, the compiler may
-% have to be run twice to clear warning/error messages.
-
-% *** CITATION PACKAGES ***
-%
-%\usepackage{cite}
-% cite.sty was written by Donald Arseneau
-% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
-% \cite{} output to follow that of the IEEE. Loading the cite package will
-% result in citation numbers being automatically sorted and properly
-% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
-% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
-% \cite will automatically add leading space, if needed. Use cite.sty's
-% noadjust option (cite.sty V3.8 and later) if you want to turn this off
-% such as if a citation ever needs to be enclosed in parenthesis.
-% cite.sty is already installed on most LaTeX systems. Be sure and use
-% version 5.0 (2009-03-20) and later if using hyperref.sty.
-% The latest version can be obtained at:
-% http://www.ctan.org/pkg/cite
-% The documentation is contained in the cite.sty file itself.
-
-% *** GRAPHICS RELATED PACKAGES ***
-%
-\ifCLASSINFOpdf
-  \usepackage[pdftex]{graphicx}
-  \usepackage{epstopdf}
-  % declare the path(s) where your graphic files are
- \graphicspath{{diagrams/}}
-  % and their extensions so you won't have to specify these with
-  % every instance of \includegraphics
-  \DeclareGraphicsExtensions{.pdf,.jpeg,.png,.eps}
-\else
-  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
-  % will default to the driver specified in the system graphics.cfg if no
-  % driver is specified.
-  \usepackage[dvips]{graphicx}
-  % declare the path(s) where your graphic files are
-  \graphicspath{{diagrams/}}
-  % and their extensions so you won't have to specify these with
-  % every instance of \includegraphics
-  \DeclareGraphicsExtensions{.eps}
-\fi
-% graphicx was written by David Carlisle and Sebastian Rahtz. It is
-% required if you want graphics, photos, etc. graphicx.sty is already
-% installed on most LaTeX systems. The latest version and documentation
-% can be obtained at:
-% http://www.ctan.org/pkg/graphicx
-% Another good source of documentation is "Using Imported Graphics in
-% LaTeX2e" by Keith Reckdahl which can be found at:
-% http://www.ctan.org/pkg/epslatex
-%
-% latex, and pdflatex in dvi mode, support graphics in encapsulated
-% postscript (.eps) format. pdflatex in pdf mode supports graphics
-% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
-% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
-% not a bitmapped formats (.jpeg, .png). The IEEE frowns on bitmapped formats
-% which can result in "jaggedy"/blurry rendering of lines and letters as
-% well as large increases in file sizes.
-%
-% You can find documentation about the pdfTeX application at:
-% http://www.tug.org/applications/pdftex
-
-% *** MATH PACKAGES ***
-%
-%\usepackage{amsmath}
-% A popular package from the American Mathematical Society that provides
-% many useful and powerful commands for dealing with mathematics.
-%
-% Note that the amsmath package sets \interdisplaylinepenalty to 10000
-% thus preventing page breaks from occurring within multiline equations. Use:
-%\interdisplaylinepenalty=2500
-% after loading amsmath to restore such page breaks as IEEEtran.cls normally
-% does. amsmath.sty is already installed on most LaTeX systems. The latest
-% version and documentation can be obtained at:
-% http://www.ctan.org/pkg/amsmath
-
-% *** SPECIALIZED LIST PACKAGES ***
-%
-%\usepackage{algorithmic}
-% algorithmic.sty was written by Peter Williams and Rogerio Brito.
-% This package provides an algorithmic environment fo describing algorithms.
-% You can use the algorithmic environment in-text or within a figure
-% environment to provide for a floating algorithm. Do NOT use the algorithm
-% floating environment provided by algorithm.sty (by the same authors) or
-% algorithm2e.sty (by Christophe Fiorio) as the IEEE does not use dedicated
-% algorithm float types and packages that provide these will not provide
-% correct IEEE style captions. The latest version and documentation of
-% algorithmic.sty can be obtained at:
-% http://www.ctan.org/pkg/algorithms
-% Also of interest may be the (relatively newer and more customizable)
-% algorithmicx.sty package by Szasz Janos:
-% http://www.ctan.org/pkg/algorithmicx
-
-% *** ALIGNMENT PACKAGES ***
-%
-%\usepackage{array}
-% Frank Mittelbach's and David Carlisle's array.sty patches and improves
-% the standard LaTeX2e array and tabular environments to provide better
-% appearance and additional user controls. As the default LaTeX2e table
-% generation code is lacking to the point of almost being broken with
-% respect to the quality of the end results, all users are strongly
-% advised to use an enhanced (at the very least that provided by array.sty)
-% set of table tools. array.sty is already installed on most systems. The
-% latest version and documentation can be obtained at:
-% http://www.ctan.org/pkg/array
-
-% IEEEtran contains the IEEEeqnarray family of commands that can be used to
-% generate multiline equations as well as matrices, tables, etc., of high
-% quality.
-
-% *** SUBFIGURE PACKAGES ***
-%\ifCLASSOPTIONcompsoc
-%  \usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
-%\else
-%  \usepackage[caption=false,font=footnotesize]{subfig}
-%\fi
-% subfig.sty, written by Steven Douglas Cochran, is the modern replacement
-% for subfigure.sty, the latter of which is no longer maintained and is
-% incompatible with some LaTeX packages including fixltx2e. However,
-% subfig.sty requires and automatically loads Axel Sommerfeldt's caption.sty
-% which will override IEEEtran.cls' handling of captions and this will result
-% in non-IEEE style figure/table captions. To prevent this problem, be sure
-% and invoke subfig.sty's "caption=false" package option (available since
-% subfig.sty version 1.3, 2005/06/28) as this is will preserve IEEEtran.cls
-% handling of captions.
-% Note that the Computer Society format requires a larger sans serif font
-% than the serif footnote size font used in traditional IEEE formatting
-% and thus the need to invoke different subfig.sty package options depending
-% on whether compsoc mode has been enabled.
-%
-% The latest version and documentation of subfig.sty can be obtained at:
-% http://www.ctan.org/pkg/subfig
-
-% *** FLOAT PACKAGES ***
-%
-%\usepackage{fixltx2e}
-% fixltx2e, the successor to the earlier fix2col.sty, was written by
-% Frank Mittelbach and David Carlisle. This package corrects a few problems
-% in the LaTeX2e kernel, the most notable of which is that in current
-% LaTeX2e releases, the ordering of single and double column floats is not
-% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
-% single column figure to be placed prior to an earlier double column
-% figure.
-% Be aware that LaTeX2e kernels dated 2015 and later have fixltx2e.sty's
-% corrections already built into the system in which case a warning will
-% be issued if an attempt is made to load fixltx2e.sty as it is no longer
-% needed.
-% The latest version and documentation can be found at:
-% http://www.ctan.org/pkg/fixltx2e
-
-%\usepackage{stfloats}
-% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
-% the ability to do double column floats at the bottom of the page as well
-% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
-% LaTeX2e). It also provides a command:
-%\fnbelowfloat
-% to enable the placement of footnotes below bottom floats (the standard
-% LaTeX2e kernel puts them above bottom floats). This is an invasive package
-% which rewrites many portions of the LaTeX2e float routines. It may not work
-% with other packages that modify the LaTeX2e float routines. The latest
-% version and documentation can be obtained at:
-% http://www.ctan.org/pkg/stfloats
-% Do not use the stfloats baselinefloat ability as the IEEE does not allow
-% \baselineskip to stretch. Authors submitting work to the IEEE should note
-% that the IEEE rarely uses double column equations and that authors should try
-% to avoid such use. Do not be tempted to use the cuted.sty or midfloat.sty
-% packages (also by Sigitas Tolusis) as the IEEE does not format its papers in
-% such ways.
-% Do not attempt to use stfloats with fixltx2e as they are incompatible.
-% Instead, use Morten Hogholm'a dblfloatfix which combines the features
-% of both fixltx2e and stfloats:
-%
-% \usepackage{dblfloatfix}
-% The latest version can be found at:
-% http://www.ctan.org/pkg/dblfloatfix
-
-% *** PDF, URL AND HYPERLINK PACKAGES ***
-
-\usepackage{url}
-% url.sty was written by Donald Arseneau. It provides better support for
-% handling and breaking URLs. url.sty is already installed on most LaTeX
-% systems. The latest version and documentation can be obtained at:
-% http://www.ctan.org/pkg/url
-% Basically, \url{my_url_here}.
-
-% *** Do not adjust lengths that control margins, column widths, etc. ***
-% *** Do not use packages that alter fonts (such as pslatex).         ***
-% There should be no need to do such things with IEEEtran.cls V1.6 and later.
-% (Unless specifically asked to do so by the journal or conference you plan
-% to submit to, of course. )
-
-% correct bad hyphenation here
-\hyphenation{op-tical net-works semi-conduc-tor}
-
-\begin{document}
-%
-% paper title
-% Titles are generally capitalized except for words such as a, an, and, as,
-% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
-% not capitalized unless they are the first or last word of the title.
-% Linebreaks \\ can be used within to get better formatting as desired.
-% Do not put math or special symbols in the title.
-\title{Visualizing Project Evolution Through Abstract Syntax Tree Analysis}
-
-% author names and affiliations
-% use a multiple column layout for up to three different
-% affiliations
-%\author{\IEEEauthorblockN{Michael D. Feist \and Eddie Antonio Santos \and
-%  Ian Watts \and Abram Hindle}
-%\IEEEauthorblockA{Department of Computing Science\\
-%University of Alberta\\
-%Edmonton, Canada\\
-%\{mdfeist,easantos,watts1,hindle1\}@ualberta.ca}}
-
-% conference papers do not typically use \thanks and this command
-% is locked out in conference mode. If really needed, such as for
-% the acknowledgment of grants, issue a \IEEEoverridecommandlockouts
-% after \documentclass
-
-% for over three affiliations, or if they all won't fit within the width
-% of the page, use this alternative format:
-%
-\author{\IEEEauthorblockN{Michael D. Feist,
-Eddie Antonio Santos,
-Ian Watts,
-Abram Hindle}
-\IEEEauthorblockA{Department of Computing Science\\
-University of Alberta\\
-Edmonton, Canada\\
-\{mdfeist,easantos,watts1,hindle1\}@ualberta.ca}}
-
-% use for special paper notices
-%\IEEEspecialpapernotice{(Invited Paper)}
-
-% make the title area
-\maketitle
-
 % As a general rule, do not put math, special symbols or citations
@@
 %The authors would like to thank...
-
-% trigger a \newpage just before the given reference
-% number - used to balance the columns on the last page
-% adjust value as needed - may need to be readjusted if
-% the document is modified later
-%\IEEEtriggeratref{8}
-% The "triggered" command can be changed if desired:
-%\IEEEtriggercmd{\enlargethispage{-5in}}
-
-% references section
-
-% can use a bibliography generated by BibTeX as a .bbl file
-% BibTeX documentation can be easily obtained at:
-% http://mirror.ctan.org/biblio/bibtex/contrib/doc/
-% The IEEEtran BibTeX style support page is at:
-% http://www.michaelshell.org/tex/ieeetran/bibtex/
-%\nocite{*}
-\bibliographystyle{IEEEtran}
-% argument is your BibTeX string definitions and bibliography database(s)
-\bibliography{bibliography,autogen}
-
-% that's all folks
-\end{document}

### 2016-04-15 paper/paper.tex
Move document body to document.tex.

--- 2016-04-10 paper/paper.tex
@@
 \usepackage{changepage}
-\usepackage{url}
-\usepackage{breakurl}
-\usepackage[breaklinks]{hyperref}
-\def\UrlBreaks{\do\/\do-}
+\usepackage{hyperref}
+%\usepackage[hyphenbreaks]{breakurl}
+%\def\UrlBreaks{\do\/\do-}
 
@@
 \else
+  \GenericError{}
+  {This document will not compile on DVI devices!}
+  {It contains images unsupported by DVI (namely, PNG images).}
+  {}
+
   % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
@@
 
-% As a general rule, do not put math, special symbols or citations
-% in the abstract
-\begin{abstract}
-What is a developer's contribution to a repository?
-By only counting commits and number of lines changed, existing tools that analyze source code repositories fall short on showing the effective contributions made by each developer.
-When many commits are viewed as a group, the details are lost.
-Commit information can be misleading since lines of code give no indication of what was actually being worked on without careful examination of the changed code.
-Providing a semantic view of this information could provide deeper insights into how software projects evolve since changes to design and features are not clearly visible from line changes alone.
-We present TypeV: a method for visualizing Java source code repositories.
-Instead of counting line changes in a commit we extract detailed type information over time by using the differences between abstract syntax trees (ASTs).
-We are then able to track the additions and deletions of declarations and invocations for each type.
-Furthermore, we can track each author's type usage over time.
-Using this tool we examine specific cases in well-known repositories where our tool reveals interesting and useful information.
-We then compare type coverage information from the AST compared to file coverage to determine if unique information is provided by type information.
-\end{abstract}
-
-% no keywords
-
-% For peer review papers, you can put extra information on the cover
-% page as needed:
-% \ifCLASSOPTIONpeerreview
-% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
-% \fi
-%
-% For peerreview papers, this IEEEtran command inserts a page break and
-% creates the second title. It will be ignored for other modes.
-\IEEEpeerreviewmaketitle
-
-\begin{figure*}[!t]
-\centering
-\includegraphics[width=\textwidth]{antlr-3-code-freq-edit}
-\caption{GitHub graph showing code frequency.
-Note that the graph was truncated to better show the details, thus the changes that occur in mid 2015 are much larger than shown.
-The additions reach over 400k and the deletions are over 300k.}
-\label{fig:GH_code_freq}
-\end{figure*}
-
-\section{Introduction}
-
-Source code is often represented as text in a source code file.
-This text is then parsed by a compiler or interpreter with the intent of eventually executing the instructions within the source code file.
-While programmers may type in individual characters, their intention is at a much higher level.
-Their intent is written in tokens that represent statements, identifiers and types that are used to solve their problems.
-Programmers solve problems in a way that is readable to other developers and executable by a computer.
-The programmer rarely operates at the character level, or even the line level, so why should our analyses of source code operate at granularities unfamiliar to developers?
-
-The history and evolution of a software project is often tracked by version control systems, such as Git, in a programming language agnostic way.
-Updates to source code, documentation, and tests are recorded in the form of commits, which show the differences in lines between each subsequent revision of the software.
-Version control systems become vital for large projects since individual changes would be impossible to track manually.
-Large projects have thousands of individual changes made by many different developers.
-With more changes to peruse, it becomes more difficult to see how a project is evolving over time since there is so much information to look through.
-Most hosting services for Git repositories provide some basic tools for analyzing projects over time~\cite{github-graphs,bitbucket-graphs}; however, as we will show, these tools often lack the kind of detail to gather insightful analysis.
-While commits may store individual changes to lines of code, they represent the change in \emph{structure} of the software.
-By visualizing lines changed, or files changed, the current state-of-the-art for software project visualization rarely capture the \emph{meaning} behind each commit; lines of code alone does not describe how a repository changed at a particular point in time.
-Thus, information regarding contributions from developers is often misleading since the number of commits and lines of code changed does not necessarily reflect meaningful changes to a repository~\cite{robles2014}.
-
-Consider tools easily available to an open source software developer: GitHub graphs~\cite{github-graphs}.
-Since GitHub tracks contributions based solely on number of commits, and numbers of lines changes, data may be lacking, or even misleading.
-In the case of the Antlr4 project (Figure~\ref{fig:GH_code_freq}), the graph creates more questions than it answers.
-What happened at those big spikes of line changes in the graph?
-One of the spikes is so large, we had to truncate the graph to be able to show it.
-How did these large changes affect the code base?
-Who committed these large changes?
-Was it a team of contributors or a one particularly prolific programmer?
-Plotting the number of lines changed may be misleading, because a huge spike in additions may be caused a number of reasons, such as the inclusion of automatically-generated documentation, the inclusion of an external library, or perhaps even a huge refactoring effort.
-
-In this paper, we propose a method that analyses the changes in type usage and method invocations over time, as well as a tool to visualize this analysis, which we call TypeV.
-We claim that type usage offers more insight than line-based, file-based, and commit-based methods of visualization and analysis.
-Of the services we have surveyed, none of the tools visualize project evolution using granular type analysis per author via commit-by-commit abstract syntax tree differencing.
-For simplicity, we analyze only Java projects, as it allowed us to quickly develop and test our software.
-Since our approach described in this paper can be generally applied to any statically-typed language, we plan to extend it to other languages.
-
-In Java, a type can be a primitive, provided by the language itself, or object references.
-Primitive types are those that are built into the language such as an \texttt{int} or \texttt{char}.
-Object references point to instances of classes.
-Classes are defined in the project itself or through an included library.
-The types in a project represent the functionality available for a developer to work with.
-Java's syntax encourages the creation of classes, thus we expect substantial Java projects to define many domain-specific types.
-Different parts of a program use different types and external dependencies.
-When writing or contributing to a project, the author's choice of which part of the program to edit will in turn affect what types they end up working with --- that is, it will affect the author's \emph{type coverage}.
-
-Type information is useful for seeing what libraries and functions are being used.
-For example, when a part of code becomes \emph{deprecated}, continuing its use may lead to design flaws or even security vulnerabilities.
-In Apache Nifi, an earlier version of the software used a compromised key-derivation functions; hence, this class was deprecated, but left in for compatibility with older versions of the software~\cite{nifi}.
-Viewing the type information of a change can show when usage of deprecated classes are added or removed from a project as well as how many remain over time.
-In a more general sense, we can see when types are being added or removed from the project as well as what types are heavily used throughout the project.
-Even better, we can see \emph{who} uses which types heavily.
-
-Having tools for analyzing repositories based on type usage and method invocation would benefit many different parts of the software development process.
-Software management requires a high level understanding of the state of a repository and the changes being made which affect the functionality of the software.
-Tracking developer contributions in a less misleading way is also important for management since important contributions may be small and focused, but nevertheless, significant.
-Line difference (or simply, \emph{diff}) only provide meaningful information when they are analyzed by someone with familiarity with the source code being changed.
-Providing tools which bring more meaning to line differences would allow people less familiar with the code to understand changes being made.
-When large changes are made that cover many aspects of a project, it would also be useful to see how the change affects the software rather than a several thousand line diff.
-Seeing how a project has changed over time is also vital for management since evaluating the state of completion of software can be very difficult.
-
-In the following sections of the paper we will introduce a tool, TypeV, that we developed that visualizes changes to types used in a project over time.
-TypeV is able to extract type information from a repository by computing commit-by-commit abstract syntax tree differencing, which we describe in section~\ref{sec:methodology}.
-We seek to answer the following questions: \\ \\
-\textbf{RQ1:} How can we extract type and method usage over time from repositories? \\ \\
-\textbf{RQ2:} Does the type information give any interesting or useful insight into the project? \\ \\
-\textbf{RQ3:} How does type coverage compare to file coverage?
-
-\section{Related Work}
-
-Analyzing source code is a popular research topic.
-Most of the focus is on how certain programming languages and features are used.
-However, there is no focus specifically on how types are used throughout a project.
-
-Dyer \textit{et al.} \cite{Dyer:2014:MBA:2568225.2568295} mined AST nodes to study the use of new Java language features over time.
-They found that all new Java features do get used, but not nearly as often as they could have been.
-New features varied greatly in popularity and adoption rate.
-Developers would also modify old code to use new features as they were released.
-The authors also found the most popular features and how teams adopted new features.
-They did not check to see how much a developer used a feature but instead only checked to see if they used it at all.
-
-Grechanik \textit{et al.} \cite{Grechanik:2010:EIL:1852786.1852801} examined the structure of Java programs mined from 2080 programs.
-This paper examined the breakdown of syntactic structures in open source repositories, however it does not consider per author statistics.
-
-Meyerovich and Rabkin \cite{Meyerovich:2013:EAP:2509136.2509515} surveyed developers and examined repositories to learn about language adoption and usage.
-Developers feel that certain language features are more important than others.
-
-Parnin \textit{et al.} \cite{Parnin:2011:JGA:1985441.1985446} mined repositories to see how Java generics have been used in open source projects and found that generic usages were often introduced by a single developer in a project.
-Generics usage was also fairly narrow, with the primary use being collecting and traversing lists of objects.
-
-Wagstrom \textit{et al.} \cite{Patrick:Wagstrom:2012} explored the roles that developers take in a networked, social development environments like GitHub.
-These role were defined as their level of contribution as well as the types of issues they dealt with.
-Developers will take on multiple roles in a project and will sometimes fulfill the same role across different projects.
-
-L\"{a}mmel \textit{et al.} \cite{Lammel:2011:LAA:1982185.1982471} used ASTs to examine the usage of APIs in Java projects.
-They find what APIs are popular as well as if they are used in a framework like manner.
-
-The aforementioned research is trying to learn how developers are using programming languages by looking at many repositories.
-This type of info can be extracted from individual repositories using our new tool.
-Type info may also provide additional information on how developers use languages.
-
-Wu \textit{et al.} \cite{wu2004a} explore punctuated changes to software repositories with a tool that generates an Evolution Spectrograph.
-Each file has a representative row in the chart, with each cell in the row representing a section of time.
-This diagram would show changes to incoming and outgoing dependencies by darkening the cell where the change happened and lightening it over time if no changes were made.
-
-In another paper, Wu \textit{et al.} \cite{wu2004} the Evolutionary Spectrogram is used but with greater control of granularity for files and software subsystems such as directory structure.
-Spectrograms are also used to display the cardinality of a developer's commits.
-The cardinality of a commit is the number of files or subsystems it affects in a system.
-All of the developers changes together show how many changes affect broad sections of the software.
-
-Voinea \textit{et al.} \cite{voinea2006} visualizes CVS repositories using CVSgrab and adding their own enhancements to show changes to files over time.
-Files can be sorted by creation time, activity, or alphabetically and colour coded based on author, file size, file type or by search strings.
-
-Kim \textit{et al.} \cite{Kim:2009:DRS:1555001.1555046} built a tool called Logical Structural Diff (LSdiff) which examines code changes and produces logical rules to represent them.
-Changes which do not match the rules are then pointed out to the developer.
-They also formed a focus group with software engineers who found that LSdiff complements existing tools well and showed promise for discovering inconsistent changes.
-
-The service EasyBi \cite{EasyBi} provides additional visualizations for Git repositories for the purpose of business decisions.
-EasyBi works by taking the git log file and generating charts and graphs from it.
-It tries to solve the issue of summarizing many commits and has greater controls for granularity and authors but still operates on commits and line differences.
-
-The Gitinspector project \cite{Gitinspector} is an open source tool for the statistical analysis of git repositories.
-It provides an advanced timeline view of changes per author and cumulative statistics but again relies upon the line differences to summarize the work of each author.
-
-Wagstrom \textit{et al.} \cite{Patrick:Wagstrom:2012} explored the roles that developers take in a networked, social development environments like GitHub.
-These role were defined as their level of contribution as well as the types of issues they dealt with.
-Developers will take on multiple roles in a project and will sometimes fulfill the same role across different projects.
-
-Lee \textit{et al.} \cite{lee2013} developed a tool for visualizing the branching structure of git repositories. This was used to analyze concurrent workflows in popular open source repositories.
-
-The GitHub tools have the strength that they do not rely on what type of code is being committed.
-It does not matter what language or if errors exist in the commit, GitHub will simply spit out the number of line additions and deletions.
-The issue with this type of analysis is that it does not give a clear breakdown of is being changed only that changes have occurred.
-
-\section{Methodology}
-\label{sec:methodology}
-
-\begin{figure}[!h]
-\centering
-\includegraphics[width=\columnwidth]{context}
-\caption{Shows the pipeline for TypeV. First we list the commits in chronological order. Then we perform AST diffs for each file. Finally, we display the information.}
-\label{fig:context}
-\end{figure}
-
-The core of our analysis is calculating the differences between abstract syntax trees of subsequent commits (Figure~\ref{fig:context}).
-
-An \emph{abstract syntax tree} (AST) is a data structure derived from source code by breaking its syntactic constructs into a tree.
-Each node in the tree represents a syntactically valid chunk of code.
-This captures the structure of the code in an abstract and tractable manner.
-Converting raw source text to an AST is often one of the first steps performed by compilers.
-When a programmer changes code in a file, the difference will appear in the AST.
-This means that the AST of revisions in code repositories can be compared to see what structures a programmer is modifying.
-
-The Java ASTs in our analysis were generated using Spoon~\cite{pawlak:hal-01169705}.
-Spoon is a tool for transforming and analyzing Java source code.
-It breaks code up into a meta-model including the AST itself.
-The model consists of three parts: structural elements, code elements and references to program elements.
-According to Spoon's authors: ''The structural part contains the declarations of the program elements, such as interface, class, variable, method, annotation, and enum declarations.
-The code part contains the executable Java code, such as the one found in method bodies.
-The reference part models the references to program elements (for instance a reference to a type)''~\cite{pawlak:hal-01169705}.
-The Spoon model is convenient because it provides the infrastructure for performing AST differencing (described later) while retaining the code elements (such as type information) for analysis afterwards.
-Spoon also preserves all type and package (library) information, information vital to our semantic analysis.
-
-To determine what has changed between two subsequent commits, we computed their \emph{AST difference}.
-An AST difference (or simply \emph{AST diff}) states which nodes were added and removed from a tree (step two in Figure~\ref{fig:context}).
-For this purpose, we used GumTree~\cite{falleri:hal-01054552}, an algorithm made to compute the difference between two ASTs generated by Spoon.
-For new files and deleted files we did not use GumTree; we were simply able to run Spoon on the files and treated everything as an addition or deletion respectively.
-
-Once we generated ASTs, we were able to traverse through the trees to count the number of added or deleted \emph{declarations} and \emph{invocations} for a given type.
-
-\begin{description}
-\item [declaration] 
-  Stating that use of a field or variable of a given type. For example declaration of a String type variable might look as follows:
-  \\
-  \begin{adjustwidth}{2.5em}{0pt}
-  String myString; \\
-  \end{adjustwidth}
-\item [invocation] 
-  A call site for a method that belongs to a type. For example an invocation of the trim method of a String type might look as follows:
-  \\
-  \begin{adjustwidth}{2.5em}{0pt}
-  myString.trim(); \\
-  \end{adjustwidth}
-\end{description}
-
-An example of what our analysis tool might produce is as follows:
-
-\begin{verbatim}
-#DECLARE | INSERT | java.net.URI | 2
-#DECLARE | DELETE | java.lang.String | 2
-\end{verbatim}
-
-This states that two declarations of type \texttt{java.net.URI} were added, and two declarations of type \texttt{java.lang.String} were deleted.
-Note the fully qualified type names.
-The source text may not literally have the characters \texttt{java.net.URI}, but Spoon was able to infer from the import statements in the Java source text that references to ''URI'' resolve to the fully-qualified type name \texttt{java.net.URI}.
-Thus, TypeV is able to keep track of the actual class being used and is not confused when two classes in separate packages have the same name.
-
-\begin{figure}[!h]
-\centering
-\includegraphics[height=2.5in]{commits_view}
-\caption{Commits increase in time from bottom up.
-Therefore, ASTs from B and C are diff'd against ASTs in A.
-ASTs from E are diff'd against D.
-Since D is a merge commit, its changes are ignored as they are recorded in C and B.}
-\label{fig:commits}
-\end{figure}
-
-In order to gather all the changes for a given repository over time, we generated ASTs for each commit and compared the difference.
-We did this by running \texttt{git log} over the Git repository and listing all the commit IDs in chronological order (Figure~\ref{fig:commits}).
-Each commit and its parent were then compared using AST differencing as described above.
-There are a few edge cases; however, if a commit has no parent, then every declaration or invocation in the commit was treated as an addition.
-Since we compared each commit to its parent, as reported by Git, we were able to account for any branching that happened outside the primary branch of the repository (usually named \texttt{master}).
-If a commit has more than one parent, this indicates a merge commit in Git.
-Merge commits contain the changes made in each parent commit combined.
-Therefore, we ignored merge commits since all changes in the merge should already have been accounted for in previous commits.
-
-\section{TypeV: The Visualization Tool}
-
-\begin{figure*}[!t]
-\centering
-\includegraphics[width=\textwidth]{Screenshot_antlr4}
-\caption{}
-\label{fig:main-view}
-\end{figure*}
-
-The data would be difficult to analyze if displayed as its raw text form; instead, we created an HTML5 web visualization application which is intended for developers, project managers, and researchers to use (Figure~\ref{fig:main-view}).
-The source code for TypeV is available online on GitHub,\footnote{\url{https://github.com/mdfeist/TypeV}}.
-
-TypeV provide a variety of views to help organize the data and provide different levels of detail.
-The main view displays a series of stacked bar graphs which show changes to types over a period of time in a project (Figure~\ref{fig:main-view}).
-The blue and orange portions account for the number of additions and deletions in a similar fashion to GitHub’s code frequency graph (Figure~\ref{fig:GH_code_freq}).
-The height of each bar shows the number of changes to a type.
-The width of each bar represents a set time period and can be changed to an hour, day, week, or month.
-A bar which fills the row vertically is the period with the largest number of changes to that type out of the visible time periods.
-The other bars in the row are proportional to that maximum number.
-The bar height is calculated by multiplying the maximum bar height by the number of changes during that time period divided by the maximum number of changes attested for this type.
-The rows are sorted so that the graph for the most used types appear the top and are in descending order.
-
-When a user mouses over a specific bar in the graph they are given simple statistics for the type and time period at that location.
-This includes the fully-qualified name of the type, the number of commits, the number of authors, the number of additions, the number of deletions, and the start and end date of that time period.
-If the user clicks on the bar, they are given more detailed statistics which list the authors who made the commits and each commit message.
-
-One issue with analyzing individual authors is that authors can commit from a variety of different computers which can have different emails and usernames attached to the Git configuration.
-This means in the \texttt{git log} one author might appear as several different users.
-There is no perfect way to deal with this problem; however, we allow the user of our software to select and view multiple users at a time.
-Thus allowing the user to view changes from an individual or group of developers.
-
-To allow for a more detailed analysis of a specific component of the repository we added some basic filtering tools.
-The first is that a user can filter by type name.
-The second allows the user to view changes that fall within a specific start and end date.
-The third allows them to specify the time span represented by each bar: either hour, day, week, or month.
-Finally, they can choose to see only types declarations, types used in declarations and invocations, or only invocations.
-
-\section{Case Study}
-
-In the following sections, we present case studies demonstrating the extra insight that the TypeV method makes available.
-
-\subsection{antlr4}
-
-\begin{figure}[!ht]
-\centering
-\includegraphics[width=\columnwidth]{ParserATNSimulator}
-\caption{Snapshot of TypeV showing changes to ParserATNSimulator in antlr4. Date ranges from approximately the begining of 2011 to the end of April in 2012.}
-\label{fig:parser1}
-\end{figure}
-
-\begin{figure}[!ht]
-\centering
-\includegraphics[width=\columnwidth]{Antlr4-GH-Lines-ParserATNSimulator}
-\caption{GitHub's line difference graph showing changes to antlr4. The time spans approximately the same length as Figure~\ref{fig:parser1}.}
-\label{fig:parser2}
-\end{figure}
-
-We chose Antlr4 because it is a substantial Java project with a commit history spanning several years. We can also see distinct changes throughout its history.
-
-For example, in December of 2011 we can see that the ParserATNSimulator class has a large amount of additions and deletions in roughly equal proportion (Figure~\ref{fig:parser1}).
-This suggests that ParserATNSimulator is not being removed from the project but instead there is some refactoring being done.
-If we take a closer look at the commit messages we can see that they mention refactoring and reorganization of code.
-Zooming in on that date,  shows that during that month they were finishing work on a new version of ParserATNSimulator.
-
-If we tried to do the same analyses with the GitHub tools it would be significantly more difficult to achieve the same conclusion (Figure~\ref{fig:parser2}).
-First if we only look at line changes this change to ParserATNSimulator is overwhelmed by other changes and hardly shows up on the GitHub tools.
-Then if we did notice an increase in activity we would not know what types were being worked on.
-Finally we could not quickly look at the commit messages for that specific time and type.
-
-At certain times we can can see major increases across all types.
-This shows us when major changes to the code were being done versus code maintenance.
-Usually this happens when a large amount of code is being pulled in from another location.
-
-\subsection{Apache Bookkeeper}
-
-For deprecation analysis, we chose Apache Bookkeeper.
-As an Apache Foundation project it is a well-curated example of an open source software project.
-It also demonstrates an instance of deprecated methods.
-
-Apache Bookkeeper deprecated the method AbstractConfiguration.getLedgerManagerType() in release 4.2.0\footnote{\url{https://bookkeeper.apache.org/docs/r4.2.0/apidocs/deprecated-list.html}} and replaced it with AbstractConfiguration.getLedgerManagerFactoryClass().
-If we look at getLedgerManagerType() we can see at the time of the start of the deprecation all instances were removed and getLedgerManagerFactoryClass() was added.
-After the deprecation, we see no more changes to getLedgerManagerType() and only some additions of getLedgerManagerFactoryClass().
-
-\section{Evaluation}
-
-\begin{table}[!t]
-\renewcommand{\arraystretch}{1.3}
-\caption{Comparision of Type Coverage and File Coverage in Projects}
-\label{tab:summary}
-\centering
-\begin{tabular}{c|ccc}
-\hline
-\bfseries Project & \bfseries Wilcoxon & \bfseries Pearson (lin) cor & \bfseries Spearman (rank) cor \\
-\hline
-antlr4 & p$\ll$0.01 & 0.943 & 0.907\\
-bookkeeper & 0.313 & 0.984 & 0.972\\
-curator & 0.007 & 0.975 & 0.980\\
-tika & 0.653 & 0.955 & 0.998\\
-\hline
-\end{tabular}
-\end{table}
-
-\begin{figure}[!h]
-\centering
-\includegraphics[width=\columnwidth]{antlr-density}
-\caption{}
-\end{figure}
-
-\begin{figure}[!h]
-\centering
-\includegraphics[width=\columnwidth]{tika-density}
-\caption{}
-\end{figure}
-
-The information regarding which types are used in each change is unique to the AST diff method, so it will be useful to see how changes affect the types in a project rather than just the files involved in a diff.
-We define type coverage as the number of types that have been touched out of the total number of types during either a certain time period or the entire project.
-File coverage is the same measurement, but using the number of source files instead.
-To determine if type coverage gives different information than file coverage, we analyzed four popular Java projects: Antlr, Apache Bookkeeper, Apache Curator and Apache Tika.
-We counted the file coverage and type coverage on a commit-by-commit basis.
-That is, each datum represents a commit with three dimensions: file coverage and type coverage, and its commit date.
-Data is sorted in ascending order of commit date.
-Table~\ref{tab:summary} lists a summary of the data provided.
-In general, all measures are strongly correlated, both for Pearson (linear) correlation, and Spearman (rank) correlation.
-As type coverage and file coverage are both cumulative measures, positive correlation is expected.
-However, even when the distributions are significantly different (as with Antlr and Curator), the correlation is still strongly positive, but not 100\%.
-This means that when the coverage for types and files of a commit increase, the relationship is not necessarily 1-to-1.
-This is remarkable in a language like Java, a classes and files are generally in a 1-to-1 relationship.We conclude from this that type coverage yields related, yet different information than file coverage alone.
-
-\section{Discussion}
-
-When managing group projects in an introductory software engineering course, it is very important to be able to see the contributions that students are making to the group repository to judge learning, progress and collaboration.
-When supervising 4-5 groups of students, it becomes unfeasible to look through all of the commits of each student.
-Students may take the wrong approach in a certain piece of code which becomes ingrained in the project and very difficult to change.
-Many groups have problems with group members who do not complete their assigned work or make contributions which end up being meaningless.
-Issues like this can be difficult to detect and prevent.
-As a manager, having a tool like TypeV allows one to quickly summarize the work done by each group member, organized by the types they have been working on.
-This can show when intervention or guidance is required, without the large time commitment of looking through commits.
-
-The purpose of deprecated software features is to give developers time to change projects to follow a new or better standard without breaking the code.
-Deprecated features can be marked for removal from a project or have some flaw that would make a different feature the better option.
-Given the time at which a feature was deprecated, seeing the progress of replacement has some important implications.
-In the case of a security flaw, that time could indicated times at which systems or users were vulnerable.
-If a feature is marked for removal, a project using a deprecated type has a limited amount of time to make the changes.
-Removing all instances of deprecated features could also mean a project can be upgraded to use new tools or features that were not possible while the deprecated features were present.
-Here we show examples where deprecated classes were removed f
-
-Bug triaging for large open source projects is a very time consuming task, with hundreds  of bugs being reported daily.
-Each bug must be processed and assigned to a developer that has the experience required to fix the bug.
-In order to be effective, the triager has to have knowledge of what each developer has worked on as well as the parts of the code may be affected by the bugs.
-Being able to directly see what types a developer has worked with over time is important information for triaging.
-If the type causing the bug is known, the developers which have recently used that type can be assigned the bug.
-The commit messages for the affected types are also visible, providing greater context about how the type has been used.
-The developer who is assigned the bug can also use type information to locate other affected pieces of code and know which developers have been using the affected types.
-
-\subsection{Threats to Validity}
-
-Some commands in git present a problem gathering the statistics.
-Authors can pull other projects into the repository and these additions will count towards the author's changes.
-These will show up as massive changes by a single author with many types that they did not actually create.
-Changes of this size are visible in the graph and can be found by inspecting the commit messages.
-
-We only look at the master branch of a repository and ignore other branches; therefore, we might miss some of the work done by an author in another branch.
-Going through all of the branches in a repository would take a very long time and contain a large amount of duplicate data.
-Being able to choose which branch to analyze is something that can be added.
-
-We ignore the merge commits so we do not double count changes; however, if there is a merge conflict and the developer makes changes then these changes will not appear in TypeV.
-We make the assumptions that if this case happens then the changes are minor.
-
-\subsection{Future work}
-
-Currently for generic types we treat every instance of a generic as a different type.
-So for example if a user declares a list of strings and a list of integers these will appear as two different types.
-However, if they declare multiple instances of a list of strings then these instances will count toward the same type.
-In the future we would nice to come up with a way to better handle generic types.
-
-Currently we are only looking at declarations and invocations, but it would be nice to also view what classes these changes were made to.
-
-\section{Conclusion}
-The conclusion goes here.
-
-% conference papers do not normally have an appendix
-
-% use section* for acknowledgment
-%\section*{Acknowledgment}
-%The authors would like to thank...
+\input{document.tex}
 

### 2016-04-24 www/js/preprocess-data.js
Extract preprocessData to its own module.

@@ -1,0 +1,69 @@
+import ASTDiff from './ast-diff.js';
+
+import assert from './assert.js';
+
+/**
+ * Converts the raw data delivered by the JSON endpoint to something more
+ * useful and malleable in JavaScript.
+ *
+ * This includes return ASTDiffs.
+ */
+export default function preprocessData(data) {
+  assert(data.types instanceof Array);
+  assert(data.commits instanceof Array);
+  assert(data.dates instanceof Array);
+
+  /* A set of types. */
+  var types = new Set(data.types);
+  /* Mapping GitSha -> Commit Metadata. */
+  var commits = createCommitMap(data.commits);
+
+  return {
+    types,
+    commits,
+    /* A copy of AST Diff data, in asscending order of date. */
+    astDiffs: createASTDiffsInAscendingOrder(data.dates, commits)
+  };
+}
+
+/* Maps Git SHA to the raw commit data. */
+function createCommitMap (commits) {
+  var commitMap = {};
+
+  commits.forEach(function (commit) {
+    var sha = commit.commitID;
+    assert(looksLikeAGitSha(sha));
+    commitMap[sha] = commit;
+  });
+
+  return commitMap;
+}
+
+/* Returns AST Diff data, in asscending order of date. */
+function createASTDiffsInAscendingOrder(astDiffsByType, commits) {
+  /* Ensure each commit has a proper Date object. */
+  for (let sha in commits) {
+    if (!commits.hasOwnProperty(sha)) {
+      continue;
+    }
+    let commit = commits[sha];
+    commit.date = new Date(commit.date);
+  }
+
+  return astDiffsByType
+    .map(ASTDiff.withCommitMap.bind(ASTDiff, commits))
+    /* Note: unary + coerces to smallint using Date#valueOf() */
+    /* See explanation in Cell#isAcceptableDiff(). */
+    .sort((a, b) => d3.ascending(+a.date, +b.date));
+}
+
+/*=== Predicates used in assertions and checks ===*/
+
+function looksLikeAGitSha(thing) {
+  if (typeof thing !== 'string') {
+    return false;
+  }
+
+  return thing.match(/^[0-9a-f]{5,}$/i);
+}
+/*global d3*/

### 2016-04-24 www/js/project_view.js

@@ -7,13 +7,14 @@
  *  - Ian Watts
  */
 
-import ASTDiff from './ast-diff.js';
+
 import JavaType from './java-type.js';
 import TimeSlice from './time-slice.js';
 
+import assert from './assert.js';
+import preprocessData from './preprocess-data.js';
 import { first, last } from './utils.js';
 
-import assert from './assert.js';
 
 var VALID_STEP_SIZES = d3.set(['hour', 'day', 'month', 'week']);
 
@@ -63,57 +64,6 @@ function createTable2(filter) {
 
 /*=== Core functions ====*/
 
-/**
- * Modifies original data.
- */
-function preprocessData(data) {
-  assert(data.types instanceof Array);
-  assert(data.commits instanceof Array);
-  assert(data.dates instanceof Array);
-
-  /* A set of types. */
-  var types = new Set(data.types);
-  /* Mapping GitSha -> Commit Metadata. */
-  var commits = createCommitMap(data.commits);
-
-  return {
-    types,
-    commits,
-    /* A copy of AST Diff data, in asscending order of date. */
-    astDiffs: createASTDiffsInAscendingOrder(data.dates, commits)
-  };
-}
-
-/* Maps Git SHA to the raw commit data. */
-function createCommitMap (commits) {
-  var commitMap = {};
-
-  commits.forEach(function (commit) {
-    var sha = commit.commitID;
-    assert(looksLikeAGitSha(sha));
-    commitMap[sha] = commit;
-  });
-
-  return commitMap;
-}
-
-/* Returns AST Diff data, in asscending order of date. */
-function createASTDiffsInAscendingOrder(astDiffsByType, commits) {
-  /* Ensure each commit has a proper Date object. */
-  for (let sha in commits) {
-    if (!commits.hasOwnProperty(sha)) {
-      continue;
-    }
-    let commit = commits[sha];
-    commit.date = new Date(commit.date);
-  }
-
-  return astDiffsByType
-    .map(ASTDiff.withCommitMap.bind(ASTDiff, commits))
-    /* Note: unary + coerces to smallint using Date#valueOf() */
-    /* See explanation in Cell#isAcceptableDiff(). */
-    .sort((a, b) => d3.ascending(+a.date, +b.date));
-}
 
 /**
  * Takes preprocessed data, and returns an array of types, in ascending order
@@ -159,6 +109,7 @@ function filterTypes(data, filters) {
[…]

### 2016-04-24 www/js/data-filter.js
Extract filterTypes.

@@ -1,0 +1,209 @@
+import JavaType from './java-type.js';
+import TimeSlice from './time-slice.js';
+
+import assert from './assert.js';
+import { first, last, union } from './utils.js';
+
+const VALID_STEP_SIZES = d3.set(['hour', 'day', 'month', 'week']);
+
+/**
+ * Takes preprocessed data, and returns an array of types, in ascending order
+ * of popularity.
+ *
+ * Each type has its Cells, with ASTDiffs.
+ */
+export function filterTypes(data, filters) {
+  /* Let filters be undefined or null. */
+  filters = filters ? filters : {};
+
+  /* Either the date provided, or the first date attested. */
+  var startDate = filters.start || first(data.astDiffs).date;
+  /* Either the date provided or the last date attested. */
+  var endDate = filters.end || last(data.astDiffs).date;
+  var numberOfTypesUpperBound = filters.limit || Infinity;
+  var stepSize = filters.stepSize || 'month';
+  var authors = filters.authors || [];
+  var typeFilter = filters.typeFilter ? filters.typeFilter : null;
+
+  assert(startDate instanceof Date);
+  assert(endDate instanceof Date);
+  assert(startDate < endDate);
+  assert(typeof numberOfTypesUpperBound === 'number');
+  assert(VALID_STEP_SIZES.has(stepSize));
+  assert(authors instanceof Array);
+
+  /* Find the range of diffs to use. */
+  var lowerIndex = d3.bisectLeft(data.astDiffs, startDate);
+  var upperIndex = d3.bisectRight(data.astDiffs, endDate);
+
+  var applicableDiffs = data.astDiffs.slice(lowerIndex, upperIndex);
+  /* Filter types, first by the types that are ACTUALLY present in the
+   * filtered ASTDiffs */
+  var typesPresent = countTypeAbsoluteFrequency(applicableDiffs);
+  /* XXX: refactor... */
+  var typeNames = typeFilter === null ? Object.keys(typesPresent) :
+    Object.keys(typesPresent).filter(typeName => {
+      return typeName.toLowerCase().includes(typeFilter.toLowerCase());
+    });
+
+  var sortedTypeNames = typeNames
+    .sort((a, b) => d3.descending(typesPresent[a], typesPresent[b]))
+    .slice(0, numberOfTypesUpperBound);
+
+  /* TODO: This is actually okay! You've just got an empty selection! */
+  assert(sortedTypeNames.length > 0);
+
[…]

### 2016-04-24 www/js/project_view.js

@@ -8,15 +8,9 @@
  */
 
 
-import JavaType from './java-type.js';
-import TimeSlice from './time-slice.js';
-
 import assert from './assert.js';
 import preprocessData from './preprocess-data.js';
-import { first, last } from './utils.js';
-
-
-var VALID_STEP_SIZES = d3.set(['hour', 'day', 'month', 'week']);
+import { filterTypes } from './data-filter.js';
 
 /* Shim the assert function in there! */
 
@@ -62,219 +56,12 @@ function createTable2(filter) {
   return data;
 }
 
-/*=== Core functions ====*/
-
-
-/**
- * Takes preprocessed data, and returns an array of types, in ascending order
- * of popularity.
- *
- * Each type has its Cells, with ASTDiffs.
- */
-function filterTypes(data, filters) {
-  /* Let filters be undefined or null. */
-  filters = filters ? filters : {};
-
-  /* Either the date provided, or the first date attested. */
-  var startDate = filters.start || first(data.astDiffs).date;
-  /* Either the date provided or the last date attested. */
-  var endDate = filters.end || last(data.astDiffs).date;
-  var numberOfTypesUpperBound = filters.limit || Infinity;
-  var stepSize = filters.stepSize || 'month';
-  var authors = filters.authors || [];
-  var typeFilter = filters.typeFilter ? filters.typeFilter : null;
-
-  assert(startDate instanceof Date);
-  assert(endDate instanceof Date);
-  assert(startDate < endDate);
-  assert(typeof numberOfTypesUpperBound === 'number');
-  assert(VALID_STEP_SIZES.has(stepSize));
-  assert(authors instanceof Array);
-
-  /* Find the range of diffs to use. */
-  var lowerIndex = d3.bisectLeft(data.astDiffs, startDate);
-  var upperIndex = d3.bisectRight(data.astDiffs, endDate);
-
-  var applicableDiffs = data.astDiffs.slice(lowerIndex, upperIndex);
-  /* Filter types, first by the types that are ACTUALLY present in the
-   * filtered ASTDiffs */
-  var typesPresent = countTypeAbsoluteFrequency(applicableDiffs);
-  /* XXX: refactor... */
-  var typeNames = typeFilter === null ? Object.keys(typesPresent) :
-    Object.keys(typesPresent).filter(typeName => {
-      return typeName.toLowerCase().includes(typeFilter.toLowerCase());
-    });
-
-  var sortedTypeNames = typeNames
[…]

### 2016-04-24 www/js/utils.js

@@ -4,7 +4,6 @@ import assert from './assert.js';
  * A bunch of random utilities I had no idea what to do with.
  */
 
-
 /**
  * Returns the first element of an array. Complains bitterly if the array is
  * empty.
@@ -27,4 +26,14 @@ export function last(array) {
     throw new Error('Cannot get the first item of an empty array; is the project empty?');
   }
   return array[array.length - 1];
+}
+
+/**
+ * Union of two sets.
+ */
+export function union(set, iterable) {
+  for (var item of iterable) {
+    set.add(item);
+  }
+  return set;
 }

### 2016-04-24 www/js/data-filter.js
Refactor filterTypes to DataView.filter()

@@ -4,7 +4,78 @@ import TimeSlice from './time-slice.js';
 import assert from './assert.js';
 import { first, last, union } from './utils.js';
 
-const VALID_STEP_SIZES = d3.set(['hour', 'day', 'month', 'week']);
+const VALID_STEP_SIZES = ['hour', 'day', 'month', 'week'];
+
+export default class DataView {
+  constructor({types, timeslices, typesPresent, authorStats, astDiffs,
+              typesOverall, filesOverall}) {
+    this.types = types;
+    this.timeslices = timeslices;
+    this.typesPresent = typesPresent;
+    this.authorStats = authorStats;
+    this.astDiffs = astDiffs;
+    this.typesOverall = typesOverall;
+    this.filesOverall = filesOverall;
+
+    assert(this.minDate <= this.maxDate);
+  }
+
+  /**
+   * Returns a new DataView with the given filters applied:
+   *
+   * # startDate
+   * # endDate
+   * # limit -- maximum number of types
+   * # stepSize -- What size to group ASTDiffs
+   * # authors -- Filter only to the given authors.
+   * # typeFilter -- only types matching this pattern are selected.
+   */
+  static filter(data, filters) {
+    const rawFilteredData = filterTypes(data, filters);
+    return new DataView(rawFilteredData);
+  }
+
+  /* TODO: Rename to allTypeNames */
+  /** All types name, without arbitrary filtering or sorting. **/
+  get allTypes() {
+    return Object.keys(this.typesPresent);
+  }
+
+  /* TODO: rename to authorNames. */
+  /** The name of every author selected in the view. */
+  get authors() {
+    return Object.keys(this.authorStats);
+  }
+
+  get minDate() {
+    return first(this.timeslices).startDate;
+  }
+
+  get maxDate() {
+    return last(this.timeslices).endDate;
+  }
+
+  get numberOfTypes() {
+    return this.typesOverall.size;
+  }
+
+  get numberOfFiles() {
+    return this.filesOverall.size;
+  }
+
+  /* The following two are used for the date display picker thing. */
+  get absoluteMinDate() {
+    return first(this.astDiffs).date;
+  }
+
+  get absoluteMaxDate() {
+    return last(this.astDiffs).date;
+  }
+
+  get numberOfColumns() {
+    return this.timeslices.length;
+  }
+}
 
 /**
  * Takes preprocessed data, and returns an array of types, in ascending order
[…]

### 2016-04-24 www/js/project_view.js

@@ -10,7 +10,7 @@
 
 import assert from './assert.js';
 import preprocessData from './preprocess-data.js';
-import { filterTypes } from './data-filter.js';
+import DataView from './data-filter.js';
 
 /* Shim the assert function in there! */
 
@@ -49,7 +49,7 @@ function createTable2(filter) {
   dnaTable.innerHTML = "";
 
   var processed = window.preprocessedData;
-  var data = window.filteredData = filterTypes(processed, filter);
+  var data = window.filteredData = DataView.filter(processed, filter);
   drawGraph(data, dnaTable.offsetWidth);
   drawStats(data, dnaTable.offsetWidth);
 

### 2016-04-25 server.py
Add force-directed graph.

@@ -461,15 +461,14 @@ def root():
 def dashboard():
     return render_template('projects.html', projects=projects)
 
-@app.route('/projects/<path:path>')
-def show_project(path):
-    name = path.replace('/','')
+@app.route('/projects/<name>/')
+def show_project(name):
     for project in projects:
         if (project.getDir() == name):
            return render_template('project_view.html', authors=list(project.getAuthors()))
-    return name
+    return 'Project not found', 404
 
-@app.route('/projects/<path:path>/get_project')
+@app.route('/projects/<path>/get_project')
 def get_project(path):
     #query_string = request.query_string
     query_type = request.args.get('type')
@@ -483,7 +482,25 @@ def get_project(path):
             options.ignore_large_commits = False
             return project.getJSON(options)
 
-    return ""
+    # Could not find project :/
+    return 'Project not found', 404
+
+
+@app.route('/projects/<path>/fdg')
+def force_directed_graph(path):
+    query_type = 'Types'
+
+    for project in projects:
+        if (project.getDir() == path):
+            options = attrdict()
+            options.get = "Project"
+            options.types = query_type
+            options.ignore_large_commits = False
+            data = project.getJSON(options)
+            return render_template('temporary_fdg.html', data=data)
+
+    return 'Project not found', 404
+
 
 @app.before_first_request
 def _run_on_start():

### 2016-04-25 www/js/fdg.js

@@ -1,0 +1,143 @@
+/**
+ * Temporary file for force-directed graphs.
+ */
+import preprocessData from "./preprocess-data.js";
+import DataView from "./data-filter.js";
+
+const TYPE = 0;
+const AUTHOR = 1;
+
+export default function doIt() {
+  /** TODO: pre-preprocess data? */
+  var data = DataView.filter(preprocessData(window.DATA), {
+    limit: 40
+  });
+
+  var graph = graphFromData(data);
+
+  var width = 960, height = 500;
+
+  var color = d3.scale.category10();
+
+  var force = d3.layout.force()
+      .charge(-120)
+      .linkDistance(100)
+      .size([width, height]);
+
+  // Insert stuff into #figure
+  var svg = d3.select("#figure").append("svg")
+      .attr("width", width)
+      .attr("height", height);
+
+  force
+      .nodes(graph.nodes)
+      .links(graph.links)
+      .start();
+
+  var link = svg.selectAll(".link")
+      .data(graph.links)
+    .enter().append("line")
+      .attr("class", "link")
+      .style("stroke-width", function(d) { return Math.sqrt(d.value); });
+
+  var node = svg.selectAll(".node")
+      .data(graph.nodes)
+    .enter().append("circle")
+      .attr("class", "node")
+      // Make authors bigger.
+      .attr("r", node => node.kind == AUTHOR ? 10 : 5)
+      .style("fill", node => color(node.kind))
+      .call(force.drag);
+
+  // This is not... this is... no.
+  node.append("title")
+      .text(node => node.label);
+
+  force.on("tick", function() {
+    link.attr("x1", function(d) { return d.source.x; })
+        .attr("y1", function(d) { return d.source.y; })
+        .attr("x2", function(d) { return d.target.x; })
+        .attr("y2", function(d) { return d.target.y; });
+
+    node.attr("cx", function(d) { return d.x; })
+        .attr("cy", function(d) { return d.y; });
+  });
+
+  document.querySelector('#loading-indicator').remove();
+}
+
+/**
+ * returns {
+ *      nodes: [ordered list of authors or types]
+ *      links: [list of links between authors and types]
+ *  }
+ */
+function graphFromData(data) {
+  var nodes = [];
+
+  for (let author of data.authors) {
+    nodes.push({label: author, data: author, kind: AUTHOR});
+  }
+
+  for (let type of data.types) {
+    nodes.push({label: type.name, data: type, kind: TYPE});
+  }
+
+  var authorRelations = new Map();
+
[…]

### 2016-04-25 www/.eslintrc.js
Fix lack of createTable2().

@@ -19,12 +19,13 @@ module.exports = {
         ],
         /* Prefer single quotes, but forgive double quotes. */
         "quotes": [
-            "warn",
-            "single",
-            {
-                "avoidEscape": true,
-                "allowTemplateLiterals": true
-            }
+            0,
+            //"warn",
+            //"single",
+            //{
+            //    "avoidEscape": true,
+            //    "allowTemplateLiterals": true
+            //}
         ],
         /* ALWAYS use semicolons! */
         "semi": [

### 2016-04-25 www/js/project_view.js

@@ -39,9 +39,10 @@ export function createTable(data, filter) {
   return createTable2(filter);
 }
 
-window.createTable = createTable;
-
-function createTable2(filter) {
+/**
+ * Draw table given filter.
+ */
+export function createTable2(filter) {
   /* Plop this in dna-table div */
   var dnaTable = document.getElementById('dna-table');
 

### 2016-04-25 paper/document.tex
Add Antlr code frequency vector.
Latest document update.

@@
-% As a general rule, do not put math, special symbols or citations
-% in the abstract
+% === above this line ===
+% preamble
+% begin document
+% maketitle
+
 \begin{abstract}
@@
 Furthermore, we can track each author's type usage over time.
-Using this tool we examine specific cases in well-known repositories where our tool reveals interesting and useful information.
+Using TypeV, we examine specific cases in well-known repositories where our tool reveals interesting and useful information.
 We then compare type coverage information from the AST compared to file coverage to determine if unique information is provided by type information.
@@
 
-% For peer review papers, you can put extra information on the cover
-% page as needed:
-% \ifCLASSOPTIONpeerreview
-% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
-% \fi
-%
-% For peerreview papers, this IEEEtran command inserts a page break and
-% creates the second title. It will be ignored for other modes.
 \IEEEpeerreviewmaketitle
@@
 \centering
-\includegraphics[width=\textwidth]{antlr-3-code-freq-edit}
+\includegraphics[width=\textwidth]{antlr4-code-freq}
 \caption{GitHub graph showing code frequency.
@@
 
-The core of our analysis is calculating the differences between abstract syntax trees of subsequent commits (Figure~\ref{fig:context}).
+Given a Git repository, we want to extract per-author type evolution over time.
+We accomplished this by computing the differences between abstract syntax trees of subsequent commits (Figure~\ref{fig:context}).
 
-An \emph{abstract syntax tree} (AST) is a data structure derived from source code by breaking its syntactic constructs into a tree.
+An \emph{abstract syntax tree} (AST) is a data structure derived from source code that breaks its syntactic constructs into a tree.
 Each node in the tree represents a syntactically valid chunk of code.
@@
 Converting raw source text to an AST is often one of the first steps performed by compilers.
-When a programmer changes code in a file, the difference will appear in the AST.
+When a programmer makes a non-trivial change in code in a file, that difference will be reflected in the AST.
 This means that the AST of revisions in code repositories can be compared to see what structures a programmer is modifying.
@@
 
+\section{Evaluation}
+
+\begin{table}[!t]
+\renewcommand{\arraystretch}{1.3}
+\caption{Comparison of Type Coverage and File Coverage in Projects}
+\label{tab:summary}
+\centering
+\begin{tabular}{c|ccc}
+\hline
+\bfseries Project & \bfseries Wilcoxon& \bfseries Pearson (linear) & \bfseries Spearman (rank) \\
+& $p$-value & Correlation & Correlation \\
+\hline
+antlr4 & $p\ll0.01$ & 0.943 & 0.907\\
+bookkeeper & $0.313$ & 0.984 & 0.972\\
+curator & $0.007$ & 0.975 & 0.980\\
+tika & $0.653$ & 0.955 & 0.998\\
+\hline
+\end{tabular}
+\end{table}
+
+\begin{figure}[!h]
+\centering
+\includegraphics[width=\columnwidth]{antlr-density}
+\caption{}
+\end{figure}
+
+\begin{figure}[!h]
+\centering
+\includegraphics[width=\columnwidth]{tika-density}
+\caption{}
+\end{figure}
+
+The information regarding which types are used in each change is unique to the AST diff method, so it will be useful to see how changes affect the types in a project rather than just the files involved in a diff.
+We define type coverage as the number of types that have been touched out of the total number of types during either a certain time period or the entire project.
+File coverage is the same measurement, but using the number of source files instead.
+To determine if type coverage gives different information than file coverage, we analyzed four popular Java projects: Antlr, Apache Bookkeeper, Apache Curator and Apache Tika.
+We counted the file coverage and type coverage on a commit-by-commit basis.
+That is, each datum represents a commit with three dimensions: file coverage and type coverage, and its commit date.
+Data is sorted in ascending order of commit date.
+Table~\ref{tab:summary} lists a summary of the data provided.
+In general, all measures are strongly correlated, both for Pearson (linear) correlation, and Spearman (rank) correlation.
+As type coverage and file coverage are both cumulative measures, positive correlation is expected.
+However, even when the distributions are significantly different (as with Antlr and Curator), the correlation is still strongly positive, but not 100\%.
+This means that when the coverage for types and files of a commit increase, the relationship is not necessarily 1-to-1.
+This is remarkable in a language like Java, a classes and files are generally in a 1-to-1 relationship.We conclude from this that type coverage yields related, yet different information than file coverage alone.
+
 \section{Case Study}
@@
 
-\section{Evaluation}
-
-\begin{table}[!t]
-\renewcommand{\arraystretch}{1.3}
-\caption{Comparison of Type Coverage and File Coverage in Projects}
-\label{tab:summary}
-\centering
-\begin{tabular}{c|ccc}
-\hline
-\bfseries Project & \bfseries Wilcoxon& \bfseries Pearson (linear) & \bfseries Spearman (rank) \\
-& p-value & Correlation & Correlation \\
-\hline
-antlr4 & p$\ll$0.01 & 0.943 & 0.907\\
-bookkeeper & 0.313 & 0.984 & 0.972\\
-curator & 0.007 & 0.975 & 0.980\\
-tika & 0.653 & 0.955 & 0.998\\
-\hline
-\end{tabular}
-\end{table}
-
-\begin{figure}[!h]
-\centering
-\includegraphics[width=\columnwidth]{antlr-density}
-\caption{}
-\end{figure}
-
-\begin{figure}[!h]
-\centering
-\includegraphics[width=\columnwidth]{tika-density}
-\caption{}
-\end{figure}
-
-The information regarding which types are used in each change is unique to the AST diff method, so it will be useful to see how changes affect the types in a project rather than just the files involved in a diff.
-We define type coverage as the number of types that have been touched out of the total number of types during either a certain time period or the entire project.
-File coverage is the same measurement, but using the number of source files instead.
-To determine if type coverage gives different information than file coverage, we analyzed four popular Java projects: Antlr, Apache Bookkeeper, Apache Curator and Apache Tika.
-We counted the file coverage and type coverage on a commit-by-commit basis.
-That is, each datum represents a commit with three dimensions: file coverage and type coverage, and its commit date.
-Data is sorted in ascending order of commit date.
-Table~\ref{tab:summary} lists a summary of the data provided.
-In general, all measures are strongly correlated, both for Pearson (linear) correlation, and Spearman (rank) correlation.
-As type coverage and file coverage are both cumulative measures, positive correlation is expected.
-However, even when the distributions are significantly different (as with Antlr and Curator), the correlation is still strongly positive, but not 100\%.
-This means that when the coverage for types and files of a commit increase, the relationship is not necessarily 1-to-1.
-This is remarkable in a language like Java, a classes and files are generally in a 1-to-1 relationship.We conclude from this that type coverage yields related, yet different information than file coverage alone.
-
 \section{Discussion}
@@
 Removing all instances of deprecated features could also mean a project can be upgraded to use new tools or features that were not possible while the deprecated features were present.
-Here we show examples where deprecated classes were removed f
+Here we show examples where deprecated classes were removed.
 
@@
 
-The methodology used in TypeV is only well-defined for statically-typed languages such as Haskell and Java.
-Dynamically-typed languages like Erlang and JavaScript.
+The methodology used in TypeV is only well-defined for statically-typed languages such as Haskell or Java; it is not well defined for dynamically-typed languages such as Erlang or JavaScript.
 However, type annotation systems such as TypEr for Erlang~\cite{typer} or statically-typed language dialects such as TypeScript for JavaScript may prove to be sufficient to enable type analysis as we have demonstrated for Java.
@@
 
-% conference papers do not normally have an appendix
-
-% use section* for acknowledgment
-%\section*{Acknowledgment}
-%The authors would like to thank...
+% === Below this line ===
+% bibliography
+% end document

### 2016-04-25 www/js/data-filter.js
Sneak in some project info.

@@ -8,11 +8,12 @@ const VALID_STEP_SIZES = ['hour', 'day', 'month', 'week'];
 
 export default class DataView {
   constructor({types, timeslices, typesPresent, authorStats, astDiffs,
-              typesOverall, filesOverall}) {
+              typesOverall, filesOverall, commits}) {
     this.types = types;
     this.timeslices = timeslices;
     this.typesPresent = typesPresent;
     this.authorStats = authorStats;
+    this.commits = commits;
     this.astDiffs = astDiffs;
     this.typesOverall = typesOverall;
     this.filesOverall = filesOverall;
@@ -74,6 +75,10 @@ export default class DataView {
 
   get numberOfColumns() {
     return this.timeslices.length;
+  }
+
+  get numberOfCommits() {
+    return Object.keys(this.commits).length;
   }
 }
 
@@ -212,6 +217,7 @@ function filterTypes(data, filters) {
   return {
     /* Filtered types. */
     types,
+    commits: data.commits,
     typesPresent,
     timeslices,
     authorStats: authorMap,

### 2016-04-25 www/js/project_view.js

@@ -72,6 +72,11 @@ function drawGraph(data, width) {
   var cellHeight = 64;
   var height = cellHeight * data.types.length;
 
+  /* XXX: Sneak in a little info about the project. */
+  d3.select('body').append('pre')
+    .text(summary(data))
+    .style('display', 'none');
+
   /* Create a scale for the types i.e., the y-axis */
   var yScale = d3.scale.ordinal()
     .domain(data.types.map(type => type.fullyQualifiedName ))
@@ -509,6 +514,10 @@ function ensureAxisIsAtGraphBottom(graph, axis) {
   } else {
     axis.classList.add('axis-floating');
   }
+}
+
+function summary({numberOfCommits, authors, numberOfFiles, numberOfTypes, astDiffs}) {
+  return `project & ${numberOfCommits} & ${authors.length} & ${numberOfFiles} & ${numberOfTypes} & ${astDiffs.length}`;
 }
 
 /**

### 2016-04-25 server.py
Change the title on pages.

@@ -465,7 +465,8 @@ def dashboard():
 def show_project(name):
     for project in projects:
         if (project.getDir() == name):
-           return render_template('project_view.html', authors=list(project.getAuthors()))
+           return render_template('project_view.html', name=name,
+                                  authors=list(project.getAuthors()))
     return 'Project not found', 404
 
 @app.route('/projects/<path>/get_project')

### 2016-04-26 paper/document.tex
Update bibliography some more.

@@
 \includegraphics[width=\textwidth]{antlr4-code-freq}
-\caption{GitHub graph showing code frequency.
-Note that the graph was truncated to better show the details, thus the changes that occur in mid 2015 are much larger than shown.
+\caption{GitHub's code frequency graph for ANTLR4.
+Note that the graph was cropped to better show the details.
+Thus, the changes that occur in mid 2015 are much larger than shown.
 The additions reach over 400k and the deletions are over 300k.}
@@
 Source code is often represented as text in a source code file.
-This text is then parsed by a compiler or interpreter with the intent of eventually executing the instructions within the source code file.
+This text is then parsed by a compiler or interpreter with the intent of eventually executing the code represented by the source text.
 While programmers may type in individual characters, their intention is at a much higher level.
@@
 The programmer rarely operates at the character level, or even the line level, so why should our analyses of source code operate at granularities unfamiliar to developers?
-
+The plain text nature of software does not stop at source code level.
 The history and evolution of a software project is often tracked by version control systems, such as Git, in a programming language agnostic way.
-Updates to source code, documentation, and tests are recorded in the form of commits, which show the differences in lines between each subsequent revision of the software.
+Updates to source code, documentation, and tests are recorded in the form of commits, which store the differences in lines of text in a file between each subsequent revision of the software.
 Version control systems become vital for large projects since individual changes would be impossible to track manually.
@@
 Most hosting services for Git repositories provide some basic tools for analyzing projects over time~\cite{github-graphs,bitbucket-graphs}; however, as we will show, these tools often lack the kind of detail to gather insightful analysis.
-While commits may store individual changes to lines of code, they represent the change in \emph{structure} of the software.
+As with source code, commits operate at the text level.
+While commits may store individual changes as lines of code, they imply a change in \emph{structure} of the software.
 By visualizing lines changed, or files changed, the current state-of-the-art for software project visualization rarely capture the \emph{meaning} behind each commit; lines of code alone does not describe how a repository changed at a particular point in time.
@@
 
+The goal of a software evolution visualization is to provide an alternate view that provides insight into the development of the software that is otherwise unavailable.
+By visualizing changes to lines and files, we are tracking the changes in text---not software.
+What would be more meaningful are the structural changes in code.
+These are changes in variable declarations, invocations of methods, and usage of types.
+
+In this paper, we propose a method that analyses the changes in type usage and method invocations over time.
+Using commits, we are able to keep track of \emph{who} is changing what types.
+We then present this data using simple visualizations.
+We claim that type usage and method invocation offers more insight than line-based, file-based, and commit-based methods of visualization and analysis.
+For simplicity, we analyze only Java projects, as it allowed us to quickly develop and test our software.
+However, since the approach described in this paper can be generally applied to any statically-typed language, it can be extended it to other languages.
+Our goal is to describe a methodology for obtaining richer, more meaningful data from source code repositories which can be used in more insightful analyses.
+
+In the following sections of the paper we will describe our tool, TypeV, that we developed to compute and visualizes changes to types used in a project over time.
+TypeV is able to extract type information from a repository by computing commit-by-commit abstract syntax tree differences, which we describe in section~\ref{sec:methodology}.
+In this paper, we answer the following questions: \\ \\
+\textbf{RQ1:} How can we extract type and method usage over time from repositories? \\ \\
+\textbf{RQ2:} Are type coverage file coverage different metrics?
+\\ \\
+\textbf{RQ3:} Does the type information give any interesting or useful insight into the project?
+
+\section{Related Work}
+
+\subsection{Analysis of Languages and Features}
+
+Analyzing source code is a popular research topic and Java is often chosen as the subject due to its popularity.
+Grechanik \textit{et al.} \cite{Grechanik:2010:EIL:1852786.1852801} examined the structure of Java programs mined from 2080 programs.
+This paper examined the breakdown of syntactic structures in open source repositories, however it does not consider per author statistics.
+Parnin \textit{et al.} \cite{Parnin:2011:JGA:1985441.1985446} mined repositories to see how Java generics have been used in open source projects and found that generic usages were often introduced by a single developer in a project and were primarily being used for collecting and traversing lists of objects.
+
+Further analysis of the Java language has been done using ASTs.
+L\"{a}mmel \textit{et al.} \cite{Lammel:2011:LAA:1982185.1982471} used ASTs to examine the usage of APIs in Java projects.
+They found what APIs are popular and if they were used in a framework like manner.
+Dyer \textit{et al.} \cite{Dyer:2014:MBA:2568225.2568295} mined AST nodes to study the use of new Java language features over time.
+The authors found the most popular features and how teams adopted new features.
+New features varied greatly in popularity and adoption rate.
+They did not check to see how much a developer used a feature but instead checked to see if they used it at all.
+In a more general analysis of programming languages, Meyerovich and Rabkin \cite{Meyerovich:2013:EAP:2509136.2509515} surveyed developers and examined repositories to learn about language adoption and usage.
+Developers feel that certain language features are more important than others.
+
+The focus of these papers is on how programming languages and features are used by looking at many repositories.
+Java is a strongly typed language, yet there is no focus specifically on how types are used throughout a project.
+Type info may provide additional information on how developers use languages but has not been explored.
+
+\subsection{Visualizing Project Evolution}
+
+A variety of visualizations tools get created in order to help with understanding the changes that get made to software over time.
+Wu \textit{et al.} \cite{wu2004a} explored punctuated changes to software repositories with a tool that generates an Evolution Spectrograph.
+In this visualization, each file has a row in the chart and each cell represents a section of time where an incoming or outgoing dependency can be changed.
+The colour is determined by the frequency of change.
+This work was continued to show file directoryies and display the cardinality of a developer's commits Wu \textit{et al.} \cite{wu2004}.
+The cardinality of a commit is the number of files or subsystems it affects in a system, indicating the scope of the change.
+Other tools also look at changes in code dependencies \cite{6980224} or files \cite{6650525, 7332421} over time.
+
+Voinea \textit{et al.} \cite{voinea2006} visualizes CVS repositories using CVSgrab and adding their own enhancements to show changes to files over time.
+Files can be sorted by creation time, activity, or alphabetically and colour coded based on author, file size, file type or by search strings.
+Kim \textit{et al.} \cite{Kim:2009:DRS:1555001.1555046} built a tool called Logical Structural Diff (LSdiff) which examines code changes and produces logical rules to represent them.
+Changes which do not match the rules are then pointed out to the developer.
+They also formed a focus group with software engineers who found that LSdiff complements existing tools well and showed promise for discovering inconsistent changes.
+
+\subsection{Project Collaboration}
+
+There is also research on how developers use software repositories to collaborate.
+Wagstrom \textit{et al.} \cite{Patrick:Wagstrom:2012} explored the roles that developers take in a networked, social development environments like GitHub.
+These role were defined as their level of contribution as well as the types of issues they dealt with.
+Developers will take on multiple roles in a project and will sometimes fulfill the same role across different projects.
+Lee \textit{et al.} \cite{lee2013} developed a tool for visualizing the branching structure of git repositories.
+This was used to analyze concurrent workflows in popular open source repositories.
+
+\subsection{Non-academic works}
+
+Some visualization tools which are used for managing or making decisions about software rather than study the changes taking place.
+The service EasyBi \cite{EasyBi} provides additional visualizations for Git repositories for the purpose of business decisions.
+EasyBi works by taking the git log file and generating charts and graphs from it.
+It tries to solve the issue of summarizing many commits and has greater controls for granularity and authors but still operates on commits and line differences.
+The Gitinspector project \cite{Gitinspector} is an open source tool for the statistical analysis of git repositories.
+It provides an advanced timeline view of changes per author and cumulative statistics but again relies upon the line differences to summarize the work of each author.
+
+The GitHub tools have the strength that they do not rely on what type of code is being committed.
+It does not matter what language or if errors exist in the commit, GitHub will simply spit out the number of line additions and deletions.
+The issue with this type of analysis is that it does not give a clear breakdown of is being changed only that changes have occurred.
+
 Consider tools easily available to an open source software developer; for example GitHub graphs (see Figure~\ref{fig:GH_code_freq})~\cite{github-graphs}.
 Since GitHub tracks contributions based solely on number of commits and numbers of lines changes, data may be lacking, or even misleading.
-In the case of the Antlr4 project (Figure~\ref{fig:GH_code_freq}), the graph creates more questions than it answers.
+In the case of the ANTRL4 project (Figure~\ref{fig:GH_code_freq}), the graph creates more questions than it answers.
 What happened at those big spikes of line changes in the graph?
@@
 
-In this paper, we propose a method that analyses the changes in type usage and method invocations over time, as well as a tool to visualize this analysis, which we call TypeV.
-We claim that type usage offers more insight than line-based, file-based, and commit-based methods of visualization and analysis.
 Of the services we have surveyed, none of the tools visualize project evolution using granular type analysis per author via commit-by-commit abstract syntax tree differencing.
-For simplicity, we analyze only Java projects, as it allowed us to quickly develop and test our software.
-Since our approach described in this paper can be generally applied to any statically-typed language, we plan to extend it to other languages.
 
-In Java, a type can be a primitive, provided by the language itself, or object references.
-Primitive types are those that are built into the language such as an \texttt{int} or \texttt{char}.
-Object references point to instances of classes.
-Classes are defined in the project itself or through an included library.
-The types in a project represent the functionality available for a developer to work with.
-Java's syntax encourages the creation of classes, thus we expect substantial Java projects to define many domain-specific types.
-Different parts of a program use different types and external dependencies.
-When writing or contributing to a project, the author's choice of which part of the program to edit will in turn affect what types they end up working with --- that is, it will affect the author's \emph{type coverage}.
-
-Type information is useful for seeing what libraries and functions are being used.
-For example, when a part of code becomes \emph{deprecated}, continuing its use may lead to design flaws or even security vulnerabilities.
-In Apache Nifi, an earlier version of the software used a compromised key-derivation functions; hence, this class was deprecated, but left in for compatibility with older versions of the software~\cite{nifi}.
-Viewing the type information of a change can show when usage of deprecated classes are added or removed from a project as well as how many remain over time.
-In a more general sense, we can see when types are being added or removed from the project as well as what types are heavily used throughout the project.
-Even better, we can see \emph{who} uses which types heavily.
-
-Having tools for analyzing repositories based on type usage and method invocation would benefit many different parts of the software development process.
-Software management requires a high level understanding of the state of a repository and the changes being made which affect the functionality of the software.
-Tracking developer contributions in a less misleading way is also important for management since important contributions may be small and focused, but nevertheless, significant.
-Line difference (or simply, \emph{diff}) only provide meaningful information when they are analyzed by someone with familiarity with the source code being changed.
-Providing tools which bring more meaning to line differences would allow people less familiar with the code to understand changes being made.
-When large changes are made that cover many aspects of a project, it would also be useful to see how the change affects the software rather than a several thousand line diff.
-Seeing how a project has changed over time is also vital for management since evaluating the state of completion of software can be very difficult.
-
-In the following sections of the paper we will introduce a tool, TypeV, that we developed that visualizes changes to types used in a project over time.
-TypeV is able to extract type information from a repository by computing commit-by-commit abstract syntax tree differencing, which we describe in section~\ref{sec:methodology}.
-We seek to answer the following questions: \\ \\
-\textbf{RQ1:} How can we extract type and method usage over time from repositories? \\ \\
-\textbf{RQ2:} Does the type information give any interesting or useful insight into the project? \\ \\
-\textbf{RQ3:} How does type coverage compare to file coverage?
-
-\section{Related Work}
-
-Analyzing source code is a popular research topic.
-Most of the focus is on how certain programming languages and features are used.
-However, there is no focus specifically on how types are used throughout a project.
-
-Dyer \textit{et al.} \cite{Dyer:2014:MBA:2568225.2568295} mined AST nodes to study the use of new Java language features over time.
-They found that all new Java features do get used, but not nearly as often as they could have been.
-New features varied greatly in popularity and adoption rate.
-Developers would also modify old code to use new features as they were released.
-The authors also found the most popular features and how teams adopted new features.
-They did not check to see how much a developer used a feature but instead only checked to see if they used it at all.
-
-Grechanik \textit{et al.} \cite{Grechanik:2010:EIL:1852786.1852801} examined the structure of Java programs mined from 2080 programs.
-This paper examined the breakdown of syntactic structures in open source repositories, however it does not consider per author statistics.
-
-Meyerovich and Rabkin \cite{Meyerovich:2013:EAP:2509136.2509515} surveyed developers and examined repositories to learn about language adoption and usage.
-Developers feel that certain language features are more important than others.
-
-Parnin \textit{et al.} \cite{Parnin:2011:JGA:1985441.1985446} mined repositories to see how Java generics have been used in open source projects and found that generic usages were often introduced by a single developer in a project.
-Generics usage was also fairly narrow, with the primary use being collecting and traversing lists of objects.
-
-Wagstrom \textit{et al.} \cite{Patrick:Wagstrom:2012} explored the roles that developers take in a networked, social development environments like GitHub.
-These role were defined as their level of contribution as well as the types of issues they dealt with.
-Developers will take on multiple roles in a project and will sometimes fulfill the same role across different projects.
-
-L\"{a}mmel \textit{et al.} \cite{Lammel:2011:LAA:1982185.1982471} used ASTs to examine the usage of APIs in Java projects.
-They find what APIs are popular as well as if they are used in a framework like manner.
-
-The aforementioned research is trying to learn how developers are using programming languages by looking at many repositories.
-This type of info can be extracted from individual repositories using our new tool.
-Type info may also provide additional information on how developers use languages.
-
-Wu \textit{et al.} \cite{wu2004a} explore punctuated changes to software repositories with a tool that generates an Evolution Spectrograph.
-Each file has a representative row in the chart, with each cell in the row representing a section of time.
-This diagram would show changes to incoming and outgoing dependencies by darkening the cell where the change happened and lightening it over time if no changes were made.
-
-In another paper, Wu \textit{et al.} \cite{wu2004} the Evolutionary Spectrogram is used but with greater control of granularity for files and software subsystems such as directory structure.
-Spectrograms are also used to display the cardinality of a developer's commits.
-The cardinality of a commit is the number of files or subsystems it affects in a system.
-All of the developers changes together show how many changes affect broad sections of the software.
-
-Voinea \textit{et al.} \cite{voinea2006} visualizes CVS repositories using CVSgrab and adding their own enhancements to show changes to files over time.
-Files can be sorted by creation time, activity, or alphabetically and colour coded based on author, file size, file type or by search strings.
-
-Kim \textit{et al.} \cite{Kim:2009:DRS:1555001.1555046} built a tool called Logical Structural Diff (LSdiff) which examines code changes and produces logical rules to represent them.
-Changes which do not match the rules are then pointed out to the developer.
-They also formed a focus group with software engineers who found that LSdiff complements existing tools well and showed promise for discovering inconsistent changes.
-
-The service EasyBi \cite{EasyBi} provides additional visualizations for Git repositories for the purpose of business decisions.
-EasyBi works by taking the git log file and generating charts and graphs from it.
-It tries to solve the issue of summarizing many commits and has greater controls for granularity and authors but still operates on commits and line differences.
-
-The Gitinspector project \cite{Gitinspector} is an open source tool for the statistical analysis of git repositories.
-It provides an advanced timeline view of changes per author and cumulative statistics but again relies upon the line differences to summarize the work of each author.
-
-Wagstrom \textit{et al.} \cite{Patrick:Wagstrom:2012} explored the roles that developers take in a networked, social development environments like GitHub.
-These role were defined as their level of contribution as well as the types of issues they dealt with.
-Developers will take on multiple roles in a project and will sometimes fulfill the same role across different projects.
-
-Lee \textit{et al.} \cite{lee2013} developed a tool for visualizing the branching structure of git repositories. This was used to analyze concurrent workflows in popular open source repositories.
-
-The GitHub tools have the strength that they do not rely on what type of code is being committed.
-It does not matter what language or if errors exist in the commit, GitHub will simply spit out the number of line additions and deletions.
-The issue with this type of analysis is that it does not give a clear breakdown of is being changed only that changes have occurred.
-
-\section{Methodology}
+\section{Data Extraction}
 \label{sec:methodology}
@@
 \includegraphics[width=\columnwidth]{context}
-\caption{Shows the pipeline for TypeV. First we list the commits in chronological order. Then we perform AST diffs for each file. Finally, we display the information.}
+\caption{Data extraction pipeline for TypeV.
+First, we list the commits in branch chronological order.
+Then we calculate the AST difference between each changed file across each pair of consecutive commits.
+Finally, we may visualize the information in a variety of ways.}
 \label{fig:context}
@@
 
-Given a Git repository, we want to extract per-author type evolution over time.
-We accomplished this by computing the differences between abstract syntax trees of subsequent commits (Figure~\ref{fig:context}).
+Given a Git repository, we want to extract per-author type changes over time.
+We accomplish this by computing the differences between abstract syntax trees of every consecutive pair of commits (Figure~\ref{fig:context}).
 
@@
 This captures the structure of the code in an abstract and tractable manner.
-Converting raw source text to an AST is often one of the first steps performed by compilers.
+Because of its ease of analysis, converting raw source text to an AST is often one of the first steps performed by a compiler, performed to facilitate semantic analysis of the source code.
 When a programmer makes a non-trivial change in code in a file, that difference will be reflected in the AST.
-This means that the AST of revisions in code repositories can be compared to see what structures a programmer is modifying.
+Non-trivial changes are those that affect how a compiler will interpret the source code, and may have effects on the executable code.
+Trivial changes are those that have no affect on how a compiler will interpret the source text, such as changing insignificant whitespace or comments in source code.
+Thus, ASTs of consecutive revisions in a code repository can be compared to see what \emph{structures} a programmer is modifying.
 
@@
 Spoon is a tool for transforming and analyzing Java source code.
-It breaks code up into a meta-model including the AST itself.
-The model consists of three parts: structural elements, code elements and references to program elements.
-According to Spoon's authors: ''The structural part contains the declarations of the program elements, such as interface, class, variable, method, annotation, and enum declarations.
+It breaks code up into a meta-model consisting of three parts: structural elements, code elements and references to program elements.
+According to Spoon's authors: ``The structural part contains the declarations of the program elements, such as interface, class, variable, method, annotation, and enum declarations.
 The code part contains the executable Java code, such as the one found in method bodies.
 The reference part models the references to program elements (for instance a reference to a type)''~\cite{pawlak:hal-01169705}.
-The Spoon model is convenient because it provides the infrastructure for performing AST differencing (described later) while retaining the code elements (such as type information) for analysis afterwards.
+The Spoon model is convenient because it provides the infrastructure for performing AST differencing (described later) while retaining code elements for analysis afterwards.
 Spoon also preserves all type and package (library) information, information vital to our semantic analysis.
@@
 To determine what has changed between two subsequent commits, we computed their \emph{AST difference}.
-An AST difference (or simply \emph{AST diff}) states which nodes were added and removed from a tree (step two in Figure~\ref{fig:context}).
-For this purpose, we used GumTree~\cite{falleri:hal-01054552}, an algorithm made to compute the difference between two ASTs generated by Spoon.
-For new files and deleted files we did not use GumTree; we were simply able to run Spoon on the files and treated everything as an addition or deletion respectively.
+Given two ASTs, an AST difference (or simply \emph{AST diff}) states which nodes were must be added and which nodes must be removed to transform one tree to the next (step two in Figure~\ref{fig:context}).
+This allows us to precisely track changes two consecutive revisions of the same file.
+For AST diffing, we used GumTree~\cite{falleri:hal-01054552}, an algorithm made to compute the difference between two ASTs generated by Spoon.
+For new files and deleted files, it is unnecessary to calculate an AST difference; we simply run Spoon on the files and treated everything as an addition or deletion respectively.
 
-Once we generated ASTs, we were able to traverse through the trees to count the number of added or deleted \emph{declarations} and \emph{invocations} for a given type.
+Once we calculate the AST diffs, we were able to traverse through the trees to count the number of added or deleted \emph{declarations} and \emph{invocations} for a given type.
 
@@
 \item [declaration] 
-  Stating that use of a field or variable of a given type. For example declaration of a String type variable might look as follows:
+  Stating that use of a field or variable of a given type. For example declaration of a \texttt{String} type variable might look as follows:
   \\
   \begin{adjustwidth}{2.5em}{0pt}
-  String myString; \\
+  \texttt{String myString;} \\
   \end{adjustwidth}
@@
   \begin{adjustwidth}{2.5em}{0pt}
-  myString.trim(); \\
+  \texttt{myString.trim();} \\
   \end{adjustwidth}
@@
 
-An example of what our analysis tool might produce is as follows:
+Each change of declaration or invocation is detected and tallied per file per commit.
+These changes are emitted in line-based file format.
+For example, the following is the output of a commit that changed two declaration of \texttt{String} variables to \texttt{URI} variables:
 
@@
 
-\begin{figure}[!h]
+\begin{figure}[h]
 \centering
@@
 We did this by running \texttt{git log} over the Git repository and listing all the commit IDs in chronological order (Figure~\ref{fig:commits}).
-Each commit and its parent were then compared using AST differencing as described above.
-There are a few edge cases; however, if a commit has no parent, then every declaration or invocation in the commit was treated as an addition.
+For each commit we dumped the following information to a file: author's name, date of commit, commit id, commit message, files the commit edited, and files in the repository at the time of the commit.
+For all the edited files we also compared the file at the time of the commit with its parent commit using the AST differencing as described above.
+We achieved this by using calling \texttt{git difftool} and specifying the tool to be used as our AST differencing tool.
+One case where \texttt{git difftool} fails is when the commit has no parent.
+We were able to handle this by calling \texttt{git show} and saving the output to a temp file to be used with the AST differencing tool.
+Here we treated everything as an addition.
 Since we compared each commit to its parent, as reported by Git, we were able to account for any branching that happened outside the primary branch of the repository (usually named \texttt{master}).
@@
 
-\section{TypeV: The Visualization Tool}
+\begin{figure}[h]
+\centering
+\includegraphics[width=\columnwidth]{data-extracted}
+\caption{Hierarchy of data extracted by TypeV.
+Each repository may have zero or more commits; each commit may have zero or more changed files; each changed files may have zero or more changes, as given by our AST differencing tool.}
+\label{fig:data-extracted}
+\end{figure}
+
+The data collected from running the AST differencing component over the commits was saved to a file to be visualized later.
+In total, the information collected from one repository includes all commits including metadata for each commit such as its author, date, SHA (or revision ID), and commit message.
+Each commit may be composed one zero or more changed files; for each changed Java file, our tool extracts a change---either a declaration or an invocation---the artefact affected---a type or a method---and whether the change is an addition or a deletion of said artefact.
+Figure~\ref{fig:data-extracted} summarizes the hierarchy of the data extracted by our tool.
+
+\section{Visualization}
 
@@
 \includegraphics[width=\textwidth]{Screenshot_antlr4}
-\caption{Screenshot showing the main view for TypeV. The data being displayed is from the antlr4 repository.}
+\caption{Screenshot showing the main view for TypeV.
+The data being displayed is from the ANTLR4 repository.
+Light blue bars \additionsbox{10pt} indicate \textbf{additions}; dark red bars \deletionsbox{10pt} indicate \textbf{deletions}.}
 \label{fig:main-view}
@@
 
-The data would be difficult to analyze if displayed as its raw text form; instead, we created an HTML5 web visualization application which is intended for developers, project managers, and researchers to use (Figure~\ref{fig:main-view}).
-The source code for TypeV is available online on GitHub,\footnote{\url{https://github.com/mdfeist/TypeV}}.
+Although rich in detail, the data is difficult to analyze if displayed in its raw text form; instead, we created an HTML5 web visualization application which is intended for developers, project managers, and researchers to use (Figure~\ref{fig:main-view}).
+The source code for TypeV---both the AST differencing component and the web application--- is available online on GitHub.\footnote{\url{https://github.com/mdfeist/TypeV}} A screencast demonstrating its use is also available online.\footnote{\url{https://archive.org/details/typev-demo}}
 
-TypeV provide a variety of views to help organize the data and provide different levels of detail.
+Since we have the date and authors attached to each commit we are able to group commits together.
+This allows us to show only changes for a specific author or group of authors.
+We can also show changes that happened in a specific time period.
+
+TypeV provides a variety of views to help organize the data and provide different levels of detail.
 The main view displays a series of stacked bar graphs which show changes to types over a period of time in a project (Figure~\ref{fig:main-view}).
-The blue and orange portions account for the number of additions and deletions in a similar fashion to GitHub’s code frequency graph (Figure~\ref{fig:GH_code_freq}).
+The light blue  \additionsbox{10pt} and dark red \deletionsbox{10pt} portions account for the number of additions and deletions respectively, in a similar fashion to GitHub's code frequency graph (Figure~\ref{fig:GH_code_freq}).
 The height of each bar shows the number of changes to a type.
@@
 A bar which fills the row vertically is the period with the largest number of changes to that type out of the visible time periods.
-The other bars in the row are proportional to that maximum number.
+The other bars in the row are proportional to that maximum number of changes.
 The bar height is calculated by multiplying the maximum bar height by the number of changes during that time period divided by the maximum number of changes attested for this type.
@@
 
-\begin{table}[!t]
+\begin{table}[t]
+\renewcommand{\arraystretch}{1.3}
+\centering
+\caption{Summary of projects used in the evaluation.}
+\label{tab:project-summary}
+\begin{tabular}{c|lllll}
+\hline
+\bfseries Project & \bfseries Commits & \bfseries Authors* & \bfseries Files &   \bfseries Types &  \bfseries  Diffs\\
+\hline
+ANTLR4 & 3930 & 27 & 1793 & 2635 & 58425\\
+BookKeeper & 660 & 12 & 758 & 1872 & 31817\\
+Curator & 1524 & 43 & 1211 & 1269 & 27957\\
+Tika & 860 & 10 & 604 & 757 & 20764\\
+\hline
+\multicolumn{6}{p{\linewidth}}{\footnotesize Totals for the four projects we analyzed.
+``Diffs'' refers to the total number of invocation and declaration changes observed from the AST.
+Note that the number given for authors may be larger than the actual number of collaborators on the project, due to some collaborators using multiple email address, and thus appearing as more than one author.}
+\end{tabular}
+\end{table}
+
+\begin{table}[t]
 \renewcommand{\arraystretch}{1.3}
@@
 \hline
-antlr4 & $p\ll0.01$ & 0.943 & 0.907\\
-bookkeeper & $0.313$ & 0.984 & 0.972\\
-curator & $0.007$ & 0.975 & 0.980\\
-tika & $0.653$ & 0.955 & 0.998\\
+ANTLR4 & $p\ll0.01$ & 0.943 & 0.907\\
+BookKeeper & $0.313$ & 0.984 & 0.972\\
+Curator & $0.007$ & 0.975 & 0.980\\
+Tika & $0.653$ & 0.955 & 0.998\\
 \hline
@@
 
+In Java, a type can be a primitive, provided by the language itself, or an object reference.
+Primitive types are those that are built into the language such as an \texttt{int} or \texttt{char}.
+Object references point to instances of classes.
+Java's syntax encourages the creation of classes, thus we expect substantial Java projects to define many domain-specific types.
+Additionally, Java places the constraint that each source file must define one class.
+When writing or contributing to a project, the author's choice of which part of the program to edit will in turn affect what types they end up working with --- that is, it will affect the author's \emph{type coverage}.
+Similarly, the files an author works with
+
 The information regarding which types are used in each change is unique to the AST diff method, so it will be useful to see how changes affect the types in a project rather than just the files involved in a diff.
@@
 File coverage is the same measurement, but using the number of source files instead.
-To determine if type coverage gives different information than file coverage, we analyzed four popular Java projects: Antlr, Apache Bookkeeper, Apache Curator and Apache Tika.
+To determine if type coverage gives different information than file coverage, we analyzed four open source Java projects: ANTLR, Apache Bookkeeper, Apache Curator and Apache Tika.
+A summary of project statistics is given in Table~\ref{tab:project-summary}.
 We counted the file coverage and type coverage on a commit-by-commit basis.
@@
 As type coverage and file coverage are both cumulative measures, positive correlation is expected.
-However, even when the distributions are significantly different (as with Antlr and Curator), the correlation is still strongly positive, but not 100\%.
+However, even when the distributions are significantly different (as with ANTLR and Curator), the correlation is still strongly positive, but not 100\%.
 This means that when the coverage for types and files of a commit increase, the relationship is not necessarily 1-to-1.
@@
 
-\section{Case Study}
+\section{Illustrative Case Studies}
 
-In the following sections we present case studies demonstrating the extra insight that the TypeV method makes available.
+In the following sections, we present case studies demonstrating the extra insight that the AST differencing provides.
 
-\subsection{antlr4}
+\subsection{ANTLR4}
 
@@
 \includegraphics[width=\columnwidth]{ParserATNSimulator}
-\caption{Snapshot of TypeV showing changes to ParserATNSimulator in antlr4. Date ranges from approximately the begining of 2011 to the end of April, 2012.}
+\caption{Snapshot of TypeV showing changes to \texttt{ParserATNSimulator} in ANTLR4.. Date ranges from approximately the beginning of 2011 to the end of April, 2012.}
 \label{fig:parser1}
@@
 \includegraphics[width=\columnwidth]{Antlr4-GH-Lines-ParserATNSimulator}
-\caption{GitHub's line difference graph showing changes to antlr4. The time spans approximately the same length as Figure~\ref{fig:parser1}.}
+\caption{GitHub's code frequency graph (based on line difference) showing changes to antlr4. The time spans approximately the same length as Figure~\ref{fig:parser1}.}
 \label{fig:parser2}
@@
 
-For example, in December of 2011 we can see that the ParserATNSimulator class has a large amount of additions and deletions in roughly equal proportion (Figure~\ref{fig:parser1}).
-This suggests that ParserATNSimulator is not being removed from the project but instead there is some refactoring being done.
+For example, in December of 2011 we can see that the \texttt{ParserATNSimulator} class has a large amount of additions and deletions in roughly equal proportion (Figure~\ref{fig:parser1}).
+This suggests that \texttt{ParserATNSimulator} is not being removed from the project but instead there is some refactoring being done.
 If we take a closer look at the commit messages we can see that they mention refactoring and reorganization of code.
-Zooming in on that date,  shows that during that month they were finishing work on a new version of ParserATNSimulator.
+Zooming in on that date,  shows that during that month they were finishing work on a new version of \texttt{ParserATNSimulator}.
 
 If we tried to do the same analysis with the GitHub tools it would be significantly more difficult to achieve the conclusion achieved using TypeV (Figure~\ref{fig:parser2}).
-First if we only look at line changes this change to ParserATNSimulator is overwhelmed by other changes and hardly shows up on the GitHub tools.
+First if we only look at line changes this change to \texttt{ParserATNSimulator} is overwhelmed by other changes and hardly shows up on the GitHub tools.
 If we did notice an increase in activity we would not know what types were being worked on.
@@
 \includegraphics[width=\columnwidth]{bookkeeper_dep}
-\caption{Shows deprecated invocation getLedgerManagerType() and replacement invocation getLedgerManagerFactoryClass() for Apache Bookkeeper.}
+\caption{Shows deprecated invocation \texttt{getLedgerManagerType()} and replacement invocation \texttt{getLedgerManagerFactoryClass()} for Apache Bookkeeper.}
 \label{fig:bookkeeper-depr}
@@
 
-Apache Bookkeeper deprecated the method AbstractConfiguration.getLedgerManagerType() in release 4.2.0\footnote{\url{https://bookkeeper.apache.org/docs/r4.2.0/apidocs/deprecated-list.html}} and replaced it with AbstractConfiguration.getLedgerManagerFactoryClass().
-If we look at getLedgerManagerType() we can see at the time of the start of the deprecation all instances were removed and getLedgerManagerFactoryClass() was added (Figure~\ref{fig:bookkeeper-depr}).
-After the deprecation, we see no more changes to getLedgerManagerType() and only some additions of getLedgerManagerFactoryClass().
+Apache Bookkeeper deprecated the method \texttt{AbstractConfiguration.getLedgerManagerType()} in release 4.2.0\footnote{\url{https://bookkeeper.apache.org/docs/r4.2.0/apidocs/deprecated-list.html}} and replaced it with \texttt{AbstractConfiguration.getLedgerManagerFactoryClass()}.
+If we look at \texttt{getLedgerManagerType()} we can see at the time of the start of the deprecation all instances were removed and \texttt{getLedgerManagerFactoryClass()} was added (Figure~\ref{fig:bookkeeper-depr}).
+After the deprecation, we see no more changes to \texttt{getLedgerManagerType()} and only some additions of \texttt{getLedgerManagerFactoryClass()}.
 
 \section{Discussion}
+
+Having tools for analyzing repositories based on type usage and method invocation benefits many parts of the software development process.
+Software management requires a high level understanding of the state of a repository and the changes being made which affect the functionality of the software.
+Tracking developer contributions in a less misleading way is also important for management since important contributions may be small and focused, but nevertheless, significant.
+Line differences only provide meaningful information when they are analyzed by someone with familiarity with the source code being changed.
+Providing tools which bring more meaning than line differences would allow people less familiar with the code to understand changes being made.
+When large changes are made that cover many aspects of a project, it is also useful to see how the change affects the software rather than a noting that several thousands of lines have changed.
+Seeing how a project has changed over time is also vital for management since evaluating the state of completion of software can be very difficult.
 
@@
 As a manager, having a tool like TypeV allows one to quickly summarize the work done by each group member, organized by the types they have been working on.
-This can show when intervention or guidance is required, without the large time commitment of looking through commits.
+This can show when intervention or guidance is required, without the large time commitment of pouring through commits.
 
-The purpose of deprecated software features is to give developers time to change projects to follow a new or better standard without breaking the code.
-Deprecated features can be marked for removal from a project or have some flaw that would make a different feature the better option.
+Type information indicates which libraries and classes are being used in a project.
+
+The purpose of a deprecated software features is to give developers time to change projects to follow a new or better standard without breaking the code.
+When a piece of code becomes deprecated, continuing its use may lead to design flaws or even security vulnerabilities.
+For example, in Apache Nifi, an earlier version of the software used a compromised key-derivation functions; hence, this class was deprecated, but left in for compatibility with older versions of the software~\cite{nifi}.
+Viewing changes in terms of types can show when usage of deprecated classes are added or removed from a project as well as how many remain over time.
 Given the time at which a feature was deprecated, seeing the progress of replacement has some important implications.
@@
 Removing all instances of deprecated features could also mean a project can be upgraded to use new tools or features that were not possible while the deprecated features were present.
-Here we show examples where deprecated classes were removed.
+In our Apache BookKeeper case study,  we demonstrated an example where deprecated classes were removed entirely.
 
@@
 
-RQ1: Using abstract syntax trees, we can determine the types affected by each commits in a repository. Type information is not directly available through regular line diffs recorded by Git.
+\textbf{RQ1:} How can we extract type and method usage over time from repositories?
 
-RQ2: The evolution of a project and developer activity can be seen through the changes to types in a project.
-Significant events can be observed with greater understanding by seeing the changes to types in addition to the existing commit information.
+\begin{framed}
+\noindent
+Using abstract syntax trees, we can determine the types affected by each commits in a repository. Type information is not directly available through regular line diffs recorded by Git.
+\end{framed}
 
-RQ3: Type coverage and file coverage are found to be highly correlated but do not always have the same distribution or a one to one ratio as might be expected in Java.
+\textbf{RQ2:} How does type coverage compare to file coverage?
+
+\begin{framed}
+\noindent
+Type coverage and file coverage are found to be highly correlated but do not always have the same distribution or a one to one ratio as might be expected in Java.
+\end{framed}
+
+\textbf{RQ3:} Does type information give any interesting or useful insight into the project?
+
+\begin{framed}
+\noindent
+The evolution of a project and developer activity can be seen through the changes to types in a project.
+Significant events can be observed with greater understanding by seeing the changes to types in addition to existing commit information.
+\end{framed}
 

### 2016-04-26 paper/document.tex
Final draft of paper.

@@
 What is a developer's contribution to a repository?
-By only counting commits and number of lines changed, existing tools that analyze source code repositories fall short on showing the effective contributions made by each developer.
+By only counting commits and number of lines changed, existing tools that visualize source code repositories (such as GitHub's graphs) fall short on showing the effective contributions made by each developer.
 When many commits are viewed as a group, the details are lost.
@@
 
-\begin{figure*}[!t]
+\begin{figure*}[t]
 \centering
@@
 While programmers may type in individual characters, their intention is at a much higher level.
-Their intent is written in tokens that represent statements, identifiers and types that are used to solve their problems.
+Their intent is written in tokens that represent statements, identifiers, and types that are used to solve their problems.
 Programmers solve problems in a way that is readable to other developers and executable by a computer.
@@
 
+Consider GitHub graphs~\cite{github-graphs}, visualizations that are easily available to open source software developers.
+GitHub's graphs use line and commit information exclusively.
+In the case of the ANTRL4 project (Figure~\ref{fig:GH_code_freq}), this code frequency chart creates more questions than it answers.
+What happened at those big spikes of line changes in the graph?
+One of the spikes is so large, we had to crop the graph to be able to show details of other changes.
+How did these large changes affect the code base?
+Who committed these large changes?
+Was it a team of contributors or one particularly prolific programmer?
+Plotting the number of lines changed may be misleading.
+A huge spike in additions may be caused by a number of reasons, such as the inclusion of automatically-generated documentation, the inclusion of an external library, or perhaps a massive refactoring effort \cite{Hindle:2008:LCT:1370750.1370773}.
+
 The goal of a software evolution visualization is to provide an alternate view that provides insight into the development of the software that is otherwise unavailable.
@@
 
-In the following sections of the paper we will describe our tool, TypeV, that we developed to compute and visualizes changes to types used in a project over time.
+In the following sections of the paper we will describe our tool, TypeV, that we developed to compute and visualize changes to types used in a project over time.
 TypeV is able to extract type information from a repository by computing commit-by-commit abstract syntax tree differences, which we describe in section~\ref{sec:methodology}.
@@
 \textbf{RQ1:} How can we extract type and method usage over time from repositories? \\ \\
-\textbf{RQ2:} Are type coverage file coverage different metrics?
+\textbf{RQ2:} How does type coverage compare to file coverage?
 \\ \\
@@
 
+We will examine research which analyses software repositories for information on language and features usage, tools for visualizing project evolution, collaboration in software repositories and other tools for understanding repositories.
+
 \subsection{Analysis of Languages and Features}
 
-Analyzing source code is a popular research topic and Java is often chosen as the subject due to its popularity.
+Analyzing source code is a popular research topic and Java is often chosen as the subject of study due to its popularity.
 Grechanik \textit{et al.} \cite{Grechanik:2010:EIL:1852786.1852801} examined the structure of Java programs mined from 2080 programs.
@@
 L\"{a}mmel \textit{et al.} \cite{Lammel:2011:LAA:1982185.1982471} used ASTs to examine the usage of APIs in Java projects.
-They found what APIs are popular and if they were used in a framework like manner.
+They found what APIs are popular and if they were used in a framework-like manner.
 Dyer \textit{et al.} \cite{Dyer:2014:MBA:2568225.2568295} mined AST nodes to study the use of new Java language features over time.
-The authors found the most popular features and how teams adopted new features.
-New features varied greatly in popularity and adoption rate.
+The authors found the most popular features and the adoption rate of new features over time.
 They did not check to see how much a developer used a feature but instead checked to see if they used it at all.
@@
 The focus of these papers is on how programming languages and features are used by looking at many repositories.
-Java is a strongly typed language, yet there is no focus specifically on how types are used throughout a project.
-Type info may provide additional information on how developers use languages but has not been explored.
+Java is a statically-typed language, yet there is no focus specifically on how types are used throughout a project.
+Type data may provide additional information on how developers use languages but this has yet to be explored.
 
@@
 
-A variety of visualizations tools get created in order to help with understanding the changes that get made to software over time.
+A variety of visualizations tools have been created in order to help with understanding the changes that get made to software over time.
 Wu \textit{et al.} \cite{wu2004a} explored punctuated changes to software repositories with a tool that generates an Evolution Spectrograph.
@@
 The colour is determined by the frequency of change.
-This work was continued to show file directoryies and display the cardinality of a developer's commits Wu \textit{et al.} \cite{wu2004}.
+This work was continued to show file directories and display the cardinality of a developer's commits in Wu \textit{et al.} \cite{wu2004}.
 The cardinality of a commit is the number of files or subsystems it affects in a system, indicating the scope of the change.
-Other tools also look at changes in code dependencies \cite{6980224} or files \cite{6650525, 7332421} over time.
+Other tools also look at changes in code dependencies \cite{6980224} or files \cite{6650525, 7332421,voinea2006} over time.
 
-Voinea \textit{et al.} \cite{voinea2006} visualizes CVS repositories using CVSgrab and adding their own enhancements to show changes to files over time.
-Files can be sorted by creation time, activity, or alphabetically and colour coded based on author, file size, file type or by search strings.
+Tools also exist which track changes over time in a more complex manner.
 Kim \textit{et al.} \cite{Kim:2009:DRS:1555001.1555046} built a tool called Logical Structural Diff (LSdiff) which examines code changes and produces logical rules to represent them.
 Changes which do not match the rules are then pointed out to the developer.
-They also formed a focus group with software engineers who found that LSdiff complements existing tools well and showed promise for discovering inconsistent changes.
+Reiss and Tarvo created an IDE which allows users to see the history of files and changes to lines of code from each author as they are editing a project \cite{6650521}.
+Van Hees and Hage use a voronoi treemap to visualize changes in class and method structure \cite{7332410}.
+Servant and Jones created CHRONOS, a tool which creates a timeline of changes to a selection of code as well as the ability to query those changes \cite{6650547}.
+
+Some tools examine the evolution of projects to learn other information about them.
+Alcocer \textit{et al.} \cite{6650523} create a Performance Evolution Blueprint, which displays the changes in performance over time as a project evolves.
+Ens \textit{et al.} created ChronoTwigger to visualize the co-evolution of source and test files \cite{6980223}.
+This allowed testing practices to be more closely examined.
+
+None of the aforementioned tools perform comparisons of ASTs or specifically examine the changes to type information for their view of project evolution.
 
@@
 These role were defined as their level of contribution as well as the types of issues they dealt with.
-Developers will take on multiple roles in a project and will sometimes fulfill the same role across different projects.
-Lee \textit{et al.} \cite{lee2013} developed a tool for visualizing the branching structure of git repositories.
+They found that developers will take on multiple roles in a project and will sometimes fulfill the same role across different projects.
+Lee \textit{et al.} \cite{lee2013} developed a tool for visualizing the branching structure of Git repositories.
 This was used to analyze concurrent workflows in popular open source repositories.
+Elsen also created a visualization for branches in Git, with the ability to expand branches to view the directory structure of a project at that point \cite{6650522}.
+Shrestha \textit{et al.} made a visualization for the geographic location of a commit and \cite{6650532} and Minelli \textit{et al.} \cite{6980226} looked at the real time changes developers made to a repository.
 
@@
 
-Some visualization tools which are used for managing or making decisions about software rather than study the changes taking place.
+Some visualization tools are used for managing or making decisions about software rather than study the changes taking place.
 The service EasyBi \cite{EasyBi} provides additional visualizations for Git repositories for the purpose of business decisions.
@@
 It tries to solve the issue of summarizing many commits and has greater controls for granularity and authors but still operates on commits and line differences.
-The Gitinspector project \cite{Gitinspector} is an open source tool for the statistical analysis of git repositories.
+The Gitinspector project \cite{Gitinspector} is an open source tool for the statistical analysis of Git repositories.
 It provides an advanced timeline view of changes per author and cumulative statistics but again relies upon the line differences to summarize the work of each author.
 
-The GitHub tools have the strength that they do not rely on what type of code is being committed.
-It does not matter what language or if errors exist in the commit, GitHub will simply spit out the number of line additions and deletions.
-The issue with this type of analysis is that it does not give a clear breakdown of is being changed only that changes have occurred.
-
-Consider tools easily available to an open source software developer; for example GitHub graphs (see Figure~\ref{fig:GH_code_freq})~\cite{github-graphs}.
-Since GitHub tracks contributions based solely on number of commits and numbers of lines changes, data may be lacking, or even misleading.
-In the case of the ANTRL4 project (Figure~\ref{fig:GH_code_freq}), the graph creates more questions than it answers.
-What happened at those big spikes of line changes in the graph?
-One of the spikes is so large, we had to truncate the graph to be able to show details of other changes.
-How did these large changes affect the code base?
-Who committed these large changes?
-Was it a team of contributors or one particularly prolific programmer?
-Plotting the number of lines changed may be misleading.
-A huge spike in additions may be caused by numerous reasons, such as the inclusion of automatically-generated documentation, the inclusion of an external library, or perhaps even a huge refactoring effort.
+The basic tools easily available to any software developer are the GitHub graphs (see Figure~\ref{fig:GH_code_freq})~\cite{github-graphs}.
+GitHub tracks contributions based only on the number of commits and number of lines changed.
+This has the advantage of not relying on the type of code which was committed.
+It does not matter what language is used or if errors exist in the commit, GitHub will simply output the number of line additions and deletions.
+The issue with this type of analysis is that it does not give a clear idea of is being changed, only that changes have occurred.
+This data may be lacking, or even misleading which contributes to the need for better tools.
 
@@
 We accomplish this by computing the differences between abstract syntax trees of every consecutive pair of commits (Figure~\ref{fig:context}).
+This extracted data can then be used in a variety of visualizations as discussed in Section \ref{sec:viz}.
 
@@
 This captures the structure of the code in an abstract and tractable manner.
-Because of its ease of analysis, converting raw source text to an AST is often one of the first steps performed by a compiler, performed to facilitate semantic analysis of the source code.
 When a programmer makes a non-trivial change in code in a file, that difference will be reflected in the AST.
@@
 \begin{description}
-\item [declaration] 
+\item [declaration:] 
   Stating that use of a field or variable of a given type. For example declaration of a \texttt{String} type variable might look as follows:
@@
   \end{adjustwidth}
-\item [invocation] 
+\item [invocation:] 
   A call site for a method that belongs to a type. For example an invocation of the trim method of a String type might look as follows:
@@
 Note the fully qualified type names.
-The source text may not literally have the characters \texttt{java.net.URI}, but Spoon was able to infer from the import statements in the Java source text that references to ''URI'' resolve to the fully-qualified type name \texttt{java.net.URI}.
+The source text may not literally have the characters \texttt{java.net.URI}, but Spoon was able to infer from the import statements in the Java source text that references to ``URI'' resolve to the fully-qualified type name \texttt{java.net.URI}.
 Thus, TypeV is able to keep track of the actual class being used and is not confused when two classes in separate packages have the same name.
@@
 We did this by running \texttt{git log} over the Git repository and listing all the commit IDs in chronological order (Figure~\ref{fig:commits}).
-For each commit we dumped the following information to a file: author's name, date of commit, commit id, commit message, files the commit edited, and files in the repository at the time of the commit.
-For all the edited files we also compared the file at the time of the commit with its parent commit using the AST differencing as described above.
+For each commit, we collected the following information, as can be seen in Figure~\ref{fig:data-extracted}: the author's name, date of commit, commit ID (SHA), commit message, files the commit edited, and files in the repository at the time of the commit.
+For all the edited Java files we also compared the file at the time of the commit with its parent commit using AST differencing as described above.
 We achieved this by using calling \texttt{git difftool} and specifying the tool to be used as our AST differencing tool.
-One case where \texttt{git difftool} fails is when the commit has no parent.
-We were able to handle this by calling \texttt{git show} and saving the output to a temp file to be used with the AST differencing tool.
-Here we treated everything as an addition.
+One case where \texttt{git difftool} fails is when the commit has no parent, such as the initial commit.
+We were able to handle this by calling \texttt{git show} and saving the output to a temporary file to be used with the AST differencing tool.
+When calling \texttt{git show} instead of \texttt{git difftool} we treat everything as an addition, as in commit A in Figure \ref{fig:commits}.
 Since we compared each commit to its parent, as reported by Git, we were able to account for any branching that happened outside the primary branch of the repository (usually named \texttt{master}).
-If a commit has more than one parent, this indicates a merge commit in Git.
+If a commit has more than one parent, like commit D in Figure \ref{fig:commits}, it indicates a merge commit in Git.
 Merge commits contain the changes made in each parent commit combined.
@@
 \caption{Hierarchy of data extracted by TypeV.
-Each repository may have zero or more commits; each commit may have zero or more changed files; each changed files may have zero or more changes, as given by our AST differencing tool.}
+Each repository may have zero or more commits; each commit may have zero or more changed files; each changed files may have zero or more changes, as computed by our AST differencing tool.}
 \label{fig:data-extracted}
@@
 
-The data collected from running the AST differencing component over the commits was saved to a file to be visualized later.
-In total, the information collected from one repository includes all commits including metadata for each commit such as its author, date, SHA (or revision ID), and commit message.
-Each commit may be composed one zero or more changed files; for each changed Java file, our tool extracts a change---either a declaration or an invocation---the artefact affected---a type or a method---and whether the change is an addition or a deletion of said artefact.
+The data collected from running the AST differencing component over the commits is saved to a file to be visualized later; this data is in a general format, and is not bound to any particular visualization.
 Figure~\ref{fig:data-extracted} summarizes the hierarchy of the data extracted by our tool.
@@
 
-\begin{figure*}[!t]
+\label{sec:viz}
+
+\begin{figure*}[!ht]
 \centering
@@
 
-Although rich in detail, the data is difficult to analyze if displayed in its raw text form; instead, we created an HTML5 web visualization application which is intended for developers, project managers, and researchers to use (Figure~\ref{fig:main-view}).
+Although rich in detail, the AST diff data is difficult to analyze if displayed in its raw form; hence, we created an HTML5 web visualization application which is intended for developers, project managers, and researchers to use (Figure~\ref{fig:main-view}).
 The source code for TypeV---both the AST differencing component and the web application--- is available online on GitHub.\footnote{\url{https://github.com/mdfeist/TypeV}} A screencast demonstrating its use is also available online.\footnote{\url{https://archive.org/details/typev-demo}}
 
-Since we have the date and authors attached to each commit we are able to group commits together.
-This allows us to show only changes for a specific author or group of authors.
-We can also show changes that happened in a specific time period.
+\subsection{Views}
 
 TypeV provides a variety of views to help organize the data and provide different levels of detail.
-The main view displays a series of stacked bar graphs which show changes to types over a period of time in a project (Figure~\ref{fig:main-view}).
+The overview displays several rows of stacked bar graphs which show changes to a type or method over a period of time in a project (Figure~\ref{fig:main-view}).
+The purpose of displaying stacked bar charts is that they are a familiar visualization, thus are somewhat easy to read, and can succinctly display the relative amounts of additions and deletions per each time period.
 The light blue \additionsbox{10pt} and dark red \deletionsbox{10pt} portions account for the number of additions and deletions respectively, in a similar fashion to GitHub's code frequency graph (Figure~\ref{fig:GH_code_freq}).
-The height of each bar shows the number of changes to a type.
-The width of each bar represents a set time period and can be changed to an hour, day, week, or month.
-A bar which fills the row vertically is the period with the largest number of changes to that type out of the visible time periods.
-The other bars in the row are proportional to that maximum number of changes.
+These colours were chosen as they are high contrast, and are distinct to those red-green colour blindness.
+
+The height of each bar shows the relative number of changes to a type.
+The width of each bar is fixed, and represents a time period for binning commits that can be configured to an hour, day, week, or month's worth of commits per each bar.
+A bar that completely fills its row vertically is the period with the largest number of changes to that type in the visible date range.
+The height of the other bars in the row are in proportion to the maximum number of changes.
 The bar height is calculated by multiplying the maximum bar height by the number of changes during that time period divided by the maximum number of changes attested for this type.
-The rows are sorted so that the graph for the most used types appear the top and are in descending order.
+The rows are sorted so that the graph for the most used types or methods appear the top and are in descending order of use.
 
-When a user mouses over a specific bar in the graph they are given information for the type and time period at that location.
+One can interact with the overview by hovering and clicking on the bars.
+When a user mouses over a specific bar in the graph they are given summary information for the type and time period at that location.
 The information includes the fully-qualified name of the type, the number of commits, the number of authors, the number of additions, the number of deletions, and the start and end date of that time period.
 If the user clicks on the bar, they are given more detailed statistics which list the authors who made the commits and each commit message.
+By adding interaction, the user is able to explore the overview, find the specific information they are interested and drill-down at their leisure.
+Although our tool extracts a great deal of information, by using interaction, users can gradual reveal more information as they need it.
 
-One issue with analyzing individual authors is that authors can commit from a variety of different computers which can have different emails and usernames attached to the Git configuration.
+To allow for a more detailed analysis of a specific component of the repository we added some basic filtering tools.
+Users can filter by type name; they can view changes that fall within a specific start and end date; users can also choose to see only declarations of types, types used in declarations and invocations, or only invocations.
+One issue with analyzing individual authors is that authors can commit from a variety of different computers which can have different emails and usernames attached to their Git configuration.
 This may result in one author appearing as several different users in the \texttt{git log}.
@@
 
-To allow for a more detailed analysis of a specific component of the repository we added some basic filtering tools.
-First a user can filter by type name.
-Second the user can view changes that fall within a specific start and end date.
-Third the user can specify the time span represented by each bar: hour, day, week, or month.
-Finally, the user can choose to see only declarations of types, types used in declarations and invocations, or only invocations.
+Knowing type and author information enables other kinds of visualizations.
+For example, Figure \ref{fig:chord} is a chord diagram which shows connections between authors and types.
+The wider the arc around the circumference of the circle, the more often that entity is connected to other entities.
+For example, ``parrt'' is the top contributor, with changes covering a large number of types.
+Determining author and type information is made possible by AST differencing.
+
+\begin{figure*}[p]
+\centering
+\includegraphics[width=\textwidth]{chord}
+\caption{Chord diagram showing connections between authors (right and bottom) to types (left).}
+\label{fig:chord}
+\end{figure*}
 
@@
 \hline
-\bfseries Project & \bfseries Commits & \bfseries Authors* & \bfseries Files &   \bfseries Types &  \bfseries  Diffs\\
+\bfseries Project & \bfseries Commits & \bfseries Authors* & \bfseries Files &   \bfseries Types &  \bfseries  Changes\\
 \hline
@@
 \multicolumn{6}{p{\linewidth}}{\footnotesize Totals for the four projects we analyzed.
-``Diffs'' refers to the total number of invocation and declaration changes observed from the AST.
+``Changes'' refers to the total number of invocation or declaration changes observed from the AST.
 Note that the number given for authors may be larger than the actual number of collaborators on the project, due to some collaborators using multiple email address, and thus appearing as more than one author.}
@@
 
-\begin{table}[t]
+\begin{table}[ht]
 \renewcommand{\arraystretch}{1.3}
@@
 
-\begin{figure}[h]
+\begin{figure}[ht]
 \centering
 \includegraphics[width=\columnwidth]{antlr-density}
-\caption{}
+\caption{Density plot of type and file coverage for commits.
+The graph shows what percentage of type and file coverage the majority of commits have.
+One can see that the majority of commits have a type and file coverage around 0\% to 20\%.
+We can also see a difference between these measurements on the higher end of the graph.
+File coverage seems to have a group of commits that had a file coverage between 20\% and 50\% whereas type coverage have a similar looking grouping but between 40\% and 60\%.}
 \end{figure}
 
-\begin{figure}[h]
+\begin{figure}[ht]
 \centering
 \includegraphics[width=\columnwidth]{tika-density}
-\caption{}
+\caption{Density of type and file coverage for commits.
+Here we can see that commits did not have much file coverage between 25\% and 75\%.
+On the other hand there were a lot of commits that have type coverage around 50\%.}
 \end{figure}
+
+Is it worth going through the effort of computing commit-by-commit AST differences if we can just simply count changes to files instead?
+We define \emph{type coverage} as the number of types that have been touched out of the total number of types during either a certain time period or the entire project.
+\emph{File coverage} is the same measurement, but using the number of source files touched instead.
+File coverage can be trivially determined from an author's commit history; Type coverage, however, is uniquely provided by AST differencing, as we have described.
 
@@
 Additionally, Java places the constraint that each source file must define one class.
-When writing or contributing to a project, the author's choice of which part of the program to edit will in turn affect what types they end up working with --- that is, it will affect the author's \emph{type coverage}.
-Similarly, the files an author works with
+When writing or contributing to a project, the author's choice of which part of the program to edit will in turn affect what types they end up working with --- that is, it will affect the author's type coverage.
+Similarly, the set of files that an author works with affects that author's file coverage.
 
-The information regarding which types are used in each change is unique to the AST diff method, so it will be useful to see how changes affect the types in a project rather than just the files involved in a diff.
-We define type coverage as the number of types that have been touched out of the total number of types during either a certain time period or the entire project.
-File coverage is the same measurement, but using the number of source files instead.
-To determine if type coverage gives different information than file coverage, we analyzed four open source Java projects: ANTLR, Apache Bookkeeper, Apache Curator and Apache Tika.
-A summary of project statistics is given in Table~\ref{tab:project-summary}.
-We counted the file coverage and type coverage on a commit-by-commit basis.
+To determine if type coverage yields different information than file coverage, we analyzed four open source Java projects: ANTLR4, Apache BookKeeper, Apache Curator and Apache Tika.
+Three of the four projects were collected from the Apache foundation, as they host a large number of well-curated open source Java projects.
+A summary of the projects studied is given in Table~\ref{tab:project-summary}.
+We counted the file coverage and type coverage on a commit-by-commit basis for each author.
 That is, each datum represents a commit with three dimensions: file coverage and type coverage, and its commit date.
 Data is sorted in ascending order of commit date.
-Table~\ref{tab:summary} lists a summary of the data provided.
-In general, all measures are strongly correlated, both for Pearson (linear) correlation, and Spearman (rank) correlation.
+Table~\ref{tab:summary} lists a summary of the statistics measured.
+In general, type coverage and file coverage are strongly correlated, both using Pearson (linear) correlation, and Spearman (rank) correlation.
 As type coverage and file coverage are both cumulative measures, positive correlation is expected.
 However, even when the distributions are significantly different (as with ANTLR and Curator), the correlation is still strongly positive, but not 100\%.
-This means that when the coverage for types and files of a commit increase, the relationship is not necessarily 1-to-1.
-This is remarkable in a language like Java, a classes and files are generally in a 1-to-1 relationship.We conclude from this that type coverage yields related, yet different information than file coverage alone.
+This means that when the coverage for types and files of a commit increase, the relationship is not necessarily one-to-one.
+This is remarkable in a language like Java, as classes and files are generally in a one-to-one relationship.
+
+We conclude from this that type coverage yields related, yet different information than file coverage alone.
 
@@
 \includegraphics[width=\columnwidth]{ParserATNSimulator}
-\caption{Snapshot of TypeV showing changes to \texttt{ParserATNSimulator} in ANTLR4.. Date ranges from approximately the beginning of 2011 to the end of April, 2012.}
+\caption{Snapshot of TypeV showing changes to \texttt{ParserATNSimulator} in ANTLR4. Date ranges from approximately the beginning of 2011 to the end of April, 2012.}
 \label{fig:parser1}
@@
 \includegraphics[width=\columnwidth]{Antlr4-GH-Lines-ParserATNSimulator}
-\caption{GitHub's code frequency graph (based on line difference) showing changes to antlr4. The time spans approximately the same length as Figure~\ref{fig:parser1}.}
+\caption{GitHub's code frequency graph (based on line difference) showing changes to ANTLR4. The time spans approximately the same length as Figure~\ref{fig:parser1}.}
 \label{fig:parser2}
@@
 
-We chose Antlr4 because it is a substantial Java project with a commit history spanning several years. We can also see distinct changes throughout its history.
+We chose ANTLR4 because it is a substantial Java project with a commit history spanning several years. We can also see distinct changes throughout its history.
 
@@
 
-\subsection{Apache Bookkeeper}
+\subsection{Apache BookKeeper}
 
@@
 
-For deprecation analysis, we chose Apache Bookkeeper.
-As an Apache Foundation project it is a well-curated example of an open source software project.
-It also demonstrates an instance of deprecated methods.
-
-Apache Bookkeeper deprecated the method \texttt{AbstractConfiguration.getLedgerManagerType()} in release 4.2.0\footnote{\url{https://bookkeeper.apache.org/docs/r4.2.0/apidocs/deprecated-list.html}} and replaced it with \texttt{AbstractConfiguration.getLedgerManagerFactoryClass()}.
-If we look at \texttt{getLedgerManagerType()} we can see at the time of the start of the deprecation all instances were removed and \texttt{getLedgerManagerFactoryClass()} was added (Figure~\ref{fig:bookkeeper-depr}).
-After the deprecation, we see no more changes to \texttt{getLedgerManagerType()} and only some additions of \texttt{getLedgerManagerFactoryClass()}.
-
-\section{Discussion}
-
-Having tools for analyzing repositories based on type usage and method invocation benefits many parts of the software development process.
-Software management requires a high level understanding of the state of a repository and the changes being made which affect the functionality of the software.
-Tracking developer contributions in a less misleading way is also important for management since important contributions may be small and focused, but nevertheless, significant.
-Line differences only provide meaningful information when they are analyzed by someone with familiarity with the source code being changed.
-Providing tools which bring more meaning than line differences would allow people less familiar with the code to understand changes being made.
-When large changes are made that cover many aspects of a project, it is also useful to see how the change affects the software rather than a noting that several thousands of lines have changed.
-Seeing how a project has changed over time is also vital for management since evaluating the state of completion of software can be very difficult.
-
-When managing group projects in an introductory software engineering course, it is very important to be able to see the contributions that students are making to the group repository to judge learning, progress and collaboration.
-When supervising 4-5 groups of students, it becomes unfeasible to look through all of the commits of each student.
-Students may take the wrong approach in a certain piece of code which becomes ingrained in the project and very difficult to change.
-Many groups have problems with group members who do not complete their assigned work or make contributions which end up being meaningless.
-Issues like this can be difficult to detect and prevent.
-As a manager, having a tool like TypeV allows one to quickly summarize the work done by each group member, organized by the types they have been working on.
-This can show when intervention or guidance is required, without the large time commitment of pouring through commits.
-
-Type information indicates which libraries and classes are being used in a project.
-
-The purpose of a deprecated software features is to give developers time to change projects to follow a new or better standard without breaking the code.
+The purpose of \emph{deprecating} a software features is to give developers time to change projects to follow a new or better standard without breaking the code.
 When a piece of code becomes deprecated, continuing its use may lead to design flaws or even security vulnerabilities.
-For example, in Apache Nifi, an earlier version of the software used a compromised key-derivation functions; hence, this class was deprecated, but left in for compatibility with older versions of the software~\cite{nifi}.
+For example, in Apache Nifi, an earlier version of the software used a compromised key-derivation function; hence, this class was deprecated, but left in for compatibility with older versions of the software~\cite{nifi}.
 Viewing changes in terms of types can show when usage of deprecated classes are added or removed from a project as well as how many remain over time.
@@
 Removing all instances of deprecated features could also mean a project can be upgraded to use new tools or features that were not possible while the deprecated features were present.
-In our Apache BookKeeper case study,  we demonstrated an example where deprecated classes were removed entirely.
 
-Bug triaging for large open source projects is a very time consuming task, with hundreds  of bugs being reported daily.
+We searched for Apache foundation projects that featured an instance of deprecation and found Apache BookKeeper.
+As an Apache Foundation project it is a well-curated example of an open source software project.
+
+Apache Bookkeeper deprecated the method AbstractConfiguration.getLedgerManagerType() in release 4.2.0\footnote{\url{https://bookkeeper.apache.org/docs/r4.2.0/apidocs/deprecated-list.html}} and replaced it with AbstractConfiguration.getLedgerManagerFactoryClass().
+If we look at getLedgerManagerType() we can see at the time of the start of the deprecation all instances were removed and getLedgerManagerFactoryClass() was added (Figure~\ref{fig:bookkeeper-depr}).
+After the deprecation, we see no more changes to getLedgerManagerType() and only some additions of getLedgerManagerFactoryClass().
+
+\section{Discussion}
+
+Having tools for analyzing repositories based on type usage and method invocation benefits many parts of the software development process.
+Software managers require a high level understanding of the state of a repository and the changes being made which affect the functionality of the software.
+Tracking developer contributions in a less misleading way is also important for management since important contributions may be small and focused, but nevertheless, significant.
+Line differences only provide meaningful information when they are analyzed by someone with familiarity with the source code being changed.
+Providing tools which bring more meaning than line differences would allow people less familiar with the code to understand changes being made.
+When large changes are made that cover many aspects of a project, it is more useful to see how the change affects the software rather than noting that several thousands of lines have changed.
+Seeing how a project has changed over time is also vital for management since evaluating the state of completion of software can be very difficult.
+
+When managing group projects in an introductory software engineering course, it is very important to be able to see the contributions that students are making to the group repository to judge learning, progress, and collaboration.
+When one of the authors was supervising 4-5 groups of students, it became infeasible to look through all of the commits of each student.
+Students may take the wrong approach in a certain piece of code that can become ingrained in the project and very difficult to change.
+Some groups had problems with group members who did not complete their assigned work or made meaningless contributions.
+Issues like this can be difficult to detect and prevent.
+TypeV was used by the author to quickly summarize the work done by each group member, organized by the types they had been working on.
+This can show when intervention or guidance is required, without the large time commitment of pouring through commits.
+
+Bug triaging for large open source projects is a very time consuming task \cite{badashian2015}, with hundreds  of bugs being reported daily.
 Each bug must be processed and assigned to a developer that has the experience required to fix the bug.
@@
 
-Some commands in git present a problem gathering the statistics.
+Some actions in Git present a problem gathering the statistics.
 Authors can pull other projects into the repository and these additions will count towards the author's changes.
@@
 We only look at the master branch of a repository and ignore other branches; therefore, we might miss some of the work done by an author in another branch.
-Going through all of the branches in a repository would take a very long time and contain a large amount of duplicate data.
-Being able to choose which branch to analyze is something that can be added.
-
-We ignore the merge commits so we do not double count changes; however, if there is a merge conflict and the developer makes changes then these changes will not appear in TypeV.
-We make the assumptions that if this case happens then the changes are minor.
+As well, we ignore merge commits such that we do not double count changes; however, if there is a merge conflict, it is possible for the developer to makes changes.
+Changes made during merge conflict resolution are missed by TypeV.
+We have made the assumptions that if this case happens then the changes are minor.
 
@@
 
-Currently we are only looking at declaration of types and invocation of methods; it would be useful to also view what classes changes are being made to.
+TypeV only extracts declaration of types and invocation of methods; it would be useful to also view what classes changes are being made to.
 Enabling this ability would also enable analysis of self-invocations.
@@
 
-In this paper, we introduced TypeV, a source code repository analysis and visualization tool.
-Its use of commit-by-commit abstract syntax tree differencing enable for more insightful visualization than tools that merely use commit counts, line counts, and file changes.
-We answered our research questions as thus:
+In this paper, we introduced TypeV, a code repository analysis and visualization tool.
+Using commit-by-commit abstract syntax tree differencing enables insightful visualization that is not captured by tools that use commit counts, line counts, or file changes alone.
+In this paper, we addressed the following research questions:
 
@@
 \noindent
-Using abstract syntax trees, we can determine the types affected by each commits in a repository. Type information is not directly available through regular line diffs recorded by Git.
+Using abstract syntax trees, we can determine the types affected by each commit in a repository. Type information is not directly available through regular line diffs recorded by Git.
 \end{framed}
@@
 \noindent
-Type coverage and file coverage are found to be highly correlated but do not always have the same distribution or a one to one ratio as might be expected in Java.
+Type coverage and file coverage are found to be highly correlated but do not always have the same distribution or a one-to-one ratio as might be expected in Java.
 \end{framed}

### 2016-04-26 paper/document.tex
Updates.

@@
 \label{tab:project-summary}
-\begin{tabular}{c|lllll}
+\begin{tabular}{c|ccccc}
 \hline
-\bfseries Project & \bfseries Commits & \bfseries Authors* & \bfseries Files &   \bfseries Types &  \bfseries  Changes\\
+\bfseries Project & \bfseries Commits & \bfseries Authors & \bfseries Files &   \bfseries Types &  \bfseries  Changes\\
 \hline
@@
 \hline
-\multicolumn{6}{p{\linewidth}}{\footnotesize Totals for the four projects we analyzed.
-``Changes'' refers to the total number of invocation or declaration changes observed from the AST.
-Note that the number given for authors may be larger than the actual number of collaborators on the project, due to some collaborators using multiple email address, and thus appearing as more than one author.}
+\multicolumn{6}{p{3.25in}}{\footnotesize Totals for the four projects we analyzed.
+``Changes'' refers to the total number of invocation or declaration changes observed from the AST diffs.
+Note that the number given for authors may be larger than the actual number of collaborators on the project, due to some collaborators using multiple email address, thus appearing as more than one author.}
 \end{tabular}
@@
 In order to be effective, the triager has to have knowledge of what each developer has worked on as well as the parts of the code may be affected by the bugs.
-Being able to directly see what types a developer has worked with over time is important information for triaging.
-If the type causing the bug is known, the developers which have recently used that type can be assigned the bug.
+With TypeV, the triager is able to directly see what types a developer has worked with over time.
+This is important information for triaging since the type causing the bug, if known, will have a list of associated developers who have recently used that type and can be assigned the bug.
 The commit messages for the affected types are also visible, providing greater context about how the type has been used.
-The developer who is assigned the bug can also use type information to locate other affected pieces of code and know which developers have been using the affected types.
+The assigned developer can also use type information to locate other pieces of code affected by a bug and know which developers have been using the affected types.
 

### 2016-06-13 paper/document.tex
Fix LaTeX warnings.

@@
 
-\IEEEpeerreviewmaketitle
+\IEEEpeerreviewmaketitle%
 
@@
 Plotting the number of lines changed may be misleading.
-A huge spike in additions may be caused by a number of reasons, such as the inclusion of automatically-generated documentation, the inclusion of an external library, or perhaps a massive refactoring effort \cite{Hindle:2008:LCT:1370750.1370773}.
+A huge spike in additions may be caused by a number of reasons, such as the inclusion of automatically-generated documentation, the inclusion of an external library, or perhaps a massive refactoring effort~\cite{Hindle:2008:LCT:1370750.1370773}.
 
@@
 In the following sections of the paper we will describe our tool, TypeV, that we developed to compute and visualize changes to types used in a project over time.
-TypeV is able to extract type information from a repository by computing commit-by-commit abstract syntax tree differences, which we describe in section~\ref{sec:methodology}.
+TypeV is able to extract type information from a repository by computing commit-by-commit abstract syntax tree differences, which we describe in Section~\ref{sec:methodology}.
 In this paper, we answer the following questions: \\ \\
@@
 Analyzing source code is a popular research topic and Java is often chosen as the subject of study due to its popularity.
-Grechanik \textit{et al.} \cite{Grechanik:2010:EIL:1852786.1852801} examined the structure of Java programs mined from 2080 programs.
+Grechanik \textit{et al.}~\cite{Grechanik:2010:EIL:1852786.1852801} examined the structure of Java programs mined from 2080 programs.
 This paper examined the breakdown of syntactic structures in open source repositories, however it does not consider per author statistics.
-Parnin \textit{et al.} \cite{Parnin:2011:JGA:1985441.1985446} mined repositories to see how Java generics have been used in open source projects and found that generic usages were often introduced by a single developer in a project and were primarily being used for collecting and traversing lists of objects.
+Parnin \textit{et al.}~\cite{Parnin:2011:JGA:1985441.1985446} mined repositories to see how Java generics have been used in open source projects and found that generic usages were often introduced by a single developer in a project and were primarily being used for collecting and traversing lists of objects.
 
 Further analysis of the Java language has been done using ASTs.
-L\"{a}mmel \textit{et al.} \cite{Lammel:2011:LAA:1982185.1982471} used ASTs to examine the usage of APIs in Java projects.
+L\"{a}mmel \textit{et al.}~\cite{Lammel:2011:LAA:1982185.1982471} used ASTs to examine the usage of APIs in Java projects.
 They found what APIs are popular and if they were used in a framework-like manner.
-Dyer \textit{et al.} \cite{Dyer:2014:MBA:2568225.2568295} mined AST nodes to study the use of new Java language features over time.
+Dyer \textit{et al.}~\cite{Dyer:2014:MBA:2568225.2568295} mined AST nodes to study the use of new Java language features over time.
 The authors found the most popular features and the adoption rate of new features over time.
@@
 A variety of visualizations tools have been created in order to help with understanding the changes that get made to software over time.
-Wu \textit{et al.} \cite{wu2004a} explored punctuated changes to software repositories with a tool that generates an Evolution Spectrograph.
+Wu \textit{et al.}~\cite{wu2004a} explored punctuated changes to software repositories with a tool that generates an Evolution Spectrograph.
 In this visualization, each file has a row in the chart and each cell represents a section of time where an incoming or outgoing dependency can be changed.
 The colour is determined by the frequency of change.
-This work was continued to show file directories and display the cardinality of a developer's commits in Wu \textit{et al.} \cite{wu2004}.
+This work was continued to show file directories and display the cardinality of a developer's commits in Wu \textit{et al.}~\cite{wu2004}.
 The cardinality of a commit is the number of files or subsystems it affects in a system, indicating the scope of the change.
-Other tools also look at changes in code dependencies \cite{6980224} or files \cite{6650525, 7332421,voinea2006} over time.
+Other tools also look at changes in code dependencies~\cite{6980224} or files~\cite{6650525, 7332421,voinea2006} over time.
 
 Tools also exist which track changes over time in a more complex manner.
-Kim \textit{et al.} \cite{Kim:2009:DRS:1555001.1555046} built a tool called Logical Structural Diff (LSdiff) which examines code changes and produces logical rules to represent them.
+Kim \textit{et al.}~\cite{Kim:2009:DRS:1555001.1555046} built a tool called Logical Structural Diff (LSdiff) which examines code changes and produces logical rules to represent them.
 Changes which do not match the rules are then pointed out to the developer.
-Reiss and Tarvo created an IDE which allows users to see the history of files and changes to lines of code from each author as they are editing a project \cite{6650521}.
-Van Hees and Hage use a voronoi treemap to visualize changes in class and method structure \cite{7332410}.
-Servant and Jones created CHRONOS, a tool which creates a timeline of changes to a selection of code as well as the ability to query those changes \cite{6650547}.
+Reiss and Tarvo created an IDE which allows users to see the history of files and changes to lines of code from each author as they are editing a project~\cite{6650521}.
+Van Hees and Hage use a voronoi treemap to visualize changes in class and method structure~\cite{7332410}.
+Servant and Jones created CHRONOS, a tool which creates a timeline of changes to a selection of code as well as the ability to query those changes~\cite{6650547}.
 
 Some tools examine the evolution of projects to learn other information about them.
-Alcocer \textit{et al.} \cite{6650523} create a Performance Evolution Blueprint, which displays the changes in performance over time as a project evolves.
-Ens \textit{et al.} created ChronoTwigger to visualize the co-evolution of source and test files \cite{6980223}.
+Alcocer \textit{et al.}~\cite{6650523} create a Performance Evolution Blueprint, which displays the changes in performance over time as a project evolves.
+Ens \textit{et al.} created ChronoTwigger to visualize the co-evolution of source and test files~\cite{6980223}.
 This allowed testing practices to be more closely examined.
@@
 There is also research on how developers use software repositories to collaborate.
-Wagstrom \textit{et al.} \cite{Patrick:Wagstrom:2012} explored the roles that developers take in a networked, social development environments like GitHub.
+Wagstrom \textit{et al.}~\cite{Patrick:Wagstrom:2012} explored the roles that developers take in a networked, social development environments like GitHub.
 These role were defined as their level of contribution as well as the types of issues they dealt with.
 They found that developers will take on multiple roles in a project and will sometimes fulfill the same role across different projects.
-Lee \textit{et al.} \cite{lee2013} developed a tool for visualizing the branching structure of Git repositories.
+Lee \textit{et al.}~\cite{lee2013} developed a tool for visualizing the branching structure of Git repositories.
 This was used to analyze concurrent workflows in popular open source repositories.
-Elsen also created a visualization for branches in Git, with the ability to expand branches to view the directory structure of a project at that point \cite{6650522}.
-Shrestha \textit{et al.} made a visualization for the geographic location of a commit and \cite{6650532} and Minelli \textit{et al.} \cite{6980226} looked at the real time changes developers made to a repository.
+Elsen also created a visualization for branches in Git, with the ability to expand branches to view the directory structure of a project at that point~\cite{6650522}.
+Shrestha \textit{et al.} made a visualization for the geographic location of a commit and~\cite{6650532} and Minelli \textit{et al.}~\cite{6980226} looked at the real time changes developers made to a repository.
 
@@
 Some visualization tools are used for managing or making decisions about software rather than study the changes taking place.
-The service EasyBi \cite{EasyBi} provides additional visualizations for Git repositories for the purpose of business decisions.
+The service EasyBi~\cite{EasyBi} provides additional visualizations for Git repositories for the purpose of business decisions.
 EasyBi works by taking the git log file and generating charts and graphs from it.
 It tries to solve the issue of summarizing many commits and has greater controls for granularity and authors but still operates on commits and line differences.
-The Gitinspector project \cite{Gitinspector} is an open source tool for the statistical analysis of Git repositories.
+The Gitinspector project~\cite{Gitinspector} is an open source tool for the statistical analysis of Git repositories.
 It provides an advanced timeline view of changes per author and cumulative statistics but again relies upon the line differences to summarize the work of each author.
@@
 We accomplish this by computing the differences between abstract syntax trees of every consecutive pair of commits (Figure~\ref{fig:context}).
-This extracted data can then be used in a variety of visualizations as discussed in Section \ref{sec:viz}.
+This extracted data can then be used in a variety of visualizations as discussed in Section~\ref{sec:viz}.
 
@@
 This captures the structure of the code in an abstract and tractable manner.
-When a programmer makes a non-trivial change in code in a file, that difference will be reflected in the AST.
+When a programmer makes a non-trivial change in code in a file, that difference will be reflected in the AST\@.
 Non-trivial changes are those that affect how a compiler will interpret the source code, and may have effects on the executable code.
@@
 We were able to handle this by calling \texttt{git show} and saving the output to a temporary file to be used with the AST differencing tool.
-When calling \texttt{git show} instead of \texttt{git difftool} we treat everything as an addition, as in commit A in Figure \ref{fig:commits}.
+When calling \texttt{git show} instead of \texttt{git difftool} we treat everything as an addition, as in commit A in Figure~\ref{fig:commits}.
 Since we compared each commit to its parent, as reported by Git, we were able to account for any branching that happened outside the primary branch of the repository (usually named \texttt{master}).
-If a commit has more than one parent, like commit D in Figure \ref{fig:commits}, it indicates a merge commit in Git.
+If a commit has more than one parent, like commit D in Figure~\ref{fig:commits}, it indicates a merge commit in Git.
 Merge commits contain the changes made in each parent commit combined.
@@
 Knowing type and author information enables other kinds of visualizations.
-For example, Figure \ref{fig:chord} is a chord diagram which shows connections between authors and types.
+For example, Figure~\ref{fig:chord} is a chord diagram which shows connections between authors and types.
 The wider the arc around the circumference of the circle, the more often that entity is connected to other entities.
@@
 When managing group projects in an introductory software engineering course, it is very important to be able to see the contributions that students are making to the group repository to judge learning, progress, and collaboration.
-When one of the authors was supervising 4-5 groups of students, it became infeasible to look through all of the commits of each student.
+When one of the authors was supervising 4--5 groups of students, it became infeasible to look through all of the commits of each student.
 Students may take the wrong approach in a certain piece of code that can become ingrained in the project and very difficult to change.
@@
 
-Bug triaging for large open source projects is a very time consuming task \cite{badashian2015}, with hundreds  of bugs being reported daily.
+Bug triaging for large open source projects is a very time consuming task~\cite{badashian2015}, with hundreds  of bugs being reported daily.
 Each bug must be processed and assigned to a developer that has the experience required to fix the bug.

### 2016-06-13 paper/document.tex
Turd polishing.

@@
 For simplicity, we analyze only Java projects, as it allowed us to quickly develop and test our software.
-However, since the approach described in this paper can be generally applied to any statically-typed language, it can be extended it to other languages.
+However, since the approach described in this paper can be generally applied to any statically-typed language, it can be extended to other languages.
 Our goal is to describe a methodology for obtaining richer, more meaningful data from source code repositories which can be used in more insightful analyses.
@@
 L\"{a}mmel \textit{et al.}~\cite{Lammel:2011:LAA:1982185.1982471} used ASTs to examine the usage of APIs in Java projects.
-They found what APIs are popular and if they were used in a framework-like manner.
+They found which APIs are popular and if they were used in a framework-like manner.
 Dyer \textit{et al.}~\cite{Dyer:2014:MBA:2568225.2568295} mined AST nodes to study the use of new Java language features over time.
@@
 It does not matter what language is used or if errors exist in the commit, GitHub will simply output the number of line additions and deletions.
-The issue with this type of analysis is that it does not give a clear idea of is being changed, only that changes have occurred.
+The issue with this type of analysis is that it does not give a clear idea of what is being changed, only that changes have occurred.
 This data may be lacking, or even misleading which contributes to the need for better tools.
@@
 To determine what has changed between two subsequent commits, we computed their \emph{AST difference}.
-Given two ASTs, an AST difference (or simply \emph{AST diff}) states which nodes were must be added and which nodes must be removed to transform one tree to the next (step two in Figure~\ref{fig:context}).
-This allows us to precisely track changes two consecutive revisions of the same file.
+Given two ASTs, an AST difference (or simply \emph{AST diff}) states which nodes must be added and which nodes must be removed to transform one tree to the next (step two in Figure~\ref{fig:context}).
+This allows us to precisely track changes between two consecutive revisions of the same file.
 For AST diffing, we used GumTree~\cite{falleri:hal-01054552}, an algorithm made to compute the difference between two ASTs generated by Spoon.
@@
 For all the edited Java files we also compared the file at the time of the commit with its parent commit using AST differencing as described above.
-We achieved this by using calling \texttt{git difftool} and specifying the tool to be used as our AST differencing tool.
+We achieved this by invoking \texttt{git difftool} and specifying the tool to be used as our AST differencing tool.
 One case where \texttt{git difftool} fails is when the commit has no parent, such as the initial commit.
@@
 When calling \texttt{git show} instead of \texttt{git difftool} we treat everything as an addition, as in commit A in Figure~\ref{fig:commits}.
-Since we compared each commit to its parent, as reported by Git, we were able to account for any branching that happened outside the primary branch of the repository (usually named \texttt{master}).
+Since we compare each commit to its parent as reported by Git, we are able to account for any branching that happened outside the primary branch of the repository (usually named \texttt{master}).
 If a commit has more than one parent, like commit D in Figure~\ref{fig:commits}, it indicates a merge commit in Git.
@@
 The light blue \additionsbox{10pt} and dark red \deletionsbox{10pt} portions account for the number of additions and deletions respectively, in a similar fashion to GitHub's code frequency graph (Figure~\ref{fig:GH_code_freq}).
-These colours were chosen as they are high contrast, and are distinct to those red-green colour blindness.
+These colours were chosen as they are high contrast, and are distinct to those experiencing red-green colour blindness.
 
@@
 
-If we tried to do the same analysis with the GitHub tools it would be significantly more difficult to achieve the conclusion achieved using TypeV (Figure~\ref{fig:parser2}).
-First if we only look at line changes this change to \texttt{ParserATNSimulator} is overwhelmed by other changes and hardly shows up on the GitHub tools.
+If we tried to do the same analysis with the GitHub tools it would be significantly more difficult to reach this conclusion (Figure~\ref{fig:parser2}).
+First, if we only look at line changes this change to \texttt{ParserATNSimulator} is overwhelmed by other changes and hardly shows up on the GitHub tools.
 If we did notice an increase in activity we would not know what types were being worked on.
@@
 Each bug must be processed and assigned to a developer that has the experience required to fix the bug.
-In order to be effective, the triager has to have knowledge of what each developer has worked on as well as the parts of the code may be affected by the bugs.
+In order to be effective, the triager has to have knowledge of what each developer has worked on as well as the parts of the code that may be affected by the bugs.
 With TypeV, the triager is able to directly see what types a developer has worked with over time.

### 2016-06-19 www/js/data-filter.ts
Typeify DataFilter.

@@ -8,30 +8,52 @@ import { first, last, union } from './utils';
 
 const VALID_STEP_SIZES = new Set(['hour', 'day', 'month', 'week']);
 
+/**
+ * Maps type names to their absolute freqeuncy.
+ */
+interface TypeFrequency {
+  [typeName: string]: number;
+};
+
+interface RawFilteringResult {
+  types: JavaType[];
+  commits: CommitMap;
+  typesPresent: TypeFrequency;
+  timeslices: TimeSlice[];
+  authorStats: AuthorStatistics;
+  typesOverall: Set<string>;
+  filesOverall: Set<string>;
+  astDiffs: ASTDiff[];
+}
+
+export interface AuthorStatistics {
+  [authorID: string]: CommitStatistics[];
+}
+
+export interface CommitStatistics {
+  type: EntitySummary;
+  file: EntitySummary;
+  date: Date;
+}
+
 interface EntitySummary {
   observed: number;
   cumulative: number;
   total: number;
 }
 
-export interface CommitStatistics {
-  type: EntitySummary;
-  file: EntitySummary;
-  date: Date
-}
-
 export default class DataView {
-  public types: Array<JavaType>;
-  public timeslices: Array<TimeSlice>;
-  public typesPresent: Array<JavaType>;
-  public authorStats: { [name: string] : CommitStatistics[]};
-  public commits: Array<Object>;
-  public astDiffs: Array<ASTDiff>;
+  public types: JavaType[];
+  public timeslices: TimeSlice[];
+  public typesPresent: TypeFrequency;
+  public authorStats: AuthorStatistics;
+  public commits: CommitMap;
+  public astDiffs: ASTDiff[];
   public typesOverall: Set<String>;
   public filesOverall: Set<String>;
 
   constructor({types, timeslices, typesPresent, authorStats, astDiffs,
-              typesOverall, filesOverall, commits}) {
+              typesOverall, filesOverall, commits}: RawFilteringResult) {
     this.types = types;
     this.timeslices = timeslices;
     this.typesPresent = typesPresent;
@@ -54,10 +76,11 @@ export default class DataView {
    * # authors -- Filter only to the given authors.
    * # typeFilter -- only types matching this pattern are selected.
    */
-  static filter(data, filters: Filter) {
+  static filter(data: PreprocessedData, filters: Filter) {
     const rawFilteredData = filterTypes(data, filters);
     return new DataView(rawFilteredData);
   }
+
 
[…]

### 2016-06-19 www/js/index.ts

@@ -54,20 +54,20 @@ window.redrawTable = function () {
 }
 
 function getFilters(): Filter {
-  var start_date = null;
-  var end_date = null;
+  var startDate: Date = null;
+  var endDate: Date = null;
 
   var timestamp = Date.parse($("#min-datepicker").val())
   if (isNaN(timestamp) == false) {
-    start_date = new Date(timestamp);
+    startDate = new Date(timestamp);
   }
 
   timestamp = Date.parse($("#max-datepicker").val())
   if (isNaN(timestamp) == false) {
-    end_date = new Date(timestamp);
+    endDate = new Date(timestamp);
   }
 
-  var _authors = [];
+  var _authors: string[] = [];
 
   $('.author-check-box').each(function() {
     if($(this).prop('checked')) {
@@ -78,8 +78,8 @@ function getFilters(): Filter {
   var typeFilter = $('#type-filter').val();
 
   return {
-    start: start_date,
-    end: end_date,
+    start: startDate,
+    end: endDate,
     stepSize: $("#step").val(),
     limit: +$("#filter").val(),
     authors: _authors,

### 2016-06-19 www/js/json.d.ts

@@ -31,8 +31,8 @@ interface CommitFromJSON {
    */
   message: string;
 
-  files: any[];
-  all_files: any[];
+  files: string[];
+  all_files: string[];
 }
 
 interface Commit extends CommitFromJSON {

### 2016-06-19 www/js/data-filter.ts
Rename union() -> addAll().

@@ -4,14 +4,14 @@ import ASTDiff from './ast-diff';
 import {PreprocessedData} from './preprocess-data';
 
 import assert from './assert';
-import { first, last, union } from './utils';
+import { first, last, addAll } from './utils';
 
 const VALID_STEP_SIZES = new Set(['hour', 'day', 'month', 'week']);
 
 /**
  * Maps type names to their absolute freqeuncy.
  */
-interface TypeFrequency {
+export interface TypeFrequency {
   [typeName: string]: number;
 };
 
@@ -36,7 +36,7 @@ export interface CommitStatistics {
   date: Date;
 }
 
-interface EntitySummary {
+export interface EntitySummary {
   observed: number;
   cumulative: number;
   total: number;
@@ -234,10 +234,10 @@ function filterTypes(data: PreprocessedData, filters: Filter) {
     }
 
     /* Add all observed files and types to their respective sets. */
-    union(filesOverall, files);
-    union(authorSets[author].files, files);
-    union(typesOverall, types);
-    union(authorSets[author].types, types);
+    addAll(filesOverall, files);
+    addAll(authorSets[author].files, files);
+    addAll(typesOverall, types);
+    addAll(authorSets[author].types, types);
 
     /* Record stats! */
     authorMap[author].push({
@@ -314,7 +314,7 @@ function forEachCommit(diffs: ASTDiff[], fn: EachCommmitCallback) {
     /* Update the types. */
     typesOverall.add(diff.type);
     /* Update the files. */
-    union(filesOverall, diff.filesModified);
+    addAll(filesOverall, diff.filesModified);
   }
 }
 

### 2016-06-19 www/js/utils.ts

@@ -27,7 +27,7 @@ export function last<T>(array: T[]): T {
 /**
  * Add all items from the iterable to the set.
  */
-export function union<T>(set: Set<T>, iterable: Iterable<T>): Set<T> {
+export function addAll<T>(set: Set<T>, iterable: Iterable<T>): Set<T> {
   for (var item of iterable) {
     set.add(item);
   }

### 2016-06-19 www/js/data-filter.ts
Data filter is now well typed.

@@ -174,14 +174,19 @@ function filterTypes(data: PreprocessedData, filters: Filter) {
 
   /* Create a list of each type. */
   var types = sortedTypeNames.map(type => new JavaType(type));
-  var typeMap = {};
+  var typeMap: { [name: string]: JavaType } = {};
   /* Map full type names to their type. */
   types.forEach(function (type) {
     typeMap[type.name] = type;
   });
 
-  var authorMap = {};
-  var authorSets = {};
+  var authorMap: AuthorStatistics = {};
+  var authorSets: {
+    [name: string]: {
+      files: Set<string>;
+      types: Set<string>;
+    }
+  } = {};
   var typesOverall = new Set<string>();
   var filesOverall = new Set<string>();
 
@@ -199,7 +204,7 @@ function filterTypes(data: PreprocessedData, filters: Filter) {
     /* For each applicable diff in the time range... */
     for (let i = lowerIndex; i < upperIndex; i++) {
       let diff = applicableDiffs[i];
-      /* Count the type, regardless if it's filtered. */
+      /* Count the type name, regardless if it's filtered. */
       typesOverall.add(diff.type);
 
       let type = typeMap[diff.type];
@@ -228,8 +233,8 @@ function filterTypes(data: PreprocessedData, filters: Filter) {
     if (!(author in authorMap)) {
       authorMap[author] = [];
       authorSets[author] = {
-        files: new Set(),
-        types: new Set()
+        files: new Set<string>(),
+        types: new Set<string>()
       };
     }
 

### 2016-06-19 www/js/project-view.ts
Remove implicit anys from project-view.

@@ -9,7 +9,7 @@
 
 
 import assert from './assert';
-import preprocessData from './preprocess-data';
+import preprocessData, {PreprocessedData} from './preprocess-data';
 import DataView, {CommitStatistics} from './data-filter';
 import JavaType from './java-type';
 import Cell from './cell';
@@ -61,7 +61,7 @@ export function createTable2(filter: Filter) {
 
 /*=== Graph ===*/
 
-function drawGraph(data: DataView, width) {
+function drawGraph(data: DataView, width: number) {
   var marginLeft = 150;
   var cellHeight = 64;
   var height = cellHeight * data.types.length;
@@ -227,7 +227,7 @@ function drawGraph(data: DataView, width) {
 
       info.append('br');
 
-      var commit_map = (<any> window).preprocessedData.commits;
+      var commit_map = (<PreprocessedData>window.preprocessedData).commits;
 
       info.append('b')
         .text("Commits:");
@@ -522,7 +522,7 @@ function ensureAxisIsAtGraphBottom(graph: SVGElement, axis: SVGElement) {
   }
 }
 
-function summary({numberOfCommits, authors, numberOfFiles, numberOfTypes, astDiffs}) {
+function summary({numberOfCommits, authors, numberOfFiles, numberOfTypes, astDiffs}: DataView) {
   return `project & ${numberOfCommits} & ${authors.length} & ${numberOfFiles} & ${numberOfTypes} & ${astDiffs.length}`;
 }
 
@@ -535,7 +535,7 @@ function viewportHeight() {
 
 /*=== Utilties ===*/
 
-function cellWidthFromScale(cell: Cell, scale: (Date) => number) {
+function cellWidthFromScale(cell: Cell, scale: (_: Date) => number) {
   var bigger = scale(cell.endDate);
   var smaller = scale(cell.startDate);
   assert(bigger > smaller);

### 2016-06-19 www/js/project-view.ts
Minor changes.

@@ -23,7 +23,7 @@ var cellInfo = d3.select("body")
   .style("position", "absolute")
   .style("z-index", "10")
   .style("visibility", "hidden")
-  .style("width", String(CELL_INFO_WIDTH) + "px")
+  .style("width", `${CELL_INFO_WIDTH}px`)
   .classed('panel panel-default', true);
 
 cellInfo.append('div')
@@ -104,9 +104,9 @@ function drawGraph(data: DataView, width: number) {
       .attr('x', marginLeft)
       .attr('width', width - marginLeft)
       .attr('height', maxCellHeight)
-      .style('fill', function (d, i) {
+      .style('fill', function (_, i) {
         /* Make alternating colour bands. */
-        return i&1 ? '#f4f4f4' : '#fafafa';
+        return i & 1 ? '#f4f4f4' : '#fafafa';
       });
 
   row.each(createCellsForType);

### 2016-06-20 paper/document.tex
Add colour perception cite.

@@
 Programmers solve problems in a way that is readable to other developers and executable by a computer.
-The programmer rarely operates at the character level, or even the line level, so why should our analyses of source code operate at granularities unfamiliar to developers?
+The programmer rarely operates at the character level, or even the line level, so why should our visualizations of source code operate at granularities unfamiliar to developers?
+
 The plain text nature of software does not stop at source code level.
@@
 With more changes to peruse, it becomes more difficult to see how a project is evolving over time since there is so much information to look through.
-Most hosting services for Git repositories provide some basic tools for analyzing projects over time~\cite{github-graphs,bitbucket-graphs}; however, as we will show, these tools often lack the kind of detail to gather insightful analysis.
+Hosting services for Git repositories often provide some basic tools for analyzing projects over time~\cite{github-graphs,bitbucket-graphs}; however, as we will show, these tools often lack the kind of detail to gather insightful analysis.
 As with source code, commits operate at the text level.
-While commits may store individual changes as lines of code, they imply a change in \emph{structure} of the software.
-By visualizing lines changed, or files changed, the current state-of-the-art for software project visualization rarely capture the \emph{meaning} behind each commit; lines of code alone does not describe how a repository changed at a particular point in time.
-Thus, information regarding contributions from developers is often misleading since the number of commits and lines of code changed does not necessarily reflect meaningful changes to a repository~\cite{robles2014}.
+While commits may store individual changes as lines of code, they imply a change in the \emph{structure} of software.
+By visualizing lines changed, or files changed, many current tools for software projects rarely capture the \emph{meaning} behind each commit; lines of code alone do not describe how a repository changed at a particular point in time.
+Thus, information regarding contributions from developers is often misleading since the number of commits and lines of code changed do not necessarily reflect meaningful changes to a code repository~\cite{robles2014}.
+In this paper, we will show that abstract syntax trees (ASTs) can be used to create \emph{semantically-aware} software evolution visualizations.
 
-Consider GitHub graphs~\cite{github-graphs}, visualizations that are easily available to open source software developers.
+Consider GitHub graphs~\cite{github-graphs}: visualizations that are easily available to the many software developers using GitHub to host their software repositories.
 GitHub's graphs use line and commit information exclusively.
@@
 What happened at those big spikes of line changes in the graph?
-One of the spikes is so large, we had to crop the graph to be able to show details of other changes.
+One of the spikes is so large, we had to crop the image in order to show the magnitude of the other changes.
 How did these large changes affect the code base?
@@
 The goal of a software evolution visualization is to provide an alternate view that provides insight into the development of the software that is otherwise unavailable.
-By visualizing changes to lines and files, we are tracking the changes in text---not software.
-What would be more meaningful are the structural changes in code.
-These are changes in variable declarations, invocations of methods, and usage of types.
+By visualizing changes to lines and files, one is tracking the changes in text---not software.
+Instead, we promote \emph{semantically-aware visualizations} that visualize the structural changes in code.
+In our case these are changes in variable declarations, invocations of methods, and usage of types.
+Others have explored semantically-aware visualization~\cite{sourcevis,6980224,Kim:2009:DRS:1555001.1555046,7332410}, however, we show that this can be done with ASTs.
 
-In this paper, we propose a method that analyses the changes in type usage and method invocations over time.
-Using commits, we are able to keep track of \emph{who} is changing what types.
-We then present this data using simple visualizations.
+In this paper, we propose a semantically-aware visualization method that tracks the changes in type usage and method invocations over time.
+Using commits, we are able to keep track of \emph{who} is using what types.
+We then present this data using time-proven visualizations.
 We claim that type usage and method invocation offers more insight than line-based, file-based, and commit-based methods of visualization and analysis.
 For simplicity, we analyze only Java projects, as it allowed us to quickly develop and test our software.
-However, since the approach described in this paper can be generally applied to any statically-typed language, it can be extended to other languages.
-Our goal is to describe a methodology for obtaining richer, more meaningful data from source code repositories which can be used in more insightful analyses.
+However, since the approach described in this paper can be generally applied to any statically-typed language, it can be adapted to other programming languages.
+Our goal is to describe a methodology for obtaining richer, more meaningful data from source code repositories which can be used for more insightful analyses.
 
-In the following sections of the paper we will describe our tool, TypeV, that we developed to compute and visualize changes to types used in a project over time.
+In the following sections of the paper we will describe our tool, TypeV, that we developed to compute and visualize changes to types by author used in a project over time.
 TypeV is able to extract type information from a repository by computing commit-by-commit abstract syntax tree differences, which we describe in Section~\ref{sec:methodology}.
 In this paper, we answer the following questions: \\ \\
-\textbf{RQ1:} How can we extract type and method usage over time from repositories? \\ \\
+\textbf{RQ1:} How can we extract type and method usage from repositories over time? \\ \\
 \textbf{RQ2:} How does type coverage compare to file coverage?
@@
 
-Analyzing source code is a popular research topic and Java is often chosen as the subject of study due to its popularity.
+Analyzing source code is a popular research topic and Java is often chosen as the subject of study due to its popularity and abundance of tools.
 Grechanik \textit{et al.}~\cite{Grechanik:2010:EIL:1852786.1852801} examined the structure of Java programs mined from 2080 programs.
@@
 
-The focus of these papers is on how programming languages and features are used by looking at many repositories.
-Java is a statically-typed language, yet there is no focus specifically on how types are used throughout a project.
+The focus of these papers was on how programming languages and features are used by looking at many repositories.
+Java is a statically-typed language---its types are checked at compile time---yet there is no focus specifically on how types are used throughout a project.
 Type data may provide additional information on how developers use languages but this has yet to be explored.
@@
 The cardinality of a commit is the number of files or subsystems it affects in a system, indicating the scope of the change.
-Other tools also look at changes in code dependencies~\cite{6980224} or files~\cite{6650525, 7332421,voinea2006} over time.
+Other tools also look at changes in code dependencies~\cite{6980224} or files~\cite{6650525, 7332421, voinea2006} over time.
 
@@
 Alcocer \textit{et al.}~\cite{6650523} create a Performance Evolution Blueprint, which displays the changes in performance over time as a project evolves.
+Ogawa and Ma created code\_swarm~\cite{codeswarm}, a generative art visualization which presents the progressive evolution of file changes made by authors in source control repository over time.
+Gource~\cite{gource} is a similar visualization, showing an animated force-directed tree of the file hierarchy as authors make changes to files in the source code repository over time.
 Ens \textit{et al.} created ChronoTwigger to visualize the co-evolution of source and test files~\cite{6980223}.
 This allowed testing practices to be more closely examined.
+Anslow \textit{et al.}~\cite{sourcevis} created SourceVis, a suite of software visualizations including visualizations for semantically-aware project evolution.
+SourceVis displays the evolution of Java source code metrics such as number of packages, classes, and methods, and fields.
 
@@
 
-The basic tools easily available to any software developer are the GitHub graphs (see Figure~\ref{fig:GH_code_freq})~\cite{github-graphs}.
-GitHub tracks contributions based only on the number of commits and number of lines changed.
-This has the advantage of not relying on the type of code which was committed.
+Developers using the massively popular GitHub website for their Git repository hosting instantly have access to basic history visualizations, known as GitHub Graphs (see Figure~\ref{fig:GH_code_freq})~\cite{github-graphs}.
+GitHub Graphs tracks contributions based only on the number of commits and number of lines changed.
+This has the advantage of not relying on the of code which was committed.
 It does not matter what language is used or if errors exist in the commit, GitHub will simply output the number of line additions and deletions.
@@
 
-Of the services we have surveyed, none of the tools visualize project evolution using granular type analysis per author via commit-by-commit abstract syntax tree differencing.
+Of the services we have surveyed, none of the tools that visualize project evolution address semantic history extraction using granular type analysis per author via commit-by-commit abstract syntax tree differencing.
 
@@
 This extracted data can then be used in a variety of visualizations as discussed in Section~\ref{sec:viz}.
+The reason for using ASTs is that they capture the semantics of the code far more than simple line counts.
 
@@
 If a commit has more than one parent, like commit D in Figure~\ref{fig:commits}, it indicates a merge commit in Git.
-Merge commits contain the changes made in each parent commit combined.
-Therefore, we ignored merge commits since all changes in the merge should already have been accounted for in previous commits.
-Thus, any branch that has been merged into the primary branch would have been analyzed and included by TypeV.
+
+We deliberately chose to ignore merge commits.
+A merge commit has two cases to consider: The first case is when the parents of the commit can be automatically merged.
+In this case all the relevant changes are already recorded in the parent of the commits so we can safely ignore the merge commit~\cite{git:merge}.
+The second case occurs when the parents cannot be merged automatically: There is a merge conflict which requires manual resolution on the part of a developer.
+In this case the Git user will try to resolve the issue by modifying code.
+If we decide to include the merge commit we run into the issue of to how to compare the merge commit with its multiple parent commits and the possible modifications made to resolve the merge conflict.
+We could consider comparing the merge commit with only one parent commit at a time, however; we will end up counting all the changes already recorded in the other parent branch.
+Additionally, all changes made during the resolution of a conflict will be attributed to the author making the merge; this may exclude the author that actually wrote the code in the first place.
+The other option is to simply ignore the merge commit such that we do not count the changes done during a merge commit.
+We made the assumptions that merge conflicts occur infrequently and that changes done to resolve a merge conflict are usually minor.
+Therefore, we chose to ignore merge commits.
+Note that fast-forwards are not \emph{true merges}, hence they do not create merge commits~\cite{git:merge}; none of the previous cases involving merges occur.
 
@@
 
+The major downside to using ASTs over a method like line count is that ASTs are more expensive to compute.
+Our experiments were run on a Intel i7 with 16GB of RAM and an SSD.
+Calculating the AST diffs varied from an hour or two for smaller projects to a couple of days for very large projects.
+For example ANTLR4 took about 8 hours while Apache BookKeeper only took a couple of hours.
+The good news for ongoing projects is that AST diffs only need to be computed for new commits; TypeV can resume calculating diffs and augmenting statistics at any time.
+New diffs can be appended to the already existing database of previous AST diffs.
+
 \section{Visualization}
+
+% CITE PAPERS THAT HAVE USED OUR APPROACH!
+% AND SAY OUR VISUALIZATIONS ARE EFFECTIVE BY PROXY
+
+% TODO: \cite{sillito2006}
 
@@
 
-\subsection{Views}
-
-TypeV provides a variety of views to help organize the data and provide different levels of detail.
-The overview displays several rows of stacked bar graphs which show changes to a type or method over a period of time in a project (Figure~\ref{fig:main-view}).
-The purpose of displaying stacked bar charts is that they are a familiar visualization, thus are somewhat easy to read, and can succinctly display the relative amounts of additions and deletions per each time period.
+\subsection{Timeline View}
+For our main demonstration of semantically-aware project evolution, we chose an interactive timeline view that displays the most popular types over time.
+The timeline displays several rows of stacked bar graphs which show changes to an individual type or method over a selectable period of time in a project (Figure~\ref{fig:main-view}).
 The light blue \additionsbox{10pt} and dark red \deletionsbox{10pt} portions account for the number of additions and deletions respectively, in a similar fashion to GitHub's code frequency graph (Figure~\ref{fig:GH_code_freq}).
-These colours were chosen as they are high contrast, and are distinct to those experiencing red-green colour blindness.
-
 The height of each bar shows the relative number of changes to a type.
@@
 The bar height is calculated by multiplying the maximum bar height by the number of changes during that time period divided by the maximum number of changes attested for this type.
-The rows are sorted so that the graph for the most used types or methods appear the top and are in descending order of use.
 
@@
 If the user clicks on the bar, they are given more detailed statistics which list the authors who made the commits and each commit message.
-By adding interaction, the user is able to explore the overview, find the specific information they are interested and drill-down at their leisure.
-Although our tool extracts a great deal of information, by using interaction, users can gradual reveal more information as they need it.
 
-To allow for a more detailed analysis of a specific component of the repository we added some basic filtering tools.
-Users can filter by type name; they can view changes that fall within a specific start and end date; users can also choose to see only declarations of types, types used in declarations and invocations, or only invocations.
+To allow for a more detailed analysis of a specific component of the repository we added filtering tools.
+Users can filter by type name, view changes that fall within a specific start and end date, and select whether to see only declarations of types, types used in declarations and invocations, or only invocations.
+
 One issue with analyzing individual authors is that authors can commit from a variety of different computers which can have different emails and usernames attached to their Git configuration.
-This may result in one author appearing as several different users in the \texttt{git log}.
-There is no perfect way to deal with this problem; however, we allow the user of our software to select and view multiple users at a time, thus allowing the user to view changes from an individual or group of developers.
+This may result in one author appearing as several different authors in the \texttt{git log}.
+There is no perfect way to deal with this problem; however, we allow the user of our software to select and view multiple authors at a time, thus allowing the user to view changes from an individual or group of developers.
 
-Knowing type and author information enables other kinds of visualizations.
-For example, Figure~\ref{fig:chord} is a chord diagram which shows connections between authors and types.
-The wider the arc around the circumference of the circle, the more often that entity is connected to other entities.
-For example, ``parrt'' is the top contributor, with changes covering a large number of types.
-Determining author and type information is made possible by AST differencing.
+\subsection{Design Decisions}
 
-\begin{figure*}[p]
-\centering
-\includegraphics[width=\textwidth]{chord}
-\caption{Chord diagram showing connections between authors (right and bottom) to types (left).}
-\label{fig:chord}
-\end{figure*}
+The purpose of displaying stacked bar charts is that they are a familiar visualization.
+They can be viewed as an enhanced version of GitHub's code frequency graph~\cite{github-graphs}, with semantic data.
+We chose to create a similar visualization to GitHub to work with a visualization people may already be familiar with and evaluate our visualization tools' density of data.
+
+The stacked bar graphs succinctly display the relative amounts of additions and deletions per each time period.
+This creates a dense view of the project's evolution.
+We chose to make the height of the bars relative to only the type itself to make it easy for a user to scan a single type and see when it was used modified.
+This makes tasks such finding a refactoring (see \ref{sec:antlr4}) very clear to see simply by browsing the line.
+We stacked the additions on top of the deletions such that users can immediately compare the relative frequency of changes of a single type throughout the chosen time span.
+
+Our choice of colours was motivated to be high contrast and culturally indicative of their meaning.
+The particular colours chosen are high contrast, and are distinct to those experiencing protanomaly or deuteranomaly---red-green colour blindness.
+The red colour suggests a negative valence~\cite{moller2009}, hence suggesting deletions.
+The blue serves as a contrast to the red, indicating the opposite action (in this case, additions).
+Since
+
+The rows are sorted so that the graph for the most used types or methods appear the top and are in descending order of use such that the most frequently used type is the first visible.
+The height allows users to see when certain types are being edited and ensures that highly used types do not overpower less used type making their changes less significant.
+
+By adding interaction, the user is able to explore the timeline to find the specific information which they are interested in and drill-down at their leisure.
+Although our tool extracts a great deal of information, by using interaction, users can gradual reveal more information as they need it, rather than be bombarded with all of the information at once.
 
 \section{Evaluation}
+
+% TODO: say all these other tools used file-based stuff:
+% gource, github, code_swarm, others...?
 
@@
 We can also see a difference between these measurements on the higher end of the graph.
-File coverage seems to have a group of commits that had a file coverage between 20\% and 50\% whereas type coverage have a similar looking grouping but between 40\% and 60\%.}
+File coverage seems to have a group of commits that had a file coverage between 20\% and 50\% whereas type coverage have a similar looking grouping but between 40\% and 60\%.\label{fig:antlr4-density}}
 \end{figure}
@@
 Here we can see that commits did not have much file coverage between 25\% and 75\%.
-On the other hand there were a lot of commits that have type coverage around 50\%.}
+On the other hand there were a lot of commits that have type coverage around 50\%.\label{fig:tika-density}}
 \end{figure}
 
-Is it worth going through the effort of computing commit-by-commit AST differences if we can just simply count changes to files instead?
+Is it worth the effort of computing commit-by-commit AST differences if we can just simply count changes to files?
 We define \emph{type coverage} as the number of types that have been touched out of the total number of types during either a certain time period or the entire project.
 \emph{File coverage} is the same measurement, but using the number of source files touched instead.
-File coverage can be trivially determined from an author's commit history; Type coverage, however, is uniquely provided by AST differencing, as we have described.
+File coverage can be trivially determined from an author's commit history.
+Type coverage, however, is uniquely provided by AST differencing, as we have described.
 
-In Java, a type can be a primitive, provided by the language itself, or an object reference.
+In Java, a type can be primitive, provided by the language itself, or an object reference.
 Primitive types are those that are built into the language such as an \texttt{int} or \texttt{char}.
@@
 We counted the file coverage and type coverage on a commit-by-commit basis for each author.
-That is, each datum represents a commit with three dimensions: file coverage and type coverage, and its commit date.
+That is, each datum represents a commit with three dimensions: file coverage, type coverage, and its commit date.
 Data is sorted in ascending order of commit date.
@@
 As type coverage and file coverage are both cumulative measures, positive correlation is expected.
+We displayed the density distributions for two different projects: ANTLR4 (Figure~\ref{fig:antlr4}) and Apache Tika (Figure~\ref{fig:tika-density}).
 However, even when the distributions are significantly different (as with ANTLR and Curator), the correlation is still strongly positive, but not 100\%.
@@
 
-\subsection{ANTLR4}
+\subsection{ANTLR4}\label{sec:antlr4}
 
@@
 
-The purpose of \emph{deprecating} a software features is to give developers time to change projects to follow a new or better standard without breaking the code.
+The purpose of \emph{deprecating} software features is to give developers time to change projects to follow a new or better standard without breaking the code.
 When a piece of code becomes deprecated, continuing its use may lead to design flaws or even security vulnerabilities.
@@
 \section{Discussion}
+
+Having semantically aware evolutionary visualizations also gives users who are unfamiliar with the project at-a-glance information on the evolution of the project.
 
@@
 
-We only look at the master branch of a repository and ignore other branches; therefore, we might miss some of the work done by an author in another branch.
-As well, we ignore merge commits such that we do not double count changes; however, if there is a merge conflict, it is possible for the developer to makes changes.
+Internal threats to validity include our treatment of merge conflicts.
+As mentioned before, we ignore all merges, so if there is a merge conflict, it is possible that a developer makes non-trivial changes.
 Changes made during merge conflict resolution are missed by TypeV.
-We have made the assumptions that if this case happens then the changes are minor.
+We have made the assumptions that if this case happens, the changes are minor.
 
@@
 In the future, we want to work on a better way to track and represent generic types.
-
-TypeV only extracts declaration of types and invocation of methods; it would be useful to also view what classes changes are being made to.
-Enabling this ability would also enable analysis of self-invocations.
 
@@
 
-\textbf{RQ1:} How can we extract type and method usage over time from repositories?
+\textbf{RQ1:} How can we extract type and method usage from repositories over time?
 
@@
 
+Although calculating ASTs can be expensive, it is not computationally infeasible, and we benefit from semantically-aware code evolution visualizations that can be used to gain better insight into our software.
+
 % === Below this line ===

### 2016-06-20 paper/document.tex
Go through 50% of the paper.

@@
 
-The plain text nature of software does not stop at source code level.
+The plain text nature of software does not stop at the source code level.
 The history and evolution of a software project is often tracked by version control systems, such as Git, in a programming language agnostic way.
@@
 
-Consider GitHub graphs~\cite{github-graphs}: visualizations that are easily available to the many software developers using GitHub to host their software repositories.
+Consider GitHub graphs~\cite{github-graphs}: visualizations that are easily available to software developers using GitHub to host their Git software repositories.
 GitHub's graphs use line and commit information exclusively.
-In the case of the ANTRL4 project (Figure~\ref{fig:GH_code_freq}), this code frequency chart creates more questions than it answers.
+In the case of the ANTLR4 project (Figure~\ref{fig:GH_code_freq}), this code frequency chart creates more questions than it answers.
 What happened at those big spikes of line changes in the graph?
@@
 
-We will examine research which analyses software repositories for information on language and features usage, tools for visualizing project evolution, collaboration in software repositories and other tools for understanding repositories.
+We examined research which analyses software repositories for information on language and feature usage, tools for visualizing project evolution, collaboration in software repositories and other tools for understanding repositories.
 
@@
 Alcocer \textit{et al.}~\cite{6650523} create a Performance Evolution Blueprint, which displays the changes in performance over time as a project evolves.
-Ogawa and Ma created code\_swarm~\cite{codeswarm}, a generative art visualization which presents the progressive evolution of file changes made by authors in source control repository over time.
+Ogawa and Ma created code\_swarm~\cite{codeswarm}, a generative art visualization which presents the progressive evolution of file changes made by authors in a source control repository over time.
 Gource~\cite{gource} is a similar visualization, showing an animated force-directed tree of the file hierarchy as authors make changes to files in the source code repository over time.
@@
 Anslow \textit{et al.}~\cite{sourcevis} created SourceVis, a suite of software visualizations including visualizations for semantically-aware project evolution.
-SourceVis displays the evolution of Java source code metrics such as number of packages, classes, and methods, and fields.
+SourceVis displays the evolution of Java source code metrics such as number of packages, classes, methods, and fields.
 
@@
 There is also research on how developers use software repositories to collaborate.
-Wagstrom \textit{et al.}~\cite{Patrick:Wagstrom:2012} explored the roles that developers take in a networked, social development environments like GitHub.
-These role were defined as their level of contribution as well as the types of issues they dealt with.
+Wagstrom \textit{et al.}~\cite{Patrick:Wagstrom:2012} explored the roles that developers take in networked, social development environments like GitHub.
+These roles were defined by their level of contribution as well as the types of issues they dealt with.
 They found that developers will take on multiple roles in a project and will sometimes fulfill the same role across different projects.
@@
 
-Some visualization tools are used for managing or making decisions about software rather than study the changes taking place.
-The service EasyBi~\cite{EasyBi} provides additional visualizations for Git repositories for the purpose of business decisions.
-EasyBi works by taking the git log file and generating charts and graphs from it.
+Some visualization tools are made for managing or making decisions about software rather than studying the changes taking place in an academic context.
+EasyBi~\cite{EasyBi} is a service that provides additional visualizations for Git repositories for the purpose of business decisions.
+EasyBi works by taking the \texttt{git log} file and generating charts and graphs from it.
 It tries to solve the issue of summarizing many commits and has greater controls for granularity and authors but still operates on commits and line differences.
@@
 
-Developers using the massively popular GitHub website for their Git repository hosting instantly have access to basic history visualizations, known as GitHub Graphs (see Figure~\ref{fig:GH_code_freq})~\cite{github-graphs}.
+Developers using the massively popular GitHub website for hosting their Git repositories instantly have access to basic visualizations, known as GitHub Graphs (see Figure~\ref{fig:GH_code_freq})~\cite{github-graphs}.
 GitHub Graphs tracks contributions based only on the number of commits and number of lines changed.
-This has the advantage of not relying on the of code which was committed.
-It does not matter what language is used or if errors exist in the commit, GitHub will simply output the number of line additions and deletions.
+This has the advantage of not relying on the code which was committed.
+It does not matter what language is used or if errors exist in the commit---GitHub will simply output the number of line additions and deletions.
 The issue with this type of analysis is that it does not give a clear idea of what is being changed, only that changes have occurred.
@@
 This extracted data can then be used in a variety of visualizations as discussed in Section~\ref{sec:viz}.
-The reason for using ASTs is that they capture the semantics of the code far more than simple line counts.
+The reason for using ASTs is that they capture the semantics of the code far better than simple line counts.
 
@@
 Non-trivial changes are those that affect how a compiler will interpret the source code, and may have effects on the executable code.
-Trivial changes are those that have no affect on how a compiler will interpret the source text, such as changing insignificant whitespace or comments in source code.
+Trivial changes are those that have no effect on how a compiler will interpret the source text, such as changing insignificant whitespace or comments in source code.
 Thus, ASTs of consecutive revisions in a code repository can be compared to see what \emph{structures} a programmer is modifying.
@@
 Each change of declaration or invocation is detected and tallied per file per commit.
-These changes are emitted in line-based file format.
+These changes are emitted in an ad hoc file format.
 For example, the following is the output of a commit that changed two declaration of \texttt{String} variables to \texttt{URI} variables:
@@
 In order to gather all the changes for a given repository over time, we generated ASTs for each commit and compared the difference.
-We did this by running \texttt{git log} over the Git repository and listing all the commit IDs in chronological order (Figure~\ref{fig:commits}).
+We did this by running \texttt{git log} over the Git repository and listed all commit IDs in chronological order (Figure~\ref{fig:commits}).
 For each commit, we collected the following information, as can be seen in Figure~\ref{fig:data-extracted}: the author's name, date of commit, commit ID (SHA), commit message, files the commit edited, and files in the repository at the time of the commit.
 For all the edited Java files we also compared the file at the time of the commit with its parent commit using AST differencing as described above.
-We achieved this by invoking \texttt{git difftool} and specifying the tool to be used as our AST differencing tool.
+We achieved this by invoking \texttt{git difftool} and specifying our AST differencing tool.
 One case where \texttt{git difftool} fails is when the commit has no parent, such as the initial commit.
@@
 A merge commit has two cases to consider: The first case is when the parents of the commit can be automatically merged.
-In this case all the relevant changes are already recorded in the parent of the commits so we can safely ignore the merge commit~\cite{git:merge}.
-The second case occurs when the parents cannot be merged automatically: There is a merge conflict which requires manual resolution on the part of a developer.
-In this case the Git user will try to resolve the issue by modifying code.
-If we decide to include the merge commit we run into the issue of to how to compare the merge commit with its multiple parent commits and the possible modifications made to resolve the merge conflict.
-We could consider comparing the merge commit with only one parent commit at a time, however; we will end up counting all the changes already recorded in the other parent branch.
-Additionally, all changes made during the resolution of a conflict will be attributed to the author making the merge; this may exclude the author that actually wrote the code in the first place.
+In this case all the relevant changes are already recorded in the parent of the commit so we can safely ignore the merge commit~\cite{git:merge}.
+The second case occurs when the parents cannot be merged automatically: this is a \emph{merge conflict} which requires manual resolution on the part of a developer.
+In this case, the developer will try to resolve the issue by modifying code.
+If we decide to include the merge commit we run into several edge case issues such as how to compare the merge commit with its multiple parent commits and how to account for the possible modifications made to resolve the merge conflict.
+We could consider comparing the merge commit with only one parent commit at a time; however, we will end up counting all the changes already recorded in the other parent branch.
+Additionally, all changes made during the resolution of a conflict will be attributed to the author making the merge.
+This may exclude the author that actually wrote the code in the first place.
 The other option is to simply ignore the merge commit such that we do not count the changes done during a merge commit.
-We made the assumptions that merge conflicts occur infrequently and that changes done to resolve a merge conflict are usually minor.
+We made the assumptions that merge conflicts occur infrequently and when they do occur, the changes made to resolve a merge conflict are usually minor.
 Therefore, we chose to ignore merge commits.
-Note that fast-forwards are not \emph{true merges}, hence they do not create merge commits~\cite{git:merge}; none of the previous cases involving merges occur.
+Note that fast-forwards are not \emph{true merges}, hence they do not create merge commits~\cite{git:merge} so none of the previous cases involving merges apply.
 
@@
 Although rich in detail, the AST diff data is difficult to analyze if displayed in its raw form; hence, we created an HTML5 web visualization application which is intended for developers, project managers, and researchers to use (Figure~\ref{fig:main-view}).
-The source code for TypeV---both the AST differencing component and the web application--- is available online on GitHub.\footnote{\url{https://github.com/mdfeist/TypeV}} A screencast demonstrating its use is also available online.\footnote{\url{https://archive.org/details/typev-demo}}
+The source code for TypeV---both the AST differencing component and the web application---is available online on GitHub.\footnote{\url{https://github.com/mdfeist/TypeV}} A screencast demonstrating its use is also available online.\footnote{\url{https://archive.org/details/typev-demo}}
 
@@
 To allow for a more detailed analysis of a specific component of the repository we added filtering tools.
-Users can filter by type name, view changes that fall within a specific start and end date, and select whether to see only declarations of types, types used in declarations and invocations, or only invocations.
+Users can filter by type name; view changes that fall within a specific start and end date; and select whether to see only declarations of types, types used in declarations and invocations, or only invocations.
 
@@
 They can be viewed as an enhanced version of GitHub's code frequency graph~\cite{github-graphs}, with semantic data.
-We chose to create a similar visualization to GitHub to work with a visualization people may already be familiar with and evaluate our visualization tools' density of data.
+We chose to create a similar visualization to GitHub to work with a layout people may already be familiar with and to evaluate our visualization tools' density of data.
 
-The stacked bar graphs succinctly display the relative amounts of additions and deletions per each time period.
+The stacked bar graphs succinctly displays the relative amounts of additions and deletions per each time period.
 This creates a dense view of the project's evolution.
-We chose to make the height of the bars relative to only the type itself to make it easy for a user to scan a single type and see when it was used modified.
-This makes tasks such finding a refactoring (see \ref{sec:antlr4}) very clear to see simply by browsing the line.
+We chose to make the height of the bars relative to only the type itself to make it easy for a user to scan a single type and see when it was modified.
+This makes tasks such as finding a refactoring (see Section~\ref{sec:antlr4}) very clear to see simply by browsing the line.
 We stacked the additions on top of the deletions such that users can immediately compare the relative frequency of changes of a single type throughout the chosen time span.
@@
 The blue serves as a contrast to the red, indicating the opposite action (in this case, additions).
-Since
 
 The rows are sorted so that the graph for the most used types or methods appear the top and are in descending order of use such that the most frequently used type is the first visible.
-The height allows users to see when certain types are being edited and ensures that highly used types do not overpower less used type making their changes less significant.
+The height allows users to see when certain types are being edited and ensures that highly used types do not overpower less used types making their changes less significant.
 
@@
 Table~\ref{tab:summary} lists a summary of the statistics measured.
+
+We displayed the density distributions for two different projects: ANTLR4 (Figure~\ref{fig:antlr4-density}) and Apache Tika (Figure~\ref{fig:tika-density}).
+These plots show the distribution of the amount of commits that have X\% of type and file coverage.
+For example, Figure~\ref{fig:tika-density} shows that there are many commits with 50\% type coverage and very few commits with 50\% file coverage.
+This shows that even though the type and file coverage plots may have a similar shape the commits are showing different levels of information.
+
 In general, type coverage and file coverage are strongly correlated, both using Pearson (linear) correlation, and Spearman (rank) correlation.
 As type coverage and file coverage are both cumulative measures, positive correlation is expected.
-We displayed the density distributions for two different projects: ANTLR4 (Figure~\ref{fig:antlr4}) and Apache Tika (Figure~\ref{fig:tika-density}).
 However, even when the distributions are significantly different (as with ANTLR and Curator), the correlation is still strongly positive, but not 100\%.
@@
 
-Using TypeV we can can see that at certain times there were major changes across all types. This shows us when major changes to the code were being done versus code maintenance.
+Using TypeV we can see that at certain times there were major changes across all types. This shows us when major changes to the code were being done versus code maintenance.
 
@@
 
-The purpose of \emph{deprecating} software features is to give developers time to change projects to follow a new or better standard without breaking the code.
+The purpose of \emph{deprecating} software features is to give developers time to change projects to follow a new or better standard without breaking existing code.
 When a piece of code becomes deprecated, continuing its use may lead to design flaws or even security vulnerabilities.
@@
 Given the time at which a feature was deprecated, seeing the progress of replacement has some important implications.
-In the case of a security flaw, that time could indicated times at which systems or users were vulnerable.
-If a feature is marked for removal, a project using a deprecated type has a limited amount of time to make the changes.
-Removing all instances of deprecated features could also mean a project can be upgraded to use new tools or features that were not possible while the deprecated features were present.
+In the case of a security flaw, that time could indicate times at which systems or its users were vulnerable.
+If a feature is marked for removal, a third party using the deprecated feature has a limited time to make the appropriate changes.
+Removing all instances of deprecated features could also mean a project can be upgraded to use new tools or features that were not possible when the deprecated features were present.
 

### 2016-06-20 paper/document.tex
Better hyphenation.

@@
 
-Apache Bookkeeper deprecated the method AbstractConfiguration.getLedgerManagerType() in release 4.2.0\footnote{\url{https://bookkeeper.apache.org/docs/r4.2.0/apidocs/deprecated-list.html}} and replaced it with AbstractConfiguration.getLedgerManagerFactoryClass().
-If we look at getLedgerManagerType() we can see at the time of the start of the deprecation all instances were removed and getLedgerManagerFactoryClass() was added (Figure~\ref{fig:bookkeeper-depr}).
-After the deprecation, we see no more changes to getLedgerManagerType() and only some additions of getLedgerManagerFactoryClass().
+Apache Bookkeeper deprecated the method Abstract\-Configuration.\-get\-Ledger\-Manager\-Type() in release 4.2.0\footnote{\url{https://bookkeeper.apache.org/docs/r4.2.0/apidocs/deprecated-list.html}} and replaced it with Abstract\-Configuration.\-get\-Ledger\-Manager\-Factory\-Class().
+If we look at get\-Ledger\-Manager\-Type() we can see at the time of the start of the deprecation all instances were removed and get\-Ledger\-Manager\-Factory\-Class() was added (Figure~\ref{fig:bookkeeper-depr}).
+After the deprecation, we see no more changes to get\-Ledger\-Manager\-Type() and only some additions of get\-Ledger\-Manager\-Factory\-Class().
 

### 2016-06-20 paper/document.tex
Latest revision of the paper

@@
 
-When managing group projects in an introductory software engineering course, it is very important to be able to see the contributions that students are making to the group repository to judge learning, progress, and collaboration.
-When one of the authors was supervising 4--5 groups of students, it became infeasible to look through all of the commits of each student.
-Students may take the wrong approach in a certain piece of code that can become ingrained in the project and very difficult to change.
-Some groups had problems with group members who did not complete their assigned work or made meaningless contributions.
-Issues like this can be difficult to detect and prevent.
-TypeV was used by the author to quickly summarize the work done by each group member, organized by the types they had been working on.
-This can show when intervention or guidance is required, without the large time commitment of pouring through commits.
-
 Bug triaging for large open source projects is a very time consuming task~\cite{badashian2015}, with hundreds  of bugs being reported daily.
@@
 
+Internal threats to validity include our treatment of merge conflicts.
+As mentioned before, we ignore all merges, so if there is a merge conflict, it is possible that a developer makes non-trivial changes.
+Changes made during merge conflict resolution are missed by TypeV.
+We have made the assumptions that if this case happens, the changes are minor.
+
 Some actions in Git present a problem gathering the statistics.
@@
 
-Internal threats to validity include our treatment of merge conflicts.
-As mentioned before, we ignore all merges, so if there is a merge conflict, it is possible that a developer makes non-trivial changes.
-Changes made during merge conflict resolution are missed by TypeV.
-We have made the assumptions that if this case happens, the changes are minor.
+An external threat to validity is the limited number of projects used in the validation.
+We picked curated projects that had a long development cycle and looked into some case studies.
+However, because of the limited number of projects we observed, our conclusions may not hold for the majority of software projects.
 
 \subsection{Future work}
-
-Currently, we treat every instance of a generic type as a different type.
-For example, considered the \texttt{List<T>} generic class in Java.
-If a user declares a List of Strings and later declares a List of Integers these will appear as two different types in TypeV.
-However, if they declare multiple instances of a List of Strings then these instances will count toward the same type.
-In the future, we want to work on a better way to track and represent generic types.
 

### 2016-06-20 www/js/project-view.ts
Minor changes.

@@ -182,21 +182,21 @@ function drawGraph(data: DataView, width: number) {
       .attr('width', function (cell) {
         return cellWidthFromScale(cell, xScale);
       })
-      .attr('transform', function (cell: Cell) {
+      .attr('transform', function (cell) {
         var proportion = cell.numberOfDeletions / cell.numberOfObservations;
         var height = cell.numberOfObservations / type.numberOfObservationsInLargestCell;
         var topHalf = proportion * maxCellHeight * height;
 
         return `translate(0, ${topHalf})`;
       })
-      .attr('height', function (cell: Cell) {
+      .attr('height', function (cell) {
         var proportion = cell.numberOfAdds / cell.numberOfObservations;
         var height = cell.numberOfObservations / type.numberOfObservationsInLargestCell;
         return proportion * maxCellHeight * height;
       });
 
     /* Mouse Click: Show Cell stats */
-    cell.on("click", function(cell_data: Cell) {
+    cell.on("click", function(cellData: Cell) {
       var stats = d3.select("#stats-body");
       stats.selectAll('.content').remove();
 
@@ -206,21 +206,21 @@ function drawGraph(data: DataView, width: number) {
       info.append('b')
         .text("Info:");
       info.append('p')
-        .text("Additions: " + cell_data.numberOfAdds);
+        .text("Additions: " + cellData.numberOfAdds);
       info.append('p')
-        .text("Deletions: " + cell_data.numberOfDeletions);
+        .text("Deletions: " + cellData.numberOfDeletions);
       /* TODO: Graphs here! */
       info.append('p')
-        .text("Authors: " + cell_data.authors.length);
+        .text("Authors: " + cellData.authors.length);
       info.append('p')
-        .text("Commits: " + cell_data.commits.length);
+        .text("Commits: " + cellData.commits.length);
 
       info.append('br');
 
       info.append('b')
         .text("Authors:");
 
-      cell_data.authors.forEach(function (author) {
+      cellData.authors.forEach(function (author) {
         info.append('p')
           .text(author);
       });
@@ -232,7 +232,7 @@ function drawGraph(data: DataView, width: number) {
       info.append('b')
         .text("Commits:");
 
-      cell_data.commits.forEach(function (commitID) {
+      cellData.commits.forEach(function (commitID) {
         var commit = commit_map[commitID];
         info.append('p')
           .text("Commit ID: " + commit.commitID);

### 2016-06-21 server.py
Minor redesign.

@@ -453,6 +453,10 @@ def send_js(path):
 def send_css(path):
     return send_from_directory(app.static_folder + '/css', path)
 
+@app.route('/fonts/<path:path>')
+def send_fonts(path):
+    return send_from_directory(app.static_folder + '/fonts', path)
+
 @app.route('/')
 def root():
     return redirect('/projects/')

### 2016-06-21 www/js/index.ts
Enable Chosen on some placeholder UI elements.

@@ -13,7 +13,8 @@ function start() {
   var type = $("#type-select").val();
   loadProject(type);
 
-  $("#type-select").on("change", function(){
+  // Automatically reload the project data when the type select is changed.
+  $("#type-select").on("change", function () {
     var type = $("#type-select").val();
     loadProject(type);
   });
@@ -30,6 +31,10 @@ function hasDatePicker() {
   return (input.value !== notADateValue);
 }
 
+function enableChosen() {
+  $('select.author-aliases').chosen();
+}
+
 function loadProject(type: 'Declarations' | 'Types' | 'Invocations') {
   var xhttp = new XMLHttpRequest();
   xhttp.onreadystatechange = function () {
@@ -38,6 +43,7 @@ function loadProject(type: 'Declarations' | 'Types' | 'Invocations') {
       var project: Project = JSON.parse(xhttp.responseText);
       var results = createTable(project, filter);
       patchInputs(results);
+
 
       var link = makeCSVLink(results);
       $('#csv-link').attr('href', link);
@@ -111,20 +117,26 @@ function patchInputs(results: DataView) {
 }
 
 window.toggleView = function () {
-  $("#dna-table").toggleClass("collapse");
-  $("#stats-table").toggleClass("collapse");
+  $("#dna-table").toggleClass('collapse');
+  $("#stats-table").toggleClass('collapse');
 }
 
 window.toggleAuthors = function () {
-  $("#authors").toggleClass( "collapse" );
+  let $authors = $("#authors");
+  $authors.toggleClass('collapse');
+
+  /* Chosen must be enabled while the panel is visible. */
+  if (!$authors.hasClass('collapse')) {
+    enableChosen();
+  }
 }
 
 window.toggleStats = function () {
-  $("#stats").toggleClass( "collapse" );
+  $("#stats").toggleClass('collapse');
 }
 
 window.toggleFilters = function () {
-  $("#filters").toggleClass( "collapse" );
+  $("#filters").toggleClass('collapse');
 }
 
 window.uncheckAuthors = function () {

### 2016-06-21 www/js/index.d.ts
Render the dummy author selector in JavaScript.

@@ -18,6 +18,7 @@ interface Window {
   DATA: Project;
   preprocessedData: any;
   filteredData: any;
+  authors: string[];
 
   redrawTable(): void;
   toggleView(): void;

### 2016-06-21 www/js/index.ts

@@ -31,7 +31,57 @@ function hasDatePicker() {
   return (input.value !== notADateValue);
 }
 
-function enableChosen() {
+/* From: http://stackoverflow.com/a/6234804 */
+function escapeHTML(unsafe: string): string {
+  return unsafe
+    .replace(/&/g, "&amp;")
+    .replace(/</g, "&lt;")
+    .replace(/>/g, "&gt;")
+    .replace(/"/g, "&quot;")
+    .replace(/'/g, "&#039;");
+}
+
+/**
+ * Renders the author panel.
+ * @return {string} HTML for the authors panel.
+ */
+function renderAuthorPanel(allAuthors: string[], authorUnsafe: string): string {
+  /* TODO: take in a data structure describing author configuration. */
+  let author = escapeHTML(authorUnsafe);
+
+  return (`
+    <li>
+    <div class="panel panel-default author-configuration">
+      <div class="panel-heading"> ${author} </div>
+      <div class="panel-body">
+        <label><input type="checkbox" checked
+                      class="author-check-box"
+                      value="${author}">
+          Show this author
+        </label>
+        <label> Author aliases <br>
+          <select multiple class="author-aliases">
+            ${allAuthors.filter(name => name != author).map(unsafe =>
+                `<option value="${unsafe}">${escapeHTML(unsafe)}</option>`
+            )}
+          </select>
+        </label>
+      </div>
+    </div>
+    </li>
+  `);
+}
+
+/**
+ * Rerenders the list inside #authors-list
+ * @return {[type]} [description]
+ */
+function renderAuthorPicker() {
+  let authors = window.authors;
+  let html = authors.map(renderAuthorPanel.bind(window, authors)).join('');
+  $('#authors-list').html(html);
+
+  /* Enable chosen: */
   $('select.author-aliases').chosen();
 }
 
@@ -127,7 +177,7 @@ window.toggleAuthors = function () {
 
   /* Chosen must be enabled while the panel is visible. */
   if (!$authors.hasClass('collapse')) {
-    enableChosen();
+    renderAuthorPicker();
   }
 }
 

### 2016-06-23 www/js/ast-diff.ts
Implement author class in really dirty way.

@@ -58,6 +58,7 @@ export default class ASTDiff {
    * NOTE: This was an enumerable property in an older version. This is
    * probably unnecessary now.
    */
+  /** TODO: Name this author ID or something. */
   get author(): string {
     return this._commit.author.shorthand;
   }

### 2016-06-23 www/js/author-identity.ts

@@ -1,10 +1,13 @@
 import assert from './assert';
 
 export default class AuthorIdentity {
-  public name: String;
-  public email: String;
+  public name: string;
+  public email: string;
 
-  constructor(alias: String) {
+  /** XXX: Fix memory leak! */
+  private static _instances = new Map<string, AuthorIdentity>();
+
+  constructor(alias: string) {
     [this.name, this.email] = parseAlias(alias);
   }
 
@@ -15,13 +18,25 @@ export default class AuthorIdentity {
   toString() {
     return this.shorthand;
   }
+
+  static get(name: string): AuthorIdentity {
+    let normalized = new AuthorIdentity(name).shorthand;
+    let id = AuthorIdentity._instances.get(normalized);
+
+    if (id === undefined) {
+      id = new AuthorIdentity(name);
+      AuthorIdentity._instances.set(id.shorthand, id);
+    }
+
+    return id;
+  }
 }
 
-function parseAlias(text: String): [String, String] {
+function parseAlias(text: string): [string, string] {
   let match = text.match(/^\s*([^<]+)<([^>]+)>\s*$/);
   if (match === null || match.length !== 3) {
     throw new Error(`Could not match author alias: ${text}`);
   }
 
-  return [match[1], match[2]];
+  return [match[1].trim(), match[2].trim()];
 }

### 2016-06-23 www/js/author.ts

@@ -1,0 +1,51 @@
+import AuthorIdentity from './author-identity';
+
+/**
+ * Represents one verified full author.
+ */
+ export default class Author {
+   private _aliases: Set<AuthorIdentity>;
+
+   constructor(
+     public primaryIdentity: AuthorIdentity,
+     aliases?: AuthorIdentity[]
+   ) {
+     this._aliases = new Set(aliases);
+   }
+
+   /**
+    * The author's name.
+    * @return {string} [description]
+    */
+   get name(): string {
+     return this.primaryIdentity.name;
+   }
+
+   /**
+    * The author's email.
+    * @return {string} [description]
+    */
+   get email(): string {
+     return this.primaryIdentity.email;
+   }
+
+   /**
+    * An idenitification string, suitable for mapping back to the author.
+    * @return {string} [description]
+    */
+   get id(): string {
+     return this.shorthand;
+   }
+
+   /**
+    * A list of secondary aliases.
+    * @return {Set<AuthorIdentity>} [description]
+    */
+   get aliases(): Set<AuthorIdentity> {
+     return new Set(this._aliases);
+   }
+
+   get shorthand(): string {
+     return this.primaryIdentity.shorthand;
+   }
+ }

### 2016-06-23 www/js/data-filter.ts

@@ -3,6 +3,8 @@ import TimeSlice from './time-slice';
 import ASTDiff from './ast-diff';
 import Commit, {CommitMap} from './commit';
 import {PreprocessedData} from './preprocess-data';
+import AuthorIdentity from './author-identity';
+import Author from './author';
 
 import assert from './assert';
 import { first, last, addAll } from './utils';
@@ -27,9 +29,7 @@ interface RawFilteringResult {
   astDiffs: ASTDiff[];
 }
 
-export interface AuthorStatistics {
-  [authorID: string]: CommitStatistics[];
-}
+export type AuthorStatistics = Map<Author, CommitStatistics[]>;
 
 export interface CommitStatistics {
   type: EntitySummary;
@@ -144,7 +144,8 @@ function filterTypes(data: PreprocessedData, filters: Filter) {
   var endDate: Date = filters.end || last(data.astDiffs).date;
   var numberOfTypesUpperBound = filters.limit || Infinity;
   var stepSize: StepSize = filters.stepSize || 'month';
-  var authors: string[] = filters.authors || [];
+  var authorNames: string[] =
+    filters.authors.length > 0 ? filters.authors : data.authorNames;
   var typeFilter = filters.typeFilter ? filters.typeFilter : null;
 
   assert(startDate < endDate);
@@ -179,13 +180,35 @@ function filterTypes(data: PreprocessedData, filters: Filter) {
     typeMap[type.name] = type;
   });
 
-  var authorMap: AuthorStatistics = {};
-  var authorSets: {
-    [name: string]: {
-      files: Set<string>;
-      types: Set<string>;
+  /* XXX: TEMPORARY! */
+  let authors = new Map<AuthorIdentity, Author>();
+  let commits = data.commits;
+
+  /* XXX: TEMPORARY! */
+  for (let name of authorNames) {
+    let id = AuthorIdentity.get(name)
+    authors.set(id, new Author(id));
+  }
+
+  /* XXX: TEMPORARY! */
+  function ensureDefined<T> (value: T): T {
+    if (value === null || value === undefined) {
+      debugger;
+      throw new Error('Got an undefined value')
     }
-  } = {};
+    return value;
+  }
+
+  /* XXX: TEMPORARY! */
+  let identityToAuthor = (id: AuthorIdentity) => ensureDefined(authors.get(id));
+
+  type CoverageCounts = {
+    files: Set<string>;
+    types: Set<string>;
+  };
+
[…]

### 2016-06-23 www/js/index.ts

@@ -1,6 +1,7 @@
 ///<reference path="index.d.ts" />
 import { createTable, createTable2, makeCSVLink } from "./project-view";
 import DataView from './data-filter';
+import Author from './author';
 
 /* Add date picker widgets on platforms that don't have them. */
 if (!hasDatePicker()) {

### 2016-06-23 www/js/preprocess-data.ts

@@ -8,6 +8,7 @@ export interface PreprocessedData {
   typeNames: Set<string>;
   commits: CommitMap;
   astDiffs: ASTDiff[];
+  authorNames: string[];
 };
 
 /**
@@ -21,6 +22,7 @@ export default function preprocessData(data: Project): PreprocessedData {
   assert(data.types instanceof Array);
   assert(data.commits instanceof Array);
   assert(data.dates instanceof Array);
+  assert(data.authors instanceof Array);
 
   /* A set of type names */
   var typeNames = new Set(data.types);
@@ -28,10 +30,11 @@ export default function preprocessData(data: Project): PreprocessedData {
   var commits = createCommitMap(commitsFromJsonToInternalFormat(data.commits));
 
   return {
-    typeNames,
-    commits,
     /* A copy of AST Diff data, in asscending order of date. */
-    astDiffs: createASTDiffsInAscendingOrder(data.dates, commits)
+    astDiffs: createASTDiffsInAscendingOrder(data.dates, commits),
+    authorNames: data.authors,
+    commits,
+    typeNames
   };
 }
 
@@ -43,7 +46,7 @@ function commitsFromJsonToInternalFormat(commits: CommitFromJSON[]): Commit[] {
   for (let commitJSON of commits) {
     let commit = <any> commitJSON as Commit;
     commit.date = new Date(commitJSON.date);
-    commit.author = new AuthorIdentity(commitJSON.author);
+    commit.author = AuthorIdentity.get(commitJSON.author);
   }
   return <any[]> commits as Commit[];
 }

### 2016-06-23 www/js/project-view.ts

@@ -7,7 +7,6 @@
  *  - Ian Watts
  */
 
-
 import assert from './assert';
 import preprocessData, {PreprocessedData} from './preprocess-data';
 import DataView, {CommitStatistics} from './data-filter';
@@ -420,8 +419,10 @@ function drawStats(data: DataView, width: number) {
    * Per author stats.
    */
 
+  let authors = Array.from(data.authorStats.keys());
+
   var authorCoverage = d3.select('#coverage-by-author').selectAll('.author-coverage')
-      .data(data.primaryAuthorAliases)
+      .data(authors)
     .enter().append('svg')
       .attr('height', rowHeight + 32)
       .attr('width', width)
@@ -429,11 +430,11 @@ function drawStats(data: DataView, width: number) {
 
   authorCoverage.append('path')
     .classed('file-coverage', true)
-    .attr('d', (authorName) => lineFunctionSmallFiles(data.authorStats[authorName]));
+    .attr('d', (author) => lineFunctionSmallFiles(data.authorStats.get(author)));
 
   authorCoverage.append('path')
     .classed('type-coverage', true)
-    .attr('d', (authorName) => lineFunctionSmallTypes(data.authorStats[authorName]));
+    .attr('d', (author) => lineFunctionSmallTypes(data.authorStats.get(author)));
 
   /* Axes. */
   authorCoverage.append('g')
@@ -442,7 +443,7 @@ function drawStats(data: DataView, width: number) {
 
   authorCoverage.append('text')
     .attr('transform', `translate(${marginLeft}, 16)`)
-    .text(authorName => authorName.toString());
+    .text((author) => author.name);
 
   /* TODO: Make independent axis for each author... */
   /*
@@ -471,17 +472,17 @@ export function makeCSVLink(data: DataView) {
   addRow('Metric', 'Author', 'Date', 'Coverage');
   type Metric = 'file' | 'type';
 
-  for (var authorName of Object.keys(data.authorStats)) {
-    for (var stats of data.authorStats[authorName]) {
+  for (var author of data.authorStats.keys()) {
+    for (var stats of data.authorStats.get(author)) {
       addRow(
         'file',
-        authorName,
+        author.id,
         +stats.date, // Coerce to Unix timestamp in ms
         stats.file.cumulative / filesTotal
       );
       addRow(
         'type',
-        authorName,
+        author.id,
         +stats.date, // Coerce to Unix timestamp in ms
         stats.type.cumulative / typesTotal
       );

### 2016-06-23 www/js/data-filter.ts
Remove debugger stuff.

@@ -191,16 +191,7 @@ function filterTypes(data: PreprocessedData, filters: Filter) {
   }
 
   /* XXX: TEMPORARY! */
-  function ensureDefined<T> (value: T): T {
-    if (value === null || value === undefined) {
-      debugger;
-      throw new Error('Got an undefined value')
-    }
-    return value;
-  }
-
-  /* XXX: TEMPORARY! */
-  let identityToAuthor = (id: AuthorIdentity) => ensureDefined(authors.get(id));
+  let identityToAuthor = (id: AuthorIdentity) => authors.get(id);
 
   type CoverageCounts = {
     files: Set<string>;

### 2016-06-23 www/js/author-identity.ts
Fake a private constructor.

@@ -1,5 +1,11 @@
 import assert from './assert';
 
+/**
+ * A private symbol used to fake a private constructor.
+ * @return {[type]} [description]
+ */
+const PRIVATE = Symbol();
+
 export default class AuthorIdentity {
   public name: string;
   public email: string;
@@ -7,7 +13,10 @@ export default class AuthorIdentity {
   /** XXX: Fix memory leak! */
   private static _instances = new Map<string, AuthorIdentity>();
 
-  constructor(alias: string) {
+  constructor(alias: string, token: any) {
+    if (token !== PRIVATE) {
+      throw new Error(`Use AuthorIdentity.get("${alias}") instead.`);
+    }
     [this.name, this.email] = parseAlias(alias);
   }
 
@@ -19,15 +28,19 @@ export default class AuthorIdentity {
     return this.shorthand;
   }
 
+  /**
+   * Get a globally unique instance.
+   * @return {AuthorIdentity}      [description]
+   */
   static get(name: string): AuthorIdentity {
-    let normalized = new AuthorIdentity(name).shorthand;
+    let normalized = new AuthorIdentity(name, PRIVATE).shorthand;
     let id = AuthorIdentity._instances.get(normalized);
 
     if (id === undefined) {
-      id = new AuthorIdentity(name);
+      id = new AuthorIdentity(name, PRIVATE);
       AuthorIdentity._instances.set(id.shorthand, id);
     }
-
+  
     return id;
   }
 }

### 2016-06-23 www/js/author/author-identity.ts
Refactor: Move author stuff to directory.

@@ -1,5 +1,3 @@
-import assert from './assert';
-
 /**
  * A private symbol used to fake a private constructor.
  * @return {[type]} [description]
@@ -40,7 +38,7 @@ export default class AuthorIdentity {
       id = new AuthorIdentity(name, PRIVATE);
       AuthorIdentity._instances.set(id.shorthand, id);
     }
-  
+
     return id;
   }
 }

### 2016-06-23 www/js/author/index.ts

@@ -1,4 +1,5 @@
 import AuthorIdentity from './author-identity';
+export { AuthorIdentity };
 
 /**
  * Represents one verified full author.

### 2016-06-23 www/js/commit.ts

@@ -1,4 +1,4 @@
-import AuthorIdentity from './author-identity';
+import {AuthorIdentity} from './author';
 
 export default Commit;
 

### 2016-06-23 www/js/data-filter.ts

@@ -3,8 +3,7 @@ import TimeSlice from './time-slice';
 import ASTDiff from './ast-diff';
 import Commit, {CommitMap} from './commit';
 import {PreprocessedData} from './preprocess-data';
-import AuthorIdentity from './author-identity';
-import Author from './author';
+import Author, {AuthorIdentity} from './author';
 
 import assert from './assert';
 import { first, last, addAll } from './utils';

### 2016-06-23 www/js/index.ts

@@ -1,7 +1,6 @@
 ///<reference path="index.d.ts" />
 import { createTable, createTable2, makeCSVLink } from "./project-view";
 import DataView from './data-filter';
-import Author from './author';
 
 /* Add date picker widgets on platforms that don't have them. */
 if (!hasDatePicker()) {

### 2016-06-23 www/js/preprocess-data.ts

@@ -1,5 +1,5 @@
 import ASTDiff from './ast-diff';
-import AuthorIdentity from './author-identity';
+import {AuthorIdentity} from './author';
 import Commit, {CommitMap} from './commit';
 
 import assert from './assert';

### 2016-06-23 www/js/author/configuration.ts
Add author configuration.

@@ -1,0 +1,83 @@
+import Author, {AuthorIdentity} from './';
+
+/**
+ * Configuration for authors.
+ */
+export default class AuthorConfiguration {
+  private _enabled = new WeakSet<Author>();
+  private _aliases = new Map<AuthorIdentity, Author>();
+
+  constructor({aliases, enabled}: ConfigurationJSON) {
+    /* Add all aliases. */
+    for (let name of Object.keys(aliases)) {
+      this._mapTo(name, aliases[name]);
+    }
+
+    /* Enable all authors explicitly. */
+    for (let name of enabled) {
+      this.enableAuthorByName(name);
+    }
+  }
+
+  getAuthorByName(name: string): Author {
+    let alias = AuthorIdentity.get(name);
+    let author = this._aliases.get(alias);
+
+    if (!author) {
+      throw new Error(`Author not known: ${alias}`);
+    }
+
+    return author;
+  }
+
+  enableAuthorByName(name: string): void {
+    let author = this.getAuthorByName(name);
+    this._enabled.add(author);
+  }
+
+  disableAuthorByName(name: string): void {
+    let author = this.getAuthorByName(name);
+    this._enabled.delete(author);
+  }
+
+  addAlias(alias: AuthorIdentity, author: Author) {
+    let existingMapping: Author = this._aliases.get(alias)
+
+    /* Check if we're remapping an existing alias. */
+    if (existingMapping && existingMapping.primaryIdentity !== alias) {
+      throw new Error(`Alias already mapped: ${alias.shorthand} => ${existingMapping.shorthand}`);
+    }
+
+    this._aliases.set(alias, author);
+  }
+
+  isEnabled(author: Author): boolean {
+    return this._enabled.has(author);
+  }
+
+  /**
+   * Maps an alias name to an author.
+   */
+  protected _mapTo(aliasName: string, authorName: string) {
+    let alias = AuthorIdentity.get(aliasName)
+    let author = this._vivifyAuthor(alias);
+
+    this.addAlias(alias, author);
+  }
+
+  protected _vivifyAuthor(id: AuthorIdentity): Author {
+    let author = this._aliases.get(id);
+
+    if (!author) {
+      author = new Author(id);
+      this.addAlias(id, author);
+    }
+
+    return author;
+  }
+}
+
+interface ConfigurationJSON {
+  aliases: { [name: string]: string };
+  enabled: string[];
+}

### 2016-06-23 www/js/author/configuration.ts
Create authors entirely in JavaScript.

@@ -21,6 +21,10 @@ export default class AuthorConfiguration {
 
   getAuthorByName(name: string): Author {
     let alias = AuthorIdentity.get(name);
+    return this.get(alias);
+  }
+
+  get(alias: AuthorIdentity): Author {
     let author = this._aliases.get(alias);
 
     if (!author) {

### 2016-06-23 www/js/author/index.ts

@@ -1,5 +1,7 @@
 import AuthorIdentity from './author-identity';
-export { AuthorIdentity };
+import AuthorConfiguration from './configuration';
+
+export { AuthorConfiguration, AuthorIdentity };
 
 /**
  * Represents one verified full author.

### 2016-06-23 www/js/index.ts

@@ -77,7 +77,10 @@ function renderAuthorPanel(allAuthors: string[], authorUnsafe: string): string {
  * @return {[type]} [description]
  */
 function renderAuthorPicker() {
-  let authors = window.authors;
+  let authors = window.authors.sort((a, b) =>
+                                    a.trim().toLowerCase() < b.trim().toLowerCase() ?
+                                      -1 : 1
+                                   );
   let html = authors.map(renderAuthorPanel.bind(window, authors)).join('');
   $('#authors-list').html(html);
 
@@ -123,8 +126,8 @@ function getFilters(): Filter {
     endDate = new Date(timestamp);
   }
 
+  /* Find enabled authors. */
   var _authors: string[] = [];
-
   $('.author-check-box').each(function() {
     if($(this).prop('checked')) {
       _authors.push($(this).val());

### 2016-06-23 www/js/project-view.ts

@@ -8,12 +8,15 @@
  */
 
 import assert from './assert';
-import preprocessData, {PreprocessedData} from './preprocess-data';
+
+import Author, {AuthorConfiguration} from './author';
+import Cell from './cell';
+import Commit, {CommitMap} from './commit';
 import DataView, {CommitStatistics} from './data-filter';
 import JavaType from './java-type';
-import Cell from './cell';
 import TimeSlice from './time-slice';
-import Commit, {CommitMap} from './commit';
+
+import preprocessData, {PreprocessedData} from './preprocess-data';
 
 var CELL_INFO_WIDTH = 500;
 
@@ -33,8 +36,12 @@ cellInfo.append('div')
 
 /* Draw table given JSON */
 export function createTable(data: Project, filter: Filter) {
+  /* TODO: Remove global! */
   window.DATA = data;
+  /* TODO: Remove global! */
   window.preprocessedData = preprocessData(data);
+  /* TODO: Remove global! */
+  window.authors = data.authors;
 
   return createTable2(filter);
 }
@@ -48,6 +55,49 @@ export function createTable2(filter: Filter) {
 
   /* Clear previous table */
   dnaTable.innerHTML = "";
+
+
+  let config = new AuthorConfiguration({
+    aliases: {
+      "Andrey Yegorov <ayegorov@salesforce.com>":       "Andrey Yegorov <ayegorov@salesforce.com>",
+      "Arun M. Krishnakumar <akrishnakumar@salesforce.com>":       "Arun M. Krishnakumar <akrishnakumar@salesforce.com>",
+      "Benjamin Reed <breed@apache.org>":       "Benjamin Reed <breed@apache.org>",
+      "eolivelli <eolivelli@gmail.com>":       "eolivelli <eolivelli@gmail.com>",
+      "Flavio Paiva Junqueira <fpj@apache.org>":       "Flavio Paiva Junqueira <fpj@apache.org>",
+      "fpj <fpj@apache.org>":       "fpj <fpj@apache.org>",
+      "Ivan Brendan Kelly <ivank@apache.org>":       "Ivan Brendan Kelly <ivank@apache.org>",
+      "Ivan Kelly <ivan@bleurgh.com>":       "Ivan Brendan Kelly <ivank@apache.org>",
[…]

### 2016-06-23 www/js/author/configuration.ts
Fix author merging logic.

@@ -63,8 +63,9 @@ export default class AuthorConfiguration {
    * Maps an alias name to an author.
    */
   protected _mapTo(aliasName: string, authorName: string) {
-    let alias = AuthorIdentity.get(aliasName)
-    let author = this._vivifyAuthor(alias);
+    let alias = AuthorIdentity.get(aliasName);
+    let primaryIdentity = AuthorIdentity.get(authorName);
+    let author = this._vivifyAuthor(primaryIdentity);
 
     this.addAlias(alias, author);
   }

### 2016-06-23 www/js/author/index.ts

@@ -6,49 +6,39 @@ export { AuthorConfiguration, AuthorIdentity };
 /**
  * Represents one verified full author.
  */
- export default class Author {
-   private _aliases: Set<AuthorIdentity>;
+export default class Author {
+  constructor(public primaryIdentity: AuthorIdentity) {
+  }
 
-   constructor(
-     public primaryIdentity: AuthorIdentity,
-     aliases?: AuthorIdentity[]
-   ) {
-     this._aliases = new Set(aliases);
-   }
+  /**
+   * The author's name.
+   * @return {string} [description]
+   */
+  get name(): string {
+    return this.primaryIdentity.name;
+  }
 
-   /**
-    * The author's name.
-    * @return {string} [description]
-    */
-   get name(): string {
-     return this.primaryIdentity.name;
-   }
+  /**
+   * The author's email.
+   * @return {string} [description]
+   */
+  get email(): string {
+    return this.primaryIdentity.email;
+  }
 
-   /**
-    * The author's email.
-    * @return {string} [description]
-    */
-   get email(): string {
-     return this.primaryIdentity.email;
-   }
+  /**
+   * An idenitification string, suitable for mapping back to the author.
+   * @return {string} [description]
+   */
+  get id(): string {
+    return this.shorthand;
+  }
 
-   /**
-    * An idenitification string, suitable for mapping back to the author.
-    * @return {string} [description]
-    */
-   get id(): string {
-     return this.shorthand;
-   }
+  get shorthand(): string {
+    return this.primaryIdentity.shorthand;
+  }
 
-   /**
-    * A list of secondary aliases.
-    * @return {Set<AuthorIdentity>} [description]
-    */
-   get aliases(): Set<AuthorIdentity> {
-     return new Set(this._aliases);
-   }
-
-   get shorthand(): string {
-     return this.primaryIdentity.shorthand;
-   }
- }
+  toString(): string {
+    return `Author(${JSON.stringify(this.id)})`;
+  }
+}

### 2016-06-23 www/js/data-filter.ts
Incorporate AuthorConfiguration to data filter.

@@ -3,7 +3,7 @@ import TimeSlice from './time-slice';
 import ASTDiff from './ast-diff';
 import Commit, {CommitMap} from './commit';
 import {PreprocessedData} from './preprocess-data';
-import Author, {AuthorIdentity} from './author';
+import Author, {AuthorIdentity, AuthorConfiguration} from './author';
 
 import assert from './assert';
 import { first, last, addAll } from './utils';
@@ -76,8 +76,8 @@ export default class DataView {
    * # authors -- Filter only to the given authors.
    * # typeFilter -- only types matching this pattern are selected.
    */
-  static filter(data: PreprocessedData, filters: Filter) {
-    const rawFilteredData = filterTypes(data, filters);
+  static filter(data: PreprocessedData, filters: Filter, config: AuthorConfiguration) {
+    const rawFilteredData = filterTypes(data, filters, config);
     return new DataView(rawFilteredData);
   }
 
@@ -133,7 +133,7 @@ export default class DataView {
  *
  * Each type has its Cells, with ASTDiffs.
  */
-function filterTypes(data: PreprocessedData, filters: Filter) {
+function filterTypes(data: PreprocessedData, filters: Filter, authors: AuthorConfiguration) {
   /* Let filters be undefined or null. */
   filters = filters ? filters : {};
 
@@ -179,18 +179,7 @@ function filterTypes(data: PreprocessedData, filters: Filter) {
     typeMap[type.name] = type;
   });
 
-  /* XXX: TEMPORARY! */
-  let authors = new Map<AuthorIdentity, Author>();
   let commits = data.commits;
-
-  /* XXX: TEMPORARY! */
-  for (let name of authorNames) {
-    let id = AuthorIdentity.get(name)
-    authors.set(id, new Author(id));
-  }
-
-  /* XXX: TEMPORARY! */
-  let identityToAuthor = (id: AuthorIdentity) => authors.get(id);
 
   type CoverageCounts = {
     files: Set<string>;
@@ -240,7 +229,7 @@ function filterTypes(data: PreprocessedData, filters: Filter) {
   /* Will need to recalculate this for this measurement: */
   typesOverall = new Set();
   forEachCommit(applicableDiffs, (commit, files, types) => {
-    let author = identityToAuthor(commit.author);
+    let author = authors.get(commit.author);
     let date = new Date(commit.date.valueOf());
 
     if (!authorMap.has(author)) {

### 2016-06-23 www/js/project-view.ts

@@ -64,7 +64,7 @@ export function createTable2(filter: Filter) {
       "Benjamin Reed <breed@apache.org>":       "Benjamin Reed <breed@apache.org>",
       "eolivelli <eolivelli@gmail.com>":       "eolivelli <eolivelli@gmail.com>",
       "Flavio Paiva Junqueira <fpj@apache.org>":       "Flavio Paiva Junqueira <fpj@apache.org>",
-      "fpj <fpj@apache.org>":       "fpj <fpj@apache.org>",
+      "fpj <fpj@apache.org>":              "Flavio Paiva Junqueira <fpj@apache.org>",
       "Ivan Brendan Kelly <ivank@apache.org>":       "Ivan Brendan Kelly <ivank@apache.org>",
       "Ivan Kelly <ivan@bleurgh.com>":       "Ivan Brendan Kelly <ivank@apache.org>",
       "Ivan Kelly <ivank@apache.org>":       "Ivan Brendan Kelly <ivank@apache.org>",
@@ -102,7 +102,7 @@ export function createTable2(filter: Filter) {
   // TODO: get rid of window
   var processed = window.preprocessedData;
   // TODO: get rid of window
-  var data = window.filteredData = DataView.filter(processed, filter);
+  var data = window.filteredData = DataView.filter(processed, filter, config);
   drawGraph(data, dnaTable.offsetWidth);
   drawStats(data, dnaTable.offsetWidth);
 

### 2016-06-23 www/js/author/configuration.ts
Render author selector using AuthorConfig.

@@ -6,6 +6,7 @@ import Author, {AuthorIdentity} from './';
 export default class AuthorConfiguration {
   private _enabled = new WeakSet<Author>();
   private _aliases = new Map<AuthorIdentity, Author>();
+  private _known = new Set<Author>();
 
   constructor({aliases, enabled}: ConfigurationJSON) {
     /* Add all aliases. */
@@ -15,8 +16,17 @@ export default class AuthorConfiguration {
 
     /* Enable all authors explicitly. */
     for (let name of enabled) {
-      this.enableAuthorByName(name);
+      let author = this.getAuthorByName(name);
+      this.enable(author);
     }
+  }
+
+  get authors(): Set<Author> {
+    return new Set(this._known);
+  }
+
+  get aliases(): Iterable<AuthorIdentity> {
+    return this._aliases.keys();
   }
 
   getAuthorByName(name: string): Author {
@@ -34,14 +44,20 @@ export default class AuthorConfiguration {
     return author;
   }
 
-  enableAuthorByName(name: string): void {
-    let author = this.getAuthorByName(name);
+  enable(author: Author): void {
     this._enabled.add(author);
   }
 
-  disableAuthorByName(name: string): void {
-    let author = this.getAuthorByName(name);
+  disable(author: Author): void {
     this._enabled.delete(author);
+  }
+
+  /**
+   * True if the ID is a primary identity of some author.
+   */
+  isPrimaryIdentity(id: AuthorIdentity): boolean {
+    let author = this.get(id);
+    return author.primaryIdentity === id;
   }
 
   addAlias(alias: AuthorIdentity, author: Author) {
@@ -75,6 +91,7 @@ export default class AuthorConfiguration {
 
     if (!author) {
       author = new Author(id);
+      this._known.add(author);
       this.addAlias(id, author);
     }
 

### 2016-06-23 www/js/index.ts

@@ -1,6 +1,7 @@
 ///<reference path="index.d.ts" />
 import { createTable, createTable2, makeCSVLink } from "./project-view";
 import DataView from './data-filter';
+import Author, {AuthorIdentity, AuthorConfiguration} from './author';
 
 /* Add date picker widgets on platforms that don't have them. */
 if (!hasDatePicker()) {
@@ -41,52 +42,89 @@ function escapeHTML(unsafe: string): string {
     .replace(/'/g, "&#039;");
 }
 
-/**
- * Renders the author panel.
- * @return {string} HTML for the authors panel.
- */
-function renderAuthorPanel(allAuthors: string[], authorUnsafe: string): string {
-  /* TODO: take in a data structure describing author configuration. */
-  let author = escapeHTML(authorUnsafe);
+type Selector = string | Node | JQuery;
 
-  return (`
-    <li>
-    <div class="panel panel-default author-configuration">
-      <div class="panel-heading"> ${author} </div>
-      <div class="panel-body">
-        <label><input type="checkbox" checked
-                      class="author-check-box"
-                      value="${author}">
-          Show this author
-        </label>
-        <label> Author aliases <br>
-          <select multiple class="author-aliases">
-            ${allAuthors.filter(name => name != author).map(unsafe =>
-                `<option value="${unsafe}">${escapeHTML(unsafe)}</option>`
-            )}
-          </select>
-        </label>
+class ManageAuthorsPanel {
+  protected $element: JQuery;
+  protected aliases: AuthorIdentity[];
+
+  constructor(protected initialConfig: AuthorConfiguration, element: Selector) {
+    this.$element = $(element);
+    this.aliases = Array.from(initialConfig.aliases).sort(byShorthand);
+
+    function byShorthand(a: AuthorIdentity, b: AuthorIdentity) {
+      return a.shorthand.toLowerCase() < b.shorthand.toLowerCase() ? -1 : 1;
+    }
+  }
+
+  get config(): AuthorConfiguration {
+    return this.initialConfig;
+  }
+
+  /**
+   * Renders the author panel.
+   * @return {string} HTML for the authors panel.
+   */
+  protected renderAuthorPanel = (id: AuthorIdentity): string => {
+    let authorName = escapeHTML(id.shorthand);
+    let author = this.config.get(id);
+    let isPrimary = id === author.primaryIdentity;
+    let constrolDisabled = !isPrimary;
+    let isEnabled = this.config.isEnabled(author);
+    
[…]

### 2016-06-23 www/js/author/configuration.ts
More UI and implementation.

@@ -1,12 +1,13 @@
 import Author, {AuthorIdentity} from './';
 
+import assert from '../assert';
+
 /**
  * Configuration for authors.
  */
 export default class AuthorConfiguration {
   private _enabled = new WeakSet<Author>();
   private _aliases = new Map<AuthorIdentity, Author>();
-  private _known = new Set<Author>();
 
   constructor({aliases, enabled}: ConfigurationJSON) {
     /* Add all aliases. */
@@ -22,7 +23,7 @@ export default class AuthorConfiguration {
   }
 
   get authors(): Set<Author> {
-    return new Set(this._known);
+    return new Set(this._aliases.values());
   }
 
   get aliases(): Iterable<AuthorIdentity> {
@@ -71,6 +72,21 @@ export default class AuthorConfiguration {
     this._aliases.set(alias, author);
   }
 
+  removeAlias(id: AuthorIdentity): void {
+    let numberAuthorsBefore = this.authors.size;
+    let originalAuthor = this.get(id);
+    this._aliases.delete(id);
+
+    let newAuthor = this._vivifyAuthor(id);
+    this.addAlias(id, newAuthor);
+
+    if (this.isEnabled(originalAuthor)) {
+      this.enable(newAuthor);
+    }
+
+    assert(this.authors.size === 1 + numberAuthorsBefore);
+  }
+
   isEnabled(author: Author): boolean {
     return this._enabled.has(author);
   }
@@ -91,7 +107,6 @@ export default class AuthorConfiguration {
 
     if (!author) {
       author = new Author(id);
-      this._known.add(author);
       this.addAlias(id, author);
     }
 

### 2016-06-23 www/js/index.ts

@@ -3,6 +3,8 @@ import { createTable, createTable2, makeCSVLink } from "./project-view";
 import DataView from './data-filter';
 import Author, {AuthorIdentity, AuthorConfiguration} from './author';
 
+import assert from './assert';
+
 /* Add date picker widgets on platforms that don't have them. */
 if (!hasDatePicker()) {
   $('input[type="date"]').datepicker();
@@ -69,24 +71,30 @@ class ManageAuthorsPanel {
     let authorName = escapeHTML(id.shorthand);
     let author = this.config.get(id);
     let isPrimary = id === author.primaryIdentity;
-    let constrolDisabled = !isPrimary;
+    let controlDisabled = !isPrimary;
     let isEnabled = this.config.isEnabled(author);
     
     /* Poor man's React: */
     return (`
       <li>
-      <div class="panel panel-default author-configuration ${constrolDisabled ? 'disabled' : ''}">
+      <div class="panel panel-default author-configuration ${controlDisabled ? 'disabled' : ''}">
         <div class="panel-heading"> ${authorName} </div>
         <div class="panel-body">
           <label><input type="checkbox"
                         ${isEnabled ? 'checked' : ''}
                         class="author-check-box"
-                        ${constrolDisabled ? 'disabled' : ''}
-                        value="${id.shorthand}">
+                        ${controlDisabled ? 'disabled' : ''}
+                        value="${escapeHTML(id.shorthand)}">
             Show this author
           </label>
           <label> Author aliases <br>
-            <select ${constrolDisabled ? 'disabled' : ''} multiple class="author-aliases">
+            <select
+              multiple
+              class="author-aliases"
+              data-placeholder="No duplicates for this author"
+              ${controlDisabled ? 'disabled' : ''}
+              name="${escapeHTML(author.id)}"
+            >
               ${this.makeOptions(author)}
             </select>
           </label>
@@ -99,15 +107,49 @@ class ManageAuthorsPanel {
   protected makeOptions(author: Author): string {
     return this.aliases
       .filter(alias => alias !== author.primaryIdentity)
-      .map(alias => (
-        `<option
-            ${this.config.get(alias) === author ? 'selected' : ''}
-            value="${alias.shorthand}"
+      .map(alias => {
+        let isForThisAuthor = this.config.get(alias) === author;
+        let isPrimaryForOther = this.config.isPrimaryIdentity(alias);
+
[…]

### 2016-06-23 www/js/index.ts
Move author panel stuff its own file.

@@ -1,9 +1,8 @@
 ///<reference path="index.d.ts" />
 import { createTable, createTable2, makeCSVLink } from "./project-view";
 import DataView from './data-filter';
-import Author, {AuthorIdentity, AuthorConfiguration} from './author';
+import ManageAuthorsPanel from './manage-author-panel';
 
-import assert from './assert';
 
 /* Add date picker widgets on platforms that don't have them. */
 if (!hasDatePicker()) {
@@ -33,134 +32,6 @@ function hasDatePicker() {
 
   return (input.value !== notADateValue);
 }
-
-/* From: http://stackoverflow.com/a/6234804 */
-function escapeHTML(unsafe: string): string {
-  return unsafe
-    .replace(/&/g, "&amp;")
-    .replace(/</g, "&lt;")
-    .replace(/>/g, "&gt;")
-    .replace(/"/g, "&quot;")
-    .replace(/'/g, "&#039;");
-}
-
-type Selector = string | Node | JQuery;
-
-class ManageAuthorsPanel {
-  protected $element: JQuery;
-  protected aliases: AuthorIdentity[];
-
-  constructor(protected initialConfig: AuthorConfiguration, element: Selector) {
-    this.$element = $(element);
-    this.aliases = Array.from(initialConfig.aliases).sort(byShorthand);
-
-    function byShorthand(a: AuthorIdentity, b: AuthorIdentity) {
-      return a.shorthand.toLowerCase() < b.shorthand.toLowerCase() ? -1 : 1;
-    }
-  }
-
-  get config(): AuthorConfiguration {
-    return this.initialConfig;
-  }
-
-  /**
-   * Renders the author panel.
-   * @return {string} HTML for the authors panel.
-   */
-  protected renderAuthorPanel = (id: AuthorIdentity): string => {
-    let authorName = escapeHTML(id.shorthand);
-    let author = this.config.get(id);
-    let isPrimary = id === author.primaryIdentity;
-    let controlDisabled = !isPrimary;
-    let isEnabled = this.config.isEnabled(author);
-    
-    /* Poor man's React: */
-    return (`
-      <li>
-      <div class="panel panel-default author-configuration ${controlDisabled ? 'disabled' : ''}">
-        <div class="panel-heading"> ${authorName} </div>
-        <div class="panel-body">
-          <label><input type="checkbox"
-                        ${isEnabled ? 'checked' : ''}
-                        class="author-check-box"
[…]

### 2016-06-23 www/js/manage-author-panel.ts

@@ -1,0 +1,133 @@
+import Author, {AuthorIdentity, AuthorConfiguration} from './author';
+
+import assert from './assert';
+
+type Selector = string | Node | JQuery;
+
+export default class ManageAuthorsPanel {
+  protected $element: JQuery;
+  protected aliases: AuthorIdentity[];
+
+  constructor(protected initialConfig: AuthorConfiguration, element: Selector) {
+    this.$element = $(element);
+    this.aliases = Array.from(initialConfig.aliases).sort(byShorthand);
+
+    function byShorthand(a: AuthorIdentity, b: AuthorIdentity) {
+      return a.shorthand.toLowerCase() < b.shorthand.toLowerCase() ? -1 : 1;
+    }
+  }
+
+  get config(): AuthorConfiguration {
+    return this.initialConfig;
+  }
+
+  /**
+   * Renders the author panel.
+   * @return {string} HTML for the authors panel.
+   */
+  protected renderAuthorPanel = (id: AuthorIdentity): string => {
+    let authorName = escapeHTML(id.shorthand);
+    let author = this.config.get(id);
+    let isPrimary = id === author.primaryIdentity;
+    let controlDisabled = !isPrimary;
+    let isEnabled = this.config.isEnabled(author);
+
+    /* Poor man's React: */
+    return (`
+      <li>
+      <div class="panel panel-default author-configuration ${controlDisabled ? 'disabled' : ''}">
+        <div class="panel-heading"> ${authorName} </div>
+        <div class="panel-body">
+          <label><input type="checkbox"
+                        ${isEnabled ? 'checked' : ''}
+                        class="author-check-box"
+                        ${controlDisabled ? 'disabled' : ''}
+                        value="${escapeHTML(id.shorthand)}">
+            Show this author
+          </label>
+          <label> Author aliases <br>
+            <select
+              multiple
+              class="author-aliases"
+              data-placeholder="No duplicates for this author"
+              ${controlDisabled ? 'disabled' : ''}
+              name="${escapeHTML(author.id)}"
+            >
+              ${this.makeOptions(author)}
+            </select>
+          </label>
+        </div>
+      </div>
+      </li>
+    `);
+  }
+
+  protected makeOptions(author: Author): string {
+    return this.aliases
+      .filter(alias => alias !== author.primaryIdentity)
+      .map(alias => {
[…]

### 2016-06-23 www/js/index.ts
A bit of refactoring.

@@ -33,23 +33,34 @@ function hasDatePicker() {
   return (input.value !== notADateValue);
 }
 
-function loadProject(type: 'Declarations' | 'Types' | 'Invocations') {
-  var xhttp = new XMLHttpRequest();
-  xhttp.onreadystatechange = function () {
-    if (xhttp.readyState == 4 && xhttp.status == 200) {
-      var filter = getFilters();
-      var project: Project = JSON.parse(xhttp.responseText);
-      var results = createTable(project, filter);
-      patchInputs(results);
+type EntityType = 'Declarations' | 'Types' | 'Invocations';
 
+function fetchProject(type: EntityType): Promise<Project> {
+  return new Promise(function (resolve, reject) {
+    let xhttp = new XMLHttpRequest();
+    xhttp.onreadystatechange = function () {
+      if (xhttp.readyState === 4 && xhttp.status == 200) {
+        let project: Project = JSON.parse(xhttp.responseText);
+        resolve(project);
+      } else if (xhttp.readyState === 4) {
+        reject(new Error(`bad http status: ${xhttp.status}`));
+      }
+    };
 
-      var link = makeCSVLink(results);
-      $('#csv-link').attr('href', link);
-    }
-  };
+    xhttp.open('GET', `get_project?type=${type}`, true);
+    xhttp.send();
+  });
+}
 
-  xhttp.open('GET', `get_project?type=${type}`, true);
-  xhttp.send();
+function loadProject(type: EntityType) {
+  fetchProject(type).then(project => {
+    let filter = getFilters();
+    let results = createTable(project, filter);
+    patchInputs(results);
+
+    var link = makeCSVLink(results);
+    $('#csv-link').attr('href', link);
+  });
 }
 
 window.redrawTable = function () {
@@ -124,10 +135,10 @@ window.toggleAuthors = function () {
   let $authors = $("#authors");
   $authors.toggleClass('collapse');
 
-  /* Chosen must be enabled while the panel is visible. */
   if (!$authors.hasClass('collapse')) {
-    new ManageAuthorsPanel((<any>window).CONFIG, $('#authors-list'))
-      .renderAuthorPicker();
+    let panel =
+      new ManageAuthorsPanel((<any>window).CONFIG, $('#authors-list'));
+    panel.renderAuthorPicker();
   }
 }
 

### 2016-06-23 www/js/manage-author-panel.ts

@@ -8,17 +8,13 @@ export default class ManageAuthorsPanel {
   protected $element: JQuery;
   protected aliases: AuthorIdentity[];
 
-  constructor(protected initialConfig: AuthorConfiguration, element: Selector) {
+  constructor(protected config: AuthorConfiguration, element: Selector) {
     this.$element = $(element);
-    this.aliases = Array.from(initialConfig.aliases).sort(byShorthand);
+    this.aliases = Array.from(config.aliases).sort(byShorthand);
 
     function byShorthand(a: AuthorIdentity, b: AuthorIdentity) {
       return a.shorthand.toLowerCase() < b.shorthand.toLowerCase() ? -1 : 1;
     }
-  }
-
-  get config(): AuthorConfiguration {
-    return this.initialConfig;
   }
 
   /**

### 2016-06-24 www/js/author/configuration.ts
Implement AuthorConfiguration#toPlainObject().

@@ -91,6 +91,23 @@ export default class AuthorConfiguration {
     return this._enabled.has(author);
   }
 
+  toPlainObject(): ConfigurationJSON {
+    let aliases: { [name: string]: string } = {};
+    let enabled: string[] = [];
+
+    for (let [alias, author] of this._aliases) {
+      aliases[alias.shorthand] = author.id;
+    }
+
+    for (let author of this.authors) {
+      if (this.isEnabled(author)) {
+        enabled.push(author.id);
+      }
+    }
+
+    return { aliases, enabled }
+  }
+
   /**
    * Maps an alias name to an author.
    */

### 2016-06-24 www/js/project-view.ts
Disable stats view.

@@ -97,6 +97,7 @@ export function createTable2(filter: Filter) {
       "Uma Maheswara Rao G <umamahesh@apache.org>"
     ]
   });
+  // TODO: get rid of window
   (<any>window).CONFIG = config;
 
   // TODO: get rid of window
@@ -104,7 +105,9 @@ export function createTable2(filter: Filter) {
   // TODO: get rid of window
   var data = window.filteredData = DataView.filter(processed, filter, config);
   drawGraph(data, dnaTable.offsetWidth);
-  drawStats(data, dnaTable.offsetWidth);
+
+  /* TODO: This view is not ready yet. */
+  //drawStats(data, dnaTable.offsetWidth);
 
   return data;
 }

### 2016-06-24 www/js/manage-author-panel.ts
Add Jaccard distance suggestions.

@@ -8,13 +8,33 @@ export default class ManageAuthorsPanel {
   protected $element: JQuery;
   protected aliases: AuthorIdentity[];
 
+  /**
+   *
+   */
+  protected similaritySet = new Map<AuthorIdentity, Set<string>>();
+
   constructor(protected config: AuthorConfiguration, element: Selector) {
     this.$element = $(element);
     this.aliases = Array.from(config.aliases).sort(byShorthand);
 
+    for (let id of this.aliases) {
+      this.similaritySet.set(
+        id,
+        new Set(id.name + id.email)
+      );
+    }
+
     function byShorthand(a: AuthorIdentity, b: AuthorIdentity) {
       return a.shorthand.toLowerCase() < b.shorthand.toLowerCase() ? -1 : 1;
     }
+  }
+
+  protected similarityTo(aId: AuthorIdentity) {
+    let a = this.similaritySet.get(aId);
+    return (bId: AuthorIdentity) => {
+      let b = this.similaritySet.get(bId);
+      return 1.0 - jaccard(a, b);
+    };
   }
 
   /**
@@ -59,8 +79,10 @@ export default class ManageAuthorsPanel {
   }
 
   protected makeOptions(author: Author): string {
-    return this.aliases
-      .filter(alias => alias !== author.primaryIdentity)
+    let options = this.aliases
+      .filter(alias => alias !== author.primaryIdentity);
+
+    return sortBy(options, this.similarityTo(author.primaryIdentity))
       .map(alias => {
         let isForThisAuthor = this.config.get(alias) === author;
         let isPrimaryForOther = this.config.isPrimaryIdentity(alias);
@@ -106,12 +128,9 @@ export default class ManageAuthorsPanel {
       })
   }
 
-  /**
-   * Rerenders the list inside #authors-list
-   * @return {[type]} [description]
-   */
-  renderAuthorPicker(): void {
+  renderAuthorPicker(): this {
     this.render();
+    return this;
   }
 }
 
@@ -126,4 +145,26 @@ function escapeHTML(unsafe: string): string {
 }
 
 
+function jaccard<T>(a: Set<T>, b: Set<T>): number {
+  let numElementsInCommon = intersectionMagnitude(a, b);
+  return numElementsInCommon / (a.size + b.size - numElementsInCommon);
+}
 
+function intersectionMagnitude<T>(a: Set<T>, b: Set<T>): number {
[…]

### 2016-06-24 www/js/manage-author-panel.ts
Minor changes.

@@ -146,11 +146,11 @@ function escapeHTML(unsafe: string): string {
 
 
 function jaccard<T>(a: Set<T>, b: Set<T>): number {
-  let numElementsInCommon = intersectionMagnitude(a, b);
-  return numElementsInCommon / (a.size + b.size - numElementsInCommon);
+  let numItemsInCommon = itemsInCommon(a, b);
+  return numItemsInCommon / (a.size + b.size - numItemsInCommon);
 }
 
-function intersectionMagnitude<T>(a: Set<T>, b: Set<T>): number {
+function itemsInCommon<T>(a: Set<T>, b: Set<T>): number {
   let count = 0;
   for (let item of a) {
     if (b.has(item)) {
@@ -163,7 +163,7 @@ function intersectionMagnitude<T>(a: Set<T>, b: Set<T>): number {
 
 function sortBy<T, U>(arr: T[], mapper: (item: T) => U): T[] {
   let paired = arr.map((item: T) => [mapper(item), item] as [U, T]);
-  paired.sort(([a, ], [b, _]) => {
+  paired.sort(([a,], [b,]) => {
     return a === b ? 0 : a < b ? -1 : 1;
   });
   return paired.map(([_, item]) => item);

### 2016-06-24 www/js/author/configuration.ts
Add AuthorConfiguration#generateDefaultConfiguration().

@@ -3,6 +3,13 @@ import Author, {AuthorIdentity} from './';
 import assert from '../assert';
 
 /**
+ * A map of name of alias to name of author.
+ */
+export interface AliasMap {
+  [name: string]: string;
+}
+
+/**
  * Configuration for authors.
  */
 export default class AuthorConfiguration {
@@ -109,7 +116,7 @@ export default class AuthorConfiguration {
   }
 
   /**
-   * Maps an alias name to an author.
+   * Maps an alias name to an author name.
    */
   protected _mapTo(aliasName: string, authorName: string) {
     let alias = AuthorIdentity.get(aliasName);
@@ -129,9 +136,28 @@ export default class AuthorConfiguration {
 
     return author;
   }
+
+  /**
+   * Generate a configuration from a list of authors.
+   * All authors are enabled, and every name maps to itself.
+   */
+  static generateDefaultConfiguration(names: string[]): AuthorConfiguration {
+    let aliases: AliasMap = {};
+
+    /* Map each name to itself. */
+    for (let name of names) {
+      aliases[name] = name;
+    }
+
+    return new AuthorConfiguration({
+      aliases,
+      /* Enable all names. */
+      enabled: names
+    });
+  }
 }
 
 interface ConfigurationJSON {
-  aliases: { [name: string]: string };
+  aliases: AliasMap;
   enabled: string[];
 }

### 2016-06-24 server.py
Generate author configuration from author JSON.

@@ -455,8 +455,7 @@ def dashboard():
 def show_project(name):
     for project in projects:
         if (project.getDir() == name):
-           return render_template('project_view.html', name=name,
-                                  authors=list(project.getAuthors()))
+           return render_template('project_view.html', name=name)
     return 'Project not found', 404
 
 

### 2016-06-24 www/js/index.d.ts

@@ -18,7 +18,6 @@ interface Window {
   DATA: Project;
   preprocessedData: any;
   filteredData: any;
-  authors: string[];
 
   redrawTable(): void;
   toggleView(): void;

### 2016-06-24 www/js/project-view.ts

@@ -40,8 +40,11 @@ export function createTable(data: Project, filter: Filter) {
   window.DATA = data;
   /* TODO: Remove global! */
   window.preprocessedData = preprocessData(data);
-  /* TODO: Remove global! */
-  window.authors = data.authors;
+
+  let config = AuthorConfiguration.generateDefaultConfiguration(data.authors);
+
+  // TODO: get rid of window
+  (<any>window).CONFIG = config;
 
   return createTable2(filter);
 }
@@ -56,54 +59,14 @@ export function createTable2(filter: Filter) {
   /* Clear previous table */
   dnaTable.innerHTML = "";
 
-
-  let config = new AuthorConfiguration({
-    aliases: {
-      "Andrey Yegorov <ayegorov@salesforce.com>":       "Andrey Yegorov <ayegorov@salesforce.com>",
-      "Arun M. Krishnakumar <akrishnakumar@salesforce.com>":       "Arun M. Krishnakumar <akrishnakumar@salesforce.com>",
-      "Benjamin Reed <breed@apache.org>":       "Benjamin Reed <breed@apache.org>",
-      "eolivelli <eolivelli@gmail.com>":       "eolivelli <eolivelli@gmail.com>",
-      "Flavio Paiva Junqueira <fpj@apache.org>":       "Flavio Paiva Junqueira <fpj@apache.org>",
-      "fpj <fpj@apache.org>":              "Flavio Paiva Junqueira <fpj@apache.org>",
-      "Ivan Brendan Kelly <ivank@apache.org>":       "Ivan Brendan Kelly <ivank@apache.org>",
-      "Ivan Kelly <ivan@bleurgh.com>":       "Ivan Brendan Kelly <ivank@apache.org>",
-      "Ivan Kelly <ivank@apache.org>":       "Ivan Brendan Kelly <ivank@apache.org>",
-      "Ivan Kelly <ivank@yahoo-inc.com>":       "Ivan Brendan Kelly <ivank@apache.org>",
-      "Matteo Merli <matteo.merli@gmail.com>":       "Matteo Merli <matteo.merli@gmail.com>",
-      "Matteo Merli <mmerli@apache.org>":       "Matteo Merli <matteo.merli@gmail.com>",
-      "Matteo Merli <mmerli@yahoo-inc.com>":       "Matteo Merli <matteo.merli@gmail.com>",
[…]

### 2016-06-24 www/js/index.ts
Redraw when author panel is closed.

@@ -136,9 +136,13 @@ window.toggleAuthors = function () {
   $authors.toggleClass('collapse');
 
   if (!$authors.hasClass('collapse')) {
+    /* Open the author mananagment panel. */
     let panel =
       new ManageAuthorsPanel((<any>window).CONFIG, $('#authors-list'));
     panel.renderAuthorPicker();
+  } else {
+    /* Close the author managment panel. This usually requires a redraw. */
+    window.redrawTable();
   }
 }
 

### 2016-06-24 www/js/data-filter.ts
Fix author number calculation.

@@ -88,8 +88,8 @@ export default class DataView {
   }
 
   /** The primary alias of every author selected in the view. */
-  get primaryAuthorAliases() {
-    return Object.keys(this.authorStats);
+  get primaryAuthorAliases(): number {
+    return this.authorStats.size;
   }
 
   get minDate() {

### 2016-06-24 www/js/project-view.ts
Fix the weird LaTeX row generator thing.

@@ -83,9 +83,13 @@ function drawGraph(data: DataView, width: number) {
   var height = cellHeight * data.types.length;
 
   /* XXX: Sneak in a little info about the project. */
-  d3.select('body').append('pre')
-    .text(summary(data))
-    .style('display', 'none');
+  let summaryText = d3.select('body').selectAll('.summary')
+      .data([summary(data)])
+      .text(d => d)
+    .enter().append('pre')
+      .classed('summary', true)
+      .text(d => d)
+      .style('display', 'none');
 
   /* Create a scale for the types i.e., the y-axis */
   var yScale = d3.scale.ordinal()
@@ -547,7 +551,7 @@ function ensureAxisIsAtGraphBottom(graph: SVGElement, axis: SVGElement) {
  * Returns LaTeX code for a table row summarizing the current project.
  */
 function summary({numberOfCommits, primaryAuthorAliases, numberOfFiles, numberOfTypes, astDiffs}: DataView) {
-  return `project & ${numberOfCommits} & ${primaryAuthorAliases.length} & ${numberOfFiles} & ${numberOfTypes} & ${astDiffs.length}`;
+  return `project & ${numberOfCommits} & ${primaryAuthorAliases} & ${numberOfFiles} & ${numberOfTypes} & ${astDiffs.length}`;
 }
 
 /**

### 2016-06-24 www/js/project-view.ts
Enable local storage for bookkeeper only!

@@ -20,6 +20,8 @@ import preprocessData, {PreprocessedData} from './preprocess-data';
 
 var CELL_INFO_WIDTH = 500;
 
+const TEMP_PROJECT = 'bookeeper';
+
 /* Cell Info */
 var cellInfo = d3.select("body")
   .append("div")
@@ -41,10 +43,8 @@ export function createTable(data: Project, filter: Filter) {
   /* TODO: Remove global! */
   window.preprocessedData = preprocessData(data);
 
-  let config = AuthorConfiguration.generateDefaultConfiguration(data.authors);
-
-  // TODO: get rid of window
-  (<any>window).CONFIG = config;
+  // TODO: Get rid of window.
+  let config = (<any>window).CONFIG = getAuthorConfiguration(TEMP_PROJECT, data.authors);
 
   return createTable2(filter);
 }
@@ -62,12 +62,11 @@ export function createTable2(filter: Filter) {
   // TODO: get rid of window
   var processed = window.preprocessedData;
   // TODO: get rid of window
-  var data = window.filteredData = DataView.filter(
-    processed,
-    filter,
-    (<any> window).CONFIG as AuthorConfiguration
-  );
+  let config = (<any> window).CONFIG as AuthorConfiguration;
+  var data = window.filteredData = DataView.filter(processed, filter, config);
+
   drawGraph(data, dnaTable.offsetWidth);
+  saveAuthorConfiguration(TEMP_PROJECT, config);
 
   /* TODO: This view is not ready yet. */
   //drawStats(data, dnaTable.offsetWidth);
@@ -571,4 +570,26 @@ function cellWidthFromScale(cell: Cell, scale: (_: Date) => number) {
   return Math.ceil(bigger - smaller) + 1;
 }
 
-/*globals d3*/
+/**
+ * Attempts to fetch configuration from LocalStorage, falling back on the
+ * provided list of author names.
+ */
+function getAuthorConfiguration(projectName: string, fallback: string[]): AuthorConfiguration {
+  const key = `${projectName}:authors`;
+
+  let configAsJSON = window.localStorage.getItem(key);
+
+  if (configAsJSON) {
+    let configObject = JSON.parse(configAsJSON);
+    /* TODO: Assert that all authors are accounted for.  */
+    return new AuthorConfiguration(configObject);
+  } else {
+    return AuthorConfiguration.generateDefaultConfiguration(fallback);
+  }
+}
+
[…]

### 2016-06-24 server.py
Add project name to a bunch of attributes.

@@ -47,6 +47,14 @@ class Project:
         self._dir = ""
         self._commits = []
 
+    @property
+    def name(self):
+        """
+        The project name. Note that this is currently synonymous with the
+        project directory.
+        """
+        return self.getDir()
+
     def setDir(self, name):
         self._dir = name
 
@@ -83,6 +91,7 @@ class Project:
         json_data = {}
         types = set()
 
+        json_data["name"] = self.name
         json_data["commits"] = list()
         json_data["dates"] = list()
         json_data["authors"] = list(self.getAuthors())

### 2016-06-24 www/js/json.d.ts

@@ -8,9 +8,27 @@
  * The JSON sent from the Flask server to the client.
  */
 declare interface Project {
+  /**
+   * The project's name.
+   */
+  name: string;
+  /**
+   * Array of all commits in the repository. Not guarenteed to be any order.
+   */
   commits: CommitFromJSON[];
+  /**
+   * Array of edits that occured in a project. This as a many-to-one
+   * relationship with commits.
+   */
   dates: Edit[];
+  /**
+   * Array of type names in the project.
+   */
   types: string[];
+  /**
+   * Array of authors in the project. Note that not all projects may have made
+   * edits.
+   */
   authors: string[];
 }
 

### 2016-06-24 www/js/preprocess-data.ts

@@ -5,6 +5,7 @@ import Commit, {CommitMap} from './commit';
 import assert from './assert';
 
 export interface PreprocessedData {
+  projectName: string;
   typeNames: Set<string>;
   commits: CommitMap;
   astDiffs: ASTDiff[];
@@ -34,6 +35,7 @@ export default function preprocessData(data: Project): PreprocessedData {
     astDiffs: createASTDiffsInAscendingOrder(data.dates, commits),
     authorNames: data.authors,
     commits,
+    projectName: data.name,
     typeNames
   };
 }

### 2016-06-24 www/js/project-view.ts

@@ -41,10 +41,11 @@ export function createTable(data: Project, filter: Filter) {
   /* TODO: Remove global! */
   window.DATA = data;
   /* TODO: Remove global! */
-  window.preprocessedData = preprocessData(data);
+  let preprocessed = window.preprocessedData = preprocessData(data);
 
   // TODO: Get rid of window.
-  let config = (<any>window).CONFIG = getAuthorConfiguration(TEMP_PROJECT, data.authors);
+  let config = (<any>window).CONFIG =
+    getAuthorConfiguration(preprocessed.projectName, data.authors);
 
   return createTable2(filter);
 }
@@ -60,13 +61,13 @@ export function createTable2(filter: Filter) {
   dnaTable.innerHTML = "";
 
   // TODO: get rid of window
-  var processed = window.preprocessedData;
+  var processed = window.preprocessedData as PreprocessedData;
   // TODO: get rid of window
   let config = (<any> window).CONFIG as AuthorConfiguration;
   var data = window.filteredData = DataView.filter(processed, filter, config);
 
   drawGraph(data, dnaTable.offsetWidth);
-  saveAuthorConfiguration(TEMP_PROJECT, config);
+  saveAuthorConfiguration(processed.projectName, config);
 
   /* TODO: This view is not ready yet. */
   //drawStats(data, dnaTable.offsetWidth);

### 2016-06-24 www/js/manage-author-panel.ts
Base similarity on author name and email without domain for better top-1 accuracy.

@@ -7,22 +7,13 @@ type Selector = string | Node | JQuery;
 export default class ManageAuthorsPanel {
   protected $element: JQuery;
   protected aliases: AuthorIdentity[];
-
-  /**
-   *
-   */
   protected similaritySet = new Map<AuthorIdentity, Set<string>>();
 
   constructor(protected config: AuthorConfiguration, element: Selector) {
     this.$element = $(element);
     this.aliases = Array.from(config.aliases).sort(byShorthand);
 
-    for (let id of this.aliases) {
-      this.similaritySet.set(
-        id,
-        new Set(id.name + id.email)
-      );
-    }
+    this.createSimilaritySets();
 
     function byShorthand(a: AuthorIdentity, b: AuthorIdentity) {
       return a.shorthand.toLowerCase() < b.shorthand.toLowerCase() ? -1 : 1;
@@ -35,6 +26,15 @@ export default class ManageAuthorsPanel {
       let b = this.similaritySet.get(bId);
       return 1.0 - jaccard(a, b);
     };
+  }
+
+  protected createSimilaritySets() {
+    assert(this.aliases instanceof Array);
+
+    for (let id of this.aliases) {
+      let characters = id.name + id.email.split('@')[0];
+      this.similaritySet.set(id, new Set(characters));
+    }
   }
 
   /**

### 2016-06-24 paper/document.tex
Update document text and add balance.

@@
+% Copyright (C) 2016 Michael D. Feist, Eddie Antonio Santos, Ian Watts, Abram Hindle
+% 
+% All TeX sources are licensed under Creative Commons Attribution-NoDerivatives 4.0 International (CC BY-ND 4.0) license.
+% See terms at <http://creativecommons.org/licenses/by-nd/4.0/>
+
 % === above this line ===
@@
 
-A variety of visualizations tools have been created in order to help with understanding the changes that get made to software over time.
-Wu \textit{et al.}~\cite{wu2004a} explored punctuated changes to software repositories with a tool that generates an Evolution Spectrograph.
+Several visualizations tools have been created to understand how a software system changes over time.
+Wu \textit{et al.}~\cite{wu2004a} explored punctuated changes of software repositories with a tool that generates an Evolution Spectrograph.
 In this visualization, each file has a row in the chart and each cell represents a section of time where an incoming or outgoing dependency can be changed.
@@
 Other tools also look at changes in code dependencies~\cite{6980224} or files~\cite{6650525, 7332421, voinea2006} over time.
-
-Tools also exist which track changes over time in a more complex manner.
-Kim \textit{et al.}~\cite{Kim:2009:DRS:1555001.1555046} built a tool called Logical Structural Diff (LSdiff) which examines code changes and produces logical rules to represent them.
-Changes which do not match the rules are then pointed out to the developer.
-Reiss and Tarvo created an IDE which allows users to see the history of files and changes to lines of code from each author as they are editing a project~\cite{6650521}.
-Van Hees and Hage use a voronoi treemap to visualize changes in class and method structure~\cite{7332410}.
-Servant and Jones created CHRONOS, a tool which creates a timeline of changes to a selection of code as well as the ability to query those changes~\cite{6650547}.
-
-Some tools examine the evolution of projects to learn other information about them.
 Alcocer \textit{et al.}~\cite{6650523} create a Performance Evolution Blueprint, which displays the changes in performance over time as a project evolves.
@@
 
-None of the aforementioned tools perform comparisons of ASTs or specifically examine the changes to type information for their view of project evolution.
+Tools also exist which track changes over time in a more complex manner.
+Kim \textit{et al.}~\cite{Kim:2009:DRS:1555001.1555046} built a tool called Logical Structural Diff (LSdiff) which examines code changes and produces logical rules to represent them.
+Changes which do not match the rules are then pointed out to the developer.
+Reiss and Tarvo created an IDE which allows users to see the history of files and changes to lines of code from each author as they are editing a project~\cite{6650521}.
+Van Hees and Hage use a voronoi treemap to visualize changes in class and method structure~\cite{7332410}.
+Servant and Jones created CHRONOS, a tool which creates a timeline of changes to a selection of code as well as the ability to query those changes~\cite{6650547}.
 
@@
 
-Developers using the massively popular GitHub website for hosting their Git repositories instantly have access to basic visualizations, known as GitHub Graphs (see Figure~\ref{fig:GH_code_freq})~\cite{github-graphs}.
+Developers using the massively popular GitHub website for hosting their Git repositories instantly have access to basic visualizations, known as GitHub Graphs (Figure~\ref{fig:GH_code_freq})~\cite{github-graphs}.
 GitHub Graphs tracks contributions based only on the number of commits and number of lines changed.
@@
 
-Given a Git repository, we want to extract per-author type changes over time.
+Given a Git repository, we want to extract per-author use of types over time.
 We accomplish this by computing the differences between abstract syntax trees of every consecutive pair of commits (Figure~\ref{fig:context}).
@@
 This captures the structure of the code in an abstract and tractable manner.
-When a programmer makes a non-trivial change in code in a file, that difference will be reflected in the AST\@.
+When a programmer makes a non-trivial change to a source code file, that difference will be reflected in the AST\@.
 Non-trivial changes are those that affect how a compiler will interpret the source code, and may have effects on the executable code.
@@
 To determine what has changed between two subsequent commits, we computed their \emph{AST difference}.
-Given two ASTs, an AST difference (or simply \emph{AST diff}) states which nodes must be added and which nodes must be removed to transform one tree to the next (step two in Figure~\ref{fig:context}).
+Given two ASTs, an AST difference (or simply as \emph{AST diff}) states which nodes must be added and which nodes must be removed to transform one tree to the next (step two in Figure~\ref{fig:context}).
 This allows us to precisely track changes between two consecutive revisions of the same file.
@@
 This states that two declarations of type \texttt{java.net.URI} were added, and two declarations of type \texttt{java.lang.String} were deleted.
-Note the fully qualified type names.
-The source text may not literally have the characters \texttt{java.net.URI}, but Spoon was able to infer from the import statements in the Java source text that references to ``URI'' resolve to the fully-qualified type name \texttt{java.net.URI}.
-Thus, TypeV is able to keep track of the actual class being used and is not confused when two classes in separate packages have the same name.
+Note that TypeV is able to find the fully qualified type names.
+Thus, the source text may not literally have the characters \texttt{java.net.URI}, but Spoon was able to infer from the import statements in the Java source text that references to ``URI'' resolve to the fully-qualified type name \texttt{java.net.URI}.
 
@@
 Therefore, we chose to ignore merge commits.
-Note that fast-forwards are not \emph{true merges}, hence they do not create merge commits~\cite{git:merge} so none of the previous cases involving merges apply.
 
@@
 
-Although rich in detail, the AST diff data is difficult to analyze if displayed in its raw form; hence, we created an HTML5 web visualization application which is intended for developers, project managers, and researchers to use (Figure~\ref{fig:main-view}).
+Although rich in detail, the AST diffs are difficult to analyze if displayed in their raw form; hence, we created an HTML5 web visualization application which is intended for developers, project managers, and researchers to use (Figure~\ref{fig:main-view}).
 The source code for TypeV---both the AST differencing component and the web application---is available online on GitHub.\footnote{\url{https://github.com/mdfeist/TypeV}} A screencast demonstrating its use is also available online.\footnote{\url{https://archive.org/details/typev-demo}}
@@
 We chose to make the height of the bars relative to only the type itself to make it easy for a user to scan a single type and see when it was modified.
-This makes tasks such as finding a refactoring (see Section~\ref{sec:antlr4}) very clear to see simply by browsing the line.
+This makes tasks such as finding a refactoring (Section~\ref{sec:antlr4}) very clear to see simply by browsing the line.
 We stacked the additions on top of the deletions such that users can immediately compare the relative frequency of changes of a single type throughout the chosen time span.
@@
 
-By adding interaction, the user is able to explore the timeline to find the specific information which they are interested in and drill-down at their leisure.
-Although our tool extracts a great deal of information, by using interaction, users can gradual reveal more information as they need it, rather than be bombarded with all of the information at once.
+By adding interaction, the user is able to explore the timeline to find the specific information that they are interested in and drill-down at their leisure.
+Although our tool extracts a great deal of information, users can gradual reveal more information as they need it, rather than be bombarded with all of the information at once.
 
 \section{Evaluation}
+\label{sec:eval}
 
@@
 \hline
-ANTLR4 & 3930 & 27 & 1793 & 2635 & 58425\\
-BookKeeper & 660 & 12 & 758 & 1872 & 31817\\
-Curator & 1524 & 43 & 1211 & 1269 & 27957\\
+ANTLR4 & 3930 & 19 & 1792 & 2635 & 58425\\
+BookKeeper & 660 & 10 & 758 & 1872 & 31817\\
+Curator & 1524 & 35 & 1211 & 1269 & 27957\\
 Tika & 860 & 10 & 604 & 757 & 20764\\
 \hline
-\multicolumn{6}{p{3.25in}}{\footnotesize Totals for the four projects we analyzed.
-``Changes'' refers to the total number of invocation or declaration changes observed from the AST diffs.
-Note that the number given for authors may be larger than the actual number of collaborators on the project, due to some collaborators using multiple email address, thus appearing as more than one author.}
+\multicolumn{6}{p{3.25in}}{\footnotesize Totals for the four projects we analyzed. ``Changes'' refers to the total number of invocation or declaration changes observed from the AST diffs.}
 \end{tabular}
@@
 
-Is it worth the effort of computing commit-by-commit AST differences if we can just simply count changes to files?
+\textbf{RQ2}: Is it worth the effort of computing commit-by-commit AST differences if we can simply count changes to files?
+We will answer this question by comparing type coverage and file coverage.
 We define \emph{type coverage} as the number of types that have been touched out of the total number of types during either a certain time period or the entire project.
@@
 
-To determine if type coverage yields different information than file coverage, we analyzed four open source Java projects: ANTLR4, Apache BookKeeper, Apache Curator and Apache Tika.
+To determine if type coverage yields different information than file coverage, we analyzed four open source Java projects: ANTLR4, Apache BookKeeper, Apache Curator and Apache Tika (Figure~\ref{tab:project-summary}).
 Three of the four projects were collected from the Apache foundation, as they host a large number of well-curated open source Java projects.
-A summary of the projects studied is given in Table~\ref{tab:project-summary}.
 We counted the file coverage and type coverage on a commit-by-commit basis for each author.
@@
 These plots show the distribution of the amount of commits that have X\% of type and file coverage.
-For example, Figure~\ref{fig:tika-density} shows that there are many commits with 50\% type coverage and very few commits with 50\% file coverage.
-This shows that even though the type and file coverage plots may have a similar shape the commits are showing different levels of information.
+For example, Figure~\ref{fig:tika-density} shows that there are many commits with 50\% type coverage and while there are relatively few commits with 50\% file coverage.
+This shows that even though the type and file coverage plots may have a similar shape, the commits are showing different levels of information.
 
@@
 
-We conclude from this that type coverage yields related, yet different information than file coverage alone.
+We conclude from these findings that type coverage yields related, yet different information than file coverage alone.
 
 \section{Illustrative Case Studies}
+\label{sec:cases}
 
@@
 \centering
-\includegraphics[width=\columnwidth]{ParserATNSimulator}
-\caption{Snapshot of TypeV showing changes to \texttt{ParserATNSimulator} in ANTLR4. Date ranges from approximately the beginning of 2011 to the end of April, 2012.}
-\label{fig:parser1}
-\end{figure}
-
-\begin{figure}[!ht]
-\centering
-\includegraphics[width=\columnwidth]{Antlr4-GH-Lines-ParserATNSimulator}
-\caption{GitHub's code frequency graph (based on line difference) showing changes to ANTLR4. The time spans approximately the same length as Figure~\ref{fig:parser1}.}
-\label{fig:parser2}
+\includegraphics[width=\columnwidth]{ParserATNSimulator-Antlr4-GH-Lines}
+\caption{Screenshots of TypeV (above) and GitHub's code frequency graph based on line difference (below).
+TypeV is showing changes to \texttt{ParserATNSimulator} in ANTLR4 while GitHub's graph is showing all line changes to ANTLR4.
+The date ranges from approximately April 2011 to April 2012.
+Note that both the TypeV and GitHub graphs are approximately lined up along the x-axis (time) and that we can see the loss of detail in the GitHub graphs.}
+\label{fig:parser}
 \end{figure}
@@
 
-For example, in December of 2011 we can see that the \texttt{ParserATNSimulator} class has a large amount of additions and deletions in roughly equal proportion (Figure~\ref{fig:parser1}).
+For example, in December of 2011, in Figure~\ref{fig:parser} we can see that the \texttt{ParserATNSimulator} class has a large amount of additions and deletions in roughly equal proportion.
 This suggests that \texttt{ParserATNSimulator} is not being removed from the project but instead there is some refactoring being done.
@@
 
-If we tried to do the same analysis with the GitHub tools it would be significantly more difficult to reach this conclusion (Figure~\ref{fig:parser2}).
+If we tried to do the same analysis with the GitHub tools it would be significantly more difficult to reach this conclusion (Figure~\ref{fig:parser}).
 First, if we only look at line changes this change to \texttt{ParserATNSimulator} is overwhelmed by other changes and hardly shows up on the GitHub tools.

### 2016-06-27 paper/document.tex
Last change before major rehaul of intro.

@@
 
-The major downside to using ASTs over a method like line count is that ASTs are more expensive to compute.
+The major downside of AST diffs over a method like line count is that ASTs are more expensive to compute.
 Our experiments were run on a Intel i7 with 16GB of RAM and an SSD.
@@
 For example ANTLR4 took about 8 hours while Apache BookKeeper only took a couple of hours.
-The good news for ongoing projects is that AST diffs only need to be computed for new commits; TypeV can resume calculating diffs and augmenting statistics at any time.
-New diffs can be appended to the already existing database of previous AST diffs.
+The good news for ongoing projects is that AST diffs only need to be computed for new commits; TypeV can resume calculating diffs and augment statistics at any time.
 
@@
 This may result in one author appearing as several different authors in the \texttt{git log}.
-There is no perfect way to deal with this problem; however, we allow the user of our software to select and view multiple authors at a time, thus allowing the user to view changes from an individual or group of developers.
+There is no perfect way to deal with this problem; however, we added an interface for managing authors.
+A user can manually specify an author's various aliases using a simple dropdown list.
+All statistics attributed to an author's various aliases are then remapped as belonging to one canonical author as specified by the user.
+
+In addition, our author managing view also provides a facility for selecting and deselecting authors in the view.
+Thus, visualization users can focus on the changes made by one author or a group of authors, rather than only having the option of view changes made by all authors cumulatively.
 
@@
 Although our tool extracts a great deal of information, users can gradual reveal more information as they need it, rather than be bombarded with all of the information at once.
+
+% TODO: Add design decisions for author merging?
 
@@
 
+In addition, the author statistics given in Table~\ref{tab:project-summary} was calculated after manual author merging.
+We are not personally familiar with the project, let alone the authors collaborating on the project.
+Therefore, our efforts to manually merge authors may have been erroneous.
+
 Some actions in Git present a problem gathering the statistics.

### 2016-06-27 paper/document.tex
New intro.

@@
 
-Source code is often represented as text in a source code file.
-This text is then parsed by a compiler or interpreter with the intent of eventually executing the code represented by the source text.
-While programmers may type in individual characters, their intention is at a much higher level.
-Their intent is written in tokens that represent statements, identifiers, and types that are used to solve their problems.
-Programmers solve problems in a way that is readable to other developers and executable by a computer.
-The programmer rarely operates at the character level, or even the line level, so why should our visualizations of source code operate at granularities unfamiliar to developers?
+Visualizations can assist our understanding of a software repository.
+The history of a code base is often tracked in a version control system such as Git or SVN.
+Popular source code repository hosting websites such as GitHub~\cite{github-graphs} and BitBucket~\cite{bitbucket-graphs} often provide a way to visualize project evolution.
+These visualizations often show changes to lines of code, as line changes are what the underlying version control system tracks.
+The advantage of visualizing line changes is that it works for any text-based file.
+There is a major disadvantage, however: by only visualizing changes to lines of code, GitHub (Figure~\ref{fig:GH_code_freq}) and similar visualizations lose the \emph{meaning} behind those changes.
+Although a programmer may type individual characters into their text editor, the lines of code represent something much deeper.
+A Java programmer, for instance, writes lines of code with the \emph{intent} of creating packages, defining classes, declaring fields, and implementing methods.
+An author commits lines of code to implement features, write tests, fix bugs, and refactor code bases.
+Lines of code merely tell us that changes were made; not the \emph{reason} for the changes.
 
-The plain text nature of software does not stop at the source code level.
-The history and evolution of a software project is often tracked by version control systems, such as Git, in a programming language agnostic way.
-Updates to source code, documentation, and tests are recorded in the form of commits, which store the differences in lines of text in a file between each subsequent revision of the software.
-Version control systems become vital for large projects since individual changes would be impossible to track manually.
-Large projects have thousands of individual changes made by many different developers.
-With more changes to peruse, it becomes more difficult to see how a project is evolving over time since there is so much information to look through.
-Hosting services for Git repositories often provide some basic tools for analyzing projects over time~\cite{github-graphs,bitbucket-graphs}; however, as we will show, these tools often lack the kind of detail to gather insightful analysis.
-As with source code, commits operate at the text level.
-While commits may store individual changes as lines of code, they imply a change in the \emph{structure} of software.
-By visualizing lines changed, or files changed, many current tools for software projects rarely capture the \emph{meaning} behind each commit; lines of code alone do not describe how a repository changed at a particular point in time.
-Thus, information regarding contributions from developers is often misleading since the number of commits and lines of code changed do not necessarily reflect meaningful changes to a code repository~\cite{robles2014}.
-In this paper, we will show that abstract syntax trees (ASTs) can be used to create \emph{semantically-aware} software evolution visualizations.
-
-Consider GitHub graphs~\cite{github-graphs}: visualizations that are easily available to software developers using GitHub to host their Git software repositories.
-GitHub's graphs use line and commit information exclusively.
-In the case of the ANTLR4 project (Figure~\ref{fig:GH_code_freq}), this code frequency chart creates more questions than it answers.
-What happened at those big spikes of line changes in the graph?
-One of the spikes is so large, we had to crop the image in order to show the magnitude of the other changes.
-How did these large changes affect the code base?
-Who committed these large changes?
-Was it a team of contributors or one particularly prolific programmer?
-Plotting the number of lines changed may be misleading.
-A huge spike in additions may be caused by a number of reasons, such as the inclusion of automatically-generated documentation, the inclusion of an external library, or perhaps a massive refactoring effort~\cite{Hindle:2008:LCT:1370750.1370773}.
-
-The goal of a software evolution visualization is to provide an alternate view that provides insight into the development of the software that is otherwise unavailable.
-By visualizing changes to lines and files, one is tracking the changes in text---not software.
-Instead, we promote \emph{semantically-aware visualizations} that visualize the structural changes in code.
-In our case these are changes in variable declarations, invocations of methods, and usage of types.
-Others have explored semantically-aware visualization~\cite{sourcevis,6980224,Kim:2009:DRS:1555001.1555046,7332410}, however, we show that this can be done with ASTs.
-
-In this paper, we propose a semantically-aware visualization method that tracks the changes in type usage and method invocations over time.
-Using commits, we are able to keep track of \emph{who} is using what types.
-We then present this data using time-proven visualizations.
-We claim that type usage and method invocation offers more insight than line-based, file-based, and commit-based methods of visualization and analysis.
-For simplicity, we analyze only Java projects, as it allowed us to quickly develop and test our software.
-However, since the approach described in this paper can be generally applied to any statically-typed language, it can be adapted to other programming languages.
-Our goal is to describe a methodology for obtaining richer, more meaningful data from source code repositories which can be used for more insightful analyses.
-
-In the following sections of the paper we will describe our tool, TypeV, that we developed to compute and visualize changes to types by author used in a project over time.
-TypeV is able to extract type information from a repository by computing commit-by-commit abstract syntax tree differences, which we describe in Section~\ref{sec:methodology}.
-In this paper, we answer the following questions: \\ \\
-\textbf{RQ1:} How can we extract type and method usage from repositories over time? \\ \\
+To address the lack of semantically-aware visualizations we created \textbf{TypeV}.
+Instead of visualization changes to lines of code, we visualize changes to type usages and invocations.
+By using types instead of lines of code, our visualization can better answer the high-level questions programmers often ask~\cite{sillito2006}.
+TypeV is composed of two parts: a data extraction application that performs commit-by-commit AST differencing to extract type changes; and an interactive web visualization that allows for the exploration of a project's evolution.
+Using TypeV, we seek to answer the following research questions:
+  \\ \\
+\textbf{RQ1:} How can we extract type and method usage from repositories over time?
+  \\ \\
 \textbf{RQ2:} How does type coverage compare to file coverage?
   \\ \\
-\textbf{RQ3:} Does the type information give any interesting or useful insight into the project?
+\textbf{RQ3:} What extra insight can we gain from type-based visualization?
 
@@
 
-One issue with analyzing individual authors is that authors can commit from a variety of different computers which can have different emails and usernames attached to their Git configuration.
+One issue with analyzing source code repositories is that authors can commit from a variety of different computers which can have different emails and usernames attached to their Git configuration.
 This may result in one author appearing as several different authors in the \texttt{git log}.
@@
 All statistics attributed to an author's various aliases are then remapped as belonging to one canonical author as specified by the user.
-
-In addition, our author managing view also provides a facility for selecting and deselecting authors in the view.
-Thus, visualization users can focus on the changes made by one author or a group of authors, rather than only having the option of view changes made by all authors cumulatively.
+The author managing view also provides a facility for selecting and deselecting authors shown in the timeline.
+Thus, visualization users can focus on the changes made by a particular author or a group of authors, rather than having to view the changes made by all authors.
 

### 2016-06-27 paper/document.tex
Add reference to RQs.
Remove reference of screencast.

@@
 
+\textbf{RQ1:} How can we extract type and method usage from repositories over time?
+
 Given a Git repository, we want to extract per-author use of types over time.
@@
 Although rich in detail, the AST diffs are difficult to analyze if displayed in their raw form; hence, we created an HTML5 web visualization application which is intended for developers, project managers, and researchers to use (Figure~\ref{fig:main-view}).
-The source code for TypeV---both the AST differencing component and the web application---is available online on GitHub.\footnote{\url{https://github.com/mdfeist/TypeV}} A screencast demonstrating its use is also available online.\footnote{\url{https://archive.org/details/typev-demo}}
+The source code for TypeV---both the AST differencing component and the web application---is available online on GitHub.\footnote{\url{https://github.com/mdfeist/TypeV}}
 
@@
 
-In the following sections, we present case studies demonstrating the extra insight that the AST differencing provides.
+\textbf{RQ3}: What extra insight can we gain from type-based visualization? We present case studies demonstrating the insight AST differencing can provide.
 

### 2016-06-27 paper/document.tex
Add author merging screenshot.

@@
 
-\begin{figure}[!h]
+\begin{figure}[t]
 \centering
@@
 
-\begin{figure}[h]
+\begin{figure}[b]
 \centering
@@
 
-\begin{figure}[h]
+\begin{figure}[tb]
 \centering
@@
 
-\begin{figure*}[!ht]
+\begin{figure*}[tb]
 \centering
@@
 
+\begin{figure}[tb]
+\centering
+\includegraphics[width=\columnwidth]{author-merging}
+\caption{Screenshot showing the author management view for Apache Bookkeeper.
+Users can use a drop-down list to select the aliases for an author, and select whether to show or hide the author in the timeline view.
+Authors in the drop-down list are sorted in decreasing order of Jaccard similarity.}
+\label{fig:author-merging}
+\end{figure}
+
 One issue with analyzing source code repositories is that authors can commit from a variety of different computers which can have different emails and usernames attached to their Git configuration.
 This may result in one author appearing as several different authors in the \texttt{git log}.
-There is no perfect way to deal with this problem; however, we added an interface for managing authors.
+There is no perfect way to deal with this problem; however, we added an interface for managing authors (Figure~\ref{fig:author-merging}.
 A user can manually specify an author's various aliases using a simple dropdown list.

### 2016-06-27 www/js/project-view.ts
Start to refactor timeline view into its own class.

@@ -77,283 +77,302 @@ export function createTable2(filter: Filter) {
 
 /*=== Graph ===*/
 
-function drawGraph(data: DataView, width: number) {
-  var marginLeft = 150;
-  var cellHeight = 64;
-  var height = cellHeight * data.types.length;
+const Timeline = class Timeline {
+  public marginLeft = 150;
+  public cellHeight = 64;
+  public height: number;
 
-  /* XXX: Sneak in a little info about the project. */
-  let summaryText = d3.select('body').selectAll('.summary')
-      .data([summary(data)])
-      .text(d => d)
-    .enter().append('pre')
-      .classed('summary', true)
-      .text(d => d)
-      .style('display', 'none');
-
-  /* Create a scale for the types i.e., the y-axis */
-  var yScale = d3.scale.ordinal()
-    .domain(data.types.map(type => type.fullyQualifiedName ))
-    .rangeBands([0, height]);
-
-  /* Create a scale for the dates i.e., the x-axis */
-  var xScale = d3.time.scale()
-    .domain([data.minDate, data.maxDate])
-    .range([marginLeft, width]);
-
-  var timeAxis = d3.svg.axis()
-    .scale(xScale)
-    .orient('bottom');
-
-  //var cellWidth = cellWidthFromScale(first(first(data.types).cells), xScale);
-  var maxCellHeight = yScale.rangeBand() - 2;
-
-  var svg = d3.select('#dna-table').append('svg')
-      .classed('main-figure', true)
-      .attr("width", width)
-      .attr("height", height);
-
-  var row = svg.selectAll('.row')
-      .data(data.types)
-    .enter().append('g')
-      .classed('row', true)
-      .attr('transform', (type) => `translate(0, ${yScale(type.fullyQualifiedName)})`);
-
-  /* The background. */
-  row.append('rect')
-      .attr('x', marginLeft)
-      .attr('width', width - marginLeft)
-      .attr('height', maxCellHeight)
-      .style('fill', function (_, i) {
-        /* Make alternating colour bands. */
-        return i & 1 ? '#f4f4f4' : '#fafafa';
-      });
-
-  /* Create a closure over data.commits. Function#bind() won't work
-   * since d3 wants to set up its own `this` to the callback! */
-  let createCell = function (type: JavaType) {
-    return createCellsForType.call(this, data.commits, type);
-  }
-  row.each(createCell);
-
-  row.append('text')
[…]

### 2016-07-04 paper/document.tex
Fix typos.

@@
 By adding interaction, the user is able to explore the timeline to find the specific information that they are interested in and drill-down at their leisure.
-Although our tool extracts a great deal of information, users can gradual reveal more information as they need it, rather than be bombarded with all of the information at once.
+Although our tool extracts a great deal of information, users can gradually reveal more information as they need it, rather than be bombarded with all of the information at once.
 
@@
 
-To determine if type coverage yields different information than file coverage, we analyzed four open source Java projects: ANTLR4, Apache BookKeeper, Apache Curator and Apache Tika (Figure~\ref{tab:project-summary}).
+To determine if type coverage yields different information than file coverage, we analyzed four open source Java projects: ANTLR4, Apache BookKeeper, Apache Curator and Apache Tika (Table~\ref{tab:project-summary}).
 Three of the four projects were collected from the Apache foundation, as they host a large number of well-curated open source Java projects.
@@
 In this paper, we introduced TypeV, a code repository analysis and visualization tool.
-Using commit-by-commit abstract syntax tree differencing enables insightful visualization that is not captured by tools that use commit counts, line counts, or file changes alone.
+Using commit-by-commit abstract syntax tree differencing enables insightful visualizations that are not captured by tools that use commit counts, line counts, or file changes alone.
 In this paper, we addressed the following research questions:

### 2016-07-06 paper/document.tex
Update paper with my TODOs.

@@
 
-% TODO: say all these other tools used file-based stuff:
-% gource, github, code_swarm, others...?
-
 \begin{table}[t]
@@
 \includegraphics[width=\columnwidth]{antlr-density}
-\caption{Density plot of type and file coverage for commits.
+\caption{R density plot of type and file coverage for commits.
 The graph shows what percentage of type and file coverage the majority of commits have.
@@
 \includegraphics[width=\columnwidth]{tika-density}
-\caption{Density of type and file coverage for commits.
+\caption{R density plot of type and file coverage for commits.
 Here we can see that commits did not have much file coverage between 25\% and 75\%.
@@
 \end{figure}
+
+% TODO: discuss density graphs more
 
@@
 
+% Discuss "well-curated"
+
 To determine if type coverage yields different information than file coverage, we analyzed four open source Java projects: ANTLR4, Apache BookKeeper, Apache Curator and Apache Tika (Table~\ref{tab:project-summary}).
-Three of the four projects were collected from the Apache foundation, as they host a large number of well-curated open source Java projects.
+We chose ANTLR4 because it is a large Java project with a rich history of commits.
+ANTLR4 is also a member of the Qualitas Corpus~\cite{qualitas}, a curated collection of Java systems curated for empirical software engineering research.
+We then collected three smaller projects from the Apache foundation, as they host a large number of well-curated open source Java projects.
+These particular Apache projects were chosen for their clear development practices and available documentation.
 We counted the file coverage and type coverage on a commit-by-commit basis for each author.
@@
 
-Having semantically aware evolutionary visualizations also gives users who are unfamiliar with the project at-a-glance information on the evolution of the project.
+% Discuss user study
 
@@
 Seeing how a project has changed over time is also vital for management since evaluating the state of completion of software can be very difficult.
+
+% Discuss viz design in comparison to code_swarm
+
+The timeline visualization presented in this paper (Section~\ref{sec:viz}) is intended to be a practical tool that enables a software developer's investigation of a project's evolution.
+Our timeline view can be seen as an evolution of simple visualizations such as GitHub's code frequency~\cite{github-graphs}.
+GitHub's graph is static, however it allows one to see an entire project's history at a glance.
+TypeV improves this by enabling one to see a particular \emph{type's} history at a glance.
+Given the greater density of information available, we made our timeline interactive such that a user is able to drill-down and explore the data at their own pace.
+Using interaction, TypeV not only shows what types have changed in project, but makes it possible to see \emph{why} a type has changed.
+We believe conventionality is a strong point in a developer's goal to make interesting discoveries in a project's evolution, but this has to be confirmed in user studies.
 
@@
 
+A weakness of the timeline view described in this paper is that details about an author's activities are buried.
+As future work, the data extracted by TypeV can be used to create a visualization similar to Gource~\cite{gource} or code\_swarm~\cite{codeswarm}.
+These visualizations animate authors scurrying back-and-forth over files they edit; we proposed to animate authors' activities over types.
+The focal point would be a particular type or group of types.
+The type itself may be represented abstractly as a shape or its UML representation that grows and shrinks as its usage increases and wanes over time.
+
 \section{Conclusion}
@@
 
+% Acknowledgments
+
+\section*{Acknowledgments}
+
+The authors are grateful to the reviewers for their helpful suggestions on improving this work.
+
+% TODO: acknowledge Craig?
+
 % === Below this line ===

### 2016-07-11 paper/document.tex
Almost there...

@@
 
-Several visualizations tools have been created to understand how a software system changes over time.
+Several visualization tools have been created to understand how a software system changes over time.
 Wu \textit{et al.}~\cite{wu2004a} explored punctuated changes of software repositories with a tool that generates an Evolution Spectrograph.
@@
 The colour is determined by the frequency of change.
-This work was continued to show file directories and display the cardinality of a developer's commits in Wu \textit{et al.}~\cite{wu2004}.
+This work was continued to show file directories and display the cardinality of a developer's commits in subsequent paper by Wu \textit{et al.}~\cite{wu2004}.
 The cardinality of a commit is the number of files or subsystems it affects in a system, indicating the scope of the change.
 Other tools also look at changes in code dependencies~\cite{6980224} or files~\cite{6650525, 7332421, voinea2006} over time.
-Alcocer \textit{et al.}~\cite{6650523} create a Performance Evolution Blueprint, which displays the changes in performance over time as a project evolves.
+Alcocer \textit{et al.}~\cite{6650523} created a Performance Evolution Blueprint, which displays the changes in performance over time as a project evolves.
 Ogawa and Ma created code\_swarm~\cite{codeswarm}, a generative art visualization which presents the progressive evolution of file changes made by authors in a source control repository over time.
@@
 Elsen also created a visualization for branches in Git, with the ability to expand branches to view the directory structure of a project at that point~\cite{6650522}.
-Shrestha \textit{et al.} made a visualization for the geographic location of a commit and~\cite{6650532} and Minelli \textit{et al.}~\cite{6980226} looked at the real time changes developers made to a repository.
+Shrestha \textit{et al.} made a visualization for the geographic location of a commit~\cite{6650532} and Minelli \textit{et al.}~\cite{6980226} looked at the real time changes developers made to a repository.
 
@@
 \item [declaration:]
-  Stating that use of a field or variable of a given type. For example declaration of a \texttt{String} type variable might look as follows:
+  Stating the use of a field or variable of a given type. For example declaration of a \texttt{String} type variable might look as follows:
   \\
@@
 The major downside of AST diffs over a method like line count is that ASTs are more expensive to compute.
-Our experiments were run on a Intel i7 with 16GB of RAM and an SSD.
+Our experiments were run on an Intel i7 with 16GB of RAM and an SSD.
 Calculating the AST diffs varied from an hour or two for smaller projects to a couple of days for very large projects.
@@
 This may result in one author appearing as several different authors in the \texttt{git log}.
-There is no perfect way to deal with this problem; however, we added an interface for managing authors (Figure~\ref{fig:author-merging}.
+There is no perfect way to deal with this problem; however, we added an interface for managing authors (Figure~\ref{fig:author-merging}).
 A user can manually specify an author's various aliases using a simple dropdown list.
@@
 We then collected three smaller projects from the Apache foundation, as they host a large number of well-curated open source Java projects.
-These particular Apache projects were chosen for their clear development practices and available documentation.
+These particular Apache projects were chosen for their clear development practices and easily available documentation.
 We counted the file coverage and type coverage on a commit-by-commit basis for each author.
@@
 We displayed the density distributions for two different projects: ANTLR4 (Figure~\ref{fig:antlr4-density}) and Apache Tika (Figure~\ref{fig:tika-density}).
-These plots show the distribution of the amount of commits that have X\% of type and file coverage.
-For example, Figure~\ref{fig:tika-density} shows that there are many commits with 50\% type coverage and while there are relatively few commits with 50\% file coverage.
-This shows that even though the type and file coverage plots may have a similar shape, the commits are showing different levels of information.
+These plots were produced using R and show the distribution of the amount of commits that have X\% of type and file coverage.
+The red plot represents file coverage and the blue plot represents type coverage.
+The x-axis shows the percent of the coverage and the y-axis shows the number of commits which have the similar percent of coverage.
+For example, Figure~\ref{fig:tika-density} shows that there are many commits with 50\% type coverage and relatively few commits with 50\% file coverage.
+Also notice how in Figure~\ref{fig:antlr4-density} the type coverage is pushed to the right, meaning commits usually touch a higher percentage of types than files.
+So if we only looked at files touched we might be missing important changes in the project.
+Figures~\ref{fig:antlr4-density} and ~\ref{fig:tika-density} show that even though the type and file coverage plots may have a similar shape, the commits are showing different levels of information.
 
@@
 
-% Discuss user study
-
 Having tools for analyzing repositories based on type usage and method invocation benefits many parts of the software development process.
@@
 
-The timeline visualization presented in this paper (Section~\ref{sec:viz}) is intended to be a practical tool that enables a software developer's investigation of a project's evolution.
+The timeline visualization presented in this paper (Section~\ref{sec:viz}) is intended to be a practical tool that enables users to investigate the evolution of a project.
 Our timeline view can be seen as an evolution of simple visualizations such as GitHub's code frequency~\cite{github-graphs}.
@@
 Given the greater density of information available, we made our timeline interactive such that a user is able to drill-down and explore the data at their own pace.
-Using interaction, TypeV not only shows what types have changed in project, but makes it possible to see \emph{why} a type has changed.
+Using interaction, TypeV not only shows what types have changed in a project, but makes it possible to see \emph{why} a type has changed.
 We believe conventionality is a strong point in a developer's goal to make interesting discoveries in a project's evolution, but this has to be confirmed in user studies.
 
+TypeV can also be useful for bug triaging.
 Bug triaging for large open source projects is a very time consuming task~\cite{badashian2015}, with hundreds  of bugs being reported daily.
@@
 \subsection{Future work}
+
+A user study is needed to test the usefulness of TypeV.
+Some questions that need to be answered are: Can users new to a project easily find when major changes took place and, more importantly, what was being changed?
+Given a specific time period, can users find what was being worked on?
+The study should test TypeV against a line-based visualization such as GitHub’s graphs~\cite{github-graphs} and a file-based visualization such as code\_swarm~\cite{codeswarm}.
 

### 2016-07-11 paper/document.tex
Cut down to 10 pages.

@@
 Visualizations can assist our understanding of a software repository.
-The history of a code base is often tracked in a version control system such as Git or SVN.
 Popular source code repository hosting websites such as GitHub~\cite{github-graphs} and BitBucket~\cite{bitbucket-graphs} often provide a way to visualize project evolution.
@@
 \caption{Hierarchy of data extracted by TypeV.
-Each repository may have zero or more commits; each commit may have zero or more changed files; each changed files may have zero or more changes, as computed by our AST differencing tool.}
+Each repository may have zero or more commits; each commit may have zero or more changed files; each changed file may have zero or more changes, as computed by our AST differencing tool.}
 \label{fig:data-extracted}
@@
 
-In Java, a type can be primitive, provided by the language itself, or an object reference.
-Primitive types are those that are built into the language such as an \texttt{int} or \texttt{char}.
-Object references point to instances of classes.
 Java's syntax encourages the creation of classes, thus we expect substantial Java projects to define many domain-specific types.
@@
 Internal threats to validity include our treatment of merge conflicts.
-As mentioned before, we ignore all merges, so if there is a merge conflict, it is possible that a developer makes non-trivial changes.
+As mentioned in section~\ref{sec:methodology}, we ignore all merges, so if there is a merge conflict, it is possible that a developer made non-trivial changes.
 Changes made during merge conflict resolution are missed by TypeV.
 We have made the assumptions that if this case happens, the changes are minor.
-
-In addition, the author statistics given in Table~\ref{tab:project-summary} was calculated after manual author merging.
-We are not personally familiar with the project, let alone the authors collaborating on the project.
-Therefore, our efforts to manually merge authors may have been erroneous.
 
@@
 Changes of this size are visible in the graph and can be found by inspecting the commit messages.
+
+In addition, the author statistics given in Table~\ref{tab:project-summary} was calculated after manual author merging.
+We are not personally familiar with the project, let alone the authors collaborating on the project.
+Therefore, our efforts to manually merge authors may have been erroneous.
 
@@
 As future work, the data extracted by TypeV can be used to create a visualization similar to Gource~\cite{gource} or code\_swarm~\cite{codeswarm}.
-These visualizations animate authors scurrying back-and-forth over files they edit; we proposed to animate authors' activities over types.
-The focal point would be a particular type or group of types.
+These visualizations animate authors moving back-and-forth over files they edit; we proposed to animate authors' activities over \emph{types}.
+The focal point would be a particular type.
 The type itself may be represented abstractly as a shape or its UML representation that grows and shrinks as its usage increases and wanes over time.
@@
 
-\section*{Acknowledgments}
+%\section*{Acknowledgments}
 
-The authors are grateful to the reviewers for their helpful suggestions on improving this work.
+%The authors are grateful to the reviewers for their helpful suggestions on improving this work.
 

### 2016-08-15 paper/document.tex
???

@@
 Alcocer \textit{et al.}~\cite{6650523} created a Performance Evolution Blueprint, which displays the changes in performance over time as a project evolves.
-Ogawa and Ma created code\_swarm~\cite{codeswarm}, a generative art visualization which presents the progressive evolution of file changes made by authors in a source control repository over time.
+Ogawa and Ma created \texttt{code\_swarm}~\cite{codeswarm}, a generative art visualization which presents the progressive evolution of file changes made by authors in a source control repository over time.
 Gource~\cite{gource} is a similar visualization, showing an animated force-directed tree of the file hierarchy as authors make changes to files in the source code repository over time.
@@
 Given a specific time period, can users find what was being worked on?
-The study should test TypeV against a line-based visualization such as GitHub’s graphs~\cite{github-graphs} and a file-based visualization such as code\_swarm~\cite{codeswarm}.
+The study should test TypeV against a line-based visualization such as GitHub’s graphs~\cite{github-graphs} and a file-based visualization such as \texttt{code\_swarm}~\cite{codeswarm}.
 
@@
 A weakness of the timeline view described in this paper is that details about an author's activities are buried.
-As future work, the data extracted by TypeV can be used to create a visualization similar to Gource~\cite{gource} or code\_swarm~\cite{codeswarm}.
+As future work, the data extracted by TypeV can be used to create a visualization similar to Gource~\cite{gource} or \texttt{code\_swarm}~\cite{codeswarm}.
 These visualizations animate authors moving back-and-forth over files they edit; we proposed to animate authors' activities over \emph{types}.
